### 2014-10-30 workshop-paper/paper.tex
Paper sources

  
\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}

\newcommand{\tuple}[1]{{\langle #1\rangle}}
\newcommand{\triple}[1]{{\langle #1\rangle}}
\newcommand{\pair}[1]{{\langle #1\rangle}}

\newcommand{\Omit}[1]{}

\newcommand{\eqdef}{\stackrel{\hbox{\tiny{def}}}{=}}
\newcommand{\IW}{{\textit{IW}}}
\newcommand{\SIW}{{\textit{SIW}}}
\newcommand{\ID}{{\textit{ID}}}
\newcommand{\BRFS}{{\textit{BrFS}}}

% AAAI (comment for ECAI)
% \newtheorem{theorem}{Theorem}

 \newtheorem{theorem}{Definition}
% \newtheorem{proposition}[theorem]{Proposition}
% \newtheorem{lemma}[theorem]{Lemma}
% \newtheorem{corollary}[theorem]{Corollary}
 \newtheorem{definition}[theorem]{Definition}

 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\pdfinfo{
/Title (Classical Planning Algorithms on the Atari Video Games: Preliminary Results)
/Author (Nir Lipovetzky, Miquel Ramirez, Hector Geffner)}

\setcounter{secnumdepth}{0}

\title{Classical Planning Algorithms ¡ on the Atari Video Games: Preliminary Results}

\author{Nir, Miquel, H}
%AAAI Press\\
%Association for the Advancement of Artificial Intelligence\\
%2275 East Bayshore Road, Suite 160\\
% Palo Alto, California 94303\\}

\maketitle

%% Pending: randomization, discount factor, iw11 ..
%% iw2 multivalued expands up to 256^2 n^nodes, while iw2 boolean 16^2 n^2 nodes; can iw2 boolean beat iw1 multivalued?
%% Important. optimization, need Dikjstra por postprocessing .. see emails

\begin{abstract}
\begin{quote}
The Atari  2600 games supported in  the Arcade Learning Environment (Bellemare et. al. 2013) 
all feature a known initial (RAM) state and actions that have deterministic effects. Classical planners, however, 
cannot be used  for selecting actions  for  two reasons: first,  no  compact PDDL-model of  the games is given, 
and  more importantly, the actions effects and the goals are  not known a priori. Moreover, in these games there is usually  
no set of goals to be achieved but rewards to be collected. These features do not preclude the use of   classical  algorithms like
breadth-first search or Dijkstra's algorithm,  but    these  algorithms are not effective over  large state spaces. We thus turn to a
different class of classical planning algorithms  introduced recently  that perform a \emph{structured exploration} of the state space; namely,  
like breadth-first search and Dijkstra's  algorithm they are ``blind'' and hence  do not require prior knowledge of state transitions, 
costs (rewards) or goals, and yet, like  heuristic search algorithms, they have been shown to be   effective for solving   problems over huge 
state spaces. The simplest such algorithm, called Iterated Width or IW, consists of  a sequence of calls  IW(1), IW(2), .., IW(k) where  IW(i) is  
a breadth-first search in which a state is  pruned when it is not the first state in the search  to make true some set of i atoms or less.
Preliminary empirical results suggest that the performance of IW with the  $k$ parameter fixed to $1$, i.e., IW(1), 
is much better than random or breadth-first search, and close to the performance of UCT. Other simple variations of IW
that combine exploration and exploitation are evaluated as well.
\end{quote}
\end{abstract}

\section{Introduction}

The Arcade Learning Environment (ALE)  provides a challenging  platform for evaluating general, domain-independent AI planners and 
learners through a convenient interface to hundreds of Atari 2600 games \cite{bellemare:jair2013}. 
Results have been reported so far for basic planning algorithms like breadth-first 
search and UCT \cite{bellemare:jair2013}, and for  reinforcement learning
and evolutionary  methods \cite{bellemare:jair2013,deep-mind-atari,risto-texas:atari-games}.
The empirical results are impressive in some cases, yet a lot remains to be done, as no method  approaches   the 
performance of  human players across a broad range of games.

While all  these games feature a known initial (RAM) state and actions that
have deterministic effects on this state, the problem of selecting the next action
to be done cannot be addressed with state-of-the-art classical planners for two 
reasons. First, there is no compact PDDL-like encoding of the domain, and more 
importantly,  the goal to be achieved in each game  is not given. Indeed, there are 
often no goals at all but rewards $r(a,s)$ that potentially depend on the action $a$ and 
the state $s$ in  which  the action is performed, and these rewards are not known a priori either. 
Thus no model of the  goals  or the  rewards can be used to bias the search, and hence, 
standard heuristic search and classical planning methods can't be used.

The action selection problem in the Atari games can be seen as a \emph{reinforcement learning} problem 
\cite{sutton:book} over a deterministic MDP where the state transitions and rewards are not known, or alternatively,
as a  \emph{net-benefit planning problem} \cite{ref-net-benefit-planning} with unknown state transitions  and  rewards. Namely, 
we seek  an action sequence $a_0, a_1, \ldots, a_m$ from the given current  state $s_0$ that generates 
a state sequence $s_0, s_1, \ldots, s_{m+1}$ with  maximum  total reward $\sum_{i=0}^{m} r(a_i,s_i)$,  
where $m$ is a sufficiently large planning horizon, and the rewards $r(a_i,s_i)$ and the  state-transitions $s_{i+1}=f(a_i,s_i)$ 
are known only after the action $a_i$ is applied in the state $s_i$.

The presence of unknown transition and  rewards in  the Atari games does not preclude the use of blind-search
methods like breadth-first search, Dijkstra's algorithm \cite{dijkstra},  or learning methods  such as 
LRTA* \cite{korf:lrta}, UCT \cite{uct}, and Q-learning  \cite{sutton:rl,bertsekas:neuro}. Indeed, 
the net-benefit planning problem with unknown state transitions and rewards over a given planning
horizon, can be mapped into a standard \emph{shortest-path problem} which can be solved optimally by Dijkstra's algorithm.
For this, we just need to  map the unknown rewards $r(a,s)$ into positive (unknown) action costs $c(a,s) = C - r(a,s)$
where  $C$ is a large constant that exceeds the maximum possible reward. The fact that the state 
transition and cost functions  $f(a,s)$ and $c(a,s)$ are not known a priori doesn't affect the applicability of 
Dijkstra's algorithm, which requires the value of these functions precisely when  the action $a$
is applied in the state $s$.

The limitation of the   basic blind search  methods is that they are not effective  over very spaces, neither
for solving problems off-line, nor for guiding  a lookahead search for solving problems on-line. 
In this work, we thus turn to a recent class of  planning algorithms that combine the scope of blind search
methods with the performance of state-of-the-art classical planning algorithms: namely, like ``blind'' search algorithms
they do not require prior knowledge of state transitions, costs, or goals, and yet like heuristic algorithms
they manage to search large  state spaces effectively. The basic algorithm in this class is 
called IW for Iterated Width search \cite{nir:ecai2012}. IW consists of a sequence of calls  IW(1), IW(2), .., IW(k), where 
IW(i) is  a  standard breadth-first search where states are pruned right away when they fail to make true some new tuple (set) 
of at most $i$ atoms. Namely,   IW(1) is a breadth-first search that keeps  a state only if  the state is   the first  one 
in the search to make some atom  true; IW(2) keeps a state  only if the state is  the first one to  make  a pair of atoms true, and 
so on.  Like plain breadth-first  and iterative deepening searches,  IW  is 
complete, while searching the state space in a way that makes use of the 
\emph{structure of states} given by the values of a finite set of state variables.
In the Atari games, the (RAM) state is given by a vector of 128 bytes, which we associate with 128 variables 
$X_i$, $i=1, \ldots, 128$, each of which may take up to $256$ values $x_j$. A state $s$ makes an atom $X_i=x_{j}$ 
true when the value of the $i$-th byte in the state vector $s$ is $x_j$.

While a normal, complete breadth-first search runs in time that is exponential in the total number of variables $X_i$, 
the procedure $IW(k)$ runs in time that is exponential in the $k$ parameter, $1 \leq k \leq n$. 
Elsewhere, it has been shown that  IW exhibits   state-of-the-art performance on most existing planning benchmark domains 
when goals are restricted to be single atoms  where IW provably runs in low polynomial time in the number of
variables \cite{nir:ecai2012}. For example, in a a block world instance with any number of blocks and any initial
configuration,  the goal of  placing a given  block on top of another, will be achievable in  quadratic time  
in the number of blocks as any such instance can be shown to  have \emph{width} 2. This means  that IW(2) will be 
not only complete then, but also optimal. This is also true for most other benchmark domains in classical planning
that turn out to be have small, bounded widths for any instance, as long as the goal is atomic. 
For dealing with the benchmarks where goals are  not single atoms, e.g., where a given \emph{tower} of blocks needs to be built, 
simple extensions of the basic IW procedure have been developed as well, such as Serialized IW, where IW is 
used to achieve the goals one at at time \cite{nir:ecai2012,nir:ecai2014}. The extension of IW for problems where there is no goal but an
additive reward measure   to be optimized is direct.

%%  Each iteration IW(k) keeps track of the accumulated reward up to each state $s$ that has been
%% generated but  not pruned, as $R(s) = R(s')+ r(a,s')$ where $s'$ and $a$ are  the single parent state
%% and action $a$ that led to $s$ in IW(k) with $R(s)=0$ for the current state $s$. The best reward

The sequence of calls IW(1), IW(2), \ldots, IW(k) manages to combine the scope of ``blind'' search algorithms with the performance of 
heuristic-search algorithms by exploring states  in an  order dictated by the ``width''  parameter $k$; 
states that have width 1 are explored first in linear time, then states that have  width two are explored in quadratic time, and so on. 
In classical planning problems, if the goal has a small width (something common in the classical planning domains when the  goal is a single atom, as
we have seen above), IW will find goal states  in low polynomial time. Similarly, in the reward-based setting, if  there are states of high reward 
that have small width, IW will find such states in low polynomial time, even if the state space is exponential in the number of variables. 
In this paper, we aim to test whether these ideas work for the Atari games too  when  the RAM state of the emulator
is used  as the system state.

The paper is organized as follows. We review the Iterated Width  algorithm and where it comes from, 
look at  some simple   variations of the basic algorithm that appear to be convenient  for the Atari
games, present and analyze the preliminary experimental results, summarize the main lessons, and
discuss related and future work.

\section{Iterated Width}

The Iterative Width  (IW) algorithm has been introduced as a classical planning algorithm
that takes a planning problem as an input, and computes an action sequence that solves the problem 
as the output \cite{nir:ecai2012}. The algorithm however
applies to a broader range of problems. We will characterize such
problems  by means  a finite and discrete set of states (the state space)
that correspond to vectors of size $n$. Namely, the 
states are  \emph{structured} or \emph{factored} , and we take each of the locations in the vector
to represent a variable $X_i$, and the value at that vector location to 
represent the value $x_j$ of variable $X_i$ in the state. 
In addition to the states space, a problem is defined by an initial state $s_0$, 
a set of actions $A(s)$ applicable in each state, a 
transition function $f$ such that  $s' = f(a,s)$ is 
the state that results from applying action $a \in A(s)$ to the state $s$, 
and rewards $r(a,s)$ represented by real numbers that result from  applying action $a$ in state $s$.
The transition and reward functions do not  need to be known a priori, yet in that case, the  
state and  reward that results from the application of an action in a state needs to be \emph{observable}.
The task is to compute an action sequence $a_0, \ldots, a_m$ for a large horizon $m$
that generates a state sequence $s_0, \ldots, s_{m+1}$ that  maximizes the accumulated 
reward $\sum_{i=0}^{m} r(a_i,s_i)$, or that provides a good approximation.

\subsection{The Algorithm}

IW consists of a sequence of calls  $\IW(i)$ for $i=0, 1, 2, \ldots$
over a problem $P$ until a termination condition is reached. 
The procedure $\IW(i)$  is a plain forward-state \emph{breadth-first search} with just one change: right after a state 
$s$ is generated, the state is  pruned if it doesn't pass a simple \emph{novelty test}.
More precisely,

\begin{itemize}
\item The \emph{novelty of a newly generate state $s$  in a  breadth-first search}
is   $1$ when $s$ is the first state generated in the search that makes true an atom $X=x$,
$2$  when $s$ is the first state that makes a pair of atoms $X=x$ and $Y=y$ true, and so on.

\item  $\IW(i)$ is a breadth-first search that prunes newly generated states when their novelty measure
is greater than $i$.

\item $\IW$  calls $\IW(i)$ sequentially for $i=1,2,\ldots$ until a termination condition is reached, 
returning then the best path found.
\end{itemize}

For classical planning, the termination condition is the achievement of the goal. In the on-line setting, 
as in the Atari games, the termination condition is given by a time window or a maximum number of generated nodes.
The  best path found by IW is the path  that has a  maximum accumulated reward. 
The accummulated reward $R(s)$  of a state $s$ reached in an iteration of IW is determined by the  unique parent state $s'$ 
and action $a$ leading to $s$ from $s'$ as $R(s) = R(s') + r(a,s')$. The best state is the state $s$ with maximum reward
$R(s)$ generated but not pruned by IW, and the best path is the one that leads to the state $s$ from the current state.
The action selected to be done next is then the first action along such a path  (** see eventually randomization **).

%% Like in standard breadth-first search, each iteration IW(i)  prunes 
%%% the  duplicate states as well.

\begin{table}
\begin{center}
%\resizebox{3in}{!}{
{\scriptsize
% \begin{tabular}{r@{}lrrrr@{}}
\begin{tabular}{llcccc}
%% table header
\hline\\[-1.3ex]
%\toprule
\# & \multicolumn{1}{c}{Domain} & \multicolumn{1}{c}{I} & \multicolumn{1}{c}{IW(1)} & \multicolumn{1}{c}{IW(2)} & \multicolumn{1}{c}{Neither} \\[.2ex]

\hline\\[-1.3ex]
%\hline
%% table body
1. & 8puzzle	&400	&55\%	&45\%	&0\%\\
2. & Barman	&232	&9\%	&0\%	&91\%\\
3. & Blocks World	&598	&26\%	&74\%	&0\%\\
4. &Cybersecure	&86	&65\%	&0\%	&35\%\\
% 5. &Depots	&189	&11\%	&66\%	&23\%\\
% 6. & Driver	&259	&45\%	&55\%	&0\%\\
% 7. & Elevators	&510	&0\%	&100\%	&0\%\\
% 8. & Ferry	&650	&36\%	&64\%	&0\%\\
% 9. & Floortile	&538	&96\%	&4\%	&0\%\\
% 10. & Freecell	&76	&8\%	&92\%	&0\%\\
% 11. & Grid	&19	&5\%	&84\%	&11\%\\
% 12. & Gripper	&1275	&0\%	&100\%	&0\%\\
% 13. & Logistics	&249	&18\%	&82\%	&0\%\\
% 14. & Miconic	&650	&0\%	&100\%	&0\%\\
% 15. & Mprime	&43	&5\%	&95\%	&0\%\\
% 16. & Mystery	&30	&7\%	&93\%	&0\%\\
% 17. & NoMystery	&210	&0\%	&100\%	&0\%\\
% 18. & OpenStacks	&630	&0\%	&0\%	&100\%\\
% 19. & OpenStacksIPC6	&1230	&5\%	&16\%	&79\%\\
\ldots & \ldots & \ldots & \ldots & \ldots & \\ 
20. & ParcPrinter	&975	&85\%	&15\%	&0\%\\
21. & Parking	&540	&77\%	&23\%	&0\%\\
22. & Pegsol	&964	&92\%	&8\%	&0\%\\
23. & Pipes-NonTan	&259	&44\%	&56\%	&0\%\\
24. & Pipes-Tan	&369	&59\%	&37\%	&3\%\\
25. & PSRsmall	&316	&92\%	&0\%	&8\%\\
26. & Rovers	&488	&47\%	&53\%	&0\%\\
27. & Satellite	&308	&11\%	&89\%	&0\%\\
28. & Scanalyzer	&624	&100\%	&0\%	&0\%\\
29. & Sokoban	&153	&37\%	&36\%	&27\%\\
30. & Storage	&240	&100\%	&0\%	&0\%\\
31. & Tidybot	&84	&12\%	&39\%	&49\%\\
32. & Tpp	&315	&0\%	&92\%	&8\%\\
33. & Transport	&330	&0\%	&100\%	&0\%\\
34. & Trucks	&345	&0\%	&100\%	&0\%\\
35. & Visitall	&21859	&100\%	&0\%	&0\%\\
36. & Woodworking	&1659	&100\%	&0\%	&0\%\\
37. & Zeno	&219	&21\%	&79\%	&0\%\\[.2ex]
%\hline
\hline\\[-1.3ex]
& Total/Avgs&37921	&37.0\%	&51.3\%	&11.7\%\\
\hline
\end{tabular}
}

\vskip .1cm

{\scriptsize
\begin{tabular}{@{}ccccc@{}}
\hline\\[-1.3ex]
\multicolumn{1}{c}{\# Instances}  &\multicolumn{1}{c}{\IW} & \multicolumn{1}{c}{\ID} & \multicolumn{1}{c}{\BRFS} &\multicolumn{1}{c}{\textit{GBFS +} $h_{add}$} \\[.2ex]
\hline\\[-1.3ex]
37921  &34627 &9010	&8762 &34849 \\
\hline
%\hline
\end{tabular}
}

\end{center}

\caption{\small 
\emph{Top:} Number of classical planning problems  and percentages solved by IW(1), IW(2), or neither. Instances obtained
by splitting benchmarks  with $N$ atomic goals into $N$ problems with atomic goals.
\emph{Bottom:} Number of instances solved by \IW\ in comparison with  Iterative Deepening (\ID),  
Breadth-First Search (\BRFS),   and Greedy Best First Search (\textit{GBFS}) guided by  additive heuristic.
Time and  memory outs  after 30 minutes and 2 GB resp. From \cite{nir:ecai2102}.
}
\label{table1}
\end{table}
%%%%
%%%

\subsection{Performance and Width}

IW is a systematic and complete blind-search algorithm like breadth-first search (BrFS) and iterative deepening (ID),
but   unlike these algorithms, it uses the factored representation of the states in terms of variables to structure the 
search in a different way. This structured exploration  has proved to be very effective in classical planning 
over most of the benchmark domains  when the goals are restricted to be single atoms. Table~\ref{table1} from \cite{nir:ecai2012} shows 
the percentage of instances that are  solved by  the first iteration of IW, i.e.  IW(1), by the second iteration 
IW(2), and by neither one.  These are instances that  have been obtained from the existing benchmarks
by splitting problems with $N$ atomic goals, into $N$ problems with one atomic goal each. As the table shows, 
$37\%$ of the 37921 instances are solved by IW(1) while $51.3\%$ are  solved by IW(2). Since IW($k$)
runs in time that is exponential in $k$, this mean that almost $90\%$ of the 37921 instances are  solved in time that
is either linear or quadratic in the number of problem variables, which in these encodings are all \emph{boolean}.  
Furthermore, when the performance of \IW\   is compared with  breadth-first search and iterative deepening, 
on the one hand,  and  with a  Greedy Best First Search guided by the additive heuritstic $h_{add}$
on the other (this algorithm is similar to one used by  the FF planner when hill climbing search fails \cite{hoffmann:ff}), 
it turns out that ``blind'' \IW\ solves as many problems as the informed search, 34627 vs. 34849, far ahead
of the other blind algorithms BrFS and ID  that solve 9010 and 8762 problems each. This is shown in the bottom
part of the Table~\ref{table1}. Moreover, \IW\ is  faster and results in shorter plans than the heuristic search  
\cite{nir:ecai2012}.

The min $k$ value for which $IW(k)$ solves an instance  is clearly bounded and small in most of
these instances. This is  actually no  accident and  has a \emph{theoretical explanation.} Lipovetzky and Geffner define
a structural parameter called the problem \emph{width} and show that for many of these domains,
\emph{any} solvable instance with atomic goals will  have a bounded and small width that is independent of the number
of variables and states in the problem. The min value $k$ for which the iteration  $IW(k)$ solves
the problem cannot exceed the problem width, so that the algorithm \emph{IW will run in time and space
that are exponential in the problem width.}

Formally, the \emph{width} $w(P)$ of a problem $P$ is $i$
iff $i$ is the minimum positive integer for which 
there is a sequence of tuples (sets) with at most $i$ atoms each, $t_0, t_1, \ldots, t_n$
such that 1)~$t_0$ is true in the initial state of $P$, 2)~any shortest plan $\pi$  that achieves $t_i$ in $P$
can be extended into a shortest plan that achieves $t_{i+1}$ by extending   $\pi$ with one action, and 
3)~any shortest plan that achieves $t_n$ is also a shortest plan for $P$.

One way to understand this definition is that the problem  width $w(P)$ is at most  $i$ if  there is a 
``trail of  stepping stones''  $t_0, t_1, \ldots, t_n$ to reach the problem goal such that 
A)~these ``stepping stones''  contain at most $i$ atoms each, B)~each  stepping stone $t_{i+1}$ is 
at distance 1 from the previous one $t_i$ when $t_i$ has been reached following the ``trail'', 
C)~the ``trail''  preserves ``optimality''; i.e., no  tuple $t_i$ can be reached in less than $i$ steps.

While this notion of width and the iterated  width algorithms that are based on it  have been designed
for problems where a goal state needs to be reached, the notions remain  relevant  in  optimization problems as well.
Indeed, if a  good path is made of states $s_i$ each of which have a low width, we would like 
IW to find such path in low polynomial time for a low value of the $k$ parameter. 
Later on we will discuss the changes required in IW to enforce this property, which
are not yet implemented in the code that is evaluated below.

\section{Width-Based Algorithms for the Atari Games}

Explain: IW(1), 2BFS, and possible IW11. Also, Dijkstra postprocessing over IW(1) graph
if ran.

The number of nodes \emph{generated} by IW(1) is $n \times D \times b$ in the worst case, 
where $n$ is the number of variables, $D$ is the size of their domains, and $b$ is maximum the number of 
actions applicable per  state.  This same number in breadth-first search is not linear in $n$
but exponential. For the Atari games,  $n=128$, $D=256$, and $b=18$, so that the product is
equal to $589,824$, which is large but feasible. On the other hand, the number
of nodes generated by IW(2) in the worst case is  $(n \times D)^2 \times b$, which is
equal to $1,073,741,824$ which is too large, forcing us to consider only a fraction of 
such states. For classical planning problems, the growth in the number of nodes from 
IW(1) to IW(2) is not that large, as the variables are boolean. Indeed, we could have
taken the state vector for the Atari games as a vector of 1024 boolean variables, 
and apply these algorithms to that representation. The results would be different.
In such a case, IW(1) would generate $n' \times D' \times b$ states, and 
IW(2), $(n' \times D') \times b$ states, which for $n'=1024$ and $D=2$
represent  $36,864$ and $7,5497,472$ states respectively.
These are both feasible numbers, but we haven't tried yet this representation,
under the assumption that the correlations among the bits in each one of 
the 128 words of the state vectors are meaningful. 
In summary, from the basic IW algorithm we are testing only the first call, IW(1).

IW is a purely exploration algorithm that does not take into account  the accumulated 
reward for selecting the next states to consider. As a simple, variant that combines
exploration and exploitation, we evaluated a \emph{best-first search} algorithm
with two queues: one queue  ordered first by  novelty measure 
(recall, novelty 1 means that the state is the first one to make some atom true),
and a second queue ordered by accumulated reward. In one iteration, the best first
search picks up the best node from one queue, and in the second iteration it picks
up the best node from the other queue. This way for combining multiple heuristics
is used in the LAMA planner \cite{richter:lama}, and was introduced in the
planner Fast Downward \cite{malte:fd,malte:alternating}. In addition, we break ties
in the first queue favoring states with largest accumulated reward, and 
in the second queue, favoring states with smallest novelty measure. Last, when
a node is expanded, it is removed from the queue, and its children are placed on 
both queues. The exception are the nodes with novelty 1 measure that are only
placed in first queue (** check and revise; but keep accurate and simple **).
We refer to this best-first algorithm below, as 2BFS.

IW11?

Randomization? Discount?

\section{Experimental Results in the Atari Games}

** Results, cold and detached, same format as those reported in jair paper, possibly with
std deviations and times.

** mention also ideas that have been tried and didn't work (IW2?, discount, randomization, ..)

\section{Related and Futue Work}

- Relation  to Novelty-based search (Stanley)
- IW probes vs UCT probes; UCT using novelty with noise in rollout???
- Structured Exploration of  Factored Spaces
- Dijkstra postprocessing, properties for optimization
- Learning in same instance, across diff instances of same game, etc.
- Learning controllers vs. on-line planning, etc

\end{document}

### 2014-10-30 workshop-paper/paper.tex
added table and aaai style

@@
 
+\begin{table*}
+%% \addtolength{\tabcolsep}{-2pt}
+{\scriptsize
+\begin{tabular}{@{}l|rr|rrr|rrr|rrr|rrr@{}}
+%% table header
+\hline\\[-3ex]
+\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Random} & \multicolumn{3}{c|}{BrFS}& \multicolumn{3}{c|}{IW1}& \multicolumn{3}{c|}{BFS}& \multicolumn{3}{c}{UCT} \\[.2ex]
+\hline\\[-3ex]
+%% table body
+\multicolumn{1}{c}{Game}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c|}{Dev}  & \multicolumn{1}{c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev}  & \multicolumn{1}{|c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev}  & \multicolumn{1}{|c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev}  & \multicolumn{1}{|c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev} \\[.2ex]
+\hline\\[-3ex]
+asterix         & 210   & 54.77     & 32.24 & 3900  & 5538.95   & 17.08 & 119000        & 4987.48   & 0.95  & 56320       & 4757.84   & 10.82 & \textbf{296300} & 6657.7    \\
+beam\_rider     & 448.8 & 84.64     & 43.48 & 519.2 & 179.81    & 6.02  & 3000          & 1150.22   & 0.92  & 3804        & 446.18    & 42.91 & \textbf{4320}   & 2175.26   \\
+breakout        & 1     & 0.71      & 51.56 & 2     & 1.41      & 1.37  & 383.6         & 9.84      & 3.95  & 421.2       & 13.66     & 26.03 & \textbf{512.4}  & 193.97    \\
+enduro          & 0     & 0         & 40.92 & 0.4   & 0.89      & 38.13 & \textbf{500}  & 0         & 13.12 & 292.8       & 6.18      & 32.14 & 306.8           & 8.14      \\
+freeway         & 0     & 0         & 71.84 & 0     & 0         & 15.19 & \textbf{29.4} & 0.89      & 4.44  & 19.6        & 3.29      & 90.64 & 1               & 1.41      \\
+pong            & -20.4 & 0.55      & 41.51 & -20.2 & 0.84      & 18.45 & 16.4          & 1.67      & 39.69 & \textbf{21} & 0         & 38.37 & \textbf{21}     & 0         \\
+qbert           & 185   & 345.33    & 42.83 & 160   & 85.88     & 1.82  & 1545          & 1666.06   & 1.76  & 14725       & 7872.1    & 28.61 & \textbf{18730}  & 1127.69   \\
+seaquest        & 192   & 38.99     & 40.01 & 200   & 14.14     & 14.72 & \textbf{4274} & 4206.81   & 1.59  & 4070        & 1939.55   & 26.16 & 3406            & 2402.47   \\
+space\_invaders & 187   & 120.86    & 46.11 & 115   & 61.44     & 12.1  & \textbf{2670} & 637.3     & 2.25  & 1327        & 598.04    & 19.66 & 2634            & 375.66   
+\end{tabular}
+}
+
+ \caption{\small
+   Random, BrFS, IW1, BFS, UCT performance over 5 runs, with
+   seeds 0 to 4.  \textit{T} is the action selection average
+   computation time, \textit{S} is the average score, and
+   \textit{Dev} is the standard deviation.  Numbers in bold show best performer in terms of average score.}
+\label{performance}
+
+\end{table*}
+
 ** Results, cold and detached, same format as those reported in jair paper, possibly with

### 2014-10-30 workshop-paper/paper.tex
Hectors last version

@@
 the 128 words of the state vectors are meaningful. 
-In summary, from the basic IW algorithm we are testing only the first call, IW(1).
+In summary, from the basic IW algorithm we are testing only the first 
+call, IW(1).
 
@@
 
-\section{Related and Futue Work}
+\section{Exploration and Exploitation: Discussion}
 
-- Relation  to Novelty-based search (Stanley)
-- IW probes vs UCT probes; UCT using novelty with noise in rollout???
-- Structured Exploration of  Factored Spaces
-- Dijkstra postprocessing, properties for optimization
-- Learning in same instance, across diff instances of same game, etc.
-- Learning controllers vs. on-line planning, etc
+The notion of width underlying the iterated width algorithm was developed in the context
+of classical planning in order to understand why most of the hundreds of existing benchmarks
+resulting from planning competitions appear to be relatively simple for current planners, even
+though classical planning is PSPACE-complete \cite{bylander}. The partial answer is that 
+most of these domains have a low width, and hence, can be solved in low polynomial time (by IW),
+when goals contain a single atom. At the same time, most problems with multiple atomic goals are 
+also easy too, as the goals can often be achieved one at a time \cite{nir:ecai2012}.
+
+In the Iterated Width algorithm, the key notion is the \emph{novelty measure} of a state in the underlying
+breadth-first search. In particular, a state gets a novelty 1 measure in the search if it makes true,
+some atom that up to that point had been false in all previously generated states. In the procedure IW(1)
+all other states are pruned right away, as if they were ``duplicate'' states. The algorithms IW and IW(k)
+are pure  exploration algorithms that make no use at all of the information about the goal, if there is a goal,
+or about the  reward function to optimize otherwise. They are  thus  pure exploration algorithms guided by the novelty
+measures, which use the structure of the states to provide a  structure to the exploration of the state space,
+which in the case of IW is guaranteed to be complete.
+
+The use of ``novelty measures'' for guiding an optimization search while ignoring the function that is being
+optimized, appears to be  common to novelty-based search methods developed independently 
+in the context of genetic algorithms \cite{lehman:novelty}. %%% bibtex for this below
+In these methods individuals in the population are not ranked according to optimization function but in terms of how 
+``novel'' thay are in relation to the rest of the population, thus encouraging diversity
+and exploration, rather than (greedy) exploitation. The actual definition of novelty
+in such a case is domain-dependent; for example, in the evolution of a controller
+for guiding a robot in a maze, an individual controller  will not be ranked by how
+close it takes the robot to the goal (the greedy measure), but how far it takes
+the robot in relation to the locations that are reached by following the other controllers
+(a diversity measure). The novelty measure used by IW, on the other hand, 
+is domain-independent and it follows from the structure of the state
+as captured by the problem variables.
+
+% @article{lehman:novelty,
+%  title={Abandoning objectives: Evolution through the search for novelty alone},
+%  author={Lehman, Joel and Stanley, Kenneth O},
+%  journal={Evolutionary computation},
+%  volume={19},
+%  number={2},
+%  pages={189--223},
+%  year={2011},
+%  publisher={MIT Press}
+%}
+
+The balance between exploration and exploitation has received considerable attention in
+reinforcement learning \cite{sutton:book}, where both are required for converging to an
+optimal behavior. In the Atari games, as in other deterministic problems, however, ``exploration'' 
+is not needed for optimality purposes, but just for improving the effectiveness of the 
+local lookahead searches. Indeed, a best-first search algorithm guided by (discounted)
+accumulated reward will deliver eventually  best moves, but it will not be as effective
+over small time windows, where like breadth-first search it's likely not to find any rewards
+at all. The UCT algorithm provides a method for balancing exploration and exploitation,
+which can thus be effective over small time windows. The 2BFS algorithm above with two queues
+that alternate, one guided by the novelty measures, and the other by the accumulated reward,
+provides a different scheme. The first converges to the optimal behavior asymptotically; the 
+second in a bounded number of steps. Alternately, novelty-guided probes, possibly with 
+some noise added, could potentially pay off as  a  base policy for UCT in replacement 
+of random rollouts, as they manage apparently to exploit the structure of many problems.
+
+\section{Summary and On-going Work}
+
+We have shown experimentally that width-based algorithms like IW(1) and 2BFS
+that originate in work in classical planning, can be used in the Atari games
+with a performance that approaches the state-of-the-art as captured by UCT 
+planning, sometimes using a fraction of the time. We are currently addressing three issues, 
+which we haven't managed to get in time for this submission. First, we want to extend the evaluation to 
+all the games featured in \cite{bellemare:jair}, and to add more runs per  game. 
+Second, we want to evaluate the algorithm for different fixed time windows. In the table above, for example, UCT
+is often taking an order-of-magnitude more time per decision as 2BFS, so the 
+comparison in not fair. Third, we need to revise the way duplicate states are handled 
+in IW and 2BFS to ensure that the best paths among the paths defined by the states that have been expanded
+are returned. This change is important from a theoretical point of view; it's not clear
+what it's practical impact will be.
+
+%%% - Learning in same instance, across diff instances of same game, etc.
+%%%% - Learning controllers vs. on-line planning, etc
 

### 2014-10-31 workshop-paper/paper.tex
Experiments draft

@@
 
-\begin{table*}
-%% \addtolength{\tabcolsep}{-2pt}
-{\scriptsize
-\begin{tabular}{@{}l|rr|rrr|rrr|rrr|rrr@{}}
-%% table header
-\hline\\[-3ex]
-\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Random} & \multicolumn{3}{c|}{BrFS}& \multicolumn{3}{c|}{IW1}& \multicolumn{3}{c|}{BFS}& \multicolumn{3}{c}{UCT} \\[.2ex]
-\hline\\[-3ex]
-%% table body
-\multicolumn{1}{c}{Game}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c|}{Dev}  & \multicolumn{1}{c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev}  & \multicolumn{1}{|c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev}  & \multicolumn{1}{|c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev}  & \multicolumn{1}{|c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev} \\[.2ex]
-\hline\\[-3ex]
-asterix         & 210   & 54.77     & 32.24 & 3900  & 5538.95   & 17.08 & 119000        & 4987.48   & 0.95  & 56320       & 4757.84   & 10.82 & \textbf{296300} & 6657.7    \\
-beam\_rider     & 448.8 & 84.64     & 43.48 & 519.2 & 179.81    & 6.02  & 3000          & 1150.22   & 0.92  & 3804        & 446.18    & 42.91 & \textbf{4320}   & 2175.26   \\
-breakout        & 1     & 0.71      & 51.56 & 2     & 1.41      & 1.37  & 383.6         & 9.84      & 3.95  & 421.2       & 13.66     & 26.03 & \textbf{512.4}  & 193.97    \\
-enduro          & 0     & 0         & 40.92 & 0.4   & 0.89      & 38.13 & \textbf{500}  & 0         & 13.12 & 292.8       & 6.18      & 32.14 & 306.8           & 8.14      \\
-freeway         & 0     & 0         & 71.84 & 0     & 0         & 15.19 & \textbf{29.4} & 0.89      & 4.44  & 19.6        & 3.29      & 90.64 & 1               & 1.41      \\
-pong            & -20.4 & 0.55      & 41.51 & -20.2 & 0.84      & 18.45 & 16.4          & 1.67      & 39.69 & \textbf{21} & 0         & 38.37 & \textbf{21}     & 0         \\
-qbert           & 185   & 345.33    & 42.83 & 160   & 85.88     & 1.82  & 1545          & 1666.06   & 1.76  & 14725       & 7872.1    & 28.61 & \textbf{18730}  & 1127.69   \\
-seaquest        & 192   & 38.99     & 40.01 & 200   & 14.14     & 14.72 & \textbf{4274} & 4206.81   & 1.59  & 4070        & 1939.55   & 26.16 & 3406            & 2402.47   \\
-space\_invaders & 187   & 120.86    & 46.11 & 115   & 61.44     & 12.1  & \textbf{2670} & 637.3     & 2.25  & 1327        & 598.04    & 19.66 & 2634            & 375.66   
-\end{tabular}
-}
+\input{performance_table}
 
- \caption{\small
-   Random, BrFS, IW1, BFS, UCT performance over 5 runs, with
-   seeds 0 to 4.  \textit{T} is the action selection average
-   computation time, \textit{S} is the average score, and
-   \textit{Dev} is the standard deviation.  Numbers in bold show best performer in terms of average score.}
-\label{performance}
+The algorithms IW1 and BFS are tested over 9 games selected from those
+considered in the training set in \cite{bellemare:jair}, and those
+considered in \cite{deep-mind-atari}. Each game has 5 runs, each with
+a different random seed $0,\ldots,4$.  We compare the performance of
+IW1 and BFS with UCT and BRFS. We also include results for a baseline random
+agent. We use the same parameters in the literature, all algorithms use a discount factor of 0.999, and each action
+simulates 5 frames. Each algorithm has a budget on the number of nodes
+it can generate per lookahead. BRFS, IW1 and BFS have a maximum budget
+of 100k. For UCT we use the same parameters as in
+\cite{bellemare:jair}, 0.1 exploration rate, and 500 rollouts of a maximum depth of 300 steps,
+which gives a maximum budget of 150k nodes. We tested UCT with a fix
+budget of 100k and max depth 300, but resulted in practically the same
+results. All algorithms reuse the tree built between lookaheads by
+setting the new root of the tree to the node of the selected action,
+and deleting the other branches. Note that this may have a different
+effect on algorithms were all branches have the same number of
+generated nodes, ex. BRFS, than those that may focus on different
+areas of the tree, ex. IW1, BFS, UCT.
 
-\end{table*}
+On Table~\ref{performance}, we can observe the performance of each
+algorithm on terms of average score with standard deviation, and
+average time per lookahead in seconds. The Random agent don't include
+average time as it just chooses an action randomly. BRFS outperform
+the Random agent in 7 out of 9 domains. BRFS rarely finds any reward,
+turning itself mostly into a random agent, but exploring 100k nodes up
+to an average depth of 4. When BRFS is turned int IW1 using novelty to
+explore the search space, the jump on performance is important. IW1 in
+most of the cases does not completely use its budget, and is able to
+achieve depths of 70 to 260 depending on the game, which first
+explains how IW1 is able to achieve a performance that is closer to
+UCT than to BRFS. In 4 out of 9 tested games it is the best
+performer. Remarkably, is the first domain independent solver that
+shows good performance on Freeway, where rewards are very sparse and
+deep in the search tree. BFS outperforms IW1 and UCT in 4 and 2
+domains respectively.  Both IW1 and BFS performance in terms of score
+is close to UCT, but they are some times one order of magnitud
+faster. The speedup is due to the bounded space in IW1, and to the
+fact that in some games, both IW1 and BFS are able to reuse an
+important part of their trees between lookaheads.
 
-** Results, cold and detached, same format as those reported in jair paper, possibly with
-std deviations and times.
+The novelty of states in IW1 and BFS may depend on the order that
+nodes are generated. Therefore, every time a node is expanded, we
+randomized the order in which nodes get genreated. Randomizing the
+order of expansion is also applied to BRFS, but has no impact on its
+performance.
 
-** mention also ideas that have been tried and didn't work (IW2?, discount, randomization, ..)
+IW2 is not included in the table due to its low performance, which is
+closer to BRFS than to the other algorithms. This is mainly due to the
+huge number of states that are novel in this algorith, thus,
+achieveing insufficient depths to gather some reward.
+
+ ** mention also ideas that have been tried and didn't work (IW2?,
+ discount, randomization, ..)
 

### 2014-10-31 workshop-paper/performance_table.tex
Experiments draft

\begin{table*}
\addtolength{\tabcolsep}{-2pt}
{\scriptsize
\begin{tabular}{@{}l|rr|rrr|rrr|rrr|rrr@{}}
%% table header
\hline\\[-3ex]
\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Random} & \multicolumn{3}{c|}{BrFS}& \multicolumn{3}{c|}{IW1}& \multicolumn{3}{c|}{BFS}& \multicolumn{3}{c}{UCT} \\[.2ex]
\hline\\[-3ex]
%% table body
\multicolumn{1}{c}{Game}  & \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Std Dev}  & \multicolumn{1}{c}{Time} & \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c}{Std Dev} & \multicolumn{1}{|c}{Time} & \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c}{Std Dev}  & \multicolumn{1}{|c}{Time} & \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c}{Std Dev}  & \multicolumn{1}{|c}{Time} & \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c}{Std Dev} \\[.2ex]
\hline\\[-3ex]
asterix         & 210   & 54.77     & 32.24 & 3900  & 5538.95   & 17.08 & 119000        & 4987.48   & 0.95  & 56320       & 4757.84   & 10.82 & \textbf{296300} & 6657.7    \\
beam\_rider     & 448.8 & 84.64     & 43.48 & 519.2 & 179.81    & 6.02  & 3000          & 1150.22   & 0.92  & 3804        & 446.18    & 42.91 & \textbf{4320}   & 2175.26   \\
breakout        & 1     & 0.71      & 51.56 & 2     & 1.41      & 1.37  & 383.6         & 9.84      & 3.95  & 421.2       & 13.66     & 26.03 & \textbf{512.4}  & 193.97    \\
enduro          & 0     & 0         & 40.92 & 0.4   & 0.89      & 38.13 & \textbf{500}  & 0         & 13.12 & 292.8       & 6.18      & 32.14 & 306.8           & 8.14      \\
freeway         & 0     & 0         & 71.84 & 0     & 0         & 15.19 & \textbf{29.4} & 0.89      & 4.44  & 19.6        & 3.29      & 90.64 & 1               & 1.41      \\
pong            & -20.4 & 0.55      & 41.51 & -20.2 & 0.84      & 18.45 & 16.4          & 1.67      & 39.69 & \textbf{21} & 0         & 38.37 & \textbf{21}     & 0         \\
qbert           & 185   & 345.33    & 42.83 & 160   & 85.88     & 1.82  & 1545          & 1666.06   & 1.76  & 14725       & 7872.1    & 28.61 & \textbf{18730}  & 1127.69   \\
seaquest        & 192   & 38.99     & 40.01 & 200   & 14.14     & 14.72 & \textbf{4274} & 4206.81   & 1.59  & 4070        & 1939.55   & 26.16 & 3406            & 2402.47   \\
space\_invaders & 187   & 120.86    & 46.11 & 115   & 61.44     & 12.1  & \textbf{2670} & 637.3     & 2.25  & 1327        & 598.04    & 19.66 & 2634            & 375.66   
\end{tabular}
}

 \caption{\small
   Random, BrFS, IW1, BFS, UCT performance over 5 runs, with
   seeds 0 to 4. \textit{Time} is the action selection average
   computation time in seconds, \textit{Avg. Score} is the average score, and
   \textit{Std Dev} is the standard deviation.  Numbers in bold show best performer in terms of average score.}
\label{performance}

\end{table*}

### 2014-10-31 workshop-paper/performance_table.tex
Table updated

@@
 \multicolumn{1}{c}{Game}  & \multicolumn{1}{c}{Avg.
-Score} & \multicolumn{1}{c|}{Std Dev}  & \multicolumn{1}{c}{Time} & \multicolumn{1}{c}{Avg.
-Score} & \multicolumn{1}{c}{Std Dev} & \multicolumn{1}{|c}{Time} & \multicolumn{1}{c}{Avg.
-Score} & \multicolumn{1}{c}{Std Dev}  & \multicolumn{1}{|c}{Time} & \multicolumn{1}{c}{Avg.
-Score} & \multicolumn{1}{c}{Std Dev}  & \multicolumn{1}{|c}{Time} & \multicolumn{1}{c}{Avg.
-Score} & \multicolumn{1}{c}{Std Dev} \\[.2ex]
+Score} & \multicolumn{1}{c|}{Std Dev}  & \multicolumn{1}{c}{Avg.
+Score} & \multicolumn{1}{c}{Std Dev} &  \multicolumn{1}{c|}{Time}  & \multicolumn{1}{c}{Avg.
+Score} & \multicolumn{1}{c}{Std Dev} & \multicolumn{1}{c|}{Time}  & \multicolumn{1}{c}{Avg.
+Score} & \multicolumn{1}{c}{Std Dev} & \multicolumn{1}{c|}{Time}  & \multicolumn{1}{c}{Avg.
+Score} & \multicolumn{1}{c}{Std Dev} & \multicolumn{1}{c}{Time}\\[.2ex]
 \hline\\[-3ex]
-asterix         & 210   & 54.77     & 32.24 & 3900  & 5538.95   & 17.08 & 119000        & 4987.48   & 0.95  & 56320       & 4757.84   & 10.82 & \textbf{296300} & 6657.7    \\
-beam\_rider     & 448.8 & 84.64     & 43.48 & 519.2 & 179.81    & 6.02  & 3000          & 1150.22   & 0.92  & 3804        & 446.18    & 42.91 & \textbf{4320}   & 2175.26   \\
-breakout        & 1     & 0.71      & 51.56 & 2     & 1.41      & 1.37  & 383.6         & 9.84      & 3.95  & 421.2       & 13.66     & 26.03 & \textbf{512.4}  & 193.97    \\
-enduro          & 0     & 0         & 40.92 & 0.4   & 0.89      & 38.13 & \textbf{500}  & 0         & 13.12 & 292.8       & 6.18      & 32.14 & 306.8           & 8.14      \\
-freeway         & 0     & 0         & 71.84 & 0     & 0         & 15.19 & \textbf{29.4} & 0.89      & 4.44  & 19.6        & 3.29      & 90.64 & 1               & 1.41      \\
-pong            & -20.4 & 0.55      & 41.51 & -20.2 & 0.84      & 18.45 & 16.4          & 1.67      & 39.69 & \textbf{21} & 0         & 38.37 & \textbf{21}     & 0         \\
-qbert           & 185   & 345.33    & 42.83 & 160   & 85.88     & 1.82  & 1545          & 1666.06   & 1.76  & 14725       & 7872.1    & 28.61 & \textbf{18730}  & 1127.69   \\
-seaquest        & 192   & 38.99     & 40.01 & 200   & 14.14     & 14.72 & \textbf{4274} & 4206.81   & 1.59  & 4070        & 1939.55   & 26.16 & 3406            & 2402.47   \\
-space\_invaders & 187   & 120.86    & 46.11 & 115   & 61.44     & 12.1  & \textbf{2670} & 637.3     & 2.25  & 1327        & 598.04    & 19.66 & 2634            & 375.66   
+asterix         & 210    & 54.77     & 3900  & 5538.95   & 32.24 & 119000        & 4987.48   & 17.08 & 56320       & 4757.84   & 0.95  & \textbf{296300 (290700)} & 6657.7    & 10.82 \\
+beam\_rider     & 448.8  & 84.64     & 519.2 & 179.81    & 43.48 & 3000          & 1150.22   & 6.02  & 3804        & 446.18    & 0.92  & \textbf{4320 (6624)}     & 2175.26   & 42.91 \\
+breakout        & 1      & 0.71      & 2     & 1.41      & 51.56 & 383.6         & 9.84      & 1.37  & 421.2       & 13.66     & 3.95  & \textbf{512.4 (364)}     & 193.97    & 26.03 \\
+enduro          & 0      & 0         & 0.4   & 0.89      & 40.92 & \textbf{500}  & 0         & 38.13 & 292.8       & 6.18      & 13.12 & 306.8 (286)              & 8.14      & 32.14 \\
+freeway         & 0      & 0         & 0     & 0         & 71.84 & \textbf{29.4} & 0.89      & 15.19 & 19.6        & 3.29      & 4.44  & 1 (0)                    & 1.41      & 90.64 \\
+pong            & -20.4  & 0.55      & -20.2 & 0.84      & 41.51 & 16.4          & 1.67      & 18.45 & \textbf{21} & 0         & 39.69 & \textbf{21 (21)}         & 0         & 38.37 \\
+qbert           & 185    & 345.33    & 160   & 85.88     & 42.83 & 1545          & 1666.06   & 1.82  & 14725       & 7872.1    & 1.76  & \textbf{18730 (17343)}   & 1127.69   & 28.61 \\
+seaquest        & 192    & 38.99     & 200   & 14.14     & 40.01 & \textbf{4274} & 4206.81   & 14.72 & 4070        & 1939.55   & 1.59  & 3406 (5132)              & 2402.47   & 26.16 \\
+space\_invaders & 187    & 120.86    & 115   & 61.44     & 46.11 & \textbf{2670} & 637.3     & 12.1  & 1327        & 598.04    & 2.25  & 2634 (2718)              & 375.66    & 19.66
 \end{tabular}
@@
 
- \caption{\small
-   Random, BrFS, IW1, BFS, UCT performance over 5 runs, with
-   seeds 0 to 4. \textit{Time} is the action selection average
-   computation time in seconds, \textit{Avg. Score} is the average score, and
-   \textit{Std Dev} is the standard deviation.  Numbers in bold show best performer in terms of average score.}
+ \caption{\small Random, BrFS, IW1, BFS, UCT performance over 5 runs,
+   with seeds 0 to 4, and maximum duration of 18.000
+   frames. \textit{Time} is the action selection average computation
+   time in seconds, \textit{Avg. Score} is the average score achieved,
+   and \textit{Std Dev} is the standard deviation.  UCT performance
+   reported in \cite{bellemare:jair} is shown in parenthesis. Numbers
+   in bold show best performer in terms of average score.}
 \label{performance}

### 2014-10-31 workshop-paper/experimental_results.tex
Bibliography and Experimental Section separated

--- 2014-10-31 workshop-paper/paper.tex
@@
-  
-\documentclass[letterpaper]{article}
-\usepackage{aaai}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\frenchspacing
-\setlength{\pdfpagewidth}{8.5in}
-\setlength{\pdfpageheight}{11in}
-
-\newcommand{\tuple}[1]{{\langle #1\rangle}}
-\newcommand{\triple}[1]{{\langle #1\rangle}}
-\newcommand{\pair}[1]{{\langle #1\rangle}}
-
-\newcommand{\Omit}[1]{}
-
-\newcommand{\eqdef}{\stackrel{\hbox{\tiny{def}}}{=}}
-\newcommand{\IW}{{\textit{IW}}}
-\newcommand{\SIW}{{\textit{SIW}}}
-\newcommand{\ID}{{\textit{ID}}}
-\newcommand{\BRFS}{{\textit{BrFS}}}
-
-% AAAI (comment for ECAI)
-% \newtheorem{theorem}{Theorem}
-
- \newtheorem{theorem}{Definition}
-% \newtheorem{proposition}[theorem]{Proposition}
-% \newtheorem{lemma}[theorem]{Lemma}
-% \newtheorem{corollary}[theorem]{Corollary}
- \newtheorem{definition}[theorem]{Definition}
-
- \begin{document}
-% The file aaai.sty is the style file for AAAI Press 
-% proceedings, working notes, and technical reports.
-%
-\pdfinfo{
-/Title (Classical Planning Algorithms on the Atari Video Games: Preliminary Results)
-/Author (Nir Lipovetzky, Miquel Ramirez, Hector Geffner)}
-
-\setcounter{secnumdepth}{0}
-
-\title{Classical Planning Algorithms ¡ on the Atari Video Games: Preliminary Results}
-
-\author{Nir, Miquel, H}
-%AAAI Press\\
-%Association for the Advancement of Artificial Intelligence\\
-%2275 East Bayshore Road, Suite 160\\
-% Palo Alto, California 94303\\}
-
-\maketitle
-
-%% Pending: randomization, discount factor, iw11 ..
-%% iw2 multivalued expands up to 256^2 n^nodes, while iw2 boolean 16^2 n^2 nodes; can iw2 boolean beat iw1 multivalued?
-%% Important. optimization, need Dikjstra por postprocessing .. see emails
-
-\begin{abstract}
-\begin{quote}
-The Atari  2600 games supported in  the Arcade Learning Environment (Bellemare et. al. 2013) 
-all feature a known initial (RAM) state and actions that have deterministic effects. Classical planners, however, 
-cannot be used  for selecting actions  for  two reasons: first,  no  compact PDDL-model of  the games is given, 
-and  more importantly, the actions effects and the goals are  not known a priori. Moreover, in these games there is usually  
-no set of goals to be achieved but rewards to be collected. These features do not preclude the use of   classical  algorithms like
-breadth-first search or Dijkstra's algorithm,  but    these  algorithms are not effective over  large state spaces. We thus turn to a
-different class of classical planning algorithms  introduced recently  that perform a \emph{structured exploration} of the state space; namely,  
-like breadth-first search and Dijkstra's  algorithm they are ``blind'' and hence  do not require prior knowledge of state transitions, 
-costs (rewards) or goals, and yet, like  heuristic search algorithms, they have been shown to be   effective for solving   problems over huge 
-state spaces. The simplest such algorithm, called Iterated Width or IW, consists of  a sequence of calls  IW(1), IW(2), .., IW(k) where  IW(i) is  
-a breadth-first search in which a state is  pruned when it is not the first state in the search  to make true some set of i atoms or less.
-Preliminary empirical results suggest that the performance of IW with the  $k$ parameter fixed to $1$, i.e., IW(1), 
-is much better than random or breadth-first search, and close to the performance of UCT. Other simple variations of IW
-that combine exploration and exploitation are evaluated as well.
-\end{quote}
-\end{abstract}
-
-\section{Introduction}
-
-The Arcade Learning Environment (ALE)  provides a challenging  platform for evaluating general, domain-independent AI planners and 
-learners through a convenient interface to hundreds of Atari 2600 games \cite{bellemare:jair2013}. 
-Results have been reported so far for basic planning algorithms like breadth-first 
-search and UCT \cite{bellemare:jair2013}, and for  reinforcement learning
-and evolutionary  methods \cite{bellemare:jair2013,deep-mind-atari,risto-texas:atari-games}.
-The empirical results are impressive in some cases, yet a lot remains to be done, as no method  approaches   the 
-performance of  human players across a broad range of games.
-
-While all  these games feature a known initial (RAM) state and actions that
-have deterministic effects on this state, the problem of selecting the next action
-to be done cannot be addressed with state-of-the-art classical planners for two 
-reasons. First, there is no compact PDDL-like encoding of the domain, and more 
-importantly,  the goal to be achieved in each game  is not given. Indeed, there are 
-often no goals at all but rewards $r(a,s)$ that potentially depend on the action $a$ and 
-the state $s$ in  which  the action is performed, and these rewards are not known a priori either. 
-Thus no model of the  goals  or the  rewards can be used to bias the search, and hence, 
-standard heuristic search and classical planning methods can't be used.
-
-The action selection problem in the Atari games can be seen as a \emph{reinforcement learning} problem 
-\cite{sutton:book} over a deterministic MDP where the state transitions and rewards are not known, or alternatively,
-as a  \emph{net-benefit planning problem} \cite{ref-net-benefit-planning} with unknown state transitions  and  rewards. Namely, 
-we seek  an action sequence $a_0, a_1, \ldots, a_m$ from the given current  state $s_0$ that generates 
-a state sequence $s_0, s_1, \ldots, s_{m+1}$ with  maximum  total reward $\sum_{i=0}^{m} r(a_i,s_i)$,  
-where $m$ is a sufficiently large planning horizon, and the rewards $r(a_i,s_i)$ and the  state-transitions $s_{i+1}=f(a_i,s_i)$ 
-are known only after the action $a_i$ is applied in the state $s_i$.
-
-The presence of unknown transition and  rewards in  the Atari games does not preclude the use of blind-search
-methods like breadth-first search, Dijkstra's algorithm \cite{dijkstra},  or learning methods  such as 
-LRTA* \cite{korf:lrta}, UCT \cite{uct}, and Q-learning  \cite{sutton:rl,bertsekas:neuro}. Indeed, 
-the net-benefit planning problem with unknown state transitions and rewards over a given planning
-horizon, can be mapped into a standard \emph{shortest-path problem} which can be solved optimally by Dijkstra's algorithm.
-For this, we just need to  map the unknown rewards $r(a,s)$ into positive (unknown) action costs $c(a,s) = C - r(a,s)$
-where  $C$ is a large constant that exceeds the maximum possible reward. The fact that the state 
-transition and cost functions  $f(a,s)$ and $c(a,s)$ are not known a priori doesn't affect the applicability of 
-Dijkstra's algorithm, which requires the value of these functions precisely when  the action $a$
-is applied in the state $s$.
-
-The limitation of the   basic blind search  methods is that they are not effective  over very spaces, neither
-for solving problems off-line, nor for guiding  a lookahead search for solving problems on-line. 
-In this work, we thus turn to a recent class of  planning algorithms that combine the scope of blind search
-methods with the performance of state-of-the-art classical planning algorithms: namely, like ``blind'' search algorithms
-they do not require prior knowledge of state transitions, costs, or goals, and yet like heuristic algorithms
-they manage to search large  state spaces effectively. The basic algorithm in this class is 
-called IW for Iterated Width search \cite{nir:ecai2012}. IW consists of a sequence of calls  IW(1), IW(2), .., IW(k), where 
-IW(i) is  a  standard breadth-first search where states are pruned right away when they fail to make true some new tuple (set) 
-of at most $i$ atoms. Namely,   IW(1) is a breadth-first search that keeps  a state only if  the state is   the first  one 
-in the search to make some atom  true; IW(2) keeps a state  only if the state is  the first one to  make  a pair of atoms true, and 
-so on.  Like plain breadth-first  and iterative deepening searches,  IW  is 
-complete, while searching the state space in a way that makes use of the 
-\emph{structure of states} given by the values of a finite set of state variables.
-In the Atari games, the (RAM) state is given by a vector of 128 bytes, which we associate with 128 variables 
-$X_i$, $i=1, \ldots, 128$, each of which may take up to $256$ values $x_j$. A state $s$ makes an atom $X_i=x_{j}$ 
-true when the value of the $i$-th byte in the state vector $s$ is $x_j$.
-
-While a normal, complete breadth-first search runs in time that is exponential in the total number of variables $X_i$, 
-the procedure $IW(k)$ runs in time that is exponential in the $k$ parameter, $1 \leq k \leq n$. 
-Elsewhere, it has been shown that  IW exhibits   state-of-the-art performance on most existing planning benchmark domains 
-when goals are restricted to be single atoms  where IW provably runs in low polynomial time in the number of
-variables \cite{nir:ecai2012}. For example, in a a block world instance with any number of blocks and any initial
-configuration,  the goal of  placing a given  block on top of another, will be achievable in  quadratic time  
-in the number of blocks as any such instance can be shown to  have \emph{width} 2. This means  that IW(2) will be 
-not only complete then, but also optimal. This is also true for most other benchmark domains in classical planning
-that turn out to be have small, bounded widths for any instance, as long as the goal is atomic. 
-For dealing with the benchmarks where goals are  not single atoms, e.g., where a given \emph{tower} of blocks needs to be built, 
-simple extensions of the basic IW procedure have been developed as well, such as Serialized IW, where IW is 
-used to achieve the goals one at at time \cite{nir:ecai2012,nir:ecai2014}. The extension of IW for problems where there is no goal but an
-additive reward measure   to be optimized is direct.
-
-%%  Each iteration IW(k) keeps track of the accumulated reward up to each state $s$ that has been
-%% generated but  not pruned, as $R(s) = R(s')+ r(a,s')$ where $s'$ and $a$ are  the single parent state
-%% and action $a$ that led to $s$ in IW(k) with $R(s)=0$ for the current state $s$. The best reward
-
-The sequence of calls IW(1), IW(2), \ldots, IW(k) manages to combine the scope of ``blind'' search algorithms with the performance of 
-heuristic-search algorithms by exploring states  in an  order dictated by the ``width''  parameter $k$; 
-states that have width 1 are explored first in linear time, then states that have  width two are explored in quadratic time, and so on. 
-In classical planning problems, if the goal has a small width (something common in the classical planning domains when the  goal is a single atom, as
-we have seen above), IW will find goal states  in low polynomial time. Similarly, in the reward-based setting, if  there are states of high reward 
-that have small width, IW will find such states in low polynomial time, even if the state space is exponential in the number of variables. 
-In this paper, we aim to test whether these ideas work for the Atari games too  when  the RAM state of the emulator
-is used  as the system state.
-
-The paper is organized as follows. We review the Iterated Width  algorithm and where it comes from, 
-look at  some simple   variations of the basic algorithm that appear to be convenient  for the Atari
-games, present and analyze the preliminary experimental results, summarize the main lessons, and
-discuss related and future work.
-
-\section{Iterated Width}
-
-The Iterative Width  (IW) algorithm has been introduced as a classical planning algorithm
-that takes a planning problem as an input, and computes an action sequence that solves the problem 
-as the output \cite{nir:ecai2012}. The algorithm however
-applies to a broader range of problems. We will characterize such
-problems  by means  a finite and discrete set of states (the state space)
-that correspond to vectors of size $n$. Namely, the 
-states are  \emph{structured} or \emph{factored} , and we take each of the locations in the vector
-to represent a variable $X_i$, and the value at that vector location to 
-represent the value $x_j$ of variable $X_i$ in the state. 
-In addition to the states space, a problem is defined by an initial state $s_0$, 
-a set of actions $A(s)$ applicable in each state, a 
-transition function $f$ such that  $s' = f(a,s)$ is 
-the state that results from applying action $a \in A(s)$ to the state $s$, 
-and rewards $r(a,s)$ represented by real numbers that result from  applying action $a$ in state $s$.
-The transition and reward functions do not  need to be known a priori, yet in that case, the  
-state and  reward that results from the application of an action in a state needs to be \emph{observable}.
-The task is to compute an action sequence $a_0, \ldots, a_m$ for a large horizon $m$
-that generates a state sequence $s_0, \ldots, s_{m+1}$ that  maximizes the accumulated 
-reward $\sum_{i=0}^{m} r(a_i,s_i)$, or that provides a good approximation.
-
-\subsection{The Algorithm}
-
-IW consists of a sequence of calls  $\IW(i)$ for $i=0, 1, 2, \ldots$
-over a problem $P$ until a termination condition is reached. 
-The procedure $\IW(i)$  is a plain forward-state \emph{breadth-first search} with just one change: right after a state 
-$s$ is generated, the state is  pruned if it doesn't pass a simple \emph{novelty test}.
-More precisely,
-
-\begin{itemize}
-\item The \emph{novelty of a newly generate state $s$  in a  breadth-first search}
-is   $1$ when $s$ is the first state generated in the search that makes true an atom $X=x$,
-$2$  when $s$ is the first state that makes a pair of atoms $X=x$ and $Y=y$ true, and so on.
-
-\item  $\IW(i)$ is a breadth-first search that prunes newly generated states when their novelty measure
-is greater than $i$.
-
-\item $\IW$  calls $\IW(i)$ sequentially for $i=1,2,\ldots$ until a termination condition is reached, 
-returning then the best path found.
-\end{itemize}
-
-For classical planning, the termination condition is the achievement of the goal. In the on-line setting, 
-as in the Atari games, the termination condition is given by a time window or a maximum number of generated nodes.
-The  best path found by IW is the path  that has a  maximum accumulated reward. 
-The accummulated reward $R(s)$  of a state $s$ reached in an iteration of IW is determined by the  unique parent state $s'$ 
-and action $a$ leading to $s$ from $s'$ as $R(s) = R(s') + r(a,s')$. The best state is the state $s$ with maximum reward
-$R(s)$ generated but not pruned by IW, and the best path is the one that leads to the state $s$ from the current state.
-The action selected to be done next is then the first action along such a path  (** see eventually randomization **).
-
-%% Like in standard breadth-first search, each iteration IW(i)  prunes 
-%%% the  duplicate states as well.
-
-\begin{table}
-\begin{center}
-%\resizebox{3in}{!}{
-{\scriptsize
-% \begin{tabular}{r@{}lrrrr@{}}
-\begin{tabular}{llcccc}
-%% table header
-\hline\\[-1.3ex]
-%\toprule
-\# & \multicolumn{1}{c}{Domain} & \multicolumn{1}{c}{I} & \multicolumn{1}{c}{IW(1)} & \multicolumn{1}{c}{IW(2)} & \multicolumn{1}{c}{Neither} \\[.2ex]
-
-\hline\\[-1.3ex]
-%\hline
-%% table body
-1. & 8puzzle	&400	&55\%	&45\%	&0\%\\
-2. & Barman	&232	&9\%	&0\%	&91\%\\
-3. & Blocks World	&598	&26\%	&74\%	&0\%\\
-4. &Cybersecure	&86	&65\%	&0\%	&35\%\\
-% 5. &Depots	&189	&11\%	&66\%	&23\%\\
-% 6. & Driver	&259	&45\%	&55\%	&0\%\\
-% 7. & Elevators	&510	&0\%	&100\%	&0\%\\
-% 8. & Ferry	&650	&36\%	&64\%	&0\%\\
-% 9. & Floortile	&538	&96\%	&4\%	&0\%\\
-% 10. & Freecell	&76	&8\%	&92\%	&0\%\\
-% 11. & Grid	&19	&5\%	&84\%	&11\%\\
-% 12. & Gripper	&1275	&0\%	&100\%	&0\%\\
-% 13. & Logistics	&249	&18\%	&82\%	&0\%\\
-% 14. & Miconic	&650	&0\%	&100\%	&0\%\\
-% 15. & Mprime	&43	&5\%	&95\%	&0\%\\
-% 16. & Mystery	&30	&7\%	&93\%	&0\%\\
-% 17. & NoMystery	&210	&0\%	&100\%	&0\%\\
-% 18. & OpenStacks	&630	&0\%	&0\%	&100\%\\
-% 19. & OpenStacksIPC6	&1230	&5\%	&16\%	&79\%\\
-\ldots & \ldots & \ldots & \ldots & \ldots & \\ 
-20. & ParcPrinter	&975	&85\%	&15\%	&0\%\\
-21. & Parking	&540	&77\%	&23\%	&0\%\\
-22. & Pegsol	&964	&92\%	&8\%	&0\%\\
-23. & Pipes-NonTan	&259	&44\%	&56\%	&0\%\\
-24. & Pipes-Tan	&369	&59\%	&37\%	&3\%\\
-25. & PSRsmall	&316	&92\%	&0\%	&8\%\\
-26. & Rovers	&488	&47\%	&53\%	&0\%\\
-27. & Satellite	&308	&11\%	&89\%	&0\%\\
-28. & Scanalyzer	&624	&100\%	&0\%	&0\%\\
-29. & Sokoban	&153	&37\%	&36\%	&27\%\\
-30. & Storage	&240	&100\%	&0\%	&0\%\\
-31. & Tidybot	&84	&12\%	&39\%	&49\%\\
-32. & Tpp	&315	&0\%	&92\%	&8\%\\
-33. & Transport	&330	&0\%	&100\%	&0\%\\
-34. & Trucks	&345	&0\%	&100\%	&0\%\\
-35. & Visitall	&21859	&100\%	&0\%	&0\%\\
-36. & Woodworking	&1659	&100\%	&0\%	&0\%\\
-37. & Zeno	&219	&21\%	&79\%	&0\%\\[.2ex]
-%\hline
-\hline\\[-1.3ex]
-& Total/Avgs&37921	&37.0\%	&51.3\%	&11.7\%\\
-\hline
-\end{tabular}
-}
-
-\vskip .1cm
-
-{\scriptsize
-\begin{tabular}{@{}ccccc@{}}
-\hline\\[-1.3ex]
-\multicolumn{1}{c}{\# Instances}  &\multicolumn{1}{c}{\IW} & \multicolumn{1}{c}{\ID} & \multicolumn{1}{c}{\BRFS} &\multicolumn{1}{c}{\textit{GBFS +} $h_{add}$} \\[.2ex]
-\hline\\[-1.3ex]
-37921  &34627 &9010	&8762 &34849 \\
-\hline
-%\hline
-\end{tabular}
-}
-
-\end{center}
-
-\caption{\small 
-\emph{Top:} Number of classical planning problems  and percentages solved by IW(1), IW(2), or neither. Instances obtained
-by splitting benchmarks  with $N$ atomic goals into $N$ problems with atomic goals.
-\emph{Bottom:} Number of instances solved by \IW\ in comparison with  Iterative Deepening (\ID),  
-Breadth-First Search (\BRFS),   and Greedy Best First Search (\textit{GBFS}) guided by  additive heuristic.
-Time and  memory outs  after 30 minutes and 2 GB resp. From \cite{nir:ecai2102}.
-}
-\label{table1}
-\end{table}
-%%%%
-%%%
-
-\subsection{Performance and Width}
-
-IW is a systematic and complete blind-search algorithm like breadth-first search (BrFS) and iterative deepening (ID),
-but   unlike these algorithms, it uses the factored representation of the states in terms of variables to structure the 
-search in a different way. This structured exploration  has proved to be very effective in classical planning 
-over most of the benchmark domains  when the goals are restricted to be single atoms. Table~\ref{table1} from \cite{nir:ecai2012} shows 
-the percentage of instances that are  solved by  the first iteration of IW, i.e.  IW(1), by the second iteration 
-IW(2), and by neither one.  These are instances that  have been obtained from the existing benchmarks
-by splitting problems with $N$ atomic goals, into $N$ problems with one atomic goal each. As the table shows, 
-$37\%$ of the 37921 instances are solved by IW(1) while $51.3\%$ are  solved by IW(2). Since IW($k$)
-runs in time that is exponential in $k$, this mean that almost $90\%$ of the 37921 instances are  solved in time that
-is either linear or quadratic in the number of problem variables, which in these encodings are all \emph{boolean}.  
-Furthermore, when the performance of \IW\   is compared with  breadth-first search and iterative deepening, 
-on the one hand,  and  with a  Greedy Best First Search guided by the additive heuritstic $h_{add}$
-on the other (this algorithm is similar to one used by  the FF planner when hill climbing search fails \cite{hoffmann:ff}), 
-it turns out that ``blind'' \IW\ solves as many problems as the informed search, 34627 vs. 34849, far ahead
-of the other blind algorithms BrFS and ID  that solve 9010 and 8762 problems each. This is shown in the bottom
-part of the Table~\ref{table1}. Moreover, \IW\ is  faster and results in shorter plans than the heuristic search  
-\cite{nir:ecai2012}.
-
-The min $k$ value for which $IW(k)$ solves an instance  is clearly bounded and small in most of
-these instances. This is  actually no  accident and  has a \emph{theoretical explanation.} Lipovetzky and Geffner define
-a structural parameter called the problem \emph{width} and show that for many of these domains,
-\emph{any} solvable instance with atomic goals will  have a bounded and small width that is independent of the number
-of variables and states in the problem. The min value $k$ for which the iteration  $IW(k)$ solves
-the problem cannot exceed the problem width, so that the algorithm \emph{IW will run in time and space
-that are exponential in the problem width.}
-
-Formally, the \emph{width} $w(P)$ of a problem $P$ is $i$
-iff $i$ is the minimum positive integer for which 
-there is a sequence of tuples (sets) with at most $i$ atoms each, $t_0, t_1, \ldots, t_n$
-such that 1)~$t_0$ is true in the initial state of $P$, 2)~any shortest plan $\pi$  that achieves $t_i$ in $P$
-can be extended into a shortest plan that achieves $t_{i+1}$ by extending   $\pi$ with one action, and 
-3)~any shortest plan that achieves $t_n$ is also a shortest plan for $P$.
-
-One way to understand this definition is that the problem  width $w(P)$ is at most  $i$ if  there is a 
-``trail of  stepping stones''  $t_0, t_1, \ldots, t_n$ to reach the problem goal such that 
-A)~these ``stepping stones''  contain at most $i$ atoms each, B)~each  stepping stone $t_{i+1}$ is 
-at distance 1 from the previous one $t_i$ when $t_i$ has been reached following the ``trail'', 
-C)~the ``trail''  preserves ``optimality''; i.e., no  tuple $t_i$ can be reached in less than $i$ steps.
-
-While this notion of width and the iterated  width algorithms that are based on it  have been designed
-for problems where a goal state needs to be reached, the notions remain  relevant  in  optimization problems as well.
-Indeed, if a  good path is made of states $s_i$ each of which have a low width, we would like 
-IW to find such path in low polynomial time for a low value of the $k$ parameter. 
-Later on we will discuss the changes required in IW to enforce this property, which
-are not yet implemented in the code that is evaluated below.
-
-\section{Width-Based Algorithms for the Atari Games}
-
-Explain: IW(1), 2BFS, and possible IW11. Also, Dijkstra postprocessing over IW(1) graph
-if ran.
-
-The number of nodes \emph{generated} by IW(1) is $n \times D \times b$ in the worst case, 
-where $n$ is the number of variables, $D$ is the size of their domains, and $b$ is maximum the number of 
-actions applicable per  state.  This same number in breadth-first search is not linear in $n$
-but exponential. For the Atari games,  $n=128$, $D=256$, and $b=18$, so that the product is
-equal to $589,824$, which is large but feasible. On the other hand, the number
-of nodes generated by IW(2) in the worst case is  $(n \times D)^2 \times b$, which is
-equal to $1,073,741,824$ which is too large, forcing us to consider only a fraction of 
-such states. For classical planning problems, the growth in the number of nodes from 
-IW(1) to IW(2) is not that large, as the variables are boolean. Indeed, we could have
-taken the state vector for the Atari games as a vector of 1024 boolean variables, 
-and apply these algorithms to that representation. The results would be different.
-In such a case, IW(1) would generate $n' \times D' \times b$ states, and 
-IW(2), $(n' \times D') \times b$ states, which for $n'=1024$ and $D=2$
-represent  $36,864$ and $7,5497,472$ states respectively.
-These are both feasible numbers, but we haven't tried yet this representation,
-under the assumption that the correlations among the bits in each one of 
-the 128 words of the state vectors are meaningful. 
-In summary, from the basic IW algorithm we are testing only the first 
-call, IW(1).
-
-IW is a purely exploration algorithm that does not take into account  the accumulated 
-reward for selecting the next states to consider. As a simple, variant that combines
-exploration and exploitation, we evaluated a \emph{best-first search} algorithm
-with two queues: one queue  ordered first by  novelty measure 
-(recall, novelty 1 means that the state is the first one to make some atom true),
-and a second queue ordered by accumulated reward. In one iteration, the best first
-search picks up the best node from one queue, and in the second iteration it picks
-up the best node from the other queue. This way for combining multiple heuristics
-is used in the LAMA planner \cite{richter:lama}, and was introduced in the
-planner Fast Downward \cite{malte:fd,malte:alternating}. In addition, we break ties
-in the first queue favoring states with largest accumulated reward, and 
-in the second queue, favoring states with smallest novelty measure. Last, when
-a node is expanded, it is removed from the queue, and its children are placed on 
-both queues. The exception are the nodes with novelty 1 measure that are only
-placed in first queue (** check and revise; but keep accurate and simple **).
-We refer to this best-first algorithm below, as 2BFS.
-
-IW11?
-
-Randomization? Discount?
-
-\section{Experimental Results in the Atari Games}
-
-\input{performance_table}
-
 The algorithms IW1 and BFS are tested over 9 games selected from those
@@
  discount, randomization, ..)
-
-\section{Exploration and Exploitation: Discussion}
-
-The notion of width underlying the iterated width algorithm was developed in the context
-of classical planning in order to understand why most of the hundreds of existing benchmarks
-resulting from planning competitions appear to be relatively simple for current planners, even
-though classical planning is PSPACE-complete \cite{bylander}. The partial answer is that 
-most of these domains have a low width, and hence, can be solved in low polynomial time (by IW),
-when goals contain a single atom. At the same time, most problems with multiple atomic goals are 
-also easy too, as the goals can often be achieved one at a time \cite{nir:ecai2012}.
-
-In the Iterated Width algorithm, the key notion is the \emph{novelty measure} of a state in the underlying
-breadth-first search. In particular, a state gets a novelty 1 measure in the search if it makes true,
-some atom that up to that point had been false in all previously generated states. In the procedure IW(1)
-all other states are pruned right away, as if they were ``duplicate'' states. The algorithms IW and IW(k)
-are pure  exploration algorithms that make no use at all of the information about the goal, if there is a goal,
-or about the  reward function to optimize otherwise. They are  thus  pure exploration algorithms guided by the novelty
-measures, which use the structure of the states to provide a  structure to the exploration of the state space,
-which in the case of IW is guaranteed to be complete.
-
-The use of ``novelty measures'' for guiding an optimization search while ignoring the function that is being
-optimized, appears to be  common to novelty-based search methods developed independently 
-in the context of genetic algorithms \cite{lehman:novelty}. %%% bibtex for this below
-In these methods individuals in the population are not ranked according to optimization function but in terms of how 
-``novel'' thay are in relation to the rest of the population, thus encouraging diversity
-and exploration, rather than (greedy) exploitation. The actual definition of novelty
-in such a case is domain-dependent; for example, in the evolution of a controller
-for guiding a robot in a maze, an individual controller  will not be ranked by how
-close it takes the robot to the goal (the greedy measure), but how far it takes
-the robot in relation to the locations that are reached by following the other controllers
-(a diversity measure). The novelty measure used by IW, on the other hand, 
-is domain-independent and it follows from the structure of the state
-as captured by the problem variables.
-
-% @article{lehman:novelty,
-%  title={Abandoning objectives: Evolution through the search for novelty alone},
-%  author={Lehman, Joel and Stanley, Kenneth O},
-%  journal={Evolutionary computation},
-%  volume={19},
-%  number={2},
-%  pages={189--223},
-%  year={2011},
-%  publisher={MIT Press}
-%}
-
-The balance between exploration and exploitation has received considerable attention in
-reinforcement learning \cite{sutton:book}, where both are required for converging to an
-optimal behavior. In the Atari games, as in other deterministic problems, however, ``exploration'' 
-is not needed for optimality purposes, but just for improving the effectiveness of the 
-local lookahead searches. Indeed, a best-first search algorithm guided by (discounted)
-accumulated reward will deliver eventually  best moves, but it will not be as effective
-over small time windows, where like breadth-first search it's likely not to find any rewards
-at all. The UCT algorithm provides a method for balancing exploration and exploitation,
-which can thus be effective over small time windows. The 2BFS algorithm above with two queues
-that alternate, one guided by the novelty measures, and the other by the accumulated reward,
-provides a different scheme. The first converges to the optimal behavior asymptotically; the 
-second in a bounded number of steps. Alternately, novelty-guided probes, possibly with 
-some noise added, could potentially pay off as  a  base policy for UCT in replacement 
-of random rollouts, as they manage apparently to exploit the structure of many problems.
-
-\section{Summary and On-going Work}
-
-We have shown experimentally that width-based algorithms like IW(1) and 2BFS
-that originate in work in classical planning, can be used in the Atari games
-with a performance that approaches the state-of-the-art as captured by UCT 
-planning, sometimes using a fraction of the time. We are currently addressing three issues, 
-which we haven't managed to get in time for this submission. First, we want to extend the evaluation to 
-all the games featured in \cite{bellemare:jair}, and to add more runs per  game. 
-Second, we want to evaluate the algorithm for different fixed time windows. In the table above, for example, UCT
-is often taking an order-of-magnitude more time per decision as 2BFS, so the 
-comparison in not fair. Third, we need to revise the way duplicate states are handled 
-in IW and 2BFS to ensure that the best paths among the paths defined by the states that have been expanded
-are returned. This change is important from a theoretical point of view; it's not clear
-what it's practical impact will be.
-
-%%% - Learning in same instance, across diff instances of same game, etc.
-%%%% - Learning controllers vs. on-line planning, etc
-
-\end{document}

### 2014-10-31 workshop-paper/paper.tex
Bibliography and Experimental Section separated

--- 2014-10-30 workshop-paper/paper.tex
@@
 
+%% Macros
 \newcommand{\tuple}[1]{{\langle #1\rangle}}
@@
 \newcommand{\BRFS}{{\textit{BrFS}}}
+\newcommand{\BFS}{{\textit{2BFS}}}
 
@@
 
-\begin{table*}
-%% \addtolength{\tabcolsep}{-2pt}
-{\scriptsize
-\begin{tabular}{@{}l|rr|rrr|rrr|rrr|rrr@{}}
-%% table header
-\hline\\[-3ex]
-\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Random} & \multicolumn{3}{c|}{BrFS}& \multicolumn{3}{c|}{IW1}& \multicolumn{3}{c|}{BFS}& \multicolumn{3}{c}{UCT} \\[.2ex]
-\hline\\[-3ex]
-%% table body
-\multicolumn{1}{c}{Game}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c|}{Dev}  & \multicolumn{1}{c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev}  & \multicolumn{1}{|c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev}  & \multicolumn{1}{|c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev}  & \multicolumn{1}{|c}{T} & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{Dev} \\[.2ex]
-\hline\\[-3ex]
-asterix         & 210   & 54.77     & 32.24 & 3900  & 5538.95   & 17.08 & 119000        & 4987.48   & 0.95  & 56320       & 4757.84   & 10.82 & \textbf{296300} & 6657.7    \\
-beam\_rider     & 448.8 & 84.64     & 43.48 & 519.2 & 179.81    & 6.02  & 3000          & 1150.22   & 0.92  & 3804        & 446.18    & 42.91 & \textbf{4320}   & 2175.26   \\
-breakout        & 1     & 0.71      & 51.56 & 2     & 1.41      & 1.37  & 383.6         & 9.84      & 3.95  & 421.2       & 13.66     & 26.03 & \textbf{512.4}  & 193.97    \\
-enduro          & 0     & 0         & 40.92 & 0.4   & 0.89      & 38.13 & \textbf{500}  & 0         & 13.12 & 292.8       & 6.18      & 32.14 & 306.8           & 8.14      \\
-freeway         & 0     & 0         & 71.84 & 0     & 0         & 15.19 & \textbf{29.4} & 0.89      & 4.44  & 19.6        & 3.29      & 90.64 & 1               & 1.41      \\
-pong            & -20.4 & 0.55      & 41.51 & -20.2 & 0.84      & 18.45 & 16.4          & 1.67      & 39.69 & \textbf{21} & 0         & 38.37 & \textbf{21}     & 0         \\
-qbert           & 185   & 345.33    & 42.83 & 160   & 85.88     & 1.82  & 1545          & 1666.06   & 1.76  & 14725       & 7872.1    & 28.61 & \textbf{18730}  & 1127.69   \\
-seaquest        & 192   & 38.99     & 40.01 & 200   & 14.14     & 14.72 & \textbf{4274} & 4206.81   & 1.59  & 4070        & 1939.55   & 26.16 & 3406            & 2402.47   \\
-space\_invaders & 187   & 120.86    & 46.11 & 115   & 61.44     & 12.1  & \textbf{2670} & 637.3     & 2.25  & 1327        & 598.04    & 19.66 & 2634            & 375.66   
-\end{tabular}
-}
+\input{performance_table}
 
- \caption{\small
-   Random, BrFS, IW1, BFS, UCT performance over 5 runs, with
-   seeds 0 to 4.  \textit{T} is the action selection average
-   computation time, \textit{S} is the average score, and
-   \textit{Dev} is the standard deviation.  Numbers in bold show best performer in terms of average score.}
-\label{performance}
-
-\end{table*}
-
-** Results, cold and detached, same format as those reported in jair paper, possibly with
-std deviations and times.
-
-** mention also ideas that have been tried and didn't work (IW2?, discount, randomization, ..)
+\input{experimental_results}
 
@@
 
+\bibliographystyle{aaai}
+\bibliography{control}
+
 \end{document}

### 2014-10-31 workshop-paper/paper.tex
algorithms changed

@@
 
-Explain: IW(1), 2BFS, and possible IW11. Also, Dijkstra postprocessing over IW(1) graph
-if ran.
+%%Explain: IW(1), 2BFS, and possible IW11. Also, Dijkstra postprocessing over IW(1) graph if ran.
 
@@
 In such a case, IW(1) would generate $n' \times D' \times b$ states, and 
-IW(2), $(n' \times D') \times b$ states, which for $n'=1024$ and $D=2$
+IW(2), $(n' \times D')^2 \times b$ states, which for $n'=1024$ and $D=2$
 represent  $36,864$ and $7,5497,472$ states respectively.
@@
 a node is expanded, it is removed from the queue, and its children are placed on 
-both queues. The exception are the nodes with novelty 1 measure that are only
-placed in first queue (** check and revise; but keep accurate and simple **).
+both queues. The exception are the nodes with accumulated reward 0 that are only
+placed in the first queue.
 We refer to this best-first algorithm below, as 2BFS.
 
-IW11?
+In the experiments below, we use IW(1) with two simple variations.
+First, in the breadth-first search, we generate the children in 
+random order. This makes the execution that results from the
+IW(1) lookahead less susceptible to fall into loops; a potential
+problem in local search algorithms with no memory or learning.
+Second, a discount factor $\gamma=0.995$ is added
+for discounting future rewards like in UCT. For this,
+each state $s$ keeps its depth d(s) in the IW(1) tree,
+and if state s' is the child of state s and action a, 
+$R(s')$ is set to $R(s) + \gamma^d(s´) r(a,s)$.
+A discount factor less than 1 results in a slight
+preference for rewards that can be reached earlier, 
+which is reasonable  in the on-line setting.
 
-Randomization? Discount?
+%% NIR: IW11 Don't need to be explained. Maybe in Future Work?
 

### 2014-10-31 workshop-paper/experimental_results.tex
Pass on Experimental Section

@@
-The algorithms IW1 and BFS are tested over 9 games selected from those
-considered in the training set in \cite{bellemare:jair}, and those
-considered in \cite{deep-mind-atari}. Each game has 5 runs, each with
-a different random seed $0,\ldots,4$.  We compare the performance of
-IW1 and BFS with UCT and BRFS. We also include results for a baseline random
-agent. We use the same parameters in the literature, all algorithms use a discount factor of 0.999, and each action
-simulates 5 frames. Each algorithm has a budget on the number of nodes
-it can generate per lookahead. BRFS, IW1 and BFS have a maximum budget
-of 100k. For UCT we use the same parameters as in
-\cite{bellemare:jair}, 0.1 exploration rate, and 500 rollouts of a maximum depth of 300 steps,
-which gives a maximum budget of 150k nodes. We tested UCT with a fix
-budget of 100k and max depth 300, but resulted in practically the same
-results. All algorithms reuse the tree built between lookaheads by
-setting the new root of the tree to the node of the selected action,
-and deleting the other branches. Note that this may have a different
-effect on algorithms were all branches have the same number of
-generated nodes, ex. BRFS, than those that may focus on different
-areas of the tree, ex. IW1, BFS, UCT.
+We have tested the $\IW(1)$ and \BFS~described in the previous Section
+over $9$ games selected amongst those considered in the \emph{training} set
+of \cite{bellemare:jair2013}, and those discussed in \cite{deep-mind-atari}. 
+In Table~\ref{performance} the performance of both $\IW(1)$ 
+and \BFS, characterized by the average score obtained over all episodes, the standard deviation
+of the score obtained and time (in seconds) per lookahead, is compared with that of \UCT, 
+\BRFS~and the baseline random action selection. Our experimental 
+setup  follows that of \cite{bellemare:jair2013}, with two exceptions. First, rather than
+considering $10$ episodes of each game, we have considered just $5$
+due to time constraints, as each game can take as much as $40$ hours of 
+CPU time. Second, for \UCT, rather than limiting the number of
+base policy rollouts to $500$, we have rather limited the number of frames
+to be simulated during each lookahead to $100,000$ as \cite{bellemare:jair2013} 
+does for \BRFS~and we do for both $\IW(1)$ and \BFS. Other \UCT~parameters such as the exploration constant $C$ and
+the planning horizon $m$ have not been modified ($C=0.1$, $m=1500$). We include
+the average score reported for \UCT~by \cite{bellemare:jair2013} in Table~\ref{performance}
+bracketed with parenthesis for reference. Also unchanged is the discount factor
+used by all the algorithms ($\gamma=0.995$) and the rate at which an algorithm is
+invoked to select an action (every $5$ frames~\footnote{We note that with this setup, the maximum depth $d$
+of the lookahead will be $m/5$, or $300$ nodes.}). All algorithms reuse the tree built between each 
+lookahead by setting the root of the tree to the best child of the root in the previous one,
+and deleting its siblings and their descendants, as \cite{bellemare:jair2013} does for \UCT. We note that 
+this may have a different effect on algorithms where all branches have the same number of
+generated nodes, as is the case of \BRFS, than in $\IW(1)$, \BFS~and \UCT, where the depth of
+the branches is uneven.
 
-On Table~\ref{performance}, we can observe the performance of each
-algorithm on terms of average score with standard deviation, and
-average time per lookahead in seconds. The Random agent don't include
-average time as it just chooses an action randomly. BRFS outperform
-the Random agent in 7 out of 9 domains. BRFS rarely finds any reward,
-turning itself mostly into a random agent, but exploring 100k nodes up
-to an average depth of 4. When BRFS is turned int IW1 using novelty to
-explore the search space, the jump on performance is important. IW1 in
-most of the cases does not completely use its budget, and is able to
-achieve depths of 70 to 260 depending on the game, which first
-explains how IW1 is able to achieve a performance that is closer to
-UCT than to BRFS. In 4 out of 9 tested games it is the best
-performer. Remarkably, is the first domain independent solver that
-shows good performance on Freeway, where rewards are very sparse and
-deep in the search tree. BFS outperforms IW1 and UCT in 4 and 2
-domains respectively.  Both IW1 and BFS performance in terms of score
-is close to UCT, but they are some times one order of magnitud
-faster. The speedup is due to the bounded space in IW1, and to the
-fact that in some games, both IW1 and BFS are able to reuse an
-important part of their trees between lookaheads.
+Table~\ref{performance} shows that both $\IW(1)$ and \BFS~outperform by a wide margin 
+\BRFS, which sometimes performs worse than the baseline, both in terms of average score attained and 
+average time per call. \BRFS~rarely collects any reward to inform the choice of action at
+the root node of the lookahead, as the depth of the \BRFS~search tree, having limited the number of
+frames allowed to be simulated to $100,000$, corresponds to a lookahead of $0.3$ seconds ($20$ 
+frames or $4$ nodes). Both $\IW(1)$ and \BFS~reach states up to $350$--$1500$ frames ahead 
+($70$--$260$ nodes or $6$--$22$ seconds), and that results in collecting enough rewards so as to
+inform the choice at the root in a way that much higher scores are attained. Interestingly, very often
+neither $\IW(1)$ or \BFS~fully consume the number of simulation frames allocated, since many
+states are pruned, in the case of $\IW(1)$, or relegated to the back of the open lists, in the 
+case of \BFS, due to having a novelty of $0$ and the planning horizon being limited to $m$.
 
-The novelty of states in IW1 and BFS may depend on the order that
-nodes are generated. Therefore, every time a node is expanded, we
-randomized the order in which nodes get genreated. Randomizing the
-order of expansion is also applied to BRFS, but has no impact on its
-performance.
+We observe that in $4$ out of $9$ games, $\IW(1)$ outperforms \UCT~limited to the same number
+of simulated frames, both in terms of average score attained and average time per call. \BFS~is
+generally faster than $\IW(1)$, but only outperforms \UCT~in $2$ out of $9$ games (\Freeway and
+\Seaquest) and $\IW(1)$ in $4$ out of $9$ games. Compared with the \UCT~results reported
+in \cite{bellemare:jair2013}, that by limiting the number of rollouts allow a $50$\% more frames
+to be simulated in each lookahead, $\IW(1)$ still outperforms \UCT~in $3$ out of $9$ games
+(\Breakout, \Enduro and \Freeway). As introduced in the previous Section, the order in which
+$\IW(1)$ and \BFS~consider actions is randomized, since the novelty of a given state may
+change depending on the order in which states are considered. We found this randomization to
+have a significant effect in (\textbf{TODO}: What game? \Qbert?) where (\textbf{TODO}: What was
+going on?). Last, we note that $\IW(1)$ and \BFS~attain similar performance as the baseline
+semi-random algorithm \emph{Perturb} from \cite{bellemare:jair2013}, which beated both \UCT~
+and the reinforcement learning algorithms reported.
 
-IW2 is not included in the table due to its low performance, which is
-closer to BRFS than to the other algorithms. This is mainly due to the
-huge number of states that are novel in this algorith, thus,
-achieveing insufficient depths to gather some reward.
+We finish the discussion of the results presented in Table~\ref{performance} by observing
+that $\IW(1)$ outperforms the average scores reported for the best algorithm reported in
+\cite{deep-mind-atari} but in $4$ out of $7$ games, and \BFS~does so too in $5$ out of $7$
+games.
 
- ** mention also ideas that have been tried and didn't work (IW2?,
- discount, randomization, ..)
+We do not report $IW(2)$ in Table~\ref{performance} since we found its performance to
+be very similar to that of \BRFS~in some preliminary tests. When considering $2$--tuples
+of state variables to evaluate the novelty of states, it tended to result in lookaheads
+$5$ times deeper than those resulting from applying \BRFS~, but still far too shallow so
+as to compete with the other planning algorithms considered.
+
+% On Table~\ref{performance}, we can observe the performance of each
+% algorithm on terms of average score with standard deviation, and
+% average time per lookahead in seconds. The Random agent don't include
+% average time as it just chooses an action randomly. BRFS outperform
+% the Random agent in 7 out of 9 domains. BRFS rarely finds any reward,
+% turning itself mostly into a random agent, but exploring 100k nodes up
+% to an average depth of 4. When BRFS is turned int IW1 using novelty to
+% explore the search space, the jump on performance is important. IW1 in
+% most of the cases does not completely use its budget, and is able to
+% achieve depths of 70 to 260 depending on the game, which first
+% explains how IW1 is able to achieve a performance that is closer to
+% UCT than to BRFS. In 4 out of 9 tested games it is the best
+% performer. Remarkably, is the first domain independent solver that
+% shows good performance on Freeway, where rewards are very sparse and
+% deep in the search tree. BFS outperforms IW1 and UCT in 4 and 2
+% domains respectively.  Both IW1 and BFS performance in terms of score
+% is close to UCT, but they are some times one order of magnitud
+% faster. The speedup is due to the bounded space in IW1, and to the
+% fact that in some games, both IW1 and BFS are able to reuse an
+% important part of their trees between lookaheads.
+% 
+% The novelty of states in IW1 and BFS may depend on the order that
+% nodes are generated. Therefore, every time a node is expanded, we
+% randomized the order in which nodes get genreated. Randomizing the
+% order of expansion is also applied to BRFS, but has no impact on its
+% performance.
+% 
+% IW2 is not included in the table due to its low performance, which is
+% closer to BRFS than to the other algorithms. This is mainly due to the
+% huge number of states that are novel in this algorith, thus,
+% achieveing insufficient depths to gather some reward.
+
+%  ** mention also ideas that have been tried and didn't work (IW2?,
+%  discount, randomization, ..)
+%
+
+% The algorithms IW1 and BFS are tested over 9 games selected from those
+% considered in the training set in \cite{bellemare:jair}, and those
+% considered in \cite{deep-mind-atari}. Each game has 5 runs, each with
+% a different random seed $0,\ldots,4$.  We compare the performance of
+% IW1 and BFS with UCT and BRFS. We also include results for a baseline random
+% agent. We use the same parameters in the literature, all algorithms use a discount factor of 0.999, and each action
+% simulates 5 frames. Each algorithm has a budget on the number of nodes
+% it can generate per lookahead. BRFS, IW1 and BFS have a maximum budget
+% of 100k. For UCT we use the same parameters as in
+% \cite{bellemare:jair}, 0.1 exploration rate, and 500 rollouts of a maximum depth of 300 steps,
+% which gives a maximum budget of 150k nodes. We tested UCT with a fix
+% budget of 100k and max depth 300, but resulted in practically the same
+% results. All algorithms reuse the tree built between lookaheads by
+% setting the new root of the tree to the node of the selected action,
+% and deleting the other branches. Note that this may have a different
+% effect on algorithms were all branches have the same number of
+% generated nodes, ex. BRFS, than those that may focus on different
+% areas of the tree, ex. IW1, BFS, UCT.
+
+% On Table~\ref{performance}, we can observe the performance of each
+% algorithm on terms of average score with standard deviation, and
+% average time per lookahead in seconds. The Random agent don't include
+% average time as it just chooses an action randomly. BRFS outperform
+% the Random agent in 7 out of 9 domains. BRFS rarely finds any reward,
+% turning itself mostly into a random agent, but exploring 100k nodes up
+% to an average depth of 4. When BRFS is turned int IW1 using novelty to
+% explore the search space, the jump on performance is important. IW1 in
+% most of the cases does not completely use its budget, and is able to
+% achieve depths of 70 to 260 depending on the game, which first
+% explains how IW1 is able to achieve a performance that is closer to
+% UCT than to BRFS. In 4 out of 9 tested games it is the best
+% performer. Remarkably, is the first domain independent solver that
+% shows good performance on Freeway, where rewards are very sparse and
+% deep in the search tree. BFS outperforms IW1 and UCT in 4 and 2
+% domains respectively.  Both IW1 and BFS performance in terms of score
+% is close to UCT, but they are some times one order of magnitud
+% faster. The speedup is due to the bounded space in IW1, and to the
+% fact that in some games, both IW1 and BFS are able to reuse an
+% important part of their trees between lookaheads.

### 2014-10-31 workshop-paper/paper.tex
Pass on Experimental Section
bib

@@
 Breadth-First Search (\BRFS),   and Greedy Best First Search (\textit{GBFS}) guided by  additive heuristic.
-Time and  memory outs  after 30 minutes and 2 GB resp. From \cite{nir:ecai2102}.
+Time and  memory outs  after 30 minutes and 2 GB resp. From \cite{nir:ecai2012}.
 }
@@
 is used in the LAMA planner \cite{richter:lama}, and was introduced in the
-planner Fast Downward \cite{malte:fd,malte:alternating}. In addition, we break ties
+planner Fast Downward \cite{helmert:fd,malte:alternating}. In addition, we break ties
 in the first queue favoring states with largest accumulated reward, and 
@@
 resulting from planning competitions appear to be relatively simple for current planners, even
-though classical planning is PSPACE-complete \cite{bylander}. The partial answer is that 
+though classical planning is PSPACE-complete \cite{bylander:complexity}. The partial answer is that 
 most of these domains have a low width, and hence, can be solved in low polynomial time (by IW),
@@
 which we haven't managed to get in time for this submission. First, we want to extend the evaluation to 
-all the games featured in \cite{bellemare:jair}, and to add more runs per  game. 
+all the games featured in \cite{bellemare:jair2013}, and to add more runs per  game. 
 Second, we want to evaluate the algorithm for different fixed time windows. In the table above, for example, UCT

### 2014-10-31 workshop-paper/performance_table.tex

--- 2014-10-31 workshop-paper/performance_table.tex
@@
 {\scriptsize
-\begin{tabular}{@{}l|rr|rrr|rrr|rrr|rrr@{}}
+\begin{tabular}{@{}l|rr|rrr|rrr|rrr|rrr|c@{}}
 %% table header
 \hline\\[-3ex]
-\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Random} & \multicolumn{3}{c|}{BrFS}& \multicolumn{3}{c|}{IW1}& \multicolumn{3}{c|}{BFS}& \multicolumn{3}{c}{UCT} \\[.2ex]
+\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Random} & \multicolumn{3}{c|}{\BRFS}& \multicolumn{3}{c|}{$\IW(1)$}& \multicolumn{3}{c|}{\BFS}& \multicolumn{3}{c|}{\UCT}
+& \multicolumn{1}{c}{\Deepmind} \\[.2ex]
 \hline\\[-3ex]
 %% table body
-\multicolumn{1}{c}{Game}  & \multicolumn{1}{c}{Avg.
-Score} & \multicolumn{1}{c|}{Std Dev}  & \multicolumn{1}{c}{Time} & \multicolumn{1}{c}{Avg.
-Score} & \multicolumn{1}{c}{Std Dev} & \multicolumn{1}{|c}{Time} & \multicolumn{1}{c}{Avg.
-Score} & \multicolumn{1}{c}{Std Dev}  & \multicolumn{1}{|c}{Time} & \multicolumn{1}{c}{Avg.
-Score} & \multicolumn{1}{c}{Std Dev}  & \multicolumn{1}{|c}{Time} & \multicolumn{1}{c}{Avg.
-Score} & \multicolumn{1}{c}{Std Dev} \\[.2ex]
+\multicolumn{1}{c}{Game}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c|}{SD}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} &  \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T} & \multicolumn{1}{c}{S}\\[.2ex]
 \hline\\[-3ex]
-asterix         & 210   & 54.77     & 32.24 & 3900  & 5538.95   & 17.08 & 119000        & 4987.48   & 0.95  & 56320       & 4757.84   & 10.82 & \textbf{296300} & 6657.7    \\
-beam\_rider     & 448.8 & 84.64     & 43.48 & 519.2 & 179.81    & 6.02  & 3000          & 1150.22   & 0.92  & 3804        & 446.18    & 42.91 & \textbf{4320}   & 2175.26   \\
-breakout        & 1     & 0.71      & 51.56 & 2     & 1.41      & 1.37  & 383.6         & 9.84      & 3.95  & 421.2       & 13.66     & 26.03 & \textbf{512.4}  & 193.97    \\
-enduro          & 0     & 0         & 40.92 & 0.4   & 0.89      & 38.13 & \textbf{500}  & 0         & 13.12 & 292.8       & 6.18      & 32.14 & 306.8           & 8.14      \\
-freeway         & 0     & 0         & 71.84 & 0     & 0         & 15.19 & \textbf{29.4} & 0.89      & 4.44  & 19.6        & 3.29      & 90.64 & 1               & 1.41      \\
-pong            & -20.4 & 0.55      & 41.51 & -20.2 & 0.84      & 18.45 & 16.4          & 1.67      & 39.69 & \textbf{21} & 0         & 38.37 & \textbf{21}     & 0         \\
-qbert           & 185   & 345.33    & 42.83 & 160   & 85.88     & 1.82  & 1545          & 1666.06   & 1.76  & 14725       & 7872.1    & 28.61 & \textbf{18730}  & 1127.69   \\
-seaquest        & 192   & 38.99     & 40.01 & 200   & 14.14     & 14.72 & \textbf{4274} & 4206.81   & 1.59  & 4070        & 1939.55   & 26.16 & 3406            & 2402.47   \\
-space\_invaders & 187   & 120.86    & 46.11 & 115   & 61.44     & 12.1  & \textbf{2670} & 637.3     & 2.25  & 1327        & 598.04    & 19.66 & 2634            & 375.66   
+\Asterix        & 210    & 54.77     & 3900  & 5538.95   & 32.24 & 119000        & 4987.48   & 17.08 & 56320       & 4757.84   & 0.95  & \textbf{296300 (290700)} & 6657.7    & 10.82 & n/a\\
+\BeamRider     & 448.8  & 84.64     & 519.2 & 179.81    & 43.48 & 3000          & 1150.22   & 6.02  & 3804        & 446.18    & 0.92  & \textbf{4320 (6624)}     & 2175.26   & 42.91 & 4092\\
+\Breakout       & 1      & 0.71      & 2     & 1.41      & 51.56 & 383.6         & 9.84      & 1.37  & 421.2       & 13.66     & 3.95  & \textbf{512.4 (364)}     & 193.97    & 26.03 & 168\\
+\Enduro          & 0      & 0         & 0.4   & 0.89      & 40.92 & \textbf{500}  & 0         & 38.13 & 292.8       & 6.18      & 13.12 & 306.8 (286)              & 8.14      & 32.14 & 470\\
+\Freeway         & 0      & 0         & 0     & 0         & 71.84 & \textbf{29.4} & 0.89      & 15.19 & 19.6        & 3.29      & 4.44  & 1 (0)                    & 1.41      & 90.64 & n/a \\
+\Pong           & -20.4  & 0.55      & -20.2 & 0.84      & 41.51 & 16.4          & 1.67      & 18.45 & \textbf{21} & 0         & 39.69 & \textbf{21 (21)}         & 0         & 38.37 & 20 \\
+\Qbert           & 185    & 345.33    & 160   & 85.88     & 42.83 & 1545          & 1666.06   & 1.82  & 14725       & 7872.1    & 1.76  & \textbf{18730 (17343)}   & 1127.69   & 28.61 & 1952\\
+\Seaquest       & 192    & 38.99     & 200   & 14.14     & 40.01 & \textbf{4274} & 4206.81   & 14.72 & 4070        & 1939.55   & 1.59  & 3406 (5132)              & 2402.47   & 26.16 & 1705 \\
+\SpaceInvaders & 187    & 120.86    & 115   & 61.44     & 46.11 & \textbf{2670} & 637.3     & 12.1  & 1327        & 598.04    & 2.25  & 2634 (2718)              & 375.66    & 19.66 & 581
 \end{tabular}
@@
 
- \caption{\small
-   Random, BrFS, IW1, BFS, UCT performance over 5 runs, with
-   seeds 0 to 4. \textit{Time} is the action selection average
-   computation time in seconds, \textit{Avg. Score} is the average score, and
-   \textit{Std Dev} is the standard deviation.  Numbers in bold show best performer in terms of average score.}
+ \caption{\small Random baseline, \BRFS, $\IW(1)$, \BFS, \UCT~performance over 5 runs,
+   setting the Atari 2600 random seeds to $0, \ldots, 4$, and maximum episode duration of $18,000$
+   frames. \textit{T} is the average CPU T per call to each action selection, \textit{S} is the average score achieved,
+   and \textit{SD} is the standard deviation. \UCT~performance
+   reported in \cite{bellemare:jair2013} is shown in parenthesis. Numbers
+   in bold show best performer in terms of average score. Average score reported in 
+   \cite{deep-mind-atari} is reported for reference under the column \Deepmind.
+   }
 \label{performance}

### 2014-10-31 workshop-paper/experimental_results.tex
Second pass on the Experimental Section

@@
 due to time constraints, as each game can take as much as $40$ hours of 
-CPU time. Second, for \UCT, rather than limiting the number of
-base policy rollouts to $500$, we have rather limited the number of frames
-to be simulated during each lookahead to $100,000$ as \cite{bellemare:jair2013} 
-does for \BRFS~and we do for both $\IW(1)$ and \BFS. Other \UCT~parameters such as the exploration constant $C$ and
-the planning horizon $m$ have not been modified ($C=0.1$, $m=1500$). We include
+CPU time. Second, while \UCT~is limited to perform up to $500$ base policy rollouts ($150,000$ frames
+are to be simulated while performing the lookahead), both $\IW(1)$ and \BFS are limited to
+simulate $100,000$ frames. \UCT~parameters such as the exploration constant $C$ and
+the planning horizon $m$ have not been modified ($C=0.1$, $m=300$), the latter being also
+used by both $\IW(1)$ and \BFS~to limit the depth of the lookahead. We include
 the average score reported for \UCT~by \cite{bellemare:jair2013} in Table~\ref{performance}
@@
 used by all the algorithms ($\gamma=0.995$) and the rate at which an algorithm is
-invoked to select an action (every $5$ frames~\footnote{We note that with this setup, the maximum depth $d$
-of the lookahead will be $m/5$, or $300$ nodes.}). All algorithms reuse the tree built between each 
+invoked to select an action (every $5$ frames). All algorithms reuse the tree built between each 
 lookahead by setting the root of the tree to the best child of the root in the previous one,
@@
 change depending on the order in which states are considered. We found this randomization to
-have a significant effect in (\textbf{TODO}: What game? \Qbert?) where (\textbf{TODO}: What was
-going on?). Last, we note that $\IW(1)$ and \BFS~attain similar performance as the baseline
-semi-random algorithm \emph{Perturb} from \cite{bellemare:jair2013}, which beated both \UCT~
-and the reinforcement learning algorithms reported.
+have a significant effect in \Qbert where not randomizing the order in which actions are considered
+resulted in $\IW(1)$ being trapped in a loop. Last, we note that both $\IW(1)$ and \BFS~attain similar 
+performance on \Freeway as the baseline semi-random algorithm \emph{Perturb} 
+from \cite{bellemare:jair2013}, which beated both \UCT~and the reinforcement learning algorithms reported.
 

### 2014-10-31 workshop-paper/performance_table.tex
Second pass on the Experimental Section

@@
    setting the Atari 2600 random seeds to $0, \ldots, 4$, and maximum episode duration of $18,000$
-   frames. \textit{T} is the average CPU T per call to each action selection, \textit{S} is the average score achieved,
+   frames. \textit{T} is the average CPU time per call to each action selection algorithm, \textit{S} is the average score achieved,
    and \textit{SD} is the standard deviation. \UCT~performance

### 2014-10-31 workshop-paper/experimental_results.tex
Third pass on the experimental results

@@
 \BRFS~and the baseline random action selection. Our experimental 
-setup  follows that of \cite{bellemare:jair2013}, with two exceptions. First, rather than
+setup  follows that of \cite{bellemare:jair2013}, with three exceptions. First, rather than
 considering $10$ episodes of each game, we have considered just $5$
@@
 are to be simulated while performing the lookahead), both $\IW(1)$ and \BFS are limited to
-simulate $100,000$ frames. \UCT~parameters such as the exploration constant $C$ and
-the planning horizon $m$ have not been modified ($C=0.1$, $m=300$), the latter being also
-used by both $\IW(1)$ and \BFS~to limit the depth of the lookahead. We include
-the average score reported for \UCT~by \cite{bellemare:jair2013} in Table~\ref{performance}
-bracketed with parenthesis for reference. Also unchanged is the discount factor
-used by all the algorithms ($\gamma=0.995$) and the rate at which an algorithm is
-invoked to select an action (every $5$ frames). All algorithms reuse the tree built between each 
+simulate $100,000$ frames. Other \UCT~parameters such as the exploration constant $C$ and
+the planning horizon $m$ have not been modified ($C=0.1$, $m=300$). Thirdly and last, both $\IW(1)$ 
+and \BFS~are limited to a planning horizon $m=1500$, where one only an action is evaluated every $5$ frames, matching
+the setup used during actual gameplay (an action is selected every $5$ frames, using a no-op
+for the remaining $4$). We include the average score reported for \UCT~by \cite{bellemare:jair2013} 
+in Table~\ref{performance} bracketed with parenthesis for reference. Also unchanged is the discount factor
+used by all the algorithms ($\gamma=0.995$). All algorithms reuse the tree built between each 
 lookahead by setting the root of the tree to the best child of the root in the previous one,
@@
 to be simulated in each lookahead, $\IW(1)$ still outperforms \UCT~in $3$ out of $9$ games
-(\Breakout, \Enduro and \Freeway). As introduced in the previous Section, the order in which
+(\Breakout, \Enduro~and \Freeway). As introduced in the previous Section, the order in which
 $\IW(1)$ and \BFS~consider actions is randomized, since the novelty of a given state may
 change depending on the order in which states are considered. We found this randomization to
-have a significant effect in \Qbert where not randomizing the order in which actions are considered
+have a significant effect in \Qbert~where not randomizing the order in which actions are considered
 resulted in $\IW(1)$ being trapped in a loop. Last, we note that both $\IW(1)$ and \BFS~attain similar 
-performance on \Freeway as the baseline semi-random algorithm \emph{Perturb} 
+performance on \Freeway~as the baseline semi-random algorithm \emph{Perturb} 
 from \cite{bellemare:jair2013}, which beated both \UCT~and the reinforcement learning algorithms reported.
@@
 be very similar to that of \BRFS~in some preliminary tests. When considering $2$--tuples
-of state variables to evaluate the novelty of states, it tended to result in lookaheads
+of state variables to evaluate the novelty of states, it tended to result in a lookahead
 $5$ times deeper than those resulting from applying \BRFS~, but still far too shallow so

### 2014-10-31 workshop-paper/paper.tex
inputs removed

@@
 \newcommand{\Omit}[1]{}
-
+%\newcommand{\propername}[1]{\text{\textsf{\small #1}}\xspace}
+\newcommand{\propername}[1]{\textsc{#1}}
 \newcommand{\eqdef}{\stackrel{\hbox{\tiny{def}}}{=}}
@@
 \newcommand{\BFS}{{\textit{2BFS}}}
+\newcommand{\UCT}{{\textit{UCT}}}
+\newcommand{\Asterix}{\propername{Asterix}}
+\newcommand{\BeamRider}{\propername{Beam Rider}}
+\newcommand{\Breakout}{\propername{Breakout}}
+\newcommand{\Enduro}{\propername{Enduro}}
+\newcommand{\Freeway}{\propername{Freeway}}
+\newcommand{\Pong}{\propername{Pong}}
+\newcommand{\Qbert}{\propername{Q*Bert}}
+\newcommand{\Seaquest}{\propername{Seaquest}}
+\newcommand{\SpaceInvaders}{\propername{Space Invaders}}
+\newcommand{\Deepmind}{\propername{Deepmind}}
 
@@
 
-\input{performance_table}
+\begin{table*}
+\addtolength{\tabcolsep}{-2pt}
+{\scriptsize
+\begin{tabular}{@{}l|rr|rrr|rrr|rrr|rrr|c@{}}
+%% table header
+\hline\\[-3ex]
+\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Random} & \multicolumn{3}{c|}{\BRFS}& \multicolumn{3}{c|}{$\IW(1)$}& \multicolumn{3}{c|}{\BFS}& \multicolumn{3}{c|}{\UCT}
+& \multicolumn{1}{c}{\Deepmind} \\[.2ex]
+\hline\\[-3ex]
+%% table body
+\multicolumn{1}{c}{Game}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c|}{SD}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} &  \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T} & \multicolumn{1}{c}{S}\\[.2ex]
+\hline\\[-3ex]
+\Asterix        & 210    & 54.77     & 3900  & 5538.95   & 32.24 & 119000        & 4987.48   & 17.08 & 56320       & 4757.84   & 0.95  & \textbf{296300 (290700)} & 6657.7    & 10.82 & n/a\\
+\BeamRider     & 448.8  & 84.64     & 519.2 & 179.81    & 43.48 & 3000          & 1150.22   & 6.02  & 3804        & 446.18    & 0.92  & \textbf{4320 (6624)}     & 2175.26   & 42.91 & 4092\\
+\Breakout       & 1      & 0.71      & 2     & 1.41      & 51.56 & 383.6         & 9.84      & 1.37  & 421.2       & 13.66     & 3.95  & \textbf{512.4 (364)}     & 193.97    & 26.03 & 168\\
+\Enduro          & 0      & 0         & 0.4   & 0.89      & 40.92 & \textbf{500}  & 0         & 38.13 & 292.8       & 6.18      & 13.12 & 306.8 (286)              & 8.14      & 32.14 & 470\\
+\Freeway         & 0      & 0         & 0     & 0         & 71.84 & \textbf{29.4} & 0.89      & 15.19 & 19.6        & 3.29      & 4.44  & 1 (0)                    & 1.41      & 90.64 & n/a \\
+\Pong           & -20.4  & 0.55      & -20.2 & 0.84      & 41.51 & 16.4          & 1.67      & 18.45 & \textbf{21} & 0         & 39.69 & \textbf{21 (21)}         & 0         & 38.37 & 20 \\
+\Qbert           & 185    & 345.33    & 160   & 85.88     & 42.83 & 1545          & 1666.06   & 1.82  & 14725       & 7872.1    & 1.76  & \textbf{18730 (17343)}   & 1127.69   & 28.61 & 1952\\
+\Seaquest       & 192    & 38.99     & 200   & 14.14     & 40.01 & \textbf{4274} & 4206.81   & 14.72 & 4070        & 1939.55   & 1.59  & 3406 (5132)              & 2402.47   & 26.16 & 1705 \\
+\SpaceInvaders & 187    & 120.86    & 115   & 61.44     & 46.11 & \textbf{2670} & 637.3     & 12.1  & 1327        & 598.04    & 2.25  & 2634 (2718)              & 375.66    & 19.66 & 581
+\end{tabular}
+}
 
-\input{experimental_results}
+ \caption{\small Random baseline, \BRFS, $\IW(1)$, \BFS, \UCT~performance over 5 runs,
+   setting the Atari 2600 random seeds to $0, \ldots, 4$, and maximum episode duration of $18,000$
+   frames. \textit{T} is the average CPU time per call to each action selection algorithm, \textit{S} is the average score achieved,
+   and \textit{SD} is the standard deviation. \UCT~performance
+   reported in \cite{bellemare:jair2013} is shown in parenthesis. Numbers
+   in bold show best performer in terms of average score. Average score reported in 
+   \cite{deep-mind-atari} is reported for reference under the column \Deepmind.
+   }
+\label{performance}
+
+\end{table*}
+
+We have tested the $\IW(1)$ and \BFS~described in the previous Section
+over $9$ games selected amongst those considered in the \emph{training} set
+of \cite{bellemare:jair2013}, and those discussed in \cite{deep-mind-atari}. 
+In Table~\ref{performance} the performance of both $\IW(1)$ 
+and \BFS, characterized by the average score obtained over all episodes, the standard deviation
+of the score obtained and time (in seconds) per lookahead, is compared with that of \UCT, 
+\BRFS~and the baseline random action selection. Our experimental 
+setup  follows that of \cite{bellemare:jair2013}, with three exceptions. First, rather than
+considering $10$ episodes of each game, we have considered just $5$
+due to time constraints, as each game can take as much as $40$ hours of 
+CPU time. Second, while \UCT~is limited to perform up to $500$ base policy rollouts ($150,000$ frames
+are to be simulated while performing the lookahead), both $\IW(1)$ and \BFS are limited to
+simulate $100,000$ frames. Other \UCT~parameters such as the exploration constant $C$ and
+the planning horizon $m$ have not been modified ($C=0.1$, $m=300$). Thirdly and last, both $\IW(1)$ 
+and \BFS~are limited to a planning horizon $m=1500$, where one only an action is evaluated every $5$ frames, matching
+the setup used during actual gameplay (an action is selected every $5$ frames, using a no-op
+for the remaining $4$). We include the average score reported for \UCT~by \cite{bellemare:jair2013} 
+in Table~\ref{performance} bracketed with parenthesis for reference. Also unchanged is the discount factor
+used by all the algorithms ($\gamma=0.995$). All algorithms reuse the tree built between each 
+lookahead by setting the root of the tree to the best child of the root in the previous one,
+and deleting its siblings and their descendants, as \cite{bellemare:jair2013} does for \UCT. We note that 
+this may have a different effect on algorithms where all branches have the same number of
+generated nodes, as is the case of \BRFS, than in $\IW(1)$, \BFS~and \UCT, where the depth of
+the branches is uneven.
+
+Table~\ref{performance} shows that both $\IW(1)$ and \BFS~outperform by a wide margin 
+\BRFS, which sometimes performs worse than the baseline, both in terms of average score attained and 
+average time per call. \BRFS~rarely collects any reward to inform the choice of action at
+the root node of the lookahead, as the depth of the \BRFS~search tree, having limited the number of
+frames allowed to be simulated to $100,000$, corresponds to a lookahead of $0.3$ seconds ($20$ 
+frames or $4$ nodes). Both $\IW(1)$ and \BFS~reach states up to $350$--$1500$ frames ahead 
+($70$--$260$ nodes or $6$--$22$ seconds), and that results in collecting enough rewards so as to
+inform the choice at the root in a way that much higher scores are attained. Interestingly, very often
+neither $\IW(1)$ or \BFS~fully consume the number of simulation frames allocated, since many
+states are pruned, in the case of $\IW(1)$, or relegated to the back of the open lists, in the 
+case of \BFS, due to having a novelty of $0$ and the planning horizon being limited to $m$.
+
+We observe that in $4$ out of $9$ games, $\IW(1)$ outperforms \UCT~limited to the same number
+of simulated frames, both in terms of average score attained and average time per call. \BFS~is
+generally faster than $\IW(1)$, but only outperforms \UCT~in $2$ out of $9$ games (\Freeway and
+\Seaquest) and $\IW(1)$ in $4$ out of $9$ games. Compared with the \UCT~results reported
+in \cite{bellemare:jair2013}, that by limiting the number of rollouts allow a $50$\% more frames
+to be simulated in each lookahead, $\IW(1)$ still outperforms \UCT~in $3$ out of $9$ games
+(\Breakout, \Enduro~and \Freeway). As introduced in the previous Section, the order in which
+$\IW(1)$ and \BFS~consider actions is randomized, since the novelty of a given state may
+change depending on the order in which states are considered. We found this randomization to
+have a significant effect in \Qbert~where not randomizing the order in which actions are considered
+resulted in $\IW(1)$ being trapped in a loop. Last, we note that both $\IW(1)$ and \BFS~attain similar 
+performance on \Freeway~as the baseline semi-random algorithm \emph{Perturb} 
+from \cite{bellemare:jair2013}, which beated both \UCT~and the reinforcement learning algorithms reported.
+
+We finish the discussion of the results presented in Table~\ref{performance} by observing
+that $\IW(1)$ outperforms the average scores reported for the best algorithm reported in
+\cite{deep-mind-atari} but in $4$ out of $7$ games, and \BFS~does so too in $5$ out of $7$
+games.
+
+We do not report $IW(2)$ in Table~\ref{performance} since we found its performance to
+be very similar to that of \BRFS~in some preliminary tests. When considering $2$--tuples
+of state variables to evaluate the novelty of states, it tended to result in a lookahead
+$5$ times deeper than those resulting from applying \BRFS~, but still far too shallow so
+as to compete with the other planning algorithms considered.
 

### 2014-10-31 src/agents/BestFirstSearch.cpp
changed depth to be same as uct, and bfs as it was before

@@ -245,11 +245,11 @@ void BestFirstSearch::expand_tree(TreeNode* start_node) {
     
     if(!start_node->v_children.empty()) {
 	    start_node->updateTreeNode();
-	    //num_simulated_steps += reuse_branch( start_node );
-	    //std::cout  << "Num_reused_steps: "<< num_simulated_steps << std::endl;
+	    num_simulated_steps += reuse_branch( start_node );
+	    std::cout  << "Num_reused_steps: "<< num_simulated_steps << std::endl;
 	    //COMMENT LINES BELOW, AND UNCOMMENT ABOVE TO WORKSHOP STYLE. ALSO CHANGE FN NOVEL 2ND QUEUE
-	     reset_branch( start_node );
-	     q_exploration->push(start_node);        
+	    //reset_branch( start_node );
+	    //q_exploration->push(start_node);        
 	     update_novelty_table( start_node->state.getRAM() );
     }
     else

### 2014-10-31 src/agents/BestFirstSearch.hpp

@@ -32,7 +32,7 @@ public:
 	{
 	    if ( b->fn < a->fn ) return true;
 	    //else if( b->fn == a->fn && b->m_depth < a->m_depth ) return true;
-	    else if( b->fn == a->fn && b->novelty > a->novelty ) return true;
+	    else if( b->fn == a->fn && b->novelty < a->novelty ) return true;
 	    return false;
 	}
     };

### 2014-10-31 src/common/Defaults.cpp

@@ -66,7 +66,7 @@ void setDefaultSettings(Settings &settings) {
 	//settings.setBool("iw1_stop_on_first_reward", true);
 	//settings.setInt("iw1_reward_horizon", 50 );
 	settings.setBool("iw1_stop_on_first_reward", false);
-	settings.setInt("iw1_reward_horizon", 300 );	
+	settings.setInt("iw1_reward_horizon", 60 );	
 	settings.setBool("randomize_successor_novelty", true );
 
 }

### 2014-10-31 workshop-paper/paper.tex
Hector updates

--- 2014-10-31 workshop-paper/paper.tex
@@
 
-%% Macros
 \newcommand{\tuple}[1]{{\langle #1\rangle}}
@@
 
+\newcommand{\propername}[1]{\textsc{#1}}
 \newcommand{\eqdef}{\stackrel{\hbox{\tiny{def}}}{=}}
-\newcommand{\IW}{{\textit{IW}}}
-\newcommand{\SIW}{{\textit{SIW}}}
-\newcommand{\ID}{{\textit{ID}}}
-\newcommand{\BRFS}{{\textit{BrFS}}}
-\newcommand{\BFS}{{\textit{2BFS}}}
+%\newcommand{\IW}{{\textit{IW}}}
+%\newcommand{\SIW}{{\textit{SIW}}}
+%\newcommand{\ID}{{\textit{ID}}}
+%\newcommand{\BRFS}{{\textit{BrFS}}}
+%\newcommand{\BFS}{{\textit{2BFS}}}
+%\newcommand{\UCT}{{\textit{UCT}}}
+\newcommand{\IW}{{\textsc{IW}}}
+\newcommand{\SIW}{{\textsct{SIW}}}
+\newcommand{\ID}{{\textsc{ID}}}
+\newcommand{\BRFS}{{\textsc{BrFS}}}
+\newcommand{\BFS}{{\textsc{2BFS}}}
+\newcommand{\UCT}{{\textsc{UCT}}}
+
+\newcommand{\Asterix}{\propername{Asterix}}
+\newcommand{\BeamRider}{\propername{Beam Rider}}
+\newcommand{\Breakout}{\propername{Breakout}}
+\newcommand{\Enduro}{\propername{Enduro}}
+\newcommand{\Freeway}{\propername{Freeway}}
+\newcommand{\Pong}{\propername{Pong}}
+\newcommand{\Qbert}{\propername{Q*Bert}}
+\newcommand{\Seaquest}{\propername{Seaquest}}
+\newcommand{\SpaceInvaders}{\propername{Space Invaders}}
+\newcommand{\Deepmind}{\propername{Deepmind}}
 
@@
 cannot be used  for selecting actions  for  two reasons: first,  no  compact PDDL-model of  the games is given, 
-and  more importantly, the actions effects and the goals are  not known a priori. Moreover, in these games there is usually  
+and  more importantly,  the action effects and goals are  not known a priori. Moreover, in these games there is usually  
 no set of goals to be achieved but rewards to be collected. These features do not preclude the use of   classical  algorithms like
-breadth-first search or Dijkstra's algorithm,  but    these  algorithms are not effective over  large state spaces. We thus turn to a
+breadth-first search or Dijkstra's algorithm,  but    these methods are not effective over  large state spaces. We thus turn to a
 different class of classical planning algorithms  introduced recently  that perform a \emph{structured exploration} of the state space; namely,  
@@
 a breadth-first search in which a state is  pruned when it is not the first state in the search  to make true some set of i atoms or less.
-Preliminary empirical results suggest that the performance of IW with the  $k$ parameter fixed to $1$, i.e., IW(1), 
-is much better than random or breadth-first search, and close to the performance of UCT. Other simple variations of IW
-that combine exploration and exploitation are evaluated as well.
+Preliminary empirical results suggest that the performance of IW with the  $k$ parameter fixed to $1$, 
+i.e., IW(1), is much better than random or breadth-first search, and approaches 
+the performance of UCT. A simple best-first variation  of IW,  that combine exploration and exploitation,
+is evaluated as well. 
 \end{quote}
@@
 While all  these games feature a known initial (RAM) state and actions that
-have deterministic effects on this state, the problem of selecting the next action
-to be done cannot be addressed with state-of-the-art classical planners for two 
-reasons. First, there is no compact PDDL-like encoding of the domain, and more 
+have deterministic effects on the state, the problem of selecting the next action
+to be done cannot be addressed with state-of-the-art classical planners \cite{geffner:2013book}.
+This is because there is  no compact PDDL-like encoding of the domain, and more 
 importantly,  the goal to be achieved in each game  is not given. Indeed, there are 
-often no goals at all but rewards $r(a,s)$ that potentially depend on the action $a$ and 
-the state $s$ in  which  the action is performed, and these rewards are not known a priori either. 
+often no goals  but rewards $r(a,s)$ that  depend on the action $a$ and 
+the state $s$ in  which  the action is performed, and these rewards are not known a priori 
+either.\footnote{Actually, in the Atari games, the rewards $r(a,s)$ depend only
+on the state $f(a,s)$ that results from doing action $a$ in the state $s$.}
 Thus no model of the  goals  or the  rewards can be used to bias the search, and hence, 
@@
 \cite{sutton:book} over a deterministic MDP where the state transitions and rewards are not known, or alternatively,
-as a  \emph{net-benefit planning problem} \cite{ref-net-benefit-planning} with unknown state transitions  and  rewards. Namely, 
+as a  \emph{net-benefit planning problem} \cite{ref-net-benefit-planning,keyder} with unknown state transitions  and  rewards. Namely, 
 we seek  an action sequence $a_0, a_1, \ldots, a_m$ from the given current  state $s_0$ that generates 
@@
 where $m$ is a sufficiently large planning horizon, and the rewards $r(a_i,s_i)$ and the  state-transitions $s_{i+1}=f(a_i,s_i)$ 
-are known only after the action $a_i$ is applied in the state $s_i$.
+are known only after  action $a_i$ is applied in the state $s_i$.
 
@@
 
-The limitation of the   basic blind search  methods is that they are not effective  over very spaces, neither
+The limitation of the   basic blind search  methods is that they are not effective  over large  spaces, neither
 for solving problems off-line, nor for guiding  a lookahead search for solving problems on-line. 
@@
 when goals are restricted to be single atoms  where IW provably runs in low polynomial time in the number of
-variables \cite{nir:ecai2012}. For example, in a a block world instance with any number of blocks and any initial
-configuration,  the goal of  placing a given  block on top of another, will be achievable in  quadratic time  
+variables \cite{nir:ecai2012}. For example, in a block world instance with any number of blocks and any initial
+configuration,  the goal of  placing a given  block on top of another will be achievable in  quadratic time  
 in the number of blocks as any such instance can be shown to  have \emph{width} 2. This means  that IW(2) will be 
 not only complete then, but also optimal. This is also true for most other benchmark domains in classical planning
-that turn out to be have small, bounded widths for any instance, as long as the goal is atomic. 
+that turn out to  have small, bounded widths for any instance, as long as the goal is atomic. 
 For dealing with the benchmarks where goals are  not single atoms, e.g., where a given \emph{tower} of blocks needs to be built, 
 simple extensions of the basic IW procedure have been developed as well, such as Serialized IW, where IW is 
-used to achieve the goals one at at time \cite{nir:ecai2012,nir:ecai2014}. The extension of IW for problems where there is no goal but an
-additive reward measure   to be optimized is direct.
+used to achieve the goals one at at time \cite{nir:ecai2012,nir:ecai2014}. 
+The extension of IW for problems where there is no goal but an
+additive reward measure  to be optimized, like in the Atari games, is direct.
 
@@
 
+\Omit{
 The sequence of calls IW(1), IW(2), \ldots, IW(k) manages to combine the scope of ``blind'' search algorithms with the performance of 
@@
 is used  as the system state. 
+}
 
-The paper is organized as follows. We review the Iterated Width  algorithm and where it comes from, 
-look at  some simple   variations of the basic algorithm that appear to be convenient  for the Atari
-games, present and analyze the preliminary experimental results, summarize the main lessons, and
-discuss related and future work.
+The paper is organized as follows. We review the iterated width  algorithm and where it comes from, 
+look at  simple   variations of the algorithm that appear to be convenient  for the Atari
+games, present and analyze the preliminary experimental results, and discuss related and future work.
 
@@
 represent the value $x_j$ of variable $X_i$ in the state. 
-In addition to the states space, a problem is defined by an initial state $s_0$, 
-a set of actions $A(s)$ applicable in each state, a 
+In addition to the state space, a problem is defined by an initial state $s_0$, 
+a set of actions applicable in each state, a 
 transition function $f$ such that  $s' = f(a,s)$ is 
@@
 \item The \emph{novelty of a newly generate state $s$  in a  breadth-first search}
-is   $1$ when $s$ is the first state generated in the search that makes true an atom $X=x$,
-$2$  when $s$ is the first state that makes a pair of atoms $X=x$ and $Y=y$ true, and so on.
+is   $1$ if  $s$ is the first state generated in the search that makes true some  atom $X=x$, 
+else it is $2$ if  $s$ is the first state that makes a \emph{pair} of atoms $X=x$ and $Y=y$ true, 
+and so on.
 
@@
 $R(s)$ generated but not pruned by IW, and the best path is the one that leads to the state $s$ from the current state.
-The action selected to be done next is then the first action along such a path  (** see eventually randomization **).
+The action selected to be done next is then the first action along such a path.
 
@@
 Breadth-First Search (\BRFS),   and Greedy Best First Search (\textit{GBFS}) guided by  additive heuristic.
-Time and  memory outs  after 30 minutes and 2 GB resp. From \cite{nir:ecai2012}.
+Time and  memory outs  after 30 minutes and 2 GB resp. From \cite{nir:ecai2102}.
 }
@@
 but   unlike these algorithms, it uses the factored representation of the states in terms of variables to structure the 
-search in a different way. This structured exploration  has proved to be very effective in classical planning 
-over most of the benchmark domains  when the goals are restricted to be single atoms. Table~\ref{table1} from \cite{nir:ecai2012} shows 
+search in a different way. This structured exploration  has proved to be very 
+effective over classical planning benchmark domains  when goals are  single atoms.\footnote{
+Any conjunctive goal can be mapped into a single dummy atomic goal by adding an action that
+achieves the dummy goal and that has the original conjunctive goal as a precondition. 
+Yet, this changes the domain.}
+Table~\ref{table1} from \cite{nir:ecai2012} shows 
 the percentage of instances that are  solved by  the first iteration of IW, i.e.  IW(1), by the second iteration 
@@
 Furthermore, when the performance of \IW\   is compared with  breadth-first search and iterative deepening, 
-on the one hand,  and  with a  Greedy Best First Search guided by the additive heuritstic $h_{add}$
-on the other (this algorithm is similar to one used by  the FF planner when hill climbing search fails \cite{hoffmann:ff}), 
+on the one hand,  and  with a  Greedy Best First Search guided by the additive heuristic $h_{add}$ 
+\cite{bonet:hsp-aij} on the other (this algorithm is similar 
+to one used by  the FF planner when hill climbing search fails \cite{hoffmann:ff}), 
 it turns out that ``blind'' \IW\ solves as many problems as the informed search, 34627 vs. 34849, far ahead
@@
 
-The min $k$ value for which $IW(k)$ solves an instance  is clearly bounded and small in most of
+The min $k$ value for which $IW(k)$ solves a problem  is clearly bounded and small in most of
 these instances. This is  actually no  accident and  has a \emph{theoretical explanation.} Lipovetzky and Geffner define
@@
 of variables and states in the problem. The min value $k$ for which the iteration  $IW(k)$ solves
-the problem cannot exceed the problem width, so that the algorithm \emph{IW will run in time and space
+the problem cannot exceed the problem width, so  the algorithm \emph{IW runs in time and space
 that are exponential in the problem width.}
@@
 While this notion of width and the iterated  width algorithms that are based on it  have been designed
-for problems where a goal state needs to be reached, the notions remain  relevant  in  optimization problems as well.
-Indeed, if a  good path is made of states $s_i$ each of which have a low width, we would like 
-IW to find such path in low polynomial time for a low value of the $k$ parameter. 
-Later on we will discuss the changes required in IW to enforce this property, which
-are not yet implemented in the code that is evaluated below.
+for problems where a goal state needs to be reached, the notions remain  
+relevant  in  optimization problems as well. Indeed, if a  good path is made of 
+states $s_i$ each of which have a low width, IW can be made 
+to find such path in low polynomial time for a small value of the $k$ parameter.
+Later on we will discuss  a slight  change  required in IW to enforce this property
+which is actually  not yet implemented in the code that is tested below.
 
-\section{Width-Based Algorithms for the Atari Games}
-
-%%Explain: IW(1), 2BFS, and possible IW11. Also, Dijkstra postprocessing over IW(1) graph if ran.
+\section{The Algorithms for the Atari Games}
 
 The number of nodes \emph{generated} by IW(1) is $n \times D \times b$ in the worst case, 
-where $n$ is the number of variables, $D$ is the size of their domains, and $b$ is maximum the number of 
-actions applicable per  state.  This same number in breadth-first search is not linear in $n$
+where $n$ is the number of variables, $D$ is the size of their domains, and $b$ is the number of 
+actions per  state.  This same number in breadth-first search is not linear in $n$
 but exponential. For the Atari games,  $n=128$, $D=256$, and $b=18$, so that the product is
@@
 and apply these algorithms to that representation. The results would be different.
-In such a case, IW(1) would generate $n' \times D' \times b$ states, and 
-IW(2), $(n' \times D')^2 \times b$ states, which for $n'=1024$ and $D=2$
+In such a case, IW(1) and IW(2)  would generate $n' \times D' \times b$ and 
+$(n' \times D')^2 \times b$ states, which for $n'=1024$ and $D=2$
 represent  $36,864$ and $7,5497,472$ states respectively.
-These are both feasible numbers, but we haven't tried yet this representation,
+These are feasible numbers, but we haven't tried  this representation yet
 under the assumption that the correlations among the bits in each one of 
@@
 In summary, from the basic IW algorithm we are testing only the first 
-call, IW(1).
+call IW(1).
 
@@
 with two queues: one queue  ordered first by  novelty measure 
-(recall, novelty 1 means that the state is the first one to make some atom true),
+(recall that novelty one means that the state is the first one to make some atom true),
 and a second queue ordered by accumulated reward. In one iteration, the best first
@@
 a node is expanded, it is removed from the queue, and its children are placed on 
-both queues. The exception are the nodes with accumulated reward 0 that are only
-placed in the first queue.
-We refer to this best-first algorithm below, as 2BFS.
+both queues. The exception are the nodes with no accumulated reward  that are 
+placed in the first queue only. We refer to this best-first algorithm 
+as 2BFS.
 
-In the experiments below, we use IW(1) with two simple variations.
-First, in the breadth-first search, we generate the children in 
-random order. This makes the execution that results from the
+For  the experiments below, we added two simple variations to  IW(1) and 2BFS.
+First, in the breadth-first search underlying IW(1), we generate the children in 
+random order. This makes the executions that results from the
 IW(1) lookahead less susceptible to fall into loops; a potential
 problem in local search algorithms with no memory or learning.
-Second, a discount factor $\gamma=0.995$ is added
-for discounting future rewards like in UCT. For this,
-each state $s$ keeps its depth d(s) in the IW(1) tree,
-and if state s' is the child of state s and action a, 
+Second, a discount factor $\gamma=0.995$ is used in both algorithms 
+for discounting future rewards like in UCT. 
+For this, each state $s$ keeps its depth $d(s)$ in the search tree,
+and if state s' is the child of state $s$ and action $a$, 
 $R(s')$ is set to $R(s) + \gamma^d(s´) r(a,s)$.
-A discount factor less than 1 results in a slight
+The  discount factor  results in a slight
 preference for rewards that can be reached earlier, 
-which is reasonable  in the on-line setting.
+which is a  reasonable heuristic in  on-line settings
+based on lookahead searches.
 
-%% NIR: IW11 Don't need to be explained. Maybe in Future Work?
+\section{Experimental Results}
 
-\section{Experimental Results in the Atari Games}
+\begin{table*}
+\addtolength{\tabcolsep}{-2pt}
+{\scriptsize
+\begin{tabular}{@{}l|rr|rrr|rrr|rrr|rrr|c@{}}
+%% table header
+\hline\\[-3ex]
+\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Random} & \multicolumn{3}{c|}{\BRFS}& \multicolumn{3}{c|}{$\IW(1)$}& \multicolumn{3}{c|}{\BFS}& \multicolumn{3}{c|}{\UCT}
+& \multicolumn{1}{c}{\Deepmind} \\[.2ex]
+\hline\\[-3ex]
+%% table body
+\multicolumn{1}{c}{Game}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c|}{SD}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} &  \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T} & \multicolumn{1}{c}{S}\\[.2ex]
+\hline\\[-3ex]
+\Asterix        & 210    & 54.77     & 3900  & 5538.95   & 32.24 & 119000        & 4987.48   & 17.08 & 56320       & 4757.84   & 0.95  & \textbf{296300 (290700)} & 6657.7    & 10.82 & n/a\\
+\BeamRider     & 448.8  & 84.64     & 519.2 & 179.81    & 43.48 & 3000          & 1150.22   & 6.02  & 3804        & 446.18    & 0.92  & \textbf{4320 (6624)}     & 2175.26   & 42.91 & 4092\\
+\Breakout       & 1      & 0.71      & 2     & 1.41      & 51.56 & 383.6         & 9.84      & 1.37  & 421.2       & 13.66     & 3.95  & \textbf{512.4 (364)}     & 193.97    & 26.03 & 168\\
+\Enduro          & 0      & 0         & 0.4   & 0.89      & 40.92 & \textbf{500}  & 0         & 38.13 & 292.8       & 6.18      & 13.12 & 306.8 (286)              & 8.14      & 32.14 & 470\\
+\Freeway         & 0      & 0         & 0     & 0         & 71.84 & \textbf{29.4} & 0.89      & 15.19 & 19.6        & 3.29      & 4.44  & 1 (0)                    & 1.41      & 90.64 & n/a \\
+\Pong           & -20.4  & 0.55      & -20.2 & 0.84      & 41.51 & 16.4          & 1.67      & 18.45 & \textbf{21} & 0         & 39.69 & \textbf{21 (21)}         & 0         & 38.37 & 20 \\
+\Qbert           & 185    & 345.33    & 160   & 85.88     & 42.83 & 1545          & 1666.06   & 1.82  & 14725       & 7872.1    & 1.76  & \textbf{18730 (17343)}   & 1127.69   & 28.61 & 1952\\
+\Seaquest       & 192    & 38.99     & 200   & 14.14     & 40.01 & \textbf{4274} & 4206.81   & 14.72 & 4070        & 1939.55   & 1.59  & 3406 (5132)              & 2402.47   & 26.16 & 1705 \\
+\SpaceInvaders & 187    & 120.86    & 115   & 61.44     & 46.11 & \textbf{2670} & 637.3     & 12.1  & 1327        & 598.04    & 2.25  & 2634 (2718)              & 375.66    & 19.66 & 581
+\end{tabular}
+}
 
-\input{performance_table}
+ \caption{\small Random baseline, \BRFS, $\IW(1)$, \BFS, \UCT~performance over 5 runs,
+   setting the Atari 2600 random seeds to $0, \ldots, 4$, and maximum episode duration of $18,000$
+   frames. \textit{T} is the average CPU time per call to each action selection algorithm, \textit{S} is the average score achieved,
+   and \textit{SD} is the standard deviation. \UCT~performance
+   reported in \cite{bellemare:jair2013} is shown in parenthesis. Numbers
+   in bold show best performer in terms of average score. Average score reported in 
+   \cite{deep-mind-atari} is reported for reference under the column \Deepmind.
+   }
+\label{performance}
 
-\input{experimental_results}
+\end{table*}
 
-\section{Exploration and Exploitation: Discussion}
+We tested  $\IW(1)$ and \BFS\  over $9$ games: those considered in the \emph{training} set
+of \cite{bellemare:jair2013} and those discussed in \cite{deep-mind-atari}. 
+In Table~\ref{performance} the performance of  $\IW(1)$ and \BFS\ 
+% , characterized by the average score obtained over all episodes, the standard deviation
+% of the score obtained and time (in seconds) per lookahead, 
+is compared with that of \UCT, breadth-first search (\BRFS), and random action selection. 
+Our experimental setup  follows the one in  \cite{bellemare:jair2013} with two exceptions. First, rather than
+considering $10$ episodes of each game, we have considered just $5$ due to time constraints, as the experiments 
+for each  game can take up to $40$ hours of CPU time. Second, while \UCT\ is still limited to perform up to $500$ base policy rollouts 
+of depth $300$, a maximum of budget of  $100,000$ simulated frames is applied to all the algorithms, UCT included. 
+%% Was not clear; hope this is what you meant, if not fix
+%%% ($150,000$ frames are to be simulated while performing the lookahead), both $\IW(1)$ and \BFS are limited to
+% The other \UCT\ parameters like the exploration constant $C$ and the horizon $m$ are the same ($C=0.1$, $m=300$). 
+%% matching the setup used during actual gameplay. %%  (an action is selected every $5$ frames, using a no-op for the remaining $4$). %% Needed?
+The average scores reported for \UCT\  \cite{bellemare:jair2013}, without this extra limitation, are included
+in  parenthesis for reference. $\IW(1)$ and \BFS\ have been  limited to a planning horizon $m=1500$ where one an action is evaluated every 
+$5$ frames. The discount factor used by all the algorithms is $\gamma=0.995$, and all of them  reuse the tree built 
+in the previous lookahead by setting the root of the tree to the best child of the root, while 
+deleting its siblings and their descendants as in \cite{bellemare:jair2013}.
+%%% really need what's below? it's not very clear .. at least to me HG
+%% We note that  this may have a different effect on algorithms where all branches have the same number of
+%% generated nodes, as is the case of \BRFS, than in $\IW(1)$, \BFS\ and \UCT, where the depth of
+%% the branches is uneven.
+
+Table~\ref{performance} shows that $\IW(1)$ and \BFS\ both outperform \BRFS\  and the random baseline.
+\BRFS\ rarely collects any reward, as the depth of the \BRFS\ search tree 
+results in  a lookahead of $0.3$ seconds ($20$  frames or $4$ nodes). 
+On the other hand, both $\IW(1)$ and \BFS\ reach states that are  up to $350$--$1500$ frames deep
+($70$--$260$ nodes or $6$--$22$ seconds), even if $\IW(1)$ doesn't fully use the number of simulation frames allocated
+in many cases, as when  the set of unpruned states becomes  particularly small. 
+%% case of \BFS, due to having a novelty of $0$ and the planning horizon being limited to $m$
+%%% (*** What is novelty $0$??? Not clear why BFS doesn't take all the budget.
+
+$\IW(1)$ outperforms \UCT\ in $4$ out of $9$ games, 
+both in terms of  average scores and average times per call. \BFS\ 
+is generally faster than $\IW(1)$ but only outperforms \UCT\ in $2$ of the games (\Freeway and
+\Seaquest). Compared with the UCT results from  \cite{bellemare:jair2013}, shown in parenthesis, 
+$\IW(1)$ still outperforms \UCT\ in $3$  of the  $9$ games (\Breakout, \Enduro~and \Freeway). 
+As mentioned in the previous section, the order in which $\IW(1)$ and \BFS\ generate the children
+of an expanded node are randomized, something which has a potential effect on the  novelty measures
+of nodes, and hence in the search graph. We found that this randomization helps, in particular 
+in \Qbert, where not randomizing results  in $\IW(1)$ being trapped into  loops sometimes.
+
+We also  note that in \Freeway,  both $\IW(1)$ and \BFS\ attain a  performance similar to the 
+baseline semi-random algorithm \emph{Perturb} in  \cite{bellemare:jair2013}, 
+which outperforms   both \UCT\ and  the reinforcement learning methods. 
+Likewise,  $\IW(1)$ outperforms the best learning algorithm in \cite{deep-mind-atari}  
+in $4$ out of the  $7$ games considered, and \BFS\ does so in $5$  of the $7$ games.
+
+We do not report the performance of $\IW(2)$ because  in our
+preliminary tests and according to the discussion in the previous section,  
+it doesn't  appear to improve much on  \BRFS, even if  it results in a lookahead
+that is $5$ times deeper, yet still far too shallow  to compete with the other 
+planning algorithms.
+
+\section{Discussion: Exploration and Exploitation}
 
@@
 of classical planning in order to understand why most of the hundreds of existing benchmarks
-resulting from planning competitions appear to be relatively simple for current planners, even
-though classical planning is PSPACE-complete \cite{bylander:complexity}. The partial answer is that 
-most of these domains have a low width, and hence, can be solved in low polynomial time (by IW),
-when goals contain a single atom. At the same time, most problems with multiple atomic goals are 
-also easy too, as the goals can often be achieved one at a time \cite{nir:ecai2012}.
+appear to be relatively simple for current planners, even
+though classical planning is PSPACE-complete \cite{bylander}. A  partial answer is that 
+most of these domains have a low width, and hence, can be solved in low polynomial time (by IW)
+when goals contain a single atom.  Benchmark problems with multiple atomic goals tend to be 
+easy too, as the goals can often be achieved one at a time \cite{nir:ecai2012}.
 
-In the Iterated Width algorithm, the key notion is the \emph{novelty measure} of a state in the underlying
-breadth-first search. In particular, a state gets a novelty 1 measure in the search if it makes true,
-some atom that up to that point had been false in all previously generated states. In the procedure IW(1)
-all other states are pruned right away, as if they were ``duplicate'' states. The algorithms IW and IW(k)
-are pure  exploration algorithms that make no use at all of the information about the goal, if there is a goal,
-or about the  reward function to optimize otherwise. They are  thus  pure exploration algorithms guided by the novelty
-measures, which use the structure of the states to provide a  structure to the exploration of the state space,
-which in the case of IW is guaranteed to be complete.
+In the iterated width algorithm, the key notion is the \emph{novelty measure} of a state in the underlying
+breadth-first search. These novelty measures make use of the factored representation of the states for
+providing a structure to the search: states that have width 1 are explored first in linear time, 
+then states that have  width two are explored in quadratic time, and so on. In classical planning problems 
+with atomic goals, this way of organizing the search  pays off both theoretically and practically.
 
@@
 in the context of genetic algorithms \cite{lehman:novelty}. %%% bibtex for this below
-In these methods individuals in the population are not ranked according to optimization function but in terms of how 
-``novel'' thay are in relation to the rest of the population, thus encouraging diversity
-and exploration, rather than (greedy) exploitation. The actual definition of novelty
+In these methods individuals in the population are not ranked according to the  optimization function
+but in terms of how ``novel'' thay are in relation to the rest of the population, thus encouraging diversity
+and exploration rather than (greedy) exploitation. The actual definition of novelty
 in such a case is domain-dependent; for example, in the evolution of a controller
@@
 close it takes the robot to the goal (the greedy measure), but how far it takes
-the robot in relation to the locations that are reached by following the other controllers
-(a diversity measure). The novelty measure used by IW, on the other hand, 
-is domain-independent and it follows from the structure of the state
-as captured by the problem variables.
+the robot in relation to the locations that are reached when following the other controllers
+(a diversity measure). The novelty measure used by IW, on the other hand, is domain-independent and it 
+is determined by the structure of the states as captured by the problem variables.
 
@@
 is not needed for optimality purposes, but just for improving the effectiveness of the 
-local lookahead searches. Indeed, a best-first search algorithm guided by (discounted)
+ lookahead search. Indeed, a best-first search algorithm guided by (discounted)
 accumulated reward will deliver eventually  best moves, but it will not be as effective
@@
 at all. The UCT algorithm provides a method for balancing exploration and exploitation,
-which can thus be effective over small time windows. The 2BFS algorithm above with two queues
-that alternate, one guided by the novelty measures, and the other by the accumulated reward,
+which  is  effective over small time windows. The 2BFS algorithm above with two queues
+that alternate, one guided by the novelty measures  and the other by the accumulated reward,
 provides a different scheme. The first converges to the optimal behavior asymptotically; the 
-second in a bounded number of steps. Alternately, novelty-guided probes, possibly with 
+second in a bounded number of steps. Alternatively, novelty-guided probes, possibly with 
 some noise added, could potentially pay off as  a  base policy for UCT in replacement 
-of random rollouts, as they manage apparently to exploit the structure of many problems.
+of random rollouts, as they apparently manage to exploit the structure of many problems.
 
-\section{Summary and On-going Work}
+\section{Summary and Future Work}
 
@@
 that originate in work in classical planning, can be used in the Atari games
-with a performance that approaches the state-of-the-art as captured by UCT 
-planning, sometimes using a fraction of the time. We are currently addressing three issues, 
-which we haven't managed to get in time for this submission. First, we want to extend the evaluation to 
-all the games featured in \cite{bellemare:jair2013}, and to add more runs per  game. 
-Second, we want to evaluate the algorithm for different fixed time windows. In the table above, for example, UCT
-is often taking an order-of-magnitude more time per decision as 2BFS, so the 
-comparison in not fair. Third, we need to revise the way duplicate states are handled 
-in IW and 2BFS to ensure that the best paths among the paths defined by the states that have been expanded
-are returned. This change is important from a theoretical point of view; it's not clear
-what it's practical impact will be.
+with a performance that approaches the state-of-the-art, sometimes using a fraction of the time. We are currently addressing three 
+issues, which we haven't managed to get in time for this submission. First, we want to extend the 
+evaluation to all the games featured in \cite{bellemare:jair}, and to add more runs per  game. 
+Second, we want to evaluate the various algorithms for different fixed time windows for decisions.
+In the table above, for example, UCT is often taking an order-of-magnitude more time per decision 
+than  2BFS. Third, we need to revise the way duplicate states are handled in IW and 2BFS to ensure
+that the best paths among the states that have been expanded are returned. Finally, 
+for the future, we also want to explore whether this type of algorithms
+can give meaningful results as well  when used on the screen state, as opposed 
+to the RAM state.
+
+%%% - Learning in same instance, across diff instances of same game, etc.
+%%%% - Learning controllers vs. on-line planning, etc
 

### 2014-10-31 workshop-paper/paper.tex
Table 1 fixed to top of page

@@
 
-\begin{table}
+\begin{table}[t!]
 \begin{center}
@@
 {\scriptsize
-\begin{tabular}{@{}l|rr|rrr|rrr|rrr|rrr|c@{}}
+\begin{tabular}{@{}l|rr|rrr|rrr|rrr|rrr@{}}
 %% table header
 \hline\\[-3ex]
-\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Random} & \multicolumn{3}{c|}{\BRFS}& \multicolumn{3}{c|}{$\IW(1)$}& \multicolumn{3}{c|}{\BFS}& \multicolumn{3}{c|}{\UCT}
-& \multicolumn{1}{c}{\Deepmind} \\[.2ex]
+\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Random} & \multicolumn{3}{c|}{\BRFS}& \multicolumn{3}{c|}{$\IW(1)$}& \multicolumn{3}{c|}{\BFS}& \multicolumn{3}{c}{\UCT} \\[.2ex]
 \hline\\[-3ex]
 %% table body
-\multicolumn{1}{c}{Game}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c|}{SD}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} &  \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T} & \multicolumn{1}{c}{S}\\[.2ex]
+\multicolumn{1}{c}{Game}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c|}{SD}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} &  \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{T}\\[.2ex]
 \hline\\[-3ex]
-\Asterix        & 210    & 54.77     & 3900  & 5538.95   & 32.24 & 119000        & 4987.48   & 17.08 & 56320       & 4757.84   & 0.95  & \textbf{296300 (290700)} & 6657.7    & 10.82 & n/a\\
-\BeamRider     & 448.8  & 84.64     & 519.2 & 179.81    & 43.48 & 3000          & 1150.22   & 6.02  & 3804        & 446.18    & 0.92  & \textbf{4320 (6624)}     & 2175.26   & 42.91 & 4092\\
-\Breakout       & 1      & 0.71      & 2     & 1.41      & 51.56 & 383.6         & 9.84      & 1.37  & 421.2       & 13.66     & 3.95  & \textbf{512.4 (364)}     & 193.97    & 26.03 & 168\\
-\Enduro          & 0      & 0         & 0.4   & 0.89      & 40.92 & \textbf{500}  & 0         & 38.13 & 292.8       & 6.18      & 13.12 & 306.8 (286)              & 8.14      & 32.14 & 470\\
-\Freeway         & 0      & 0         & 0     & 0         & 71.84 & \textbf{29.4} & 0.89      & 15.19 & 19.6        & 3.29      & 4.44  & 1 (0)                    & 1.41      & 90.64 & n/a \\
-\Pong           & -20.4  & 0.55      & -20.2 & 0.84      & 41.51 & 16.4          & 1.67      & 18.45 & \textbf{21} & 0         & 39.69 & \textbf{21 (21)}         & 0         & 38.37 & 20 \\
-\Qbert           & 185    & 345.33    & 160   & 85.88     & 42.83 & 1545          & 1666.06   & 1.82  & 14725       & 7872.1    & 1.76  & \textbf{18730 (17343)}   & 1127.69   & 28.61 & 1952\\
-\Seaquest       & 192    & 38.99     & 200   & 14.14     & 40.01 & \textbf{4274} & 4206.81   & 14.72 & 4070        & 1939.55   & 1.59  & 3406 (5132)              & 2402.47   & 26.16 & 1705 \\
-\SpaceInvaders & 187    & 120.86    & 115   & 61.44     & 46.11 & \textbf{2670} & 637.3     & 12.1  & 1327        & 598.04    & 2.25  & 2634 (2718)              & 375.66    & 19.66 & 581
+\Asterix        & 210    & 54.77     & 3900  & 5538.95   & 32.24 & 119000        & 4987.48   & 17.08 & 56320       & 4757.84   & 0.95  & \textbf{296300 (290700)} & 6657.7    & 10.82\\
+\BeamRider     & 448.8  & 84.64     & 519.2 & 179.81    & 43.48 & 3000          & 1150.22   & 6.02  & 3804        & 446.18    & 0.92  & \textbf{4320 (6624)}     & 2175.26   & 42.91\\
+\Breakout       & 1      & 0.71      & 2     & 1.41      & 51.56 & 383.6         & 9.84      & 1.37  & 421.2       & 13.66     & 3.95  & \textbf{512.4 (364)}     & 193.97    & 26.03\\
+\Enduro          & 0      & 0         & 0.4   & 0.89      & 40.92 & \textbf{500}  & 0         & 38.13 & 292.8       & 6.18      & 13.12 & 306.8 (286)              & 8.14      & 32.14\\
+\Freeway         & 0      & 0         & 0     & 0         & 71.84 & \textbf{29.4} & 0.89      & 15.19 & 19.6        & 3.29      & 4.44  & 1 (0)                    & 1.41      & 90.64\\
+\Pong           & -20.4  & 0.55      & -20.2 & 0.84      & 41.51 & 16.4          & 1.67      & 18.45 & \textbf{21} & 0         & 39.69 & \textbf{21 (21)}         & 0         & 38.37\\
+\Qbert           & 185    & 345.33    & 160   & 85.88     & 42.83 & 1545          & 1666.06   & 1.82  & 14725       & 7872.1    & 1.76  & \textbf{18730 (17343)}   & 1127.69   & 28.61\\
+\Seaquest       & 192    & 38.99     & 200   & 14.14     & 40.01 & \textbf{4274} & 4206.81   & 14.72 & 4070        & 1939.55   & 1.59  & 3406 (5132)              & 2402.47   & 26.16\\
+\SpaceInvaders & 187    & 120.86    & 115   & 61.44     & 46.11 & \textbf{2670} & 637.3     & 12.1  & 1327        & 598.04    & 2.25  & 2634 (2718)              & 375.66    & 19.66
 \end{tabular}

### 2014-10-31 workshop-paper/paper.tex
Table 2 Changes
* Removed the Deepmind column
* Removed the Standard Deviation column
* Column headers now use "long names"
* Synced the caption with the changes made to the table

@@
 \begin{table*}
+\centering
 \addtolength{\tabcolsep}{-2pt}
 {\scriptsize
-\begin{tabular}{@{}l|rr|rrr|rrr|rrr|rrr@{}}
+\begin{tabular}{@{}l|r|rr|rr|rr|rr@{}}
 %% table header
 \hline\\[-3ex]
-\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Random} & \multicolumn{3}{c|}{\BRFS}& \multicolumn{3}{c|}{$\IW(1)$}& \multicolumn{3}{c|}{\BFS}& \multicolumn{3}{c}{\UCT} \\[.2ex]
+\multicolumn{1}{c|}{} & \multicolumn{1}{c|}{Random} & \multicolumn{2}{c|}{\BRFS}& \multicolumn{2}{c|}{$\IW(1)$}& \multicolumn{2}{c|}{\BFS}& \multicolumn{2}{c}{\UCT} \\[.2ex]
 \hline\\[-3ex]
 %% table body
-\multicolumn{1}{c}{Game}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c|}{SD}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} &  \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c|}{T}  & \multicolumn{1}{c}{S} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{T}\\[.2ex]
+\multicolumn{1}{c}{Game}  & 
+\multicolumn{1}{c|}{Avg. Score} & 
+\multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  & 
+\multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  & 
+\multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  & 
+\multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c}{Avg. Time}\\[.2ex]
 \hline\\[-3ex]
-\Asterix        & 210    & 54.77     & 3900  & 5538.95   & 32.24 & 119000        & 4987.48   & 17.08 & 56320       & 4757.84   & 0.95  & \textbf{296300 (290700)} & 6657.7    & 10.82\\
-\BeamRider     & 448.8  & 84.64     & 519.2 & 179.81    & 43.48 & 3000          & 1150.22   & 6.02  & 3804        & 446.18    & 0.92  & \textbf{4320 (6624)}     & 2175.26   & 42.91\\
-\Breakout       & 1      & 0.71      & 2     & 1.41      & 51.56 & 383.6         & 9.84      & 1.37  & 421.2       & 13.66     & 3.95  & \textbf{512.4 (364)}     & 193.97    & 26.03\\
-\Enduro          & 0      & 0         & 0.4   & 0.89      & 40.92 & \textbf{500}  & 0         & 38.13 & 292.8       & 6.18      & 13.12 & 306.8 (286)              & 8.14      & 32.14\\
-\Freeway         & 0      & 0         & 0     & 0         & 71.84 & \textbf{29.4} & 0.89      & 15.19 & 19.6        & 3.29      & 4.44  & 1 (0)                    & 1.41      & 90.64\\
-\Pong           & -20.4  & 0.55      & -20.2 & 0.84      & 41.51 & 16.4          & 1.67      & 18.45 & \textbf{21} & 0         & 39.69 & \textbf{21 (21)}         & 0         & 38.37\\
-\Qbert           & 185    & 345.33    & 160   & 85.88     & 42.83 & 1545          & 1666.06   & 1.82  & 14725       & 7872.1    & 1.76  & \textbf{18730 (17343)}   & 1127.69   & 28.61\\
-\Seaquest       & 192    & 38.99     & 200   & 14.14     & 40.01 & \textbf{4274} & 4206.81   & 14.72 & 4070        & 1939.55   & 1.59  & 3406 (5132)              & 2402.47   & 26.16\\
-\SpaceInvaders & 187    & 120.86    & 115   & 61.44     & 46.11 & \textbf{2670} & 637.3     & 12.1  & 1327        & 598.04    & 2.25  & 2634 (2718)              & 375.66    & 19.66
+\Asterix        & 210    & 3900  & 32.24 & 119000  & 17.08 & 56320  & 0.95  & \textbf{296300} (290700) & 10.82\\
+\BeamRider     & 448.8 & 519.2 & 43.48 & 3000          & 6.02  & 3804        & 0.92  & \textbf{4320} (6624)   & 42.91\\
+\Breakout       & 1    & 2     & 51.56 & 383.6         & 1.37  & 421.2       & 3.95  & \textbf{512.4} (364)   & 26.03\\
+\Enduro          & 0   & 0.4  & 40.92 & \textbf{500}   & 38.13 & 292.8      & 13.12 & 306.8 (286)             & 32.14\\
+\Freeway         & 0   & 0     & 71.84 & \textbf{29.4} & 15.19 & 19.6       & 4.44  & 1 (0)    & 90.64\\
+\Pong           & -20.4 & -20.2 & 41.51 & 16.4         & 18.45 & \textbf{21} & 39.69 & \textbf{21} (21) & 38.37\\
+\Qbert           & 185   & 160  & 42.83 & 1545         & 1.82  & 14725       & 1.76  & \textbf{18730} (17343) & 28.61\\
+\Seaquest       & 192   & 200   & 40.01 & \textbf{4274}& 14.72 & 4070        & 1.59  & 3406 (5132)      & 26.16\\
+\SpaceInvaders & 187    & 115   & 46.11 & \textbf{2670} & 12.1  & 1327       & 2.25  & 2634 (2718)  & 19.66
 \end{tabular}
@@
    setting the Atari 2600 random seeds to $0, \ldots, 4$, and maximum episode duration of $18,000$
-   frames. \textit{T} is the average CPU time per call to each action selection algorithm, \textit{S} is the average score achieved,
-   and \textit{SD} is the standard deviation. \UCT~performance
+   frames. For every algorithm, minus the trivial random baseline algorithm, we report the average CPU time per call 
+   to each action selection algorithm, along with the average score achieved. \UCT~performance
    reported in \cite{bellemare:jair2013} is shown in parenthesis. Numbers

### 2014-10-31 workshop-paper/paper.tex
Changes to Experimental Section
* Made changes to the text as per Hector's suggestions

@@
 both in terms of  average scores and average times per call. \BFS\ 
-is generally faster than $\IW(1)$ but only outperforms \UCT\ in $2$ of the games (\Freeway and
-\Seaquest). Compared with the UCT results from  \cite{bellemare:jair2013}, shown in parenthesis, 
-$\IW(1)$ still outperforms \UCT\ in $3$  of the  $9$ games (\Breakout, \Enduro~and \Freeway). 
+is generally faster than $\IW(1)$ but only outperforms \UCT\ in $2$ of the games (\Freeway\ and
+\Seaquest). Compared with the \UCT\ results from  \cite{bellemare:jair2013}, shown in parenthesis, 
+both $\IW(1)$ and \BFS\ still outperform \UCT\ in $3$  of the  $9$ games (\Breakout, \Enduro\ and \Freeway),
+with \BFS\ and \UCT\ tied in another game, \Pong. 
 As mentioned in the previous section, the order in which $\IW(1)$ and \BFS\ generate the children

### 2014-10-31 workshop-paper/paper.tex
Macros
* Usage of macros for IW, BFS and UCT enforced

@@
 %\newcommand{\BRFS}{{\textit{BrFS}}}
-%\newcommand{\BFS}{{\textit{2BFS}}}
+%\newcommand{\BFS}{{\textit{\BFS}}}
 %\newcommand{\UCT}{{\textit{UCT}}}
@@
 costs (rewards) or goals, and yet, like  heuristic search algorithms, they have been shown to be   effective for solving   problems over huge 
-state spaces. The simplest such algorithm, called Iterated Width or IW, consists of  a sequence of calls  IW(1), IW(2), .., IW(k) where  IW(i) is  
+state spaces. The simplest such algorithm, called Iterated Width or \IW, consists of  a sequence of calls  $\IW(1)$, $\IW(2)$, 
+.., $\IW(k)$ where  $\IW(i)$ is  
 a breadth-first search in which a state is  pruned when it is not the first state in the search  to make true some set of i atoms or less.
-Preliminary empirical results suggest that the performance of IW with the  $k$ parameter fixed to $1$, 
-i.e., IW(1), is much better than random or breadth-first search, and approaches 
-the performance of UCT. A simple best-first variation  of IW,  that combine exploration and exploitation,
+Preliminary empirical results suggest that the performance of \IW\ with the  $k$ parameter fixed to $1$, 
+i.e., $\IW(1)$, is much better than random or breadth-first search, and approaches 
+the performance of \UCT. A simple best-first variation  of \IW,  that combine exploration and exploitation,
 is evaluated as well. 
@@
 Results have been reported so far for basic planning algorithms like breadth-first 
-search and UCT \cite{bellemare:jair2013}, and for  reinforcement learning
+search and \UCT\ \cite{bellemare:jair2013}, and for  reinforcement learning
 and evolutionary  methods \cite{bellemare:jair2013,deep-mind-atari,risto-texas:atari-games}.
@@
 methods like breadth-first search, Dijkstra's algorithm \cite{dijkstra},  or learning methods  such as 
-LRTA* \cite{korf:lrta}, UCT \cite{uct}, and Q-learning  \cite{sutton:rl,bertsekas:neuro}. Indeed, 
+LRTA* \cite{korf:lrta}, \UCT\ \cite{uct}, and Q-learning  \cite{sutton:rl,bertsekas:neuro}. Indeed, 
 the net-benefit planning problem with unknown state transitions and rewards over a given planning
@@
 they manage to search large  state spaces effectively. The basic algorithm in this class is 
-called IW for Iterated Width search \cite{nir:ecai2012}. IW consists of a sequence of calls  IW(1), IW(2), .., IW(k), where 
-IW(i) is  a  standard breadth-first search where states are pruned right away when they fail to make true some new tuple (set) 
-of at most $i$ atoms. Namely,   IW(1) is a breadth-first search that keeps  a state only if  the state is   the first  one 
-in the search to make some atom  true; IW(2) keeps a state  only if the state is  the first one to  make  a pair of atoms true, and 
-so on.  Like plain breadth-first  and iterative deepening searches,  IW  is 
+called \IW\ for Iterated Width search \cite{nir:ecai2012}. \IW\ consists of a sequence of calls  $\IW(1)$, $\IW(2)$, .., $\IW(k)$, where 
+$\IW(i)$ is  a  standard breadth-first search where states are pruned right away when they fail to make true some new tuple (set) 
+of at most $i$ atoms. Namely,   $\IW(1)$ is a breadth-first search that keeps  a state only if  the state is   the first  one 
+in the search to make some atom  true; $\IW(2)$ keeps a state  only if the state is  the first one to  make  a pair of atoms true, and 
+so on.  Like plain breadth-first  and iterative deepening searches,  \IW\  is 
 complete, while searching the state space in a way that makes use of the 
@@
 While a normal, complete breadth-first search runs in time that is exponential in the total number of variables $X_i$, 
-the procedure $IW(k)$ runs in time that is exponential in the $k$ parameter, $1 \leq k \leq n$. 
-Elsewhere, it has been shown that  IW exhibits   state-of-the-art performance on most existing planning benchmark domains 
-when goals are restricted to be single atoms  where IW provably runs in low polynomial time in the number of
+the procedure $\IW(k)$ runs in time that is exponential in the $k$ parameter, $1 \leq k \leq n$. 
+Elsewhere, it has been shown that  \IW\ exhibits   state-of-the-art performance on most existing planning benchmark domains 
+when goals are restricted to be single atoms  where \IW\ provably runs in low polynomial time in the number of
 variables \cite{nir:ecai2012}. For example, in a block world instance with any number of blocks and any initial
 configuration,  the goal of  placing a given  block on top of another will be achievable in  quadratic time  
-in the number of blocks as any such instance can be shown to  have \emph{width} 2. This means  that IW(2) will be 
+in the number of blocks as any such instance can be shown to  have \emph{width} 2. This means  that $\IW(2)$ will be 
 not only complete then, but also optimal. This is also true for most other benchmark domains in classical planning
@@
 For dealing with the benchmarks where goals are  not single atoms, e.g., where a given \emph{tower} of blocks needs to be built, 
-simple extensions of the basic IW procedure have been developed as well, such as Serialized IW, where IW is 
+simple extensions of the basic \IW\ procedure have been developed as well, such as Serialized \IW, where \IW\ is 
 used to achieve the goals one at at time \cite{nir:ecai2012,nir:ecai2014}. 
-The extension of IW for problems where there is no goal but an
+The extension of \IW\ for problems where there is no goal but an
 additive reward measure  to be optimized, like in the Atari games, is direct.
@@
 
-The Iterative Width  (IW) algorithm has been introduced as a classical planning algorithm
+The Iterative Width  (\IW) algorithm has been introduced as a classical planning algorithm
 that takes a planning problem as an input, and computes an action sequence that solves the problem 
@@
 
-IW consists of a sequence of calls  $\IW(i)$ for $i=0, 1, 2, \ldots$
+\IW consists of a sequence of calls  $\IW(i)$ for $i=0, 1, 2, \ldots$
 over a problem $P$ until a termination condition is reached. 
@@
 as in the Atari games, the termination condition is given by a time window or a maximum number of generated nodes.
-The  best path found by IW is the path  that has a  maximum accumulated reward. 
-The accummulated reward $R(s)$  of a state $s$ reached in an iteration of IW is determined by the  unique parent state $s'$ 
+The  best path found by \IW is the path  that has a  maximum accumulated reward. 
+The accummulated reward $R(s)$  of a state $s$ reached in an iteration of \IW is determined by the  unique parent state $s'$ 
 and action $a$ leading to $s$ from $s'$ as $R(s) = R(s') + r(a,s')$. The best state is the state $s$ with maximum reward
-$R(s)$ generated but not pruned by IW, and the best path is the one that leads to the state $s$ from the current state.
+$R(s)$ generated but not pruned by \IW, and the best path is the one that leads to the state $s$ from the current state.
 The action selected to be done next is then the first action along such a path.
@@
 %\toprule
-\# & \multicolumn{1}{c}{Domain} & \multicolumn{1}{c}{I} & \multicolumn{1}{c}{IW(1)} & \multicolumn{1}{c}{IW(2)} & \multicolumn{1}{c}{Neither} \\[.2ex]
+\# & \multicolumn{1}{c}{Domain} & \multicolumn{1}{c}{I} & \multicolumn{1}{c}{$\IW(1)$} & \multicolumn{1}{c}{$\IW(2)$} & \multicolumn{1}{c}{Neither} \\[.2ex]
 
@@
 \caption{\small 
-\emph{Top:} Number of classical planning problems  and percentages solved by IW(1), IW(2), or neither. Instances obtained
+\emph{Top:} Number of classical planning problems  and percentages solved by $\IW(1)$, $\IW(2)$, or neither. Instances obtained
 by splitting benchmarks  with $N$ atomic goals into $N$ problems with atomic goals.
@@
 
-IW is a systematic and complete blind-search algorithm like breadth-first search (BrFS) and iterative deepening (ID),
+\IW\ is a systematic and complete blind-search algorithm like breadth-first search (\BRFS) and iterative deepening (ID),
 but   unlike these algorithms, it uses the factored representation of the states in terms of variables to structure the 
@@
 Table~\ref{table1} from \cite{nir:ecai2012} shows 
-the percentage of instances that are  solved by  the first iteration of IW, i.e.  IW(1), by the second iteration 
-IW(2), and by neither one.  These are instances that  have been obtained from the existing benchmarks
+the percentage of instances that are  solved by  the first iteration of \IW, i.e.  $\IW(1)$, by the second iteration 
+$\IW(2)$, and by neither one.  These are instances that  have been obtained from the existing benchmarks
 by splitting problems with $N$ atomic goals, into $N$ problems with one atomic goal each. As the table shows, 
-$37\%$ of the 37921 instances are solved by IW(1) while $51.3\%$ are  solved by IW(2). Since IW($k$)
+$37\%$ of the 37921 instances are solved by $\IW(1)$ while $51.3\%$ are  solved by $\IW(2)$. Since $\IW(k)$
 runs in time that is exponential in $k$, this mean that almost $90\%$ of the 37921 instances are  solved in time that
@@
 
-The min $k$ value for which $IW(k)$ solves a problem  is clearly bounded and small in most of
+The min $k$ value for which $\IW(k)$ solves a problem  is clearly bounded and small in most of
 these instances. This is  actually no  accident and  has a \emph{theoretical explanation.} Lipovetzky and Geffner define
@@
 \emph{any} solvable instance with atomic goals will  have a bounded and small width that is independent of the number
-of variables and states in the problem. The min value $k$ for which the iteration  $IW(k)$ solves
-the problem cannot exceed the problem width, so  the algorithm \emph{IW runs in time and space
+of variables and states in the problem. The min value $k$ for which the iteration  $\IW(k)$ solves
+the problem cannot exceed the problem width, so  the algorithm \emph{\IW\ runs in time and space
 that are exponential in the problem width.}
@@
 relevant  in  optimization problems as well. Indeed, if a  good path is made of 
-states $s_i$ each of which have a low width, IW can be made 
+states $s_i$ each of which have a low width, \IW\ can be made 
 to find such path in low polynomial time for a small value of the $k$ parameter.
-Later on we will discuss  a slight  change  required in IW to enforce this property
+Later on we will discuss  a slight  change  required in \IW\ to enforce this property
 which is actually  not yet implemented in the code that is tested below.
@@
 
-The number of nodes \emph{generated} by IW(1) is $n \times D \times b$ in the worst case, 
+The number of nodes \emph{generated} by $\IW(1)$ is $n \times D \times b$ in the worst case, 
 where $n$ is the number of variables, $D$ is the size of their domains, and $b$ is the number of 
@@
 equal to $589,824$, which is large but feasible. On the other hand, the number
-of nodes generated by IW(2) in the worst case is  $(n \times D)^2 \times b$, which is
+of nodes generated by $\IW(2)$ in the worst case is  $(n \times D)^2 \times b$, which is
 equal to $1,073,741,824$ which is too large, forcing us to consider only a fraction of 
 such states. For classical planning problems, the growth in the number of nodes from 
-IW(1) to IW(2) is not that large, as the variables are boolean. Indeed, we could have
+$\IW(1)$ to $\IW(2)$ is not that large, as the variables are boolean. Indeed, we could have
 taken the state vector for the Atari games as a vector of 1024 boolean variables, 
 and apply these algorithms to that representation. The results would be different.
-In such a case, IW(1) and IW(2)  would generate $n' \times D' \times b$ and 
+In such a case, $\IW(1)$ and $\IW(2)$  would generate $n' \times D' \times b$ and 
 $(n' \times D')^2 \times b$ states, which for $n'=1024$ and $D=2$
@@
 the 128 words of the state vectors are meaningful. 
-In summary, from the basic IW algorithm we are testing only the first 
-call IW(1).
+In summary, from the basic \IW\ algorithm we are testing only the first 
+call $\IW(1)$.
 
-IW is a purely exploration algorithm that does not take into account  the accumulated 
+\IW\ is a purely exploration algorithm that does not take into account  the accumulated 
 reward for selecting the next states to consider. As a simple, variant that combines
@@
 placed in the first queue only. We refer to this best-first algorithm 
-as 2BFS.
+as \BFS.
 
-For  the experiments below, we added two simple variations to  IW(1) and 2BFS.
-First, in the breadth-first search underlying IW(1), we generate the children in 
+For  the experiments below, we added two simple variations to  $\IW(1)$ and \BFS.
+First, in the breadth-first search underlying $\IW(1)$, we generate the children in 
 random order. This makes the executions that results from the
-IW(1) lookahead less susceptible to fall into loops; a potential
+$\IW(1)$ lookahead less susceptible to fall into loops; a potential
 problem in local search algorithms with no memory or learning.
 Second, a discount factor $\gamma=0.995$ is used in both algorithms 
-for discounting future rewards like in UCT. 
+for discounting future rewards like in \UCT. 
 For this, each state $s$ keeps its depth $d(s)$ in the search tree,
@@
 for each  game can take up to $40$ hours of CPU time. Second, while \UCT\ is still limited to perform up to $500$ base policy rollouts 
-of depth $300$, a maximum of budget of  $100,000$ simulated frames is applied to all the algorithms, UCT included. 
+of depth $300$, a maximum of budget of  $100,000$ simulated frames is applied to all the algorithms, \UCT\ included. 
 %% Was not clear; hope this is what you meant, if not fix
@@
 though classical planning is PSPACE-complete \cite{bylander}. A  partial answer is that 
-most of these domains have a low width, and hence, can be solved in low polynomial time (by IW)
+most of these domains have a low width, and hence, can be solved in low polynomial time (by \IW)
 when goals contain a single atom.  Benchmark problems with multiple atomic goals tend to be 
@@
 the robot in relation to the locations that are reached when following the other controllers
-(a diversity measure). The novelty measure used by IW, on the other hand, is domain-independent and it 
+(a diversity measure). The novelty measure used by \IW, on the other hand, is domain-independent and it 
 is determined by the structure of the states as captured by the problem variables.
-
-% @article{lehman:novelty,
-%  title={Abandoning objectives: Evolution through the search for novelty alone},
-%  author={Lehman, Joel and Stanley, Kenneth O},
-%  journal={Evolutionary computation},
-%  volume={19},
-%  number={2},
-%  pages={189--223},
-%  year={2011},
-%  publisher={MIT Press}
-%}
 
@@
 over small time windows, where like breadth-first search it's likely not to find any rewards
-at all. The UCT algorithm provides a method for balancing exploration and exploitation,
-which  is  effective over small time windows. The 2BFS algorithm above with two queues
+at all. The \UCT\ algorithm provides a method for balancing exploration and exploitation,
+which  is  effective over small time windows. The \BFS\ algorithm above with two queues
 that alternate, one guided by the novelty measures  and the other by the accumulated reward,
@@
 second in a bounded number of steps. Alternatively, novelty-guided probes, possibly with 
-some noise added, could potentially pay off as  a  base policy for UCT in replacement 
+some noise added, could potentially pay off as  a  base policy for \UCT\ in replacement 
 of random rollouts, as they apparently manage to exploit the structure of many problems.
@@
 
-We have shown experimentally that width-based algorithms like IW(1) and 2BFS
+We have shown experimentally that width-based algorithms like $\IW(1)$ and \BFS\ 
 that originate in work in classical planning, can be used in the Atari games
@@
 Second, we want to evaluate the various algorithms for different fixed time windows for decisions.
-In the table above, for example, UCT is often taking an order-of-magnitude more time per decision 
-than  2BFS. Third, we need to revise the way duplicate states are handled in IW and 2BFS to ensure
+In the table above, for example, \UCT\ is often taking an order-of-magnitude more time per decision 
+than  \BFS. Third, we need to revise the way duplicate states are handled in \IW\ and \BFS\ to ensure
 that the best paths among the states that have been expanded are returned. Finally, 

### 2014-11-07 src/agents/BestFirstSearch.cpp
Added the correct reuse, like uct that doesnt discount budget

@@ -47,7 +47,7 @@ int BestFirstSearch::expand_node( TreeNode* curr_node )
 
 	
 	}
-
+	
 	for (int a = 0; a < num_actions; a++) {
 		Action act = curr_node->available_actions[a];
 		
@@ -72,6 +72,8 @@ int BestFirstSearch::expand_node( TreeNode* curr_node )
 			    m_gen_count_novelty2++;
 			}
 			child->fn += ( m_max_reward - child->discounted_accumulated_reward ); // Miquel: add this to obtain Hector's BFS + m_max_reward * (720 - child->depth()) ;
+
+			child->num_nodes_reusable = curr_node->num_nodes_reusable + num_actions;
 
 			if (child->depth() > m_max_depth ) m_max_depth = child->depth();
 			num_simulated_steps += child->num_simulated_steps;
@@ -156,7 +158,7 @@ int BestFirstSearch::reuse_branch(TreeNode* node) {
 		if ( curr_node->depth() > m_reward_horizon - 1 ) continue;
 		if (!node->v_children.empty()) {
 			for(size_t c = 0; c < node->v_children.size(); c++) {			
-				TreeNode* child = curr_node->v_children[c];
+				TreeNode* child = curr_node->v_children[c];				
 				
 				// This recreates the novelty table (which gets resetted every time
 				// we change the root of the search tree)
@@ -177,6 +179,11 @@ int BestFirstSearch::reuse_branch(TreeNode* node) {
 				child->updateTreeNode();
 				child->fn += ( m_max_reward - child->discounted_accumulated_reward ); // Miquel: add this to obtain Hector's BFS + m_max_reward * (720 - child->depth()) ;
 				
+				//First applicable action
+				if(child->depth() == 1)
+				    child->num_nodes_reusable = child->num_nodes();
+				else
+				    child->num_nodes_reusable = curr_node->num_nodes_reusable;
 				if (child->depth() > m_max_depth ) m_max_depth = child->depth();
 				
 				
@@ -199,10 +206,10 @@ int BestFirstSearch::reuse_branch(TreeNode* node) {
 				
 			}
 		}
-		// Stop once we have simulated a maximum number of steps
-		if (num_simulated_steps >= max_sim_steps_per_frame) {
-			break;
-		}
+		// // Stop once we have simulated a maximum number of steps
[…]

### 2014-11-07 src/agents/IW1Search.cpp

@@ -96,6 +96,7 @@ int IW1Search::expand_node( TreeNode* curr_node, queue<TreeNode*>& q )
 	int num_simulated_steps =0;
 	int num_actions = available_actions.size();
 	bool leaf_node = (curr_node->v_children.empty());
+	static     int max_nodes_per_frame = max_sim_steps_per_frame / sim_steps_per_node;
 	m_expanded_nodes++;
 	// Expand all of its children (simulates the result)
 	if(leaf_node){
@@ -127,7 +128,7 @@ int IW1Search::expand_node( TreeNode* curr_node, queue<TreeNode*>& q )
 				curr_node->v_children[a] = child;
 				child->is_terminal = true;
 				m_pruned_nodes++;
-				continue;				
+				//continue;				
 			}
 			if (child->depth() > m_max_depth ) m_max_depth = child->depth();
 			num_simulated_steps += child->num_simulated_steps;
@@ -140,27 +141,33 @@ int IW1Search::expand_node( TreeNode* curr_node, queue<TreeNode*>& q )
 			// This recreates the novelty table (which gets resetted every time
 			// we change the root of the search tree)
 			if ( m_novelty_pruning )
+			{
+			    if( child->is_terminal )
+			    {
 				if ( check_novelty_1( child->state.getRAM() ) ){
-					update_novelty_table( child->state.getRAM() );
-					child->is_terminal = false;
+				    update_novelty_table( child->state.getRAM() );
+				    child->is_terminal = false;
 				}
 				else{
-					child->is_terminal = true;
-					m_pruned_nodes++;
+				    child->is_terminal = true;
+				    m_pruned_nodes++;
 				}
-			
+			    }
+			}
 			child->updateTreeNode();
 
 			if (child->depth() > m_max_depth ) m_max_depth = child->depth();
 
-			if ( !child->is_terminal )
-				num_simulated_steps += child->num_simulated_steps;
+			// DONT COUNT REUSED NODES
+			//if ( !child->is_terminal )
+			//	num_simulated_steps += child->num_simulated_steps;
 
 		}
 	
 		// Don't expand duplicate nodes, or terminal nodes
 		if (!child->is_terminal) {
 			if (! (ignore_duplicates && test_duplicate(child)) )
+			    if( child->num_nodes_reusable < max_nodes_per_frame )
 				q.push(child);
 		}
 	}
[…]

### 2014-11-07 src/agents/SearchAgent.cpp

@@ -145,12 +145,13 @@ Action SearchAgent::act() {
 		return m_curr_action;
 	
 	std::cout << "Search Agent action selection: frame=" << frame_number << std::endl;
+	std::cout << "Is Terminal Before Lookahead? " << m_rom_settings->isTerminal() << std::endl;
 	std::cout << "Evaluating actions: " << std::endl;
 
 	float t0 = aptk::time_used();
 
 	state = m_env->cloneState();
-
+	
 	if (search_tree->is_built ) {
 		// Re-use the old tree
 		search_tree->move_to_best_sub_branch();
@@ -178,6 +179,7 @@ Action SearchAgent::act() {
 
 	m_env->restoreState( state );
 	
+	std::cout << "Is Terminal After Lookahead? " << m_rom_settings->isTerminal() << std::endl;
 
 	float tf = aptk::time_used();
 

### 2014-11-07 src/agents/TreeNode.cpp

@@ -39,7 +39,9 @@ TreeNode::TreeNode(	TreeNode* parent, ALEState &parentState):
   discounted_accumulated_reward(0),
   discount(1.0),
   act(Action::PLAYER_A_NOOP),
-  already_expanded(false)
+  already_expanded(false),
+  num_nodes_reusable(0)
+
 {
 }
 
@@ -62,7 +64,8 @@ TreeNode::TreeNode(	TreeNode* parent, ALEState &parentState,
 	discount(disc),
 	original_discount(disc),
 	act(a),
-	already_expanded(false)
+	already_expanded(false),
+	num_nodes_reusable(0)
 {
 	if(parent == NULL){
 		m_depth = 0;

### 2014-11-07 src/agents/TreeNode.hpp

@@ -96,6 +96,8 @@ class TreeNode {
 	Action                  act;
 	ActionVect              available_actions;
 	bool                    already_expanded;
+    
+        unsigned                num_nodes_reusable;
 };
 
 

### 2014-11-07 src/common/Defaults.cpp

@@ -31,7 +31,7 @@ void setDefaultSettings(Settings &settings) {
     settings.setBool("run_length_encoding", true);
 
     // Environment customization settings
-    settings.setBool("record_trajectory", true);
+    settings.setBool("record_trajectory", false);
     //settings.setBool("restricted_action_set", false);
     settings.setBool("restricted_action_set", false);
     settings.setBool("use_starting_actions", true);
@@ -66,7 +66,7 @@ void setDefaultSettings(Settings &settings) {
 	//settings.setBool("iw1_stop_on_first_reward", true);
 	//settings.setInt("iw1_reward_horizon", 50 );
 	settings.setBool("iw1_stop_on_first_reward", false);
-	settings.setInt("iw1_reward_horizon", 60 );	
+	settings.setInt("iw1_reward_horizon", 300 );	
 	settings.setBool("randomize_successor_novelty", true );
 
 }

### 2014-11-07 workshop-paper/paper.tex

@@
 cannot be used  for selecting actions  for  two reasons: first,  no  compact PDDL-model of  the games is given, 
-and  more importantly,  the action effects and goals are  not known a priori. Moreover, in these games there is usually  
+and  more importantly,  the action effects and goals are  not known \emph{a priori}. Moreover, in these games there is usually  
 no set of goals to be achieved but rewards to be collected. These features do not preclude the use of   classical  algorithms like
@@
 have deterministic effects on the state, the problem of selecting the next action
-to be done cannot be addressed with state-of-the-art classical planners \cite{geffner:2013book}.
+to be done cannot be addressed with state-of-the-art classical planners \cite{geffner:book}.
 This is because there is  no compact PDDL-like encoding of the domain, and more 
@@
 \cite{sutton:book} over a deterministic MDP where the state transitions and rewards are not known, or alternatively,
-as a  \emph{net-benefit planning problem} \cite{ref-net-benefit-planning,keyder} with unknown state transitions  and  rewards. Namely, 
+as a  \emph{net-benefit planning problem} \cite{coles2012survey,keyder:jair09} with unknown state transitions  and  rewards. Namely, 
 we seek  an action sequence $a_0, a_1, \ldots, a_m$ from the given current  state $s_0$ that generates 
@@
 methods like breadth-first search, Dijkstra's algorithm \cite{dijkstra},  or learning methods  such as 
-LRTA* \cite{korf:lrta}, \UCT\ \cite{uct}, and Q-learning  \cite{sutton:rl,bertsekas:neuro}. Indeed, 
+LRTA* \cite{korf:lrta}, \UCT\ \cite{uct}, and Q-learning  \cite{sutton:book,bertsekas:neuro}. Indeed, 
 the net-benefit planning problem with unknown state transitions and rewards over a given planning
@@
 
-\Omit{
-The sequence of calls IW(1), IW(2), \ldots, IW(k) manages to combine the scope of ``blind'' search algorithms with the performance of 
-heuristic-search algorithms by exploring states  in an  order dictated by the ``width''  parameter $k$; 
-states that have width 1 are explored first in linear time, then states that have  width two are explored in quadratic time, and so on. 
-In classical planning problems, if the goal has a small width (something common in the classical planning domains when the  goal is a single atom, as
-we have seen above), IW will find goal states  in low polynomial time. Similarly, in the reward-based setting, if  there are states of high reward 
-that have small width, IW will find such states in low polynomial time, even if the state space is exponential in the number of variables. 
-In this paper, we aim to test whether these ideas work for the Atari games too  when  the RAM state of the emulator
-is used  as the system state. 
-}
+% \Omit{
+% The sequence of calls IW(1), IW(2), \ldots, IW(k) manages to combine the scope of ``blind'' search algorithms with the performance of 
+% heuristic-search algorithms by exploring states  in an  order dictated by the ``width''  parameter $k$; 
+% states that have width 1 are explored first in linear time, then states that have  width two are explored in quadratic time, and so on. 
+% In classical planning problems, if the goal has a small width (something common in the classical planning domains when the  goal is a single atom, as
+% we have seen above), IW will find goal states  in low polynomial time. Similarly, in the reward-based setting, if  there are states of high reward 
+% that have small width, IW will find such states in low polynomial time, even if the state space is exponential in the number of variables. 
+% In this paper, we aim to test whether these ideas work for the Atari games too  when  the RAM state of the emulator
+% is used  as the system state. 
+% }
 
@@
 The  best path found by \IW is the path  that has a  maximum accumulated reward. 
-The accummulated reward $R(s)$  of a state $s$ reached in an iteration of \IW is determined by the  unique parent state $s'$ 
+The accumulated reward $R(s)$  of a state $s$ reached in an iteration of \IW is determined by the  unique parent state $s'$ 
 and action $a$ leading to $s$ from $s'$ as $R(s) = R(s') + r(a,s')$. The best state is the state $s$ with maximum reward
@@
 $37\%$ of the 37921 instances are solved by $\IW(1)$ while $51.3\%$ are  solved by $\IW(2)$. Since $\IW(k)$
-runs in time that is exponential in $k$, this mean that almost $90\%$ of the 37921 instances are  solved in time that
-is either linear or quadratic in the number of problem variables, which in these encodings are all \emph{boolean}.  
+runs in time that is exponential in $k$, this mean that almost $90\%$ of the 37,921 instances are  solved in time that
+is either linear or quadratic in the number of problem variables, which in these encoding are all \emph{boolean}.  
 Furthermore, when the performance of \IW\   is compared with  breadth-first search and iterative deepening, 
 on the one hand,  and  with a  Greedy Best First Search guided by the additive heuristic $h_{add}$ 
-\cite{bonet:hsp-aij} on the other (this algorithm is similar 
+\cite{bonet:aij-hsp} on the other (this algorithm is similar 
 to one used by  the FF planner when hill climbing search fails \cite{hoffmann:ff}), 
-it turns out that ``blind'' \IW\ solves as many problems as the informed search, 34627 vs. 34849, far ahead
-of the other blind algorithms BrFS and ID  that solve 9010 and 8762 problems each. This is shown in the bottom
+it turns out that ``blind'' \IW\ solves as many problems as the informed search, 34,627 vs. 34,849, far ahead
+of the other blind algorithms \BRFS\ and \ID\  that solve 9,010 and 8,762 problems each. This is shown in the bottom
 part of the Table~\ref{table1}. Moreover, \IW\ is  faster and results in shorter plans than the heuristic search  
@@
 is used in the LAMA planner \cite{richter:lama}, and was introduced in the
-planner Fast Downward \cite{helmert:fd,malte:alternating}. In addition, we break ties
+planner Fast Downward \cite{helmert:fd}. In addition, we break ties
 in the first queue favoring states with largest accumulated reward, and 
@@
    reported in \cite{bellemare:jair2013} is shown in parenthesis. Numbers
-   in bold show best performer in terms of average score. Average score reported in 
-   \cite{deep-mind-atari} is reported for reference under the column \Deepmind.
+   in bold show best performer in terms of average score.
    }
@@
 appear to be relatively simple for current planners, even
-though classical planning is PSPACE-complete \cite{bylander}. A  partial answer is that 
+though classical planning is PSPACE-complete \cite{bylander:complexity}. A  partial answer is that 
 most of these domains have a low width, and hence, can be solved in low polynomial time (by \IW)
@@
 In these methods individuals in the population are not ranked according to the  optimization function
-but in terms of how ``novel'' thay are in relation to the rest of the population, thus encouraging diversity
+but in terms of how ``novel'' they are in relation to the rest of the population, thus encouraging diversity
 and exploration rather than (greedy) exploitation. The actual definition of novelty
@@
 some noise added, could potentially pay off as  a  base policy for \UCT\ in replacement 
-of random rollouts, as they apparently manage to exploit the structure of many problems.
+of random roll outs, as they apparently manage to exploit the structure of many problems.
 
@@
 issues, which we haven't managed to get in time for this submission. First, we want to extend the 
-evaluation to all the games featured in \cite{bellemare:jair}, and to add more runs per  game. 
+evaluation to all the games featured in \cite{bellemare:jair2013}, and to add more runs per  game. 
 Second, we want to evaluate the various algorithms for different fixed time windows for decisions.

### 2014-11-07 src/agents/IW1Search.cpp
IW1Search dont have a preference for noop

@@ -103,7 +103,7 @@ int IW1Search::expand_node( TreeNode* curr_node, queue<TreeNode*>& q )
 		curr_node->v_children.resize( num_actions );
 		curr_node->available_actions = available_actions;
 		if(m_randomize_successor)
-		    std::random_shuffle ( curr_node->available_actions.begin()+1, curr_node->available_actions.end() );
+		    std::random_shuffle ( curr_node->available_actions.begin(), curr_node->available_actions.end() );
 	
 	}
 	for (int a = 0; a < num_actions; a++) {

### 2014-11-10 src/agents/IW1Search.cpp
Hector's Post-Processing
* Duplicates are now taken into account as suggested by Hector's mail on
"open theoretical issues"
* Implementing Dijkstra isn't as easy as it seems

@@ -174,6 +174,21 @@ int IW1Search::expand_node( TreeNode* curr_node, queue<TreeNode*>& q )
 	return num_simulated_steps;
 }
 
+
+class TreeNodeComparerExploitation
+{
+public:
+	
+	bool operator()( TreeNode* a, TreeNode* b ) const 
+	{
+		if ( b->fn < a->fn ) return true;
+		return false;
+	}
+};
+
+typedef std::priority_queue<TreeNode*, std::vector<TreeNode*>, TreeNodeComparerExploitation >
+	Dijkstra_Queue;
+
 /* *********************************************************************
    Expands the tree from the given node until i_max_sim_steps_per_frame
    is reached
@@ -196,69 +211,117 @@ void IW1Search::expand_tree(TreeNode* start_node) {
 		}
 	    }
 	}
-
 	queue<TreeNode*> q;
-	std::list< TreeNode* > pivots;
-	
-	//q.push(start_node);
-	pivots.push_back( start_node );
-
+	q.push(start_node);
+	reward_t R_max = 0;
 	update_novelty_table( start_node->state.getRAM() );
 	int num_simulated_steps = 0;
 
 	m_expanded_nodes = 0;
 	m_generated_nodes = 0;
-
+	m_dup_nodes = 0;
+	m_dup_nodes_better_reward = 0;
 	m_pruned_nodes = 0;
 
-	do {
-		
-		std::cout << "# Pivots: " << pivots.size() << std::endl;
-		std::cout << "First pivot reward: " << pivots.front()->node_reward << std::endl;
-		pivots.front()->m_depth = 0;
-		int steps = expand_node( pivots.front(), q );
+	while(!q.empty()) {
+		// Pop a node to expand
+		TreeNode* curr_node = q.front();
+		q.pop();
+		TreeNode* other = m_closed.retrieve(curr_node);
+		if ( other != nullptr ) {
+			// We got a duplicated node!
+			m_dup_nodes++;
+			if ( curr_node->accumulated_reward > other->accumulated_reward ) {
+				// Replace node in parent children array
+				for (int a = 0; a < available_actions.size(); a++) {
+					TreeNode* child = other->p_parent->v_children[a];
+					if ( child == other ) {
+						other->p_parent->v_children[a] = curr_node;
+						break;
+					}
+				}
+				// Copy children into new node, and connect children to parent
+				for (int a = 0; a < available_actions.size(); a++) {
[…]

### 2014-11-10 src/agents/IW1Search.hpp

@@ -4,6 +4,7 @@
 #include "SearchTree.hpp"
 #include "bit_matrix.hxx"
 #include "../environment/ale_ram.hpp"
+#include "closed_list.hxx"
 
 #include <queue>
 
@@ -45,12 +46,14 @@ protected:
 	virtual void	clear();
 	virtual void	move_to_best_sub_branch();
 	
-	ALERAM 			m_ram;
-	aptk::Bit_Matrix*	m_ram_novelty_table;
-	unsigned		m_pruned_nodes;
-	bool			m_stop_on_first_reward;
-	unsigned		m_reward_horizon;	
-
+	ALERAM 				m_ram;
+	aptk::Bit_Matrix*		m_ram_novelty_table;
+	unsigned			m_pruned_nodes;
+	unsigned			m_dup_nodes;
+	unsigned			m_dup_nodes_better_reward;
+	bool				m_stop_on_first_reward;
+	unsigned			m_reward_horizon;	
+	aptk::Closed_List< TreeNode >	m_closed;		
 
 };
 

### 2014-11-10 src/agents/TreeNode.cpp

@@ -40,7 +40,8 @@ TreeNode::TreeNode(	TreeNode* parent, ALEState &parentState):
   discount(1.0),
   act(Action::PLAYER_A_NOOP),
   already_expanded(false),
-  num_nodes_reusable(0)
+  num_nodes_reusable(0),
+  m_dijkstra_depth(0)
 
 {
 }
@@ -65,7 +66,9 @@ TreeNode::TreeNode(	TreeNode* parent, ALEState &parentState,
 	original_discount(disc),
 	act(a),
 	already_expanded(false),
-	num_nodes_reusable(0)
+	num_nodes_reusable(0),
+  m_dijkstra_depth(0)
+
 {
 	if(parent == NULL){
 		m_depth = 0;

### 2014-11-10 src/agents/TreeNode.hpp

@@ -86,6 +86,7 @@ class TreeNode {
     	// Whether this node was flagged as a duplicate
     	bool duplicate;
         unsigned m_depth;
+	unsigned m_dijkstra_depth;
 
 	unsigned long long	fn; // evaluation function
         unsigned                novelty;

### 2014-11-10 src/environment/ale_state.cpp

@@ -16,20 +16,24 @@
 
 /** Default constructor - loads settings from system */ 
 ALEState::ALEState():
-  m_left_paddle(PADDLE_DEFAULT_VALUE),
-  m_right_paddle(PADDLE_DEFAULT_VALUE),
-  m_frame_number(0),
-  m_episode_frame_number(0),
-  m_screen(NULL){
+  	m_left_paddle(PADDLE_DEFAULT_VALUE),
+  	m_right_paddle(PADDLE_DEFAULT_VALUE),
+  	m_frame_number(0),
+  	m_episode_frame_number(0),
+  	m_screen(NULL),
+	m_hash(0),
+	m_hash_ready(false){
 }
 
 ALEState::ALEState(const ALEState &rhs, std::string serialized):
-  m_left_paddle(rhs.m_left_paddle),
-  m_right_paddle(rhs.m_right_paddle),
-  m_frame_number(rhs.m_frame_number),
-  m_episode_frame_number(rhs.m_episode_frame_number),
-  m_serialized_state(serialized),
-  m_screen(NULL) {
+  	m_left_paddle(rhs.m_left_paddle),
+  	m_right_paddle(rhs.m_right_paddle),
+  	m_frame_number(rhs.m_frame_number),
+ 	m_episode_frame_number(rhs.m_episode_frame_number),
+  	m_serialized_state(serialized),
+  	m_screen(NULL),
+	m_hash( 0 ),
+	m_hash_ready(false) {
 }
 
 ALEState::ALEState(ALEState &rhs, std::string serialized):
@@ -38,7 +42,9 @@ ALEState::ALEState(ALEState &rhs, std::string serialized):
   m_frame_number(rhs.m_frame_number),
   m_episode_frame_number(rhs.m_episode_frame_number),
   m_serialized_state(serialized),
-  m_screen(NULL){
+  m_screen(NULL),
+	m_hash( 0 ),
+	m_hash_ready(false) {
 }
 
 ALEState::ALEState( const ALEState& other )
@@ -55,6 +61,8 @@ ALEState::ALEState( const ALEState& other )
 		m_screen = new ALEScreen( *(other.m_screen) );
 	}
 	m_ram = other.m_ram;
+	m_hash = other.m_hash;
+	m_hash_ready = other.m_hash_ready;
 }
 
 const ALEState&	ALEState::operator=( const ALEState& other )
[…]

### 2014-11-10 src/environment/ale_state.hpp

@@ -24,6 +24,7 @@
 #include "../games/RomSettings.hpp"
 #include "ale_ram.hpp"
 #include "ale_screen.hpp"
+#include "jenkins_12bit.hxx"
 
 #define PADDLE_DELTA 23000
 // MGB Values taken from Paddles.cxx (Stella 3.3) - 1400000 * [5,235] / 255
@@ -50,8 +51,14 @@ class ALEState {
     void reset(int numResetSteps = 1);
 
     /** Returns true if the two states contain the same saved information */
-    bool equals(ALEState &state);
+    bool equals(const ALEState &state) const;
 
+	bool operator==( const ALEState& s ) const {
+		return equals( s );
+	} 
+	
+	size_t	hash();
+	
     void resetPaddles(Event*);
 
     /** Applies paddle actions. This actually modifies the game state by updating the paddle
@@ -109,8 +116,9 @@ class ALEState {
     string m_serialized_state; // The stored environment state, if this is a saved state
 
     ALERAM m_ram; // The current ALE RAM
-	ALEScreen* m_screen;
-
+	ALEScreen* 	m_screen;
+	size_t		m_hash;
+	bool		m_hash_ready;
 };
 
 #endif // __ALE_STATE_HPP__

### 2014-11-17 src/agents/BFS_UCB.cpp
Bug fix
* Fixed infinite loop when queue maximizing score and bonus is empty

@@ -354,7 +354,7 @@ void BFS_UCB::expand_tree(TreeNode* start_node) {
 	
 	if ( queues_empty() ) std::cout << "Search Space Exhausted!" << std::endl;
 	for ( unsigned k = 0; k < Q.size(); k++ ) {
-		std::cout << "Size(Q[" << k << "]) = " << Q[k]->size() << std::endl;
+		std::cout << "Size(Q[" << k << "]) = " << Q[k]->size()  << "V=" << V[k] << "N=" << N[k] << std::endl;
 	}
 		
 	update_branch_return(start_node);

### 2014-11-17 src/agents/BFS_UCB.hpp

@@ -45,10 +45,11 @@ public:
 	}
 
 	void		select_queue() {
-		m_current_queue = 0;
-		float best_value = V[m_current_queue] + exploration_bonus( m_current_queue );
+		m_current_queue = Q.size();
+		float best_value = -1.0f;
 
-		for ( unsigned k = 1; k < V.size(); k++ ) {
+		for ( unsigned k = 0; k < V.size(); k++ ) {
+			if ( Q[k]->empty() ) continue;
 			float f_k = V[k] + exploration_bonus(k);
 			if ( f_k > best_value ) {
 				m_current_queue = k;

### 2014-11-17 evaluate_agents_iw.py
specific script for each alg

@@ -62,8 +62,8 @@ def main() :
 		        #( 'random', '-player_agent random_agent' ),
 			#( 'brfs', '-player_agent search_agent -search_method brfs' ),
 		        ##( 'iw11', '-player_agent search_agent -search_method iw11' )
-			#( 'iw1', '-player_agent search_agent -search_method iw1' ),
-		        ( 'bfs', '-player_agent search_agent -search_method bfs' )
+			( 'iw1', '-player_agent search_agent -search_method iw1' )
+		        #( 'bfs', '-player_agent search_agent -search_method bfs' )
 		        #( 'uct', '-player_agent search_agent -search_method uct' )
 		 ]
 

### 2014-11-17 evaluate_agents_iw.py
added secret for parallel python cluster

@@ -40,7 +40,7 @@ def main() :
 
 	# tuple of all parallel python servers to connect with
 	ppservers = ()#address of server, default=localhost
-	job_server = pp.Server(num_cpus,ppservers=ppservers,socket_timeout=2) 
+	job_server = pp.Server(num_cpus,ppservers=ppservers,socket_timeout=2, secret="password") 
 
 	print "Starting pp with", job_server.get_ncpus(), "workers"
 

### 2014-11-18 icaps15-paper/paper.tex
ICAPS-15 paper sources

--- 2014-11-07 workshop-paper/paper.tex
@@
-  
 \documentclass[letterpaper]{article}
@@
 \usepackage{courier}
+\usepackage{url}
 \frenchspacing
@@
 /Title (Classical Planning Algorithms on the Atari Video Games: Preliminary Results)
-/Author (Nir Lipovetzky, Miquel Ramirez, Hector Geffner)}
+/Author ( )}
+% (Nir Lipovetzky, Miquel Ramirez, Hector Geffner)}
 
@@
 
-\author{Nir, Miquel, H}
+\author{\textbf{ID:} 52 \ \ ; \ \ \textbf{Keywords:} classical planning, video games}
+
+% \author{Nir Lipovetzky \\
+%       The University of Melbourne \\
+%    Melbourne, Australia\\
+%    {\normalsize\url{@unimelb.edu.au}}
+%    \And
+%    Miquel Ramirez \\
+%    RMIT University \\
+%    Melbourne, Australia\\
+%   {\normalsize\url{@rmit.edu.au}}
+%    \And
+%    Hector Geffner \\
+%    ICREA \&  Universitat Pompeu Fabra \\
+%    Barcelona, SPAIN \\
+%    {\normalsize\url{@upf.edu}}\thanks{firstname.lastname}
+% }
+
 %AAAI Press\\
@@
 costs (rewards) or goals, and yet, like  heuristic search algorithms, they have been shown to be   effective for solving   problems over huge
-state spaces. The simplest such algorithm, called Iterated Width or \IW, consists of  a sequence of calls  $\IW(1)$, $\IW(2)$, 
-.., $\IW(k)$ where  $\IW(i)$ is  
-a breadth-first search in which a state is  pruned when it is not the first state in the search  to make true some set of i atoms or less.
-Preliminary empirical results suggest that the performance of \IW\ with the  $k$ parameter fixed to $1$, 
-i.e., $\IW(1)$, is much better than random or breadth-first search, and approaches 
-the performance of \UCT. A simple best-first variation  of \IW,  that combine exploration and exploitation,
-is evaluated as well. 
+state spaces. The simplest such algorithm, called Iterated Width or \IW, consists of  a sequence of calls  $\IW(1)$, $\IW(2)$, \ldots,
+$\IW(k)$ where  $\IW(i)$ is  a breadth-first search in which a state is  pruned when it is not the
+first state in the search  to make true some set of $i$ atoms. 
+The  empirical results suggest that the performance of \IW\ with the  $k$ parameter fixed to $1$, 
+i.e., $\IW(1)$, approaches the performance of the state of the art represented by 
+ \UCT, and that a simple best-first variation  of \IW\  that combines exploration 
+and exploitation is even more competitive.
+
 \end{quote}
@@
 Results have been reported so far for basic planning algorithms like breadth-first
-search and \UCT\ \cite{bellemare:jair2013}, and for  reinforcement learning
-and evolutionary  methods \cite{bellemare:jair2013,deep-mind-atari,risto-texas:atari-games}.
+search and  \UCT,  reinforcement learning algorithms, and evolutionary  methods \cite{bellemare:jair2013,deep-mind-atari,risto-texas:atari-games}.
 The empirical results are impressive in some cases, yet a lot remains to be done, as no method  approaches   the
@@
 While all  these games feature a known initial (RAM) state and actions that
-have deterministic effects on the state, the problem of selecting the next action
+have deterministic effects, the problem of selecting the next action
 to be done cannot be addressed with state-of-the-art classical planners \cite{geffner:book}.
@@
 often no goals  but rewards $r(a,s)$ that  depend on the action $a$ and
-the state $s$ in  which  the action is performed, and these rewards are not known a priori 
-either.\footnote{Actually, in the Atari games, the rewards $r(a,s)$ depend only
+the state $s$ in  which  the action is performed, and these rewards are not known.\footnote{Actually, in the Atari games, the rewards $r(a,s)$ depend only
 on the state $f(a,s)$ that results from doing action $a$ in the state $s$.}
-Thus no model of the  goals  or the  rewards can be used to bias the search, and hence, 
-standard heuristic search and classical planning methods can't be used.
+Thus no model of the  goals  or the  rewards can be used to bias the search.
 
@@
 \cite{sutton:book} over a deterministic MDP where the state transitions and rewards are not known, or alternatively,
-as a  \emph{net-benefit planning problem} \cite{coles2012survey,keyder:jair09} with unknown state transitions  and  rewards. Namely, 
-we seek  an action sequence $a_0, a_1, \ldots, a_m$ from the given current  state $s_0$ that generates 
+as a  \emph{net-benefit planning problem} \cite{coles2012survey,keyder:jair09} with unknown state transitions  and  rewards. 
+Namely, we seek  an action sequence $a_0, a_1, \ldots, a_m$ from the  current  state $s_0$ that generates
 a state sequence $s_0, s_1, \ldots, s_{m+1}$ with  maximum  total reward $\sum_{i=0}^{m} r(a_i,s_i)$,
@@
 In this work, we thus turn to a recent class of  planning algorithms that combine the scope of blind search
-methods with the performance of state-of-the-art classical planning algorithms: namely, like ``blind'' search algorithms
+methods with the performance of state-of-the-art classical planners: namely, like ``blind'' search algorithms
 they do not require prior knowledge of state transitions, costs, or goals, and yet like heuristic algorithms
@@
 Elsewhere, it has been shown that  \IW\ exhibits   state-of-the-art performance on most existing planning benchmark domains
-when goals are restricted to be single atoms  where \IW\ provably runs in low polynomial time in the number of
-variables \cite{nir:ecai2012}. For example, in a block world instance with any number of blocks and any initial
+when goals are restricted to single atoms  where \IW\ provably runs in low polynomial time in the number of
+variables \cite{nir:ecai2012}. For example, in a blocks world instance with any number of blocks and any initial
 configuration,  the goal of  placing a given  block on top of another will be achievable in  quadratic time
-in the number of blocks as any such instance can be shown to  have \emph{width} 2. This means  that $\IW(2)$ will be 
-not only complete then, but also optimal. This is also true for most other benchmark domains in classical planning
-that turn out to  have small, bounded widths for any instance, as long as the goal is atomic. 
+in the number of blocks as any such instance can be shown to  have \emph{width} 2. This means  that $\IW(2)$ will
+not only be complete then, but also optimal. This is also true for most other benchmark domains in classical planning
+that turn out to  have small, bounded widths for \emph{any} instance as long as the goal is atomic.
 For dealing with the benchmarks where goals are  not single atoms, e.g., where a given \emph{tower} of blocks needs to be built,
-simple extensions of the basic \IW\ procedure have been developed as well, such as Serialized \IW, where \IW\ is 
-used to achieve the goals one at at time \cite{nir:ecai2012,nir:ecai2014}. 
+simple extensions of the basic \IW\ procedure have been developed  such as Serialized \IW, where \IW\ is
+used to achieve the goals one at at time following a simple goal ordering  \cite{nir:ecai2012,nir:ecai2014}.
 The extension of \IW\ for problems where there is no goal but an
@@
 look at  simple   variations of the algorithm that appear to be convenient  for the Atari
-games, present and analyze the preliminary experimental results, and discuss related and future work.
+games, present and analyze the  experimental results, and discuss related and future work.
 
@@
 
-The Iterative Width  (\IW) algorithm has been introduced as a classical planning algorithm
+The Iterated Width  (\IW) algorithm has been introduced as a classical planning algorithm
 that takes a planning problem as an input, and computes an action sequence that solves the problem
@@
 applies to a broader range of problems. We will characterize such
-problems  by means  a finite and discrete set of states (the state space)
+problems  by means of  a finite and discrete set of states (the state space)
 that correspond to vectors of size $n$. Namely, the
@@
 transition function $f$ such that  $s' = f(a,s)$ is
-the state that results from applying action $a \in A(s)$ to the state $s$, 
+the state that results from applying action $a$  to the state $s$,
 and rewards $r(a,s)$ represented by real numbers that result from  applying action $a$ in state $s$.
@@
 
-\IW consists of a sequence of calls  $\IW(i)$ for $i=0, 1, 2, \ldots$
+\IW \ consists of a sequence of calls  $\IW(i)$ for $i=0, 1, 2, \ldots$
 over a problem $P$ until a termination condition is reached.
@@
 \begin{itemize}
-\item The \emph{novelty of a newly generate state $s$  in a  breadth-first search}
+\item The \emph{novelty of a newly generate state $s$  in a  search algorithm}
 is   $1$ if  $s$ is the first state generated in the search that makes true some  atom $X=x$,
@@
 as in the Atari games, the termination condition is given by a time window or a maximum number of generated nodes.
-The  best path found by \IW is the path  that has a  maximum accumulated reward. 
-The accumulated reward $R(s)$  of a state $s$ reached in an iteration of \IW is determined by the  unique parent state $s'$ 
+The  best path found by \IW \ is the path  that has a  maximum accumulated reward.
+The accumulated reward $R(s)$  of a state $s$ reached in an iteration of \IW \ is determined by the  unique parent state $s'$
 and action $a$ leading to $s$ from $s'$ as $R(s) = R(s') + r(a,s')$. The best state is the state $s$ with maximum reward
 $R(s)$ generated but not pruned by \IW, and the best path is the one that leads to the state $s$ from the current state.
-The action selected to be done next is then the first action along such a path.
+The action selected to be done next is  the first action along  such a path.
 
@@
 \caption{\small
-\emph{Top:} Number of classical planning problems  and percentages solved by $\IW(1)$, $\IW(2)$, or neither. Instances obtained
-by splitting benchmarks  with $N$ atomic goals into $N$ problems with atomic goals.
+\emph{Top:} Number of classical planning instances per domain  and percentages solved by $\IW(1)$, $\IW(2)$, or neither.
+Problems  obtained by splitting benchmarks  with $N$ atomic goals into $N$ problems with atomic goals.
 \emph{Bottom:} Number of instances solved by \IW\ in comparison with  Iterative Deepening (\ID),
 Breadth-First Search (\BRFS),   and Greedy Best First Search (\textit{GBFS}) guided by  additive heuristic.
-Time and  memory outs  after 30 minutes and 2 GB resp. From \cite{nir:ecai2012}.
+Time and  memory outs  after 30 minutes or 2 GB. Table from \cite{nir:ecai2012}.
 }
@@
 achieves the dummy goal and that has the original conjunctive goal as a precondition.
-Yet, this changes the domain.}
+Yet, this changes the definition of the domain.}
 Table~\ref{table1} from \cite{nir:ecai2012} shows
@@
 \cite{bonet:aij-hsp} on the other (this algorithm is similar
-to one used by  the FF planner when hill climbing search fails \cite{hoffmann:ff}), 
+to the one used by  the FF planner when hill climbing search fails \cite{hoffmann:ff}),
 it turns out that ``blind'' \IW\ solves as many problems as the informed search, 34,627 vs. 34,849, far ahead
 of the other blind algorithms \BRFS\ and \ID\  that solve 9,010 and 8,762 problems each. This is shown in the bottom
-part of the Table~\ref{table1}. Moreover, \IW\ is  faster and results in shorter plans than the heuristic search  
+part of  Table~\ref{table1}. Moreover, \IW\ is  faster and results in shorter plans than a heuristic search algotrithm
 \cite{nir:ecai2012}.
 
-The min $k$ value for which $\IW(k)$ solves a problem  is clearly bounded and small in most of
+The min $k$ value for which $\IW(k)$ solves a problem  is bounded and small in most of
 these instances. This is  actually no  accident and  has a \emph{theoretical explanation.} Lipovetzky and Geffner define
@@
 relevant  in  optimization problems as well. Indeed, if a  good path is made of
-states $s_i$ each of which have a low width, \IW\ can be made 
+states $s_i$ each of which has a low width, \IW\ can be made
 to find such path in low polynomial time for a small value of the $k$ parameter.
-Later on we will discuss  a slight  change  required in \IW\ to enforce this property
-which is actually  not yet implemented in the code that is tested below.
+Later on we will discuss  a slight  change  required in \IW\ to enforce this property.
+% which is actually  not yet implemented in the code that is tested below.
 
@@
 represent  $36,864$ and $7,5497,472$ states respectively.
-These are feasible numbers, but we haven't tried  this representation yet
+These are feasible numbers, but we haven't used  this representation 
 under the assumption that the correlations among the bits in each one of
@@
 \IW\ is a purely exploration algorithm that does not take into account  the accumulated
-reward for selecting the next states to consider. As a simple, variant that combines
+reward for selecting the states to consider. As a simple variant that combines
 exploration and exploitation, we evaluated a \emph{best-first search} algorithm
@@
 \centering
-\addtolength{\tabcolsep}{-2pt}
-{\scriptsize
+% \addtolength{\tabcolsep}{-2pt}
+% {\scriptsize
+{\small
 \begin{tabular}{@{}l|r|rr|rr|rr|rr@{}}
 %% table header
-\hline\\[-3ex]
-\multicolumn{1}{c|}{} & \multicolumn{1}{c|}{Random} & \multicolumn{2}{c|}{\BRFS}& \multicolumn{2}{c|}{$\IW(1)$}& \multicolumn{2}{c|}{\BFS}& \multicolumn{2}{c}{\UCT} \\[.2ex]
-\hline\\[-3ex]
+% \hline\\[-3ex]
+%\hline
+\multicolumn{1}{c|}{} & \multicolumn{1}{c|}{Random} & \multicolumn{2}{c|}{\BRFS}& \multicolumn{2}{c|}{$\IW(1)$}& \multicolumn{2}{c|}{\BFS}&
+   \multicolumn{2}{c}{\UCT}
+% \\[.2ex]
+\\
+% \hline\\\\[-3ex]
+\hline
 %% table body
-\multicolumn{1}{c}{Game}  & 
-\multicolumn{1}{c|}{Avg. Score} & 
-\multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  & 
-\multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  & 
-\multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  & 
-\multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c}{Avg. Time}\\[.2ex]
-\hline\\[-3ex]
+\multicolumn{1}{c|}{Game}  &
+% \multicolumn{1}{c|}{Avg. Score} &
+% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
+% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
+% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
+% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c}{Avg. Time}\\[.2ex]
+ \multicolumn{1}{c|}{Score} &
+ \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
+ \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
+ \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
+ \multicolumn{1}{c}{Score} & \multicolumn{1}{c}{Time}
+% \\[.2ex]
+\\[.2ex]
+% \hline\\
+\hline
 \Asterix        & 210    & 3900  & 32.24 & 119000  & 17.08 & 56320  & 0.95  & \textbf{296300} (290700) & 10.82\\
@@
 
- \caption{\small Random baseline, \BRFS, $\IW(1)$, \BFS, \UCT~performance over 5 runs,
-   setting the Atari 2600 random seeds to $0, \ldots, 4$, and maximum episode duration of $18,000$
-   frames. For every algorithm, minus the trivial random baseline algorithm, we report the average CPU time per call 
-   to each action selection algorithm, along with the average score achieved. \UCT~performance
-   reported in \cite{bellemare:jair2013} is shown in parenthesis. Numbers
-   in bold show best performer in terms of average score.
-   }
+ \caption{\small Performance that results from random policy and various  lookahead algorithms in 9 Atari games.
+The algorithms  are \BRFS, $\IW(1)$, \BFS, and \UCT. Averages  over 5 runs shown with ``random'' seeds
+$0$ to $4$, and maximum episode duration of $18,000$  frames. Every algorithm limited to a budget of
+100,000 simulated frames, except for UCT limited to 500 rollouts, 300 frames deep instead. Average scores and CPU times per action selection shown.
+Performance of UCT results  reported in \cite{bellemare:jair2013} shown in parenthesis.
+Numbers  in bold show best performer in terms of average score. }
 \label{performance}
@@
 We tested  $\IW(1)$ and \BFS\  over $9$ games: those considered in the \emph{training} set
-of \cite{bellemare:jair2013} and those discussed in \cite{deep-mind-atari}. 
-In Table~\ref{performance} the performance of  $\IW(1)$ and \BFS\ 
+of \cite{bellemare:jair2013}, abbreviated  as BNVB, and those discussed in \cite{deep-mind-atari}.
+In Table~\ref{performance},  the performance of  $\IW(1)$ and \BFS\
 % , characterized by the average score obtained over all episodes, the standard deviation
@@
 is compared with that of \UCT, breadth-first search (\BRFS), and random action selection.
-Our experimental setup  follows the one in  \cite{bellemare:jair2013} with two exceptions. First, rather than
-considering $10$ episodes of each game, we have considered just $5$ due to time constraints, as the experiments 
-for each  game can take up to $40$ hours of CPU time. Second, while \UCT\ is still limited to perform up to $500$ base policy rollouts 
-of depth $300$, a maximum of budget of  $100,000$ simulated frames is applied to all the algorithms, \UCT\ included. 
-%% Was not clear; hope this is what you meant, if not fix
-%%% ($150,000$ frames are to be simulated while performing the lookahead), both $\IW(1)$ and \BFS are limited to
-% The other \UCT\ parameters like the exploration constant $C$ and the horizon $m$ are the same ($C=0.1$, $m=300$). 
-%% matching the setup used during actual gameplay. %%  (an action is selected every $5$ frames, using a no-op for the remaining $4$). %% Needed?
-The average scores reported for \UCT\  \cite{bellemare:jair2013}, without this extra limitation, are included
-in  parenthesis for reference. $\IW(1)$ and \BFS\ have been  limited to a planning horizon $m=1500$ where one an action is evaluated every 
-$5$ frames. The discount factor used by all the algorithms is $\gamma=0.995$, and all of them  reuse the tree built 
-in the previous lookahead by setting the root of the tree to the best child of the root, while 
-deleting its siblings and their descendants as in \cite{bellemare:jair2013}.
-%%% really need what's below? it's not very clear .. at least to me HG
-%% We note that  this may have a different effect on algorithms where all branches have the same number of
-%% generated nodes, as is the case of \BRFS, than in $\IW(1)$, \BFS\ and \UCT, where the depth of
-%% the branches is uneven.
+The discount factor used by all the algorithms is $\gamma=0.995$ and the average scores
+reported for \UCT\ in  BNVB  are included in  parenthesis for reference. 
+Our experimental setup  follows the one in BNVB with two exceptions: rather than
+considering $10$ runs per games, we considered just $5$ due to time constraints (*** ???),
+and  a maximum budget of  $150,000$ simulated frames is applied to all the  algorithms, 
+including  \UCT\ that is limited to  $500$ base policy rollouts of depth $300$.
+% \footnote{In \UCT,  $500$ base policy rollouts
+% of depth $300$ is equivalent to a budget of $150,000$ simulated frames.}.
+The bound on the number of simulated frames is like a bound on lookahead time, 
+as most of the time in the lookahead search is spent in the calls to the emulator
+for computing the next RAM state.  Also in order to save time, as in BNVB,   all of the algorithms 
+reuse the subtree from  the previous lookahead that is rooted in the selected
+child, deleting its siblings and their descendants. More precisely, no calls to the
+emulator are done  for transitions that are cached in that subtree. As a result
+the number of nodes in the lookahead tree may be larger actually than the budget.
+In addition, in  \IW, the states that are ``reused''  from the  previous searches
+are ignored in the computation of the novelty of new states so that more states 
+can  escape pruning. Otherwise, \IW\ often doesn't use the full budget. This is not 
+needed in \BFS\ which does no pruning.  $\IW(1)$ and \BFS\ have been  limited 
+search up to a depth of $m=1500$  frames and up to $m=20,000$ frames per root branch.
+This is to avoid the search for going too deep or being too committed to a single
+root action. (**** Please revise ****).
 
@@
 \BRFS\ rarely collects any reward, as the depth of the \BRFS\ search tree
-results in  a lookahead of $0.3$ seconds ($20$  frames or $4$ nodes). 
+results in  a lookahead of $0.3$ seconds ($20$  frames or $4$ nodes deep).
 On the other hand, both $\IW(1)$ and \BFS\ reach states that are  up to $350$--$1500$ frames deep
-($70$--$260$ nodes or $6$--$22$ seconds), even if $\IW(1)$ doesn't fully use the number of simulation frames allocated
-in many cases, as when  the set of unpruned states becomes  particularly small. 
+($70$--$260$ nodes or $6$--$22$ seconds), even if $\IW(1)$ doesn't always use all the simulation frames
+allocated due to its agressive pruning.
 %% case of \BFS, due to having a novelty of $0$ and the planning horizon being limited to $m$
@@
 
-$\IW(1)$ outperforms \UCT\ in $4$ out of $9$ games, 
-both in terms of  average scores and average times per call. \BFS\ 
-is generally faster than $\IW(1)$ but only outperforms \UCT\ in $2$ of the games (\Freeway\ and
-\Seaquest). Compared with the \UCT\ results from  \cite{bellemare:jair2013}, shown in parenthesis, 
-both $\IW(1)$ and \BFS\ still outperform \UCT\ in $3$  of the  $9$ games (\Breakout, \Enduro\ and \Freeway),
-with \BFS\ and \UCT\ tied in another game, \Pong. 
+Interestingly, simple $\IW(1)$ outperforms \UCT\ in $4$ out of $9$ games,
+both in terms of  average scores and average times per call. \BFS\ is
+faster than $\IW(1)$ and an order-of-magnitude faster than \UCT\ in general,
+and it  outscores \UCT\ in $2$ of the games (\Freeway\ and
+\Seaquest). Compared with the \UCT\ results in   \cite{bellemare:jair2013}, shown in parenthesis,
+both $\IW(1)$ and \BFS\  outperform \UCT\ in $3$ of the  $9$ games (\Breakout, \Enduro\ and \Freeway),
+with \BFS\ and \UCT\ tied in  \Pong. It'd  be very interesting to see the results of
+\BFS\ when given more time per action, but we don't have those results yet.
 As mentioned in the previous section, the order in which $\IW(1)$ and \BFS\ generate the children
@@
 
-We also  note that in \Freeway,  both $\IW(1)$ and \BFS\ attain a  performance similar to the 
+We also  note that in \Freeway,  both $\IW(1)$ and \BFS\ attain a score similar to the
 baseline semi-random algorithm \emph{Perturb} in  \cite{bellemare:jair2013},
@@
 Likewise,  $\IW(1)$ outperforms the best learning algorithm in \cite{deep-mind-atari}
-in $4$ out of the  $7$ games considered, and \BFS\ does so in $5$  of the $7$ games.
+in $4$ out of the  $7$ games considered, and \BFS\ does so in $5$  of
+the $7$ games.
 
@@
 it doesn't  appear to improve much on  \BRFS, even if  it results in a lookahead
-that is $5$ times deeper, yet still far too shallow  to compete with the other 
+that is $5$ times deeper,  still far too shallow  to compete with the other
 planning algorithms.
 
-\section{Discussion: Exploration and Exploitation}
+\section{Exploration and Exploitation}
 
@@
 when goals contain a single atom.  Benchmark problems with multiple atomic goals tend to be
-easy too, as the goals can often be achieved one at a time \cite{nir:ecai2012}.
+easy too, as the goals can often be achieved one at a time after a simple goal ordering \cite{nir:ecai2012}.
 
@@
 providing a structure to the search: states that have width 1 are explored first in linear time,
-then states that have  width two are explored in quadratic time, and so on. In classical planning problems 
+then states that have  width 2  are explored in quadratic time, and so on. In classical planning problems
 with atomic goals, this way of organizing the search  pays off both theoretically and practically.
@@
 The use of ``novelty measures'' for guiding an optimization search while ignoring the function that is being
-optimized, appears to be  common to novelty-based search methods developed independently 
+optimized is   common to the  novelty-based search methods developed independently
 in the context of genetic algorithms \cite{lehman:novelty}. %%% bibtex for this below
@@
 for guiding a robot in a maze, an individual controller  will not be ranked by how
-close it takes the robot to the goal (the greedy measure), but how far it takes
-the robot in relation to the locations that are reached when following the other controllers
-(a diversity measure). The novelty measure used by \IW, on the other hand, is domain-independent and it 
+close it takes the robot to the goal (the greedy measure), but by the distance between
+the locations that are  reachable with it, and the locations reachable with
+the  other controllers in the population (a diversity measure). The novelty measure used by \IW, on the other hand, is domain-independent and it
 is determined by the structure of the states as captured by the problem variables.
@@
 is not needed for optimality purposes, but just for improving the effectiveness of the
- lookahead search. Indeed, a best-first search algorithm guided by (discounted)
+ lookahead search. Indeed, a best-first search algorithm guided only  by (discounted)
 accumulated reward will deliver eventually  best moves, but it will not be as effective
@@
 provides a different scheme. The first converges to the optimal behavior asymptotically; the
-second in a bounded number of steps. Alternatively, novelty-guided probes, possibly with 
-some noise added, could potentially pay off as  a  base policy for \UCT\ in replacement 
-of random roll outs, as they apparently manage to exploit the structure of many problems.
+second in a bounded number of steps, with the caveat below.
+
+% Alternatively, novelty-guided probes, possibly with
+% some noise added, could potentially pay off as  a  base policy for \UCT\ in replacement
+% of random roll outs, as they apparently manage to exploit the structure of many problems.
+
+\section{Duplicates and Optimality}
+
+The notions of width and the \IW\  algorithm guarantee that  states with low width will be generated
+in low polynomial time through \emph{shortest} paths. In the presence of rewards like the Atari games,
+however, the  interest is not  in the shortest paths but in the \emph{best} paths; i.e, the paths with
+maximum reward.  \IW\ may actually  fail to find such paths even when calling \IW($k$) with a high 
+$k$ parameter. Optimality  could be  achieved then  by replacing the breadth-first search
+underlying \IW($k$) by Dijkstra's algorithm yet such a move would make the relation between
+\IW\ and the notion of width less clear. Another option is to change  \IW\ to comply with 
+a different property; namely, that if there is a ``rewarding'' path made up of states of 
+low width, then \IW\ will find such paths or better ones in time that is exponential 
+in their width. For this actually a simple change in \IW\ suffices: when generating a state 
+$s$ that is a \emph{duplicate} of a state $s'$ that has been previously generated and not pruned, 
+$s'$ is replaced by $s$ if $R(s) > R(s')$. Moreover in such a case, the change in reward 
+must be propagated to the descendants of $R(s')$ that are in memory. This is similar to the change required 
+in A* for preserving optimality when moving from consistent to inconsistent heuristics \cite{pearl:heuristics}. 
+The alternative is to ``reopen'' such nodes.  The same change is actually in \BFS\ to ensure that, 
+if given enough time, \BFS\ will actually find the optimal paths. The code used for \IW\ and \BFS\ in the
+experiments above does not implement this change, however, as the overhead involved in checking for duplicates 
+in some test cases did not appear to pay off. More experimentation is probably needed.
 
@@
 We have shown experimentally that width-based algorithms like $\IW(1)$ and \BFS\
-that originate in work in classical planning, can be used in the Atari games
-with a performance that approaches the state-of-the-art, sometimes using a fraction of the time. We are currently addressing three 
+that originate in work in classical planning,  can be used to play the Atari video
+games and to achieve  state-of-the-art performance. This is the result of a 
+structured exploration of the state that manages to combine the scope of blind
+search algorithms with the efficiency of heuristic search methods. 
+%
+\Omit{
+We are currently addressing three
 issues, which we haven't managed to get in time for this submission. First, we want to extend the
@@
 than  \BFS. Third, we need to revise the way duplicate states are handled in \IW\ and \BFS\ to ensure
-that the best paths among the states that have been expanded are returned. Finally, 
-for the future, we also want to explore whether this type of algorithms
-can give meaningful results as well  when used on the screen state, as opposed 
-to the RAM state.
+that the best paths among the states that have been expanded are returned.Finally,}
+%
+There are many issues worth studying, we just mention two of them.
+One problem of all planning methods in Atari  games is that by using the 
+RAM state of the Atari console they ``cheat'' a bit, as they use information that
+is not available to the human player. In addition, a limitation of our
+algorithms is that they do not get better with experience.  
+In the future, we would like to explore whether  variations of these algorithms 
+can be used to play  from the information in the  screen state as opposed to the RAM state, 
+and the potential uses of the  ideas underlying these algorithms for learning.
 

### 2014-11-18 icaps15-paper/paper.tex
Experimental Section
* Corrected remark about experimental setup

@@
 reported for \UCT\ in  BNVB  are included in  parenthesis for reference. 
-Our experimental setup  follows the one in BNVB with two exceptions: rather than
-considering $10$ runs per games, we considered just $5$ due to time constraints (*** ???),
-and  a maximum budget of  $150,000$ simulated frames is applied to all the  algorithms, 
-including  \UCT\ that is limited to  $500$ base policy rollouts of depth $300$.
+Our experimental setup  follows the one in BNVB with one exception: a maximum budget of  $150,000$ simulated frames 
+is applied to all the  algorithms, including  \UCT\ that is limited to  $500$ base policy rollouts of depth $300$.
 % \footnote{In \UCT,  $500$ base policy rollouts

### 2014-11-18 src/agents/BFS_UCB.cpp
Synced with Nir's mods in main branch

@@ -28,6 +28,7 @@ int BFS_UCB::expand_node( TreeNode* curr_node )
 {
 	int num_simulated_steps =0;
 	int num_actions = available_actions.size();
+	static     int max_nodes_per_frame = max_sim_steps_per_frame / sim_steps_per_node;
 	bool leaf_node = (curr_node->v_children.empty());
 	m_expanded_nodes++;
 	if(curr_node->novelty == 1)
@@ -107,8 +108,9 @@ int BFS_UCB::expand_node( TreeNode* curr_node )
 	
 		// Don't expand duplicate nodes, or terminal nodes
 		if (!child->is_terminal) {
-		    if (! (ignore_duplicates && test_duplicate(child)) ){				
-				Q[m_current_queue]->push(child);
+		    if (! (ignore_duplicates && test_duplicate(child)) ){
+				if( child->num_nodes_reusable < max_nodes_per_frame )
+					Q[m_current_queue]->push(child);
 
 		    }
 		}
@@ -284,12 +286,6 @@ void BFS_UCB::expand_tree(TreeNode* start_node) {
 		if ( curr_node->depth() > m_max_depth ) m_max_depth = curr_node->depth();
 
 		/**
-		* check nodes that have been expanded by other queue
-		*/
-		if(  curr_node->already_expanded ) 
-			continue;
-		
-		/**
 		* check if subtree is bigger than max_budget of nodes
 		*/
 		if(  curr_node->num_nodes_reusable > max_nodes_per_frame  ) {
@@ -354,7 +350,7 @@ void BFS_UCB::expand_tree(TreeNode* start_node) {
 	
 	if ( queues_empty() ) std::cout << "Search Space Exhausted!" << std::endl;
 	for ( unsigned k = 0; k < Q.size(); k++ ) {
-		std::cout << "Size(Q[" << k << "]) = " << Q[k]->size()  << "V=" << V[k] << "N=" << N[k] << std::endl;
+		std::cout << "Size(Q[" << k << "]) = " << Q[k]->size()  << " V[" << k << "]=" << V[k] << " N[" << k << "]=" << N[k] << std::endl;
 	}
 		
 	update_branch_return(start_node);

### 2014-11-18 src/games/supported/MontezumaRevenge.cpp
Tweaking the reward function

@@ -42,6 +42,7 @@ void MontezumaRevengeSettings::step(const System& system) {
     int new_lives = readRam(&system, 0xBA);
     int some_byte = readRam(&system, 0xFE);
     m_terminal = new_lives == 0 && some_byte == 0x60;
+    //m_reward += ( new_lives > 0 ? 1 : 0 );
 }
 
 

### 2014-11-18 icaps15-paper/paper.tex
Edits to Optimality and Duplicates Section
* Clarified the sentence where we observe that accounting for alternative
  paths to the same state with differring rewards, guarantees that 2BFS will
  be optimal if given enought time.
Table added to repo

@@
 {\small
-\begin{tabular}{@{}l|r|rr|rr|rr|rr@{}}
+\begin{tabular}{@{}l|rr|rr|r|r@{}}
 %% table header
@@
 %\hline
-\multicolumn{1}{c|}{} & \multicolumn{1}{c|}{Random} & \multicolumn{2}{c|}{\BRFS}& \multicolumn{2}{c|}{$\IW(1)$}& \multicolumn{2}{c|}{\BFS}&
-   \multicolumn{2}{c}{\UCT}
+\multicolumn{1}{c|}{}& \multicolumn{2}{c|}{$\IW(1)$}& \multicolumn{2}{c|}{\BFS} & \multicolumn{1}{c|}{\BRFS} &  \multicolumn{1}{c}{\UCT}
 % \\[.2ex]
@@
 % \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c}{Avg. Time}\\[.2ex]
- \multicolumn{1}{c|}{Score} &
  \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
  \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
- \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
- \multicolumn{1}{c}{Score} & \multicolumn{1}{c}{Time}
+ \multicolumn{1}{c|}{Score} &
+ \multicolumn{1}{c}{Score} 
 % \\[.2ex]
@@
 \hline
-\Asterix        & 210    & 3900  & 32.24 & 119000  & 17.08 & 56320  & 0.95  & \textbf{296300} (290700) & 10.82\\
-\BeamRider     & 448.8 & 519.2 & 43.48 & 3000          & 6.02  & 3804        & 0.92  & \textbf{4320} (6624)   & 42.91\\
-\Breakout       & 1    & 2     & 51.56 & 383.6         & 1.37  & 421.2       & 3.95  & \textbf{512.4} (364)   & 26.03\\
-\Enduro          & 0   & 0.4  & 40.92 & \textbf{500}   & 38.13 & 292.8      & 13.12 & 306.8 (286)             & 32.14\\
-\Freeway         & 0   & 0     & 71.84 & \textbf{29.4} & 15.19 & 19.6       & 4.44  & 1 (0)    & 90.64\\
-\Pong           & -20.4 & -20.2 & 41.51 & 16.4         & 18.45 & \textbf{21} & 39.69 & \textbf{21} (21) & 38.37\\
-\Qbert           & 185   & 160  & 42.83 & 1545         & 1.82  & 14725       & 1.76  & \textbf{18730} (17343) & 28.61\\
-\Seaquest       & 192   & 200   & 40.01 & \textbf{4274}& 14.72 & 4070        & 1.59  & 3406 (5132)      & 26.16\\
-\SpaceInvaders & 187    & 115   & 46.11 & \textbf{2670} & 12.1  & 1327       & 2.25  & 2634 (2718)  & 19.66
+
+\propername{Assault} & 806.4 & 18.21 & \textbf{852.6} & 25.11 & 413.7 & 1512.2 \\ \hline
+\propername{Asterix} & 153400 & 24.33 & 77200 & 26.58 & 2135.7 & \textbf{290700} \\ \hline
+\propername{Asteroids} & \textbf{51338} & 65.92 & 22168 & 64.53 & 3127.4 & 4660.6 \\ \hline
+\propername{Atlantis} & 159420 & 12.8 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 30460 & \textbf{193858} \\ \hline
+\propername{Bank\_heist} & \textbf{717} & 39.29 & 362 & 63.83 & 21.5 & 497.8 \\ \hline
+\propername{Battle\_zone} & 11600 & 85.96 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 6312.5 & \textbf{70333.3} \\ \hline
+\propername{Beam\_rider} & 9108 & 22.53 & \textbf{9297.6} & 29.03 & 693.5 & 6624.6 \\ \hline
+\propername{Berzerk} & \textbf{2096} & 58 & 802 & 73.47 & 195 & 553.5 \\ \hline
+\propername{Bowling} & \textbf{68.6} & 9.84 & 50.4 & 59.52 & 25.5 & 25.1 \\ \hline
+\propername{Boxing} & 100 & 15.2 & 100 & 22.3 & 100 & 100 \\ \hline
+\propername{Breakout} & 384.2 & 4.22 & \textbf{771.6} & 39.47 & 1.1 & 364.4 \\ \hline
+\propername{Centipede} & 99207.2 & 39.14 & 93067 & 65.49 & \textbf{125123} & 110422 \\ \hline
+\propername{Chopper\_command} & 2100 & 41.06 & 3800 & 40.94 & 1827.3 & \textbf{34018.8} \\ \hline
+\propername{Crazy\_climber} & 36160 & 3.61 & 36940 & 58.12 & 37110 & \textbf{98172.2} \\ \hline
+\propername{Enduro} & \textbf{500} & 65.69 & 359.4 & 37.66 & 0.6 & 286.3 \\ \hline
+\propername{Freeway} & \textbf{30.6} & 31.69 & 22.6 & 60.77 & 0 & 0.4 \\ \hline
+\propername{Frostbite} & \textbf{3582} & 12.25 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 137.2 & 270.5 \\ \hline
+\propername{Gravitar} & 3920 & 61.85 & \textbf{5980} & 62.45 & 395 & 2850 \\ \hline
+\propername{Hero} & \textbf{12887} & 36.63 & 7096.6 & 51.83 & 1323.8 & 12859.5 \\ \hline
+\propername{Jamesbond} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \textbf{8550} & 25.68 & 25 & 330 \\ \hline
+\propername{Kangaroo} & \textbf{5700} & 7.9 & 5320 & 31 & 90 & 1990 \\ \hline
+\propername{Krull} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \textbf{5124} & 36.82 & 3089.2 & 5037 \\ \hline
+\propername{Montezuma\_revenge} & 0 & 17.36 & \textbf{80} & 34.72 & 0 & 0 \\ \hline
+\propername{Ms\_pacman} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 14781 & 23.97 & 1708.5 & \textbf{22336} \\ \hline
+\propername{Pong} & \textbf{21} & 16.67 & \textbf{21} & 34.92 & -20.7 & \textbf{21} \\ \hline
+\propername{Pooyan} & 11225 & 7.97 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 909.7 & \textbf{17763.4} \\ \hline
+\propername{Qbert} & 3705 & 11.07 & 11680 & 34.61 & 132.8 & \textbf{17343.4} \\ \hline
+\propername{Riverraid} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \textbf{5062} & 36.53 & 2178.5 & 4449 \\ \hline
+\propername{Seaquest} & \textbf{14272} & 24.91 & 6138 & 33.38 & 288 & 5132.4 \\ \hline
+\propername{Space\_invaders} & 2877 & 21.37 & \textbf{3974} & 34.21 & 112.2 & 2718 \\ \hline
+\propername{Star\_gunner} & 1540 & 18.89 & \textbf{4660} & 17.62 & 1345 & 1207.1 \\ \hline
+\propername{Tennis} & \textbf{23.8} & 21.05 & 23.6 & 36.33 & -23.8 & 2.8 \\ \hline
+\propername{Venture} & 775 & 23.41 & \textbf{1080} & 27.78 & 0 & 0 \\ \hline
+\propername{Video\_pinball} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 62074.6 & 43.13 & 55567.3 & \textbf{254748} \\ \hline
+\propername{Zaxxon} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 15680 & 31.01 & 0 & \textbf{22610} \\ \hline
+\multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\ \hline
+Times Best (35 games)& \multicolumn{2}{c|}{12} & \multicolumn{2}{c|}{12} & 1 & 11 \\ \hline
+Times Better than UCT (35 games) & \multicolumn{2}{c|}{23} & \multicolumn{2}{c|}{22} & 1  & -- \\ \hline
+
 \end{tabular}
@@
 
- \caption{\small Performance that results from random policy and various  lookahead algorithms in 9 Atari games.
+ \caption{\small Performance that results from random policy and various  lookahead algorithms in 35 Atari games.
 The algorithms  are \BRFS, $\IW(1)$, \BFS, and \UCT. Averages  over 5 runs shown with ``random'' seeds
 $0$ to $4$, and maximum episode duration of $18,000$  frames. Every algorithm limited to a budget of
-100,000 simulated frames, except for UCT limited to 500 rollouts, 300 frames deep instead. Average scores and CPU times per action selection shown.
-Performance of UCT results  reported in \cite{bellemare:jair2013} shown in parenthesis.
-Numbers  in bold show best performer in terms of average score. }
+150,000 simulated frames, except for UCT limited to 500 rollouts, 300 frames deep instead. Average scores and CPU times per action selection shown.
+Performance of UCT taken from \cite{bellemare:jair2013}.
+Numbers  in bold show best performer in terms of average score.
+--Domains containing (--) are still computing,  each run can take 24 hours of computation. All domains from \cite{bellemare:jair2013}  will be ready for the camera ready--. }
 \label{performance}
@@
 in A* for preserving optimality when moving from consistent to inconsistent heuristics \cite{pearl:heuristics}. 
-The alternative is to ``reopen'' such nodes.  The same change is actually in \BFS\ to ensure that, 
-if given enough time, \BFS\ will actually find the optimal paths. The code used for \IW\ and \BFS\ in the
+The alternative is to ``reopen'' such nodes. 
+%The same change is actually in \BFS\ to ensure that, 
+%if given enough time, \BFS\ will actually find the optimal paths.
+Changing \BFS\ to account for such paths guarantees that, if given enough time, \BFS\ will actually find
+the optimal paths. 
+The code used for \IW\ and \BFS\ in the
 experiments above does not implement this change, however, as the overhead involved in checking for duplicates 

### 2014-11-18 create_small_table.py

@@ -72,7 +72,7 @@ def check_trace( score_file, trace_file, game, algorithm  ):
         return e
 def retrieve_episodes( experiments_folder ) :
 	episodes = []
-	for root, dirs, files in os.walk('experiments_branching') :
+	for root, dirs, files in os.walk('experiments_300_reuse') :
 		game = os.path.split( root )[-2]
                 if "/" in game:
                         game = game.split("/")[1]               
@@ -293,7 +293,7 @@ class Algorithm_Performance :
 
 if __name__ == '__main__':
 
-	episodes = retrieve_episodes('experiments_branching')
+	episodes = retrieve_episodes('experiments_300_reuse')
 	print >> sys.stdout, len(episodes), 'episodes retrieved'
 	# the experimental data set is represented as a nested dictionary
 	# { game, { algorithm, instance of Algorithm Performance } }

### 2014-11-18 src/agents/BestFirstSearch.cpp

@@ -164,7 +164,7 @@ int BestFirstSearch::reuse_branch(TreeNode* node) {
 				
 				// This recreates the novelty table (which gets resetted every time
 				// we change the root of the search tree)
-				if ( m_novelty_pruning ){				
+			        if ( m_novelty_pruning ){				
 
 						if ( check_novelty_1( child->state.getRAM() ) ){
 							update_novelty_table( child->state.getRAM() );

### 2014-11-18 src/agents/BestFirstSearch.hpp

@@ -18,7 +18,6 @@ public:
 	{
 		if ( b->novelty < a->novelty ) return true;
 		else if( b->novelty == a->novelty && b->fn < a->fn ) return true;
-		
 		return false;
 	}
     };

### 2014-11-18 icaps15-paper/paper.tex
Table and experiments changed

@@
 \newcommand{\Deepmind}{\propername{Deepmind}}
+\newcommand{\CrazyClimber}{\propername{Crazy\_climber}}
+\newcommand{\Kangaroo}{\propername{Kangaroo}}
+\newcommand{\Pooyan}{\propername{Pooyan}}
+\newcommand{\Berzerk}{\propername{Berzerk}}
+\newcommand{\Bowling}{\propername{Bowling}}
+\newcommand{\Montezuma}{\propername{Montezuma\_revenge}}
+\newcommand{\StarGunner}{\propername{Star\_gunner}}
+\newcommand{\Venture}{\propername{Venture}}
 
@@
 \end{table*}
-
-We tested  $\IW(1)$ and \BFS\  over $9$ games: those considered in the \emph{training} set
-of \cite{bellemare:jair2013}, abbreviated  as BNVB, and those discussed in \cite{deep-mind-atari}.
-In Table~\ref{performance},  the performance of  $\IW(1)$ and \BFS\
+ We tested $\IW(1)$ and \BFS\ over $35$ games: those considered in
+\cite{bellemare:jair2013}, abbreviated as BNVB \footnote{The full set
+  of games is made of 55 games, 5 runs each. As each run takes around
+  24 hours of CPU time, we luck the resources to submit the full table
+  by the submission date, but the full table will be ready for the
+  camera ready version.}. In Table~\ref{performance}, the
+performance of $\IW(1)$ and \BFS\
 % , characterized by the average score obtained over all episodes, the standard deviation
 % of the score obtained and time (in seconds) per lookahead,
-is compared with that of \UCT, breadth-first search (\BRFS), and random action selection.
+is compared with that of \UCT\ and breadth-first search (\BRFS).
 The discount factor used by all the algorithms is $\gamma=0.995$ and the average scores
-reported for \UCT\ in  BNVB  are included in  parenthesis for reference. 
+reported for \BRFS\ and \UCT\ in  BNVB  are included for reference. 
 Our experimental setup  follows the one in BNVB with one exception: a maximum budget of  $150,000$ simulated frames 
-is applied to all the algorithms, including  \UCT\ that is limited to  $500$ base policy rollouts of depth $300$.
+is applied to \textit{all the algorithms}, including  \UCT\ that is limited to  $500$ base policy rollouts of depth $300$.
 % \footnote{In \UCT,  $500$ base policy rollouts
 % of depth $300$ is equivalent to a budget of $150,000$ simulated frames.}.
-The bound on the number of simulated frames is like a bound on lookahead time, 
-as most of the time in the lookahead search is spent in the calls to the emulator
-for computing the next RAM state.  Also in order to save time, as in BNVB,   all of the algorithms 
-reuse the subtree from  the previous lookahead that is rooted in the selected
-child, deleting its siblings and their descendants. More precisely, no calls to the
-emulator are done  for transitions that are cached in that subtree. As a result
-the number of nodes in the lookahead tree may be larger actually than the budget.
-In addition, in  \IW, the states that are ``reused''  from the  previous searches
-are ignored in the computation of the novelty of new states so that more states 
-can  escape pruning. Otherwise, \IW\ often doesn't use the full budget. This is not 
-needed in \BFS\ which does no pruning.  $\IW(1)$ and \BFS\ have been  limited 
-search up to a depth of $m=1500$  frames and up to $m=20,000$ frames per root branch.
-This is to avoid the search for going too deep or being too committed to a single
-root action. (**** Please revise ****).
+The bound on the number of simulated frames is like a bound on
+lookahead time, as most of the time in the lookahead search is spent
+in the calls to the emulator for computing the next RAM state.  Also,
+as in BNVB, all of the algorithms reuse the subtree from the previous
+lookahead that is rooted in the selected child, deleting its siblings
+and their descendants. More precisely, no calls to the emulator are
+done for transitions that are cached in that subtree. As a result the
+number of nodes in the lookahead tree may be larger actually than the
+budget. In addition, in \IW, the states that are ``reused'' from the
+previous searches are ignored in the computation of the novelty of new
+states so that more states can escape pruning. Otherwise, \IW\ often
+doesn't use the full budget. This is not needed in \BFS\ which does no
+pruning.  $\IW(1)$ and \BFS\ have been limited the search to a depth of
+$m=1500$ frames and up to $m=150,000$ frames per root branch. This is
+to avoid the search for going too deep or being too committed to a
+single root action. Similarly UCT defines a limit of $m=300$ frames
+for each rollout. Note that UCT is more likely to have a performance
+loss if the depth is increased, as the number of rollouts would
+decrease in order to maintain the same budget of $150,000$ simulated
+frames. The last difference inherited from BNVB between all algorithms
+and UCT, is that UCT selects an action randomly in the rollout every
+single frame, while the others select an action every 5 frames, and
+the remain ones are $no-op$, as is done in the simulation
+environment. Therefore, a node in all algorithms represent $5$ frames,
+but in UCT represents $1$ frame. As a result, a UCT rollout of depth
+$300$ frames involve the selection of 300 actions, decreasing the
+budget of simulated frames by $300$, while the other algorithms need a
+branch of $1500$ frames to select $300$ actions, decreasing the budget
+by $1500$ frames instead.
 
-Table~\ref{performance} shows that $\IW(1)$ and \BFS\ both outperform \BRFS\  and the random baseline.
-\BRFS\ rarely collects any reward, as the depth of the \BRFS\ search tree
-results in  a lookahead of $0.3$ seconds ($20$  frames or $4$ nodes deep).
-On the other hand, both $\IW(1)$ and \BFS\ reach states that are  up to $350$--$1500$ frames deep
-($70$--$260$ nodes or $6$--$22$ seconds), even if $\IW(1)$ doesn't always use all the simulation frames
-allocated due to its agressive pruning.
-%% case of \BFS, due to having a novelty of $0$ and the planning horizon being limited to $m$
-%%% (*** What is novelty $0$??? Not clear why BFS doesn't take all the budget.
+Table~\ref{performance} shows that $\IW(1)$ and \BFS\ both outperform
+\BRFS, which rarely collects any reward in many domains, as the
+depth of the \BRFS\ search tree results in a lookahead of $0.3$
+seconds ($20$ frames or $4$ nodes deep). On the other hand, both
+$\IW(1)$ and \BFS\ reach states that are up to $350$--$1500$ frames
+deep ($70$--$260$ nodes or $6$--$22$ seconds), even if $\IW(1)$
+doesn't always use all the simulation frames allocated due to its
+agressive pruning. This behaviour can be observed in games such as
+\Breakout, \CrazyClimber, \Kangaroo, and \Pooyan, where the average
+CPU time for each lookahead is up to $10$ times faster than \BFS. Note
+that computation time for \UCT\ and \BRFS\ is similar to \BFS, as the
+most expensive computation is to generate a frame through the
+simulator and all algorithms use always the full budget.
 
-Interestingly, simple $\IW(1)$ outperforms \UCT\ in $4$ out of $9$ games,
-both in terms of  average scores and average times per call. \BFS\ is
-faster than $\IW(1)$ and an order-of-magnitude faster than \UCT\ in general,
-and it  outscores \UCT\ in $2$ of the games (\Freeway\ and
-\Seaquest). Compared with the \UCT\ results in   \cite{bellemare:jair2013}, shown in parenthesis,
-both $\IW(1)$ and \BFS\  outperform \UCT\ in $3$ of the  $9$ games (\Breakout, \Enduro\ and \Freeway),
-with \BFS\ and \UCT\ tied in  \Pong. It'd  be very interesting to see the results of
-\BFS\ when given more time per action, but we don't have those results yet.
-As mentioned in the previous section, the order in which $\IW(1)$ and \BFS\ generate the children
-of an expanded node are randomized, something which has a potential effect on the  novelty measures
-of nodes, and hence in the search graph. We found that this randomization helps, in particular
-in \Qbert, where not randomizing results  in $\IW(1)$ being trapped into  loops sometimes.
+Interestingly, in terms of average scores out of $35$ games, simple $\IW(1)$ outperforms
+\BFS\ and \UCT\ in $18$ and $23$ games respectively,
+most often with lower average times per call. \BFS\ outscores \UCT\ in
+$22$ of the games \footnote{Note that both \IW\ and \BFS\ may improve, as 6 and 4 games  respectively are not reported in the table
+  yet.}. As mentioned in the
+previous section, the order in which $\IW(1)$ and \BFS\ generate the
+children of an expanded node is randomized, something which has a
+potential effect on the novelty measures of nodes, and hence in the
+search graph. We found that this randomization helps, in particular in
+\Qbert, where not randomizing results in $\IW(1)$ being trapped into
+loops sometimes.
 
-We also  note that in \Freeway,  both $\IW(1)$ and \BFS\ attain a score similar to the
-baseline semi-random algorithm \emph{Perturb} in  \cite{bellemare:jair2013},
-which outperforms   both \UCT\ and  the reinforcement learning methods.
-Likewise,  $\IW(1)$ outperforms the best learning algorithm in \cite{deep-mind-atari}
-in $4$ out of the  $7$ games considered, and \BFS\ does so in $5$  of
-the $7$ games.
+Likewise, $\IW(1)$ outperforms the best learning algorithm in
+\cite{deep-mind-atari} in $7$ out of the $7$ games considered, and
+\BFS\ does so in $6$ of the $7$ games.
+
+We also note that in \Freeway, \Berzerk, \Bowling, \Montezuma,
+\StarGunner, and \Venture, both $\IW(1)$ and \BFS\ attain a better
+score than the baseline semi-random algorithm \emph{Perturb} in
+\cite{bellemare:jair2013}, which outperforms both \UCT\ and the
+reinforcement learning methods. \emph{Perturb} performance helps
+understanding these domains. \emph{Perturb} selects a fixed action with probability
+0.95 and otherwise acts uniformly randomly, and the best run out of 10
+episodes is reported in BNVB. Note that all the other algorithms
+report average score instead. The behaviour of \emph{Perturb} hinder
+the nature of these domains: rewards are very deep, but tipically
+repeating the same action may lead to some reward.
+
+Consider \Freeway, where a chicken has to run across a ten lane
+highway filled with traffic. Every time it gets across, we
+receive a point, i.e. a reward of 1. If the chicken is hit by a car,
+the chicken is forced back. 12 out of 18 possible actions have an
+effect: move up ( up-right, up-left, up-fire, up-right-fire,
+up-left-fire, all have same effect as up ) or move down ( down-right,
+down-left, down-fire, down-right-fire, down-left-fire, all have same
+effect as down ), the remaining 6 actions are like no-ops. It's easy
+to understand that \emph{Perturb} would achieve a high score if the
+selected action is move up with 0.95 probability. In the following link
+\url{http://www.youtube.com/playlist?list=PLXpQcXUQ_CwenUazUivhXyYvjuS6KQOI0}
+we can observe videos of \IW\ and \UCT\ playing \Freeway. As reported
+in Table~\ref{performance}, \UCT\ does not manage to take any chicken
+across the highway, while \IW\ is the best performer out of all tested
+algorithms. The main reason why \UCT\ performs so erratically is that
+the rollouts do not manage to collect any reward, as there is a
+$3^{-1}$ probability to choose any move up action, and given that the
+depth of the closest reward is about 200 frames deep, the probability
+of a rollout achieveing any reward is $3^{-200}$, which implies
+choosing 200 move-up actions. On the other hand, \IW\ manages to find
+rewards without exhausting the full budget. The main reason that
+explains this behaviour is that at most only 3 generated states from
+the same parent node will pass the novelty test, as in any expanded
+node there is at most 3 different new possible bytes in the RAM state
+resulting from moving up, down, or do-nothing.
+
+Another challenging game is \Montezuma. In this game rewards are very
+sparse, deep, and most of the actions lead to loosing a life, which
+only becomes effective after hundreds of frames, i.e. up-left may
+enforce jumping and falling until we loose a life. All algorithms achieve 0 score, wheter RL-based or Planning-based. But, as
+shown in Table~\ref{performance}, \BFS\ achieves an average score of
+80, which is still a very low score, as the first possible reward
+increases the score in 100. The best run by \BFS\ achieves a score of
+400, pointing out that \BFS\ still is not able to consistently find rewards.
+One way to improve the performance of any algorithm in \Montezuma\ is to encode the lost of a life as a negative reward.
+Capturing the lost of lifes in the reward function is relevant to many games where not loosing a life is important in later stages of the game.
+Currently,
+rewards are encoded directly as the change of score, which does not distinguish between the benefits of keeping lifes. This
+beahaviour is also observed in other games like \Breakout\ and \SpaceInvaders, where \BFS\
+is the best performer by far, although initially it looses all lifes but 1 and then plays extremely well. Videos on the link above illustrate
+this behaviour.
 

### 2014-11-18 icaps15-paper/paper.tex
changed title

@@
 \pdfinfo{
-/Title (Classical Planning Algorithms on the Atari Video Games: Preliminary Results)
+/Title (Classical Planning Algorithms on the Atari Video Games)
 /Author ( )}
@@
 
-\title{Classical Planning Algorithms ¡ on the Atari Video Games: Preliminary Results}
+\title{Classical Planning Algorithms ¡ on the Atari Video Games}
 

### 2014-11-18 icaps15-paper/paper.tex
Nir iteration over experimental Section

@@
 \setlength{\pdfpageheight}{11in}
+
+%Table stuff
+\usepackage{xcolor,colortbl}
+\definecolor{Gray}{gray}{0.85}
+\definecolor{LightCyan}{rgb}{0.88,1,1}
+
+\newcolumntype{a}{>{\columncolor{Gray}}c}
+\newcolumntype{b}{>{\columncolor{LightCyan}}c}
+
+\newcommand{\mc}[2]{\multicolumn{#1}{b}{#2}}
 
@@
 {\small
-\begin{tabular}{@{}l|rr|rr|r|r@{}}
+\begin{tabular}{@{}l|ra|ra|r|r@{}}
 %% table header
@@
 
-\propername{Assault} & 806.4 & 18.21 & \textbf{852.6} & 25.11 & 413.7 & 1512.2 \\ \hline
+\propername{Assault} & 806.4 & 18.21 & 852.6 & 25.11 & 413.7 & \textbf{1512.2} \\ \hline
 \propername{Asterix} & 153400 & 24.33 & 77200 & 26.58 & 2135.7 & \textbf{290700} \\ \hline
-\propername{Asteroids} & \textbf{51338} & 65.92 & 22168 & 64.53 & 3127.4 & 4660.6 \\ \hline
+\propername{Asteroids} & \mc{1}{\textbf{51338}} & 65.92 & \mc{1}{22168} & 64.53 & 3127.4 & 4660.6 \\ \hline
 \propername{Atlantis} & 159420 & 12.8 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 30460 & \textbf{193858} \\ \hline
-\propername{Bank\_heist} & \textbf{717} & 39.29 & 362 & 63.83 & 21.5 & 497.8 \\ \hline
+\propername{Bank\_heist} & \mc{1}{\textbf{717}} & 39.29 & 362 & 63.83 & 21.5 & 497.8 \\ \hline
 \propername{Battle\_zone} & 11600 & 85.96 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 6312.5 & \textbf{70333.3} \\ \hline
-\propername{Beam\_rider} & 9108 & 22.53 & \textbf{9297.6} & 29.03 & 693.5 & 6624.6 \\ \hline
-\propername{Berzerk} & \textbf{2096} & 58 & 802 & 73.47 & 195 & 553.5 \\ \hline
-\propername{Bowling} & \textbf{68.6} & 9.84 & 50.4 & 59.52 & 25.5 & 25.1 \\ \hline
+\propername{Beam\_rider} & \mc{1}{9108} & 22.53 & \mc{1}{\textbf{9297.6}} & 29.03 & 693.5 & 6624.6 \\ \hline
+\propername{Berzerk} & \mc{1}{\textbf{2096}} & 58 & \mc{1}{802} & 73.47 & 195 & 553.5 \\ \hline
+\propername{Bowling} & \mc{1}{\textbf{68.6}} & 9.84 & \mc{1}{50.4} & 59.52 & 25.5 & 25.1 \\ \hline
 \propername{Boxing} & 100 & 15.2 & 100 & 22.3 & 100 & 100 \\ \hline
-\propername{Breakout} & 384.2 & 4.22 & \textbf{771.6} & 39.47 & 1.1 & 364.4 \\ \hline
+\propername{Breakout} &\mc{1}{384.2} & 4.22 & \mc{1}{\textbf{771.6}} & 39.47 & 1.1 & 364.4 \\ \hline
 \propername{Centipede} & 99207.2 & 39.14 & 93067 & 65.49 & \textbf{125123} & 110422 \\ \hline
@@
 \propername{Crazy\_climber} & 36160 & 3.61 & 36940 & 58.12 & 37110 & \textbf{98172.2} \\ \hline
-\propername{Enduro} & \textbf{500} & 65.69 & 359.4 & 37.66 & 0.6 & 286.3 \\ \hline
-\propername{Freeway} & \textbf{30.6} & 31.69 & 22.6 & 60.77 & 0 & 0.4 \\ \hline
-\propername{Frostbite} & \textbf{3582} & 12.25 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 137.2 & 270.5 \\ \hline
-\propername{Gravitar} & 3920 & 61.85 & \textbf{5980} & 62.45 & 395 & 2850 \\ \hline
-\propername{Hero} & \textbf{12887} & 36.63 & 7096.6 & 51.83 & 1323.8 & 12859.5 \\ \hline
-\propername{Jamesbond} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \textbf{8550} & 25.68 & 25 & 330 \\ \hline
-\propername{Kangaroo} & \textbf{5700} & 7.9 & 5320 & 31 & 90 & 1990 \\ \hline
-\propername{Krull} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \textbf{5124} & 36.82 & 3089.2 & 5037 \\ \hline
-\propername{Montezuma\_revenge} & 0 & 17.36 & \textbf{80} & 34.72 & 0 & 0 \\ \hline
+\propername{Enduro} & \mc{1}{\textbf{500}} & 65.69 & \mc{1}{359.4} & 37.66 & 0.6 & 286.3 \\ \hline
+\propername{Freeway} & \mc{1}{\textbf{30.6}} & 31.69 & \mc{1}{22.6} & 60.77 & 0 & 0.4 \\ \hline
+\propername{Frostbite} & \mc{1}{\textbf{3582}} & 12.25 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 137.2 & 270.5 \\ \hline
+\propername{Gravitar} & \mc{1}{3920} & 61.85 & \mc{1}{\textbf{5980}} & 62.45 & 395 & 2850 \\ \hline
+\propername{Hero} & \mc{1}{\textbf{12887}} & 36.63 & 7096.6 & 51.83 & 1323.8 & 12859.5 \\ \hline
+\propername{Jamesbond} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \mc{1}{\textbf{8550}} & 25.68 & 25 & 330 \\ \hline
+\propername{Kangaroo} & \mc{1}{\textbf{5700}} & 7.9 & \mc{1}{5320} & 31 & 90 & 1990 \\ \hline
+\propername{Krull} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \mc{1}{\textbf{5124}} & 36.82 & 3089.2 & 5037 \\ \hline
+\propername{Montezuma\_revenge} & 0 & 17.36 & \mc{1}{\textbf{80}} & 34.72 & 0 & 0 \\ \hline
 \propername{Ms\_pacman} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 14781 & 23.97 & 1708.5 & \textbf{22336} \\ \hline
@@
 \propername{Qbert} & 3705 & 11.07 & 11680 & 34.61 & 132.8 & \textbf{17343.4} \\ \hline
-\propername{Riverraid} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \textbf{5062} & 36.53 & 2178.5 & 4449 \\ \hline
-\propername{Seaquest} & \textbf{14272} & 24.91 & 6138 & 33.38 & 288 & 5132.4 \\ \hline
-\propername{Space\_invaders} & 2877 & 21.37 & \textbf{3974} & 34.21 & 112.2 & 2718 \\ \hline
-\propername{Star\_gunner} & 1540 & 18.89 & \textbf{4660} & 17.62 & 1345 & 1207.1 \\ \hline
-\propername{Tennis} & \textbf{23.8} & 21.05 & 23.6 & 36.33 & -23.8 & 2.8 \\ \hline
-\propername{Venture} & 775 & 23.41 & \textbf{1080} & 27.78 & 0 & 0 \\ \hline
+\propername{Riverraid} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \mc{1}{\textbf{5062}} & 36.53 & 2178.5 & 4449 \\ \hline
+\propername{Seaquest} & \mc{1}{\textbf{14272}} & 24.91 & \mc{1}{6138} & 33.38 & 288 & 5132.4 \\ \hline
+\propername{Space\_invaders} & \mc{1}{2877} & 21.37 & \mc{1}{\textbf{3974}} & 34.21 & 112.2 & 2718 \\ \hline
+\propername{Star\_gunner} & \mc{1}{1540} & 18.89 & \mc{1}{\textbf{4660}} & 17.62 & 1345 & 1207.1 \\ \hline
+\propername{Tennis} & \mc{1}{\textbf{23.8}} & 21.05 & \mc{1}{23.6} & 36.33 & -23.8 & 2.8 \\ \hline
+\propername{Venture} & \mc{1}{775} & 23.41 & \mc{1}{\textbf{1080}} & 27.78 & 0 & 0 \\ \hline
 \propername{Video\_pinball} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 62074.6 & 43.13 & 55567.3 & \textbf{254748} \\ \hline
@@
 \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\ \hline
-Times Best (35 games)& \multicolumn{2}{c|}{12} & \multicolumn{2}{c|}{12} & 1 & 11 \\ \hline
-Times Better than UCT (35 games) & \multicolumn{2}{c|}{23} & \multicolumn{2}{c|}{22} & 1  & -- \\ \hline
+Times Best (35 games)& \multicolumn{2}{c}{\textbf{12}} & \multicolumn{2}{c}{11} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{\textbf{12}} \\ \hline
+Times Better than \IW\ (35 games) & \multicolumn{2}{c}{--} & \multicolumn{2}{c}{\textbf{11}} & \multicolumn{1}{c}{1}  & \multicolumn{1}{c}{9} \\ \hline
+Times Better than \BFS\ (35 games) & \multicolumn{2}{c}{\textbf{12}} & \multicolumn{2}{c}{--} & \multicolumn{1}{c}{1}  & \multicolumn{1}{c}{11} \\ \hline
+Times Better than \UCT\ (35 games) & \multicolumn{2}{c}{17} & \multicolumn{2}{c}{\textbf{18}} & \multicolumn{1}{c}{1}  & \multicolumn{1}{c}{--} \\ \hline
 
@@
 Performance of UCT taken from \cite{bellemare:jair2013}.
-Numbers  in bold show best performer in terms of average score.
+Numbers  in bold show best performer in terms of average score. Numbers with light blue shade show that an algorithm outperformed \UCT.
 --Domains containing (--) are still computing,  each run can take 24 hours of computation. All domains from \cite{bellemare:jair2013}  will be ready for the camera ready--. }
@@
 Interestingly, in terms of average scores out of $35$ games, simple $\IW(1)$ outperforms
-\BFS\ and \UCT\ in $18$ and $23$ games respectively,
-most often with lower average times per call. \BFS\ outscores \UCT\ in
-$22$ of the games \footnote{Note that both \IW\ and \BFS\ may improve, as 6 and 4 games  respectively are not reported in the table
-  yet.}. As mentioned in the
+\BFS\ and \UCT\ in $12$ and $17$ games respectively,
+most often with lower average times per call. \BFS\ outscores  \IW\ and \UCT\ in
+$11$ and $18$ of the games respectively \footnote{Note that both \IW\ and \BFS\ may improve, as 6 and 4 games  respectively are not reported in the table
+  yet.}, and in overall, \IW\ and \UCT\ are the best performer in $12$ games, \BFS\ in $11$ games, and \BRFS in $1$ game. As mentioned in the
 previous section, the order in which $\IW(1)$ and \BFS\ generate the

### 2014-11-18 icaps15-paper/paper.tex
Table extracted from main text

--- 2014-11-18 icaps15-paper/paper.tex
@@
+ 
 \documentclass[letterpaper]{article}
@@
 $\IW(k)$ where  $\IW(i)$ is  a breadth-first search in which a state is  pruned when it is not the
-first state in the search  to make true some set of $i$ atoms. 
-The  empirical results suggest that the performance of \IW\ with the  $k$ parameter fixed to $1$, 
-i.e., $\IW(1)$, approaches the performance of the state of the art represented by 
- \UCT, and that a simple best-first variation  of \IW\  that combines exploration 
-and exploitation is even more competitive.
-
+first state in the search  to make true some subset of $i$ atoms. 
+The  empirical results over 35 games suggest that the performance 
+of \IW\ with the  $k$ parameter fixed to $1$, i.e., $\IW(1)$, is at the level of the state of the art represented by 
+\UCT.  A simple best-first variation  of \IW\  that combines exploration and exploitation proves
+to be very competitive as well. 
 \end{quote}
@@
 of nodes generated by $\IW(2)$ in the worst case is  $(n \times D)^2 \times b$, which is
-equal to $1,073,741,824$ which is too large, forcing us to consider only a fraction of
+equal to $1,073,741,824$ which is too large, forcing us to consider only a tiny fraction of
 such states. For classical planning problems, the growth in the number of nodes from
@@
 and if state s' is the child of state $s$ and action $a$,
-$R(s')$ is set to $R(s) + \gamma^d(s´) r(a,s)$.
+$R(s')$ is set to $R(s) + \gamma^{d(s)+1} r(a,s)$.
 The  discount factor  results in a slight
@@
 
-\begin{table*}
-\centering
-% \addtolength{\tabcolsep}{-2pt}
-% {\scriptsize
-{\small
-\begin{tabular}{@{}l|rr|rr|r|r@{}}
-%% table header
-% \hline\\[-3ex]
-%\hline
-\multicolumn{1}{c|}{}& \multicolumn{2}{c|}{$\IW(1)$}& \multicolumn{2}{c|}{\BFS} & \multicolumn{1}{c|}{\BRFS} &  \multicolumn{1}{c}{\UCT}
-% \\[.2ex]
-\\
-% \hline\\\\[-3ex]
-\hline
-%% table body
-\multicolumn{1}{c|}{Game}  &
-% \multicolumn{1}{c|}{Avg. Score} &
-% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
-% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
-% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
-% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c}{Avg. Time}\\[.2ex]
- \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
- \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
- \multicolumn{1}{c|}{Score} &
- \multicolumn{1}{c}{Score} 
-% \\[.2ex]
-\\[.2ex]
-% \hline\\
-\hline
-
-\propername{Assault} & 806.4 & 18.21 & \textbf{852.6} & 25.11 & 413.7 & 1512.2 \\ \hline
-\propername{Asterix} & 153400 & 24.33 & 77200 & 26.58 & 2135.7 & \textbf{290700} \\ \hline
-\propername{Asteroids} & \textbf{51338} & 65.92 & 22168 & 64.53 & 3127.4 & 4660.6 \\ \hline
-\propername{Atlantis} & 159420 & 12.8 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 30460 & \textbf{193858} \\ \hline
-\propername{Bank\_heist} & \textbf{717} & 39.29 & 362 & 63.83 & 21.5 & 497.8 \\ \hline
-\propername{Battle\_zone} & 11600 & 85.96 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 6312.5 & \textbf{70333.3} \\ \hline
-\propername{Beam\_rider} & 9108 & 22.53 & \textbf{9297.6} & 29.03 & 693.5 & 6624.6 \\ \hline
-\propername{Berzerk} & \textbf{2096} & 58 & 802 & 73.47 & 195 & 553.5 \\ \hline
-\propername{Bowling} & \textbf{68.6} & 9.84 & 50.4 & 59.52 & 25.5 & 25.1 \\ \hline
-\propername{Boxing} & 100 & 15.2 & 100 & 22.3 & 100 & 100 \\ \hline
-\propername{Breakout} & 384.2 & 4.22 & \textbf{771.6} & 39.47 & 1.1 & 364.4 \\ \hline
-\propername{Centipede} & 99207.2 & 39.14 & 93067 & 65.49 & \textbf{125123} & 110422 \\ \hline
-\propername{Chopper\_command} & 2100 & 41.06 & 3800 & 40.94 & 1827.3 & \textbf{34018.8} \\ \hline
-\propername{Crazy\_climber} & 36160 & 3.61 & 36940 & 58.12 & 37110 & \textbf{98172.2} \\ \hline
-\propername{Enduro} & \textbf{500} & 65.69 & 359.4 & 37.66 & 0.6 & 286.3 \\ \hline
-\propername{Freeway} & \textbf{30.6} & 31.69 & 22.6 & 60.77 & 0 & 0.4 \\ \hline
-\propername{Frostbite} & \textbf{3582} & 12.25 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 137.2 & 270.5 \\ \hline
-\propername{Gravitar} & 3920 & 61.85 & \textbf{5980} & 62.45 & 395 & 2850 \\ \hline
-\propername{Hero} & \textbf{12887} & 36.63 & 7096.6 & 51.83 & 1323.8 & 12859.5 \\ \hline
-\propername{Jamesbond} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \textbf{8550} & 25.68 & 25 & 330 \\ \hline
-\propername{Kangaroo} & \textbf{5700} & 7.9 & 5320 & 31 & 90 & 1990 \\ \hline
-\propername{Krull} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \textbf{5124} & 36.82 & 3089.2 & 5037 \\ \hline
-\propername{Montezuma\_revenge} & 0 & 17.36 & \textbf{80} & 34.72 & 0 & 0 \\ \hline
-\propername{Ms\_pacman} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 14781 & 23.97 & 1708.5 & \textbf{22336} \\ \hline
-\propername{Pong} & \textbf{21} & 16.67 & \textbf{21} & 34.92 & -20.7 & \textbf{21} \\ \hline
-\propername{Pooyan} & 11225 & 7.97 & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 909.7 & \textbf{17763.4} \\ \hline
-\propername{Qbert} & 3705 & 11.07 & 11680 & 34.61 & 132.8 & \textbf{17343.4} \\ \hline
-\propername{Riverraid} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & \textbf{5062} & 36.53 & 2178.5 & 4449 \\ \hline
-\propername{Seaquest} & \textbf{14272} & 24.91 & 6138 & 33.38 & 288 & 5132.4 \\ \hline
-\propername{Space\_invaders} & 2877 & 21.37 & \textbf{3974} & 34.21 & 112.2 & 2718 \\ \hline
-\propername{Star\_gunner} & 1540 & 18.89 & \textbf{4660} & 17.62 & 1345 & 1207.1 \\ \hline
-\propername{Tennis} & \textbf{23.8} & 21.05 & 23.6 & 36.33 & -23.8 & 2.8 \\ \hline
-\propername{Venture} & 775 & 23.41 & \textbf{1080} & 27.78 & 0 & 0 \\ \hline
-\propername{Video\_pinball} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 62074.6 & 43.13 & 55567.3 & \textbf{254748} \\ \hline
-\propername{Zaxxon} & \multicolumn{1}{l}{--} & \multicolumn{1}{l|}{--} & 15680 & 31.01 & 0 & \textbf{22610} \\ \hline
-\multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\ \hline
-Times Best (35 games)& \multicolumn{2}{c|}{12} & \multicolumn{2}{c|}{12} & 1 & 11 \\ \hline
-Times Better than UCT (35 games) & \multicolumn{2}{c|}{23} & \multicolumn{2}{c|}{22} & 1  & -- \\ \hline
-
-\end{tabular}
-}
-
- \caption{\small Performance that results from random policy and various  lookahead algorithms in 35 Atari games.
-The algorithms  are \BRFS, $\IW(1)$, \BFS, and \UCT. Averages  over 5 runs shown with ``random'' seeds
-$0$ to $4$, and maximum episode duration of $18,000$  frames. Every algorithm limited to a budget of
-150,000 simulated frames, except for UCT limited to 500 rollouts, 300 frames deep instead. Average scores and CPU times per action selection shown.
-Performance of UCT taken from \cite{bellemare:jair2013}.
-Numbers  in bold show best performer in terms of average score.
---Domains containing (--) are still computing,  each run can take 24 hours of computation. All domains from \cite{bellemare:jair2013}  will be ready for the camera ready--. }
-\label{performance}
-
-\end{table*}
- We tested $\IW(1)$ and \BFS\ over $35$ games: those considered in
-\cite{bellemare:jair2013}, abbreviated as BNVB \footnote{The full set
-  of games is made of 55 games, 5 runs each. As each run takes around
-  24 hours of CPU time, we luck the resources to submit the full table
-  by the submission date, but the full table will be ready for the
-  camera ready version.}. In Table~\ref{performance}, the
-performance of $\IW(1)$ and \BFS\
+\input{games-table}
+We tested \IW(1) and \BFS\ over $35$ of the 55 games considered in
+\cite{bellemare:jair2013}, from now on  abbreviated as BNVB.\footnote{
+Entries marked as  ``--'' in the table are   still running at the time of submission. 
+Each run can take up to 24 hours of CPU time and there are 5 runs per game and algorithm. 
+Eventually all of  the  55 games will  be included in the evaluation.}
+Table~\ref{performance} shows the performance of $\IW(1)$ and \BFS\
 % , characterized by the average score obtained over all episodes, the standard deviation
 % of the score obtained and time (in seconds) per lookahead,
-is compared with that of \UCT\ and breadth-first search (\BRFS).
-The discount factor used by all the algorithms is $\gamma=0.995$ and the average scores
-reported for \BRFS\ and \UCT\ in  BNVB  are included for reference. 
-Our experimental setup  follows the one in BNVB with one exception: a maximum budget of  $150,000$ simulated frames 
-is applied to \textit{all the algorithms}, including  \UCT\ that is limited to  $500$ base policy rollouts of depth $300$.
-% \footnote{In \UCT,  $500$ base policy rollouts
+in comparison  with breadth-first search (\BRFS) and \UCT.
+Videos of selected games played by \IW(1), \BFS, and \UCT\ can be 
+seen in Youtube.\footnote{\url{http://www.youtube.com/playlist?list=PLXpQcXUQ_CwenUazUivhXyYvjuS6KQOI0}.}
+The discount factor used by  all the algorithms is $\gamma=0.995$.
+The scores for \BRFS\ and \UCT\ are from BNVB.  
+Our experimental setup  follows theirs except that a maximum budget of  $150,000$ simulated frames 
+is applied to \textit{all the algorithms}. \UCT\ uses this budget by running
+ $500$   rollouts of depth $300$. % \footnote{In \UCT,  $500$ base policy rollouts
 % of depth $300$ is equivalent to a budget of $150,000$ simulated frames.}.
 The bound on the number of simulated frames is like a bound on
-lookahead time, as most of the time in the lookahead search is spent
-in the calls to the emulator for computing the next RAM state.  Also,
-as in BNVB, all of the algorithms reuse the subtree from the previous
-lookahead that is rooted in the selected child, deleting its siblings
-and their descendants. More precisely, no calls to the emulator are
-done for transitions that are cached in that subtree. As a result the
-number of nodes in the lookahead tree may be larger actually than the
-budget. In addition, in \IW, the states that are ``reused'' from the
-previous searches are ignored in the computation of the novelty of new
-states so that more states can escape pruning. Otherwise, \IW\ often
-doesn't use the full budget. This is not needed in \BFS\ which does no
-pruning.  $\IW(1)$ and \BFS\ have been limited the search to a depth of
+lookahead time, as most of the time in the lookahead  is spent
+in calls to the emulator for computing the next RAM state. 
+This is why the average time per action is similar to all the algorithms
+except \IW(1), that due to its agressive pruning does not always use
+the full budget and takes less time per action on average.
+
+Also,  as reported by  BNVB, all of the algorithms 
+reuse the subtree from the previous lookahead that is rooted in the selected 
+child, deleting its siblings and their descendants. More precisely, no calls 
+to the emulator are done for transitions that are cached in that subtree, 
+and such reused frames are not discounted from the budget that is thus a bound 
+on the number of ``new'' frames per lookahead. In addition, in \IW(1), the states that 
+are ``reused'' from the previous searches are ignored in the computation of the novelty of 
+new states so that more states can escape pruning. Otherwise, \IW(1) often
+uses a fraction of the  budget.  This is not needed in \BFS\ which does 
+no pruning.  \IW(1) and \BFS\ are  limited to  search up to a depth of
 $m=1500$ frames and up to $m=150,000$ frames per root branch. This is
-to avoid the search for going too deep or being too committed to a
-single root action. Similarly UCT defines a limit of $m=300$ frames
-for each rollout. Note that UCT is more likely to have a performance
-loss if the depth is increased, as the number of rollouts would
-decrease in order to maintain the same budget of $150,000$ simulated
-frames. The last difference inherited from BNVB between all algorithms
+to avoid the search from  going too deep or being too committed to a
+single root action. 
+% Similarly UCT defines a limit of $m=300$ frames for each rollout. 
+% Note that UCT is more likely to have a performance
+% loss if the depth is increased, as the number of rollouts would
+% decrease in order to maintain the same budget of $150,000$ simulated
+% frames.
+
+Last, \IW(1) and \BFS\ select an action every 5 frames, while UCT  selects 
+an action in every frame.  This means that in order to explore a branch 
+300 frames deep, UCT gets to choose 300 actions, while \IW(1) and \BFS\ 
+get to choose 60 actions, both however using the same 300 frames from the
+budget. For this, we  followed the setup  of \BRFS\ in BNVB 
+that also selects  actions every 5 frames. Since the lookahead budget 
+is given by a maximum number of (new) frames, and the time is mostly taken by 
+calls to the emulator, this may not be the best choice  for \IW(1)\ 
+and \BFS\ that are therefore not exploiting all the options afforded
+by the budget. We hope to fix this gap in the future.
+
+\Omit{
+The last difference inherited from BNVB between all algorithms
 and UCT, is that UCT selects an action randomly in the rollout every
@@
 by $1500$ frames instead.
+}
 
-Table~\ref{performance} shows that $\IW(1)$ and \BFS\ both outperform
+Table~\ref{performance} shows that both $\IW(1)$ and \BFS\  outperform
 \BRFS, which rarely collects any reward in many domains, as the
@@
 doesn't always use all the simulation frames allocated due to its
-agressive pruning. This behaviour can be observed in games such as
+pruning. This behaviour can be observed in games such as
 \Breakout, \CrazyClimber, \Kangaroo, and \Pooyan, where the average
-CPU time for each lookahead is up to $10$ times faster than \BFS. Note
-that computation time for \UCT\ and \BRFS\ is similar to \BFS, as the
-most expensive computation is to generate a frame through the
-simulator and all algorithms use always the full budget.
+CPU time for each lookahead is up to $10$ times faster than \BFS. 
+Computation time for \UCT\ and \BRFS\ are  similar to \BFS, as the
+most expensive part of the computation is the  generation of 
+frames through the simulator,  and these three  algorithms 
+always use  the full budget.
 
-Interestingly, in terms of average scores out of $35$ games, simple $\IW(1)$ outperforms
-\BFS\ and \UCT\ in $18$ and $23$ games respectively,
-most often with lower average times per call. \BFS\ outscores \UCT\ in
-$22$ of the games \footnote{Note that both \IW\ and \BFS\ may improve, as 6 and 4 games  respectively are not reported in the table
-  yet.}. As mentioned in the
+More interestingly,  \IW(1) outscores \UCT\  in 17 of the 35 games,
+while \BFS\ outscores \UCT\ in 18.\footnote{These numbers can actually
+improve further as 6  games for \IW(1) and 4 games for \BFS\ are still running.}
+On the other, \UCT\ does better than \IW(1) and \BFS\ in 9 and 11 games respectively.
+The relative performance between \IW(1)\ and \BFS\ is closer with \IW(1) being  the
+best of the two in 12 games, and  \BFS\  in 11. In terms of the number of games 
+where an algorithm is the overall winner, the situation is  even: 
+\IW(1) is the overall best in 12 games, \BFS\ in 11 games, and \UCT\ in 12 games. 
+Also, \BRFS\ is best in 1 game (Centipede), while the other three algorithms
+are tied in another game (\Pong).
+
+\Omit{As mentioned in the
 previous section, the order in which $\IW(1)$ and \BFS\ generate the
@@
 \Qbert, where not randomizing results in $\IW(1)$ being trapped into
-loops sometimes.
+loops sometimes.}
 
-Likewise, $\IW(1)$ outperforms the best learning algorithm in
-\cite{deep-mind-atari} in $7$ out of the $7$ games considered, and
-\BFS\ does so in $6$ of the $7$ games.
-
-We also note that in \Freeway, \Berzerk, \Bowling, \Montezuma,
+$\IW(1)$ also outperforms the best learning algorithm in
+\cite{deep-mind-atari} in the $7$ games considered, and
+\BFS\ does so in $6$ of the $7$ games. Likewise, in \Freeway, \Berzerk, \Bowling, \Montezuma,
 \StarGunner, and \Venture, both $\IW(1)$ and \BFS\ attain a better
 score than the baseline semi-random algorithm \emph{Perturb} in
-\cite{bellemare:jair2013}, which outperforms both \UCT\ and the
-reinforcement learning methods. \emph{Perturb} performance helps
-understanding these domains. \emph{Perturb} selects a fixed action with probability
-0.95 and otherwise acts uniformly randomly, and the best run out of 10
-episodes is reported in BNVB. Note that all the other algorithms
-report average score instead. The behaviour of \emph{Perturb} hinder
-the nature of these domains: rewards are very deep, but tipically
-repeating the same action may lead to some reward.
+\cite{bellemare:jair2013}.
+%% to say the 2 lines below must add context: .. eg
+%%% ``which in *these games* outperforms both \UCT\ and the reinforcement learning methods.''
+%% add it if that's the case ..
+\emph{Perturb} is a simple algorithm
+that selects  a fixed  action with probability 0.95, and a random
+action with probability 0.05.  For \emph{Perturb}, BNVB do not report 
+the average score  but the  best score.  \emph{Perturb} manages to do well in domains where rewards
+are deep but can be  reached by  repeating the same action.
+This is the case of \Freeway, where a chicken has to run across a ten lane
+highway filled with traffic. Every time the chicken gets across, there is one
+unit of reward. If the chicken is hit by a car, it has to start again. 
+In \Freeway, 12 out of 18 possible actions have an effect:
+6 actions behave as move-up (up-right, up-left, up-fire, up-right-fire, up-left-fire) , 
+6 as  move-down (down-right, down-left, down-fire, down-right-fire, down-left-fire), 
+and 6 actions do nothing.  \emph{Perturb}  does well in this domain 
+when the selected fixed action is move-up.  As seen 
+in Table~\ref{performance}, \UCT\ does not manage to take any chicken
+across the highway. The reason that  \UCT\ performs so erratically 
+and the rollouts do not manage to collect any reward, is that there  is a $1/3$ probability of  
+choosing a move up action,  and   200 move up actions are required  in order to collect a reward. 
+%% the previous number was not correct; no-op actions do not count, and thus needs
+%%% 200 more move up actions than move down actions, although this is not accurate either
+%%%% as move down actions that take chicken below the screen count as no-ops too ..
+\IW\ does not have this limitation and is best in \Freeway.
 
-Consider \Freeway, where a chicken has to run across a ten lane
-highway filled with traffic. Every time it gets across, we
-receive a point, i.e. a reward of 1. If the chicken is hit by a car,
-the chicken is forced back. 12 out of 18 possible actions have an
-effect: move up ( up-right, up-left, up-fire, up-right-fire,
-up-left-fire, all have same effect as up ) or move down ( down-right,
-down-left, down-fire, down-right-fire, down-left-fire, all have same
-effect as down ), the remaining 6 actions are like no-ops. It's easy
-to understand that \emph{Perturb} would achieve a high score if the
-selected action is move up with 0.95 probability. In the following link
-\url{http://www.youtube.com/playlist?list=PLXpQcXUQ_CwenUazUivhXyYvjuS6KQOI0}
-we can observe videos of \IW\ and \UCT\ playing \Freeway. As reported
-in Table~\ref{performance}, \UCT\ does not manage to take any chicken
-across the highway, while \IW\ is the best performer out of all tested
-algorithms. The main reason why \UCT\ performs so erratically is that
-the rollouts do not manage to collect any reward, as there is a
-$3^{-1}$ probability to choose any move up action, and given that the
-depth of the closest reward is about 200 frames deep, the probability
-of a rollout achieveing any reward is $3^{-200}$, which implies
-choosing 200 move-up actions. On the other hand, \IW\ manages to find
-rewards without exhausting the full budget. The main reason that
-explains this behaviour is that at most only 3 generated states from
-the same parent node will pass the novelty test, as in any expanded
-node there is at most 3 different new possible bytes in the RAM state
-resulting from moving up, down, or do-nothing.
+%%
+%%% Sobre texto below; no me convence; no sabemos como cambia el RAM aunque podamos suponer
+%%% obvio: si un byte codifica donde esta el chicken en la pantalla .. tendriamos una explicación
+%% On the other hand, \IW\ manages to find rewards without exhausting the  budget. The main reason that
+%% explains this behaviour is that at most only 3 generated states from
+%% the same parent node will pass the novelty test, as in any expanded
+%% node there is at most 3 different new possible bytes in the RAM state
+%% resulting from moving up, down, or do-nothing.
 
 Another challenging game is \Montezuma. In this game rewards are very
-sparse, deep, and most of the actions lead to loosing a life, which
-only becomes effective after hundreds of frames, i.e. up-left may
-enforce jumping and falling until we loose a life. All algorithms achieve 0 score, wheter RL-based or Planning-based. But, as
-shown in Table~\ref{performance}, \BFS\ achieves an average score of
-80, which is still a very low score, as the first possible reward
-increases the score in 100. The best run by \BFS\ achieves a score of
-400, pointing out that \BFS\ still is not able to consistently find rewards.
-One way to improve the performance of any algorithm in \Montezuma\ is to encode the lost of a life as a negative reward.
-Capturing the lost of lifes in the reward function is relevant to many games where not loosing a life is important in later stages of the game.
-Currently,
-rewards are encoded directly as the change of score, which does not distinguish between the benefits of keeping lifes. This
-beahaviour is also observed in other games like \Breakout\ and \SpaceInvaders, where \BFS\
-is the best performer by far, although initially it looses all lifes but 1 and then plays extremely well. Videos on the link above illustrate
-this behaviour.
+sparse, deep, and most of the actions lead to losing a  life with
+no immediate penalty or consequence.  All algorithms achieve 0 score, 
+except for \BFS\ that achieves an average score of 80, and a score
+of 400 in one of the runs. This means that even \BFS\  is not able to 
+consistently find rewards in this game. This game and several other games 
+like \Breakout, and \SpaceInvaders\ could be much simpler by adding
+negative rewards for losing a life. We have indeed observed that
+our planning algorithms don't care much about losing lifes
+until there is just one life left, when they play much better.
+This can be seen in the videos mentioned above, and points
+to a simple form of learning that would be useful both
+for planners and reinforcement learning approaches.
 
-We do not report the performance of $\IW(2)$ because  in our
+We are not  reporting  the performance of $\IW(k)$ with parameter $k=2$ because  in our
 preliminary tests and according to the discussion in the previous section,
 it doesn't  appear to improve much on  \BRFS, even if  it results in a lookahead
-that is $5$ times deeper,  still far too shallow  to compete with the other
+that is $5$ times deeper,  but  still too shallow  to compete with the other
 planning algorithms.
@@
 underlying \IW($k$) by Dijkstra's algorithm yet such a move would make the relation between
-\IW\ and the notion of width less clear. Another option is to change  \IW\ to comply with 
+\IW\ and the notion of width less clear. A better  option is to change  \IW\ to comply with 
 a different property; namely, that if there is a ``rewarding'' path made up of states of 
 low width, then \IW\ will find such paths or better ones in time that is exponential 
-in their width. For this actually a simple change in \IW\ suffices: when generating a state 
+in their width. For this,  a simple change in \IW\ suffices: when generating a state 
 $s$ that is a \emph{duplicate} of a state $s'$ that has been previously generated and not pruned, 
@@
 must be propagated to the descendants of $R(s')$ that are in memory. This is similar to the change required 
-in A* for preserving optimality when moving from consistent to inconsistent heuristics \cite{pearl:heuristics}. 
-The alternative is to ``reopen'' such nodes. 
-%The same change is actually in \BFS\ to ensure that, 
-%if given enough time, \BFS\ will actually find the optimal paths.
-Changing \BFS\ to account for such paths guarantees that, if given enough time, \BFS\ will actually find
-the optimal paths. 
+in the  A* search algorithm for preserving optimality when moving from consistent to inconsistent heuristics \cite{pearl:heuristics}. 
+The alternative is to ``reopen'' such nodes. The same change is actually needed in \BFS\ to ensure that, 
+if given enough time, \BFS\ will actually find optimal paths.
+% Changing \BFS\ to account for such paths guarantees that, if given enough time, \BFS\ will actually find
+% the optimal paths. 
 The code used for \IW\ and \BFS\ in the
-experiments above does not implement this change, however, as the overhead involved in checking for duplicates 
-in some test cases did not appear to pay off. More experimentation is probably needed.
+experiments above does not implement this change  as the overhead involved in checking for duplicates 
+in some test cases did not appear to pay off. More experiments  however are  needed to
+find out if this is actually the most effective option.
 
@@
 that originate in work in classical planning,  can be used to play  the Atari video
-games and to achieve  state-of-the-art performance. This is the result of a 
-structured exploration of the state that manages to combine the scope of blind
-search algorithms with the efficiency of heuristic search methods. 
-%
-\Omit{
-We are currently addressing three
-issues, which we haven't managed to get in time for this submission. First, we want to extend the
-evaluation to all the games featured in \cite{bellemare:jair2013}, and to add more runs per  game.
-Second, we want to evaluate the various algorithms for different fixed time windows for decisions.
-In the table above, for example, \UCT\ is often taking an order-of-magnitude more time per decision
-than  \BFS. Third, we need to revise the way duplicate states are handled in \IW\ and \BFS\ to ensure
-that the best paths among the states that have been expanded are returned.Finally,}
-%
-There are many issues worth studying, we just mention two of them.
-One problem of all planning methods in Atari  games is that by using the 
-RAM state of the Atari console they ``cheat'' a bit, as they use information that
-is not available to the human player. In addition, a limitation of our
-algorithms is that they do not get better with experience.  
-In the future, we would like to explore whether  variations of these algorithms 
-can be used to play  from the information in the  screen state as opposed to the RAM state, 
-and the potential uses of the  ideas underlying these algorithms for learning.
+games where they  achieve  state-of-the-art performance. This level of play 
+is the result of a structured exploration of the state space that manages to combine 
+the scope of blind search algorithms with the efficiency of heuristic search methods. 
+There are several  issues worth studying, we  mention two of them.
+Regarding planning, the results show the potential of width-based algorithms
+for a broader class of planning problems. As shown, indeed, these algorithms
+do not require PDDL encodings and can deal with state successor functions
+encoded procedurally (as in simulators) as long as the structure of the states 
+is known. Regarding the Atari games, two problems of  all the  planning methods
+in ALE are first,  that by using the RAM state of the Atari console they are ``cheating'', 
+as they use information that is not available to the human player, and second, 
+that they do not get better with experience. In the future, we would like to explore 
+whether  variations of these algorithms  can be used to play  from the information in the  
+screen state as opposed to the RAM state, and potential uses of the  new ideas
+for learning to improve performance.
 

### 2014-11-19 icaps15-paper/paper.tex
Miquel's iteration

@@
 
+\newcommand{\numgames}{$35$}
+
+\newcommand{\Assault}{\propername{Assault}}
 \newcommand{\Asterix}{\propername{Asterix}}
+\newcommand{\Asteroids}{\propername{Asteroids}}
+\newcommand{\Atlantis}{\propername{Atlantis}}
+\newcommand{\BankHeist}{\propername{Bank Heist}}
+\newcommand{\BattleZone}{\propername{Battle Zone}}
 \newcommand{\BeamRider}{\propername{Beam Rider}}
+\newcommand{\Berzerk}{\propername{Berzerk}}
+\newcommand{\Bowling}{\propername{Bowling}}
+\newcommand{\Boxing}{\propername{Boxing}}
 \newcommand{\Breakout}{\propername{Breakout}}
+\newcommand{\Centipede}{\propername{Centipede}}
+\newcommand{\ChopperCommand}{\propername{Chopper Command}}
+\newcommand{\CrazyClimber}{\propername{Crazy Climber}}
 \newcommand{\Enduro}{\propername{Enduro}}
 \newcommand{\Freeway}{\propername{Freeway}}
+\newcommand{\Frostbite}{\propername{Frostbite}}
+\newcommand{\Gravitar}{\propername{Gravitar}}
+\newcommand{\Hero}{\propername{Hero}}
+\newcommand{\JamesBond}{\propername{James Bond}}
+\newcommand{\Kangaroo}{\propername{Kangaroo}}
+\newcommand{\Krull}{\propername{Krull}}
+\newcommand{\Montezuma}{\propername{Montezuma Revenge}}
+\newcommand{\MsPacman}{\propername{Ms Pacman}}
 \newcommand{\Pong}{\propername{Pong}}
+\newcommand{\Pooyan}{\propername{Pooyan}}
 \newcommand{\Qbert}{\propername{Q*Bert}}
+\newcommand{\RiverRaid}{\propername{Riverraid}}
 \newcommand{\Seaquest}{\propername{Seaquest}}
 \newcommand{\SpaceInvaders}{\propername{Space Invaders}}
+\newcommand{\StarGunner}{\propername{Star Gunner}}
+\newcommand{\Tennis}{\propername{Tennis}}
+\newcommand{\Venture}{\propername{Venture}}
+\newcommand{\VideoPinball}{\propername{Video Pinball}}
+\newcommand{\Zaxxon}{\propername{Zaxxon}}
+
 \newcommand{\Deepmind}{\propername{Deepmind}}
-\newcommand{\CrazyClimber}{\propername{Crazy\_climber}}
-\newcommand{\Kangaroo}{\propername{Kangaroo}}
-\newcommand{\Pooyan}{\propername{Pooyan}}
-\newcommand{\Berzerk}{\propername{Berzerk}}
-\newcommand{\Bowling}{\propername{Bowling}}
-\newcommand{\Montezuma}{\propername{Montezuma\_revenge}}
-\newcommand{\StarGunner}{\propername{Star\_gunner}}
-\newcommand{\Venture}{\propername{Venture}}
 
@@
 %    Miquel Ramirez \\
-%    RMIT University \\
-%    Melbourne, Australia\\
-%   {\normalsize\url{@rmit.edu.au}}
+%    Australian National University \\
+%    Canberra, Australia\\
+%   {\normalsize\url{@gmail.com}}
 %    \And
@@
 \begin{quote}
-The Atari  2600 games supported in  the Arcade Learning Environment (Bellemare et. al. 2013)
+The Atari 2600 games supported in  the Arcade Learning Environment~\cite{bellemare:jair2013}
 all feature a known initial (RAM) state and actions that have deterministic effects. Classical planners, however,
@@
 and rewards $r(a,s)$ represented by real numbers that result from  applying action $a$ in state $s$.
-The transition and reward functions do not need to be known a priori, yet in that case, the
+The transition and reward functions do not need to be known \emph{a priori}, yet in that case, the
 state and  reward that results from the application of an action in a state needs to be \emph{observable}.
@@
 
-\IW\ is a systematic and complete blind-search algorithm like breadth-first search (\BRFS) and iterative deepening (ID),
+\IW\ is a systematic and complete blind-search algorithm like breadth-first search (\BRFS) and iterative deepening (\ID),
 but   unlike these algorithms, it uses the factored representation of the states in terms of variables to structure the
@@
 of the other blind algorithms \BRFS\ and \ID\  that solve 9,010 and 8,762 problems each. This is shown in the bottom
-part of  Table~\ref{table1}. Moreover, \IW\ is  faster and results in shorter plans than a heuristic search algotrithm
+part of  Table~\ref{table1}. Moreover, \IW\ is  faster and results in shorter plans than a heuristic search algorithm
 \cite{nir:ecai2012}.
@@
 \input{games-table}
-We tested \IW(1) and \BFS\ over $35$ of the 55 games considered in
+We tested \IW(1) and \BFS\ over \numgames\
+ of the $55$ games considered in
 \cite{bellemare:jair2013}, from now on  abbreviated as BNVB.\footnote{
 Entries marked as  ``--'' in the table are   still running at the time of submission. 
-Each run can take up to 24 hours of CPU time and there are 5 runs per game and algorithm. 
-Eventually all of  the  55 games will  be included in the evaluation.}
+Each run can take up to $24$ hours of CPU time and there are $10$ runs per game and algorithm. 
+Eventually all of  the  $55$ games will  be included in the evaluation.}
 Table~\ref{performance} shows the performance of $\IW(1)$ and \BFS\
@@
 in comparison  with breadth-first search (\BRFS) and \UCT.
-Videos of selected games played by \IW(1), \BFS, and \UCT\ can be 
+Videos of selected games played by $\IW(1)$, \BFS, and \UCT\ can be 
 seen in Youtube.\footnote{\url{http://www.youtube.com/playlist?list=PLXpQcXUQ_CwenUazUivhXyYvjuS6KQOI0}.}
 The discount factor used by  all the algorithms is $\gamma=0.995$.
-The scores for \BRFS\ and \UCT\ are from BNVB.  
+The scores reported for \BRFS\ and \UCT\ are taken from BNVB.  
 Our experimental setup  follows theirs except that a maximum budget of  $150,000$ simulated frames 
@@
 This is why the average time per action is similar to all the algorithms
-except \IW(1), that due to its agressive pruning does not always use
+except $\IW(1)$, that due to its aggressive pruning does not always use
 the full budget and takes less time per action on average.
@@
 Also,  as reported by  BNVB, all of the algorithms 
-reuse the subtree from the previous lookahead that is rooted in the selected 
+reuse the sub-tree from the previous lookahead that is rooted in the selected 
 child, deleting its siblings and their descendants. More precisely, no calls 
-to the emulator are done for transitions that are cached in that subtree, 
+to the emulator are done for transitions that are cached in that sub-tree, 
 and such reused frames are not discounted from the budget that is thus a bound 
-on the number of ``new'' frames per lookahead. In addition, in \IW(1), the states that 
+on the number of ``new'' frames per lookahead. In addition, in $\IW(1)$, the states that 
 are ``reused'' from the previous searches are ignored in the computation of the novelty of 
-new states so that more states can escape pruning. Otherwise, \IW(1) often
+new states so that more states can escape pruning. Otherwise, $\IW(1)$ often
 uses a fraction of the  budget.  This is not needed in \BFS\ which does 
-no pruning.  \IW(1) and \BFS\ are  limited to  search up to a depth of
-$m=1500$ frames and up to $m=150,000$ frames per root branch. This is
+no pruning.  $\IW(1)$ and \BFS\ are  limited to  search up to a depth of
+$m=1,500$ frames and up to $m=150,000$ frames per root branch. This is
 to avoid the search from  going too deep or being too committed to a
@@
 
-Last, \IW(1) and \BFS\ select an action every 5 frames, while UCT  selects 
+Last, \IW(1) and \BFS\ select an action every $5$ frames, while \UCT\  selects 
 an action in every frame.  This means that in order to explore a branch 
-300 frames deep, UCT gets to choose 300 actions, while \IW(1) and \BFS\ 
-get to choose 60 actions, both however using the same 300 frames from the
+$300$ frames deep, UCT gets to choose $300$ actions, while \IW($1$) and \BFS\ 
+get to choose $60$ actions, both however using the same $300$ frames from the
 budget. For this, we  followed the setup  of \BRFS\ in BNVB 
-that also selects  actions every 5 frames. Since the lookahead budget 
-is given by a maximum number of (new) frames, and the time is mostly taken by 
-calls to the emulator, this may not be the best choice  for \IW(1)\ 
-and \BFS\ that are therefore not exploiting all the options afforded
-by the budget. We hope to fix this gap in the future.
+that also selects actions every $5$ frames, matching the behavior
+of the emulator that requests an action also every $5$ frames. Since the 
+lookahead budget is given by a maximum number of (new) frames, and the time 
+is mostly taken by calls to the emulator, this may not be the best choice  
+for $\IW(1)$ and \BFS\ that may therefore not be exploiting all the 
+options afforded by the budget. We hope to investigate the implications of
+this gap in the future.
 
@@
 The last difference inherited from BNVB between all algorithms
-and UCT, is that UCT selects an action randomly in the rollout every
+and \UCT\, is that \UCT\ selects an action randomly in the rollout every
 single frame, while the others select an action every 5 frames, and
@@
 
-Table~\ref{performance} shows that both $\IW(1)$ and \BFS\  outperform
-\BRFS, which rarely collects any reward in many domains, as the
-depth of the \BRFS\ search tree results in a lookahead of $0.3$
-seconds ($20$ frames or $4$ nodes deep). On the other hand, both
-$\IW(1)$ and \BFS\ reach states that are up to $350$--$1500$ frames
-deep ($70$--$260$ nodes or $6$--$22$ seconds), even if $\IW(1)$
-doesn't always use all the simulation frames allocated due to its
+Table~\ref{performance} shows that both $\IW(1)$ and \BFS\ generally
+ outperform \BRFS, which rarely collects any reward in many domains, as 
+ the depth of the \BRFS\ search tree results in a lookahead of $0.3$
+seconds ($20$ frames or $4$ nodes deep). The notable exception to this
+is \Centipede\ where abundant reward can be collected with a shallow
+lookahead. On the other hand, both $\IW(1)$ and \BFS\ reach states that 
+are up to $350$--$1,500$ frames deep ($70$--$260$ nodes or $6$--$22$ seconds), 
+even if $\IW(1)$ does not always use all the simulation frames allocated due to its
 pruning. This behaviour can be observed in games such as
@@
 
-More interestingly,  \IW(1) outscores \UCT\  in 17 of the 35 games,
-while \BFS\ outscores \UCT\ in 18.\footnote{These numbers can actually
-improve further as 6  games for \IW(1) and 4 games for \BFS\ are still running.}
-On the other, \UCT\ does better than \IW(1) and \BFS\ in 9 and 11 games respectively.
-The relative performance between \IW(1)\ and \BFS\ is closer with \IW(1) being  the
-best of the two in 12 games, and  \BFS\  in 11. In terms of the number of games 
+More interestingly,  \IW(1) outscores \UCT\  in $17$ of the \numgames\ games,
+while \BFS\ outscores \UCT\ in $18$.
+%\footnote{These numbers can actually
+%improve further as 6  games for \IW(1) and 4 games for \BFS\ are still running.}
+On the other, \UCT\ does better than $\IW(1)$ and \BFS\ in $9$ and $11$ games respectively.
+The relative performance between $\IW(1)$ and \BFS\ is closer with $\IW(1)$ being  the
+best of the two in $12$ games, and  \BFS\  in $11$. In terms of the number of games 
 where an algorithm is the overall winner, the situation is  even: 
-\IW(1) is the overall best in 12 games, \BFS\ in 11 games, and \UCT\ in 12 games. 
-Also, \BRFS\ is best in 1 game (Centipede), while the other three algorithms
+$\IW(1)$ is the overall best in $12$ games, \BFS\ in $11$ games, and \UCT\ in $12$ games. 
+Also, \BRFS\ is best in $1$ game (\Centipede), while the other three algorithms
 are tied in another game (\Pong).
@@
 
-$\IW(1)$ also outperforms the best learning algorithm in
-\cite{deep-mind-atari} in the $7$ games considered, and
-\BFS\ does so in $6$ of the $7$ games. Likewise, in \Freeway, \Berzerk, \Bowling, \Montezuma,
-\StarGunner, and \Venture, both $\IW(1)$ and \BFS\ attain a better
+Likewise, in \Freeway\ and \Berzerk\ both $\IW(1)$ and \BFS\ attain a better
 score than the baseline semi-random algorithm \emph{Perturb} in
-\cite{bellemare:jair2013}.
+\cite{bellemare:jair2013}, that beats \UCT\ on those games.
 %% to say the 2 lines below must add context: .. eg
@@
 \emph{Perturb} is a simple algorithm
-that selects  a fixed  action with probability 0.95, and a random
-action with probability 0.05.  For \emph{Perturb}, BNVB do not report 
+that selects  a fixed  action with probability $0.95$, and a random
+action with probability $0.05$.  For \emph{Perturb}, BNVB do not report 
 the average score  but the  best score.  \emph{Perturb} manages to do well in domains where rewards
@@
 unit of reward. If the chicken is hit by a car, it has to start again. 
-In \Freeway, 12 out of 18 possible actions have an effect:
-6 actions behave as move-up (up-right, up-left, up-fire, up-right-fire, up-left-fire) , 
-6 as  move-down (down-right, down-left, down-fire, down-right-fire, down-left-fire), 
-and 6 actions do nothing.  \emph{Perturb}  does well in this domain 
-when the selected fixed action is move-up.  As seen 
-in Table~\ref{performance}, \UCT\ does not manage to take any chicken
-across the highway. The reason that  \UCT\ performs so erratically 
-and the rollouts do not manage to collect any reward, is that there  is a $1/3$ probability of  
-choosing a move up action,  and   200 move up actions are required  in order to collect a reward. 
+In \Freeway, only $12$ out of $18$ possible actions have an effect:
+$6$ actions in moving the chicken up one lane (up-right, up-left, up-fire, up-right-fire, up-left-fire), 
+$6$ in moving the chicken down one lane (down-right, down-left, down-fire, down-right-fire, down-left-fire), 
+and $6$ actions do nothing.  \emph{Perturb}  does well in this domain 
+when the selected fixed action is one of the $6$ that moves the chicken up.  As noted 
+in Table~\ref{performance} and seen in the provided video, \UCT\ does not manage to take the chicken
+across the highway at all. The reason for \UCT\ to perform erratically 
+is that the roll-outs do not manage to collect any reward. A sequence of actions 
+with at least $10$ actions embedded that move the chicken up are required in order 
+to collect a reward. Since the probability for the roll-out to do an action that moves the chicken up 
+is $1/3$, the likelihood of \UCT\ generating such a sequence of actions with the parameters chosen
+by BNVB is rather low.
 %% the previous number was not correct; no-op actions do not count, and thus needs
@@
 %%%% as move down actions that take chicken below the screen count as no-ops too ..
-\IW\ does not have this limitation and is best in \Freeway.
+$\IW(1)$ does not have this limitation and is best in \Freeway.
 
@@
 
-Another challenging game is \Montezuma. In this game rewards are very
+$\IW(1)$ also outperforms the best learning algorithm in
+\cite{deep-mind-atari} in the $7$ games considered there, and
+\BFS\ does so in $6$ of the $7$ games. Comparing 
+with the scores reported for the best Reinforcement Learning algorithm
+in \cite{bellemare:jair2013}, we note that either $\IW(1)$ and \BFS\ 
+are outperforming significantly the best learning algorithm, in those games
+where it is reported to perform better than \UCT\ (\Montezuma, \Venture\
+and \Bowling). In \Montezuma\ rewards are very
 sparse, deep, and most of the actions lead to losing a life with
-no immediate penalty or consequence.  All algorithms achieve 0 score, 
-except for \BFS\ that achieves an average score of 80, and a score
-of 400 in one of the runs. This means that even \BFS\  is not able to 
+no immediate penalty or consequence.  All algorithms achieve $0$ score, 
+except for \BFS\ that achieves an average score of $80$, and a score
+of $400$ in one of the runs. This means that even \BFS\  is not able to 
 consistently find rewards in this game. This game and several other games 
@@
 negative rewards for losing a life. We have indeed observed that
-our planning algorithms don't care much about losing lifes
-until there is just one life left, when they play much better.
-This can be seen in the videos mentioned above, and points
-to a simple form of learning that would be useful both
+our planning algorithms do not care much about losing lives
+until there is just one life left, when their play noticeably improves.
+This can be seen in the videos mentioned above, and suggests that
+ some simple form of learning would be useful both
 for planners and reinforcement learning approaches.
@@
 do not require PDDL encodings and can deal with state successor functions
-encoded procedurally (as in simulators) as long as the structure of the states 
-is known. Regarding the Atari games, two problems of  all the  planning methods
-in ALE are first,  that by using the RAM state of the Atari console they are ``cheating'', 
-as they use information that is not available to the human player, and second, 
-that they do not get better with experience. In the future, we would like to explore 
-whether  variations of these algorithms  can be used to play  from the information in the  
-screen state as opposed to the RAM state, and potential uses of the  new ideas
-for learning to improve performance.
+encoded \emph{procedurally} (as in simulators) as long as the structure of the states 
+is known. This suggests to us that the algorithms discussed in the 
+present paper could be applied in a broad variety of settings where only on-line 
+action selection is feasible or suitable.
+
+Regarding the Atari games, we identify two problems in all the planning methods
+discussed. First, by using the RAM state of the Atari console one can consider
+that they are ``cheating'', as they use information that is not readily available to the human player, 
+and second, they do not get better with experience. In the future, we would like to explore 
+whether variations of these algorithms  can be used to play from the information conveyed by the  
+ state of screen pixels as opposed to the RAM state, as well as exploring the potential uses of the novel
+ideas presented for learning, in order to improve performance.
 

### 2014-11-19 icaps15-paper/paper.tex
Corrected sentence in Experimental Section

@@
 with at least $10$ actions embedded that move the chicken up are required in order 
-to collect a reward. Since the probability for the roll-out to do an action that moves the chicken up 
-is $1/3$, the likelihood of \UCT\ generating such a sequence of actions with the parameters chosen
-by BNVB is rather low.
+to collect a reward. Since the probability for the roll-out to choose an action that moves the chicken up 
+is $1/3$, with the parameters chosen by BNVB, the likelihood of \UCT\ generating such a sequence of actions 
+that \emph{also} avoids incoming traffic in a timely fashion is rather low.
 %% the previous number was not correct; no-op actions do not count, and thus needs

### 2014-11-19 icaps15-paper/paper.tex
Experimental section edits
* The number of actions moving the chicken up is 48 (it takes 4 seconds of
  consistently doing such actions to cross the highway).

@@
 is that the roll-outs do not manage to collect any reward. A sequence of actions 
-with at least $10$ actions embedded that move the chicken up are required in order 
+with at least $48$~\footnote{One needs to move up the chicken up for at least $4$ seconds ($240$ frames) in order to get across the 
+highway.} actions embedded that move the chicken up are required in order 
 to collect a reward. Since the probability for the roll-out to choose an action that moves the chicken up 

### 2014-11-19 workshop-paper/paper.tex
Submitted Workshop version

@@
 
-\author{Nir, Miquel, H}
+\author{Nir Lipovetzky \\
+      The University of Melbourne \\
+   Melbourne, Australia\\
+   {\normalsize\url{@unimelb.edu.au}}
+   \And
+   Miquel Ramirez \\
+   RMIT University \\
+   Melbourne, Australia\\
+   {\normalsize\url{@rmit.edu.au}}
+   \And
+   Hector Geffner \\
+   ICREA \&  Universitat Pompeu Fabra \\
+   Barcelona, SPAIN \\
+   {\normalsize\url{@upf.edu}}\thanks{firstname.lastname}
+}
+
 %AAAI Press\\
@@
 costs (rewards) or goals, and yet, like  heuristic search algorithms, they have been shown to be   effective for solving   problems over huge
-state spaces. The simplest such algorithm, called Iterated Width or \IW, consists of  a sequence of calls  $\IW(1)$, $\IW(2)$, 
-.., $\IW(k)$ where  $\IW(i)$ is  
-a breadth-first search in which a state is  pruned when it is not the first state in the search  to make true some set of i atoms or less.
+state spaces. The simplest such algorithm, called Iterated Width or \IW, consists of  a sequence of calls  $\IW(1)$, $\IW(2)$, \ldots,
+$\IW(k)$ where  $\IW(i)$ is  a breadth-first search in which a state is  pruned when it is not the
+first state in the search  to make true some set of i atoms.
 Preliminary empirical results suggest that the performance of \IW\ with the  $k$ parameter fixed to $1$,
 i.e., $\IW(1)$, is much better than random or breadth-first search, and approaches
-the performance of \UCT. A simple best-first variation  of \IW,  that combine exploration and exploitation,
+the performance of \UCT. A simple best-first variation  of \IW\  that combines exploration and exploitation
 is evaluated as well.
@@
 as in the Atari games, the termination condition is given by a time window or a maximum number of generated nodes.
-The  best path found by \IW is the path  that has a  maximum accumulated reward. 
-The accumulated reward $R(s)$  of a state $s$ reached in an iteration of \IW is determined by the  unique parent state $s'$ 
+The  best path found by \IW \ is the path  that has a  maximum accumulated reward.
+The accumulated reward $R(s)$  of a state $s$ reached in an iteration of \IW \ is determined by the  unique parent state $s'$
 and action $a$ leading to $s$ from $s'$ as $R(s) = R(s') + r(a,s')$. The best state is the state $s$ with maximum reward
 $R(s)$ generated but not pruned by \IW, and the best path is the one that leads to the state $s$ from the current state.
-The action selected to be done next is then the first action along such a path.
+The action selected to be done next is  the first action along such a path.
 
@@
 \cite{bonet:aij-hsp} on the other (this algorithm is similar
-to one used by  the FF planner when hill climbing search fails \cite{hoffmann:ff}), 
+to the one used by  the FF planner when hill climbing search fails \cite{hoffmann:ff}),
 it turns out that ``blind'' \IW\ solves as many problems as the informed search, 34,627 vs. 34,849, far ahead
 of the other blind algorithms \BRFS\ and \ID\  that solve 9,010 and 8,762 problems each. This is shown in the bottom
-part of the Table~\ref{table1}. Moreover, \IW\ is  faster and results in shorter plans than the heuristic search  
+part of the Table~\ref{table1}. Moreover, \IW\ is  faster and results in shorter plans than a heuristic search algotrithm
 \cite{nir:ecai2012}.
 
-The min $k$ value for which $\IW(k)$ solves a problem  is clearly bounded and small in most of
+The min $k$ value for which $\IW(k)$ solves a problem  is bounded and small in most of
 these instances. This is  actually no  accident and  has a \emph{theoretical explanation.} Lipovetzky and Geffner define
@@
 A)~these ``stepping stones''  contain at most $i$ atoms each, B)~each  stepping stone $t_{i+1}$ is
-at distance 1 from the previous one $t_i$ when $t_i$ has been reached following the ``trail'', 
+at distance 1 from the previous one $t_i$ when $t_i$ has been reached by following the ``trail'',
 C)~the ``trail''  preserves ``optimality''; i.e., no  tuple $t_i$ can be reached in less than $i$ steps.
@@
 relevant  in  optimization problems as well. Indeed, if a  good path is made of
-states $s_i$ each of which have a low width, \IW\ can be made 
+states $s_i$ each of which has a low width, \IW\ can be made
 to find such path in low polynomial time for a small value of the $k$ parameter.
@@
 
- \caption{\small Random baseline, \BRFS, $\IW(1)$, \BFS, \UCT~performance over 5 runs,
-   setting the Atari 2600 random seeds to $0, \ldots, 4$, and maximum episode duration of $18,000$
-   frames. For every algorithm, minus the trivial random baseline algorithm, we report the average CPU time per call 
-   to each action selection algorithm, along with the average score achieved. \UCT~performance
-   reported in \cite{bellemare:jair2013} is shown in parenthesis. Numbers
-   in bold show best performer in terms of average score.
-   }
+ \caption{\small Performance that results from random policy and various  lookahead algorithms in 9 Atari games.
+The algorithms  are \BRFS, $\IW(1)$, \BFS, and \UCT. Averages  over 5 runs shown with ``random'' seeds
+$0$ to $4$, and maximum episode duration of $18,000$  frames. Every algorithm limited to a budget of
+100,000 simulated frames, except for UCT which is limited to 500 rollouts of 300 frames depth instead. Average scores and CPU times per action selection shown.
+Performance of UCT results  reported in \cite{bellemare:jair2013} shown in parenthesis.
+Numbers  in bold show best performer in terms of average score. }
 \label{performance}
@@
 is compared with that of \UCT, breadth-first search (\BRFS), and random action selection.
-Our experimental setup  follows the one in  \cite{bellemare:jair2013} with two exceptions. First, rather than
-considering $10$ episodes of each game, we have considered just $5$ due to time constraints, as the experiments 
-for each  game can take up to $40$ hours of CPU time. Second, while \UCT\ is still limited to perform up to $500$ base policy rollouts 
-of depth $300$, a maximum of budget of  $100,000$ simulated frames is applied to all the algorithms, \UCT\ included. 
+Our experimental setup  follows the one in  \cite{bellemare:jair2013} with one exception: rather than
+considering $10$ runs per games, we considered just $5$ due to time constraints. Following the setup in  \cite{bellemare:jair2013}, while \UCT\ is limited to perform up to $500$ base policy rollouts
+of depth $300$\footnote{In \UCT,  $500$ base policy rollouts
+of depth $300$ is equivalent to a budget of $150,000$ simulated frames}, a maximum budget of  $100,000$ simulated frames is applied to all the other algorithms.
 %% Was not clear; hope this is what you meant, if not fix
@@
 %% matching the setup used during actual gameplay. %%  (an action is selected every $5$ frames, using a no-op for the remaining $4$). %% Needed?
-The average scores reported for \UCT\  \cite{bellemare:jair2013}, without this extra limitation, are included
-in  parenthesis for reference. $\IW(1)$ and \BFS\ have been  limited to a planning horizon $m=1500$ where one an action is evaluated every 
-$5$ frames. The discount factor used by all the algorithms is $\gamma=0.995$, and all of them  reuse the tree built 
-in the previous lookahead by setting the root of the tree to the best child of the root, while 
-deleting its siblings and their descendants as in \cite{bellemare:jair2013}.
+The average scores reported for \UCT\  \cite{bellemare:jair2013}, are included
+in  parenthesis for reference. $\IW(1)$ and \BFS\ have been  limited to a planning horizon $m=1500$ with an action selected
+every $5$ frames. The discount factor used by all the algorithms is $\gamma=0.995$, and all of them  reuse the tree built
+in the previous lookahead as in \cite{bellemare:jair2013}, by setting the root of the tree to the best child of the root,
+while deleting its siblings and their descendants
 %%% really need what's below? it's not very clear .. at least to me HG
@@
 it doesn't  appear to improve much on  \BRFS, even if  it results in a lookahead
-that is $5$ times deeper, yet still far too shallow  to compete with the other 
+that is $5$ times deeper,  still far too shallow  to compete with the other
 planning algorithms.
@@
 provides a different scheme. The first converges to the optimal behavior asymptotically; the
-second in a bounded number of steps. Alternatively, novelty-guided probes, possibly with 
-some noise added, could potentially pay off as  a  base policy for \UCT\ in replacement 
-of random roll outs, as they apparently manage to exploit the structure of many problems.
+second in a bounded number of steps.
+% Alternatively, novelty-guided probes, possibly with
+% some noise added, could potentially pay off as  a  base policy for \UCT\ in replacement
+% of random roll outs, as they apparently manage to exploit the structure of many problems.
 
@@
 that the best paths among the states that have been expanded are returned. Finally,
-for the future, we also want to explore whether this type of algorithms
-can give meaningful results as well  when used on the screen state, as opposed 
-to the RAM state.
+for the future, we  want to explore whether this type of algorithms
+can give meaningful results   when used on the screen state, as opposed
+to the RAM state, and their potential uses for learning.
 

### 2014-11-19 icaps15-paper/paper.tex
added new games

@@
 
-\newcommand{\mc}[2]{\multicolumn{#1}{b}{#2}}
+\newcommand{\mc}[2]{\multicolumn{#1}{a}{#2}}
 
@@
 
-\newcommand{\numgames}{$35$}
+\newcommand{\numgames}{$48$}
 
@@
 \newcommand{\Breakout}{\propername{Breakout}}
+\newcommand{\Carnival}{\propername{Carnival}}
 \newcommand{\Centipede}{\propername{Centipede}}
@@
 \newcommand{\CrazyClimber}{\propername{Crazy Climber}}
+\newcommand{\DoubleDunk}{\propername{Double Dunk}}
 \newcommand{\Enduro}{\propername{Enduro}}
+\newcommand{\FishingDerby}{\propername{Fishing Derby}}
 \newcommand{\Freeway}{\propername{Freeway}}
@@
 \newcommand{\JamesBond}{\propername{James Bond}}
+\newcommand{\JourneyEscape}{\propername{Journey Escape}}
 \newcommand{\Kangaroo}{\propername{Kangaroo}}
 \newcommand{\Krull}{\propername{Krull}}
+\newcommand{\KungFuMaster}{\propername{Kung Fu Master}}
 \newcommand{\Montezuma}{\propername{Montezuma Revenge}}
 \newcommand{\MsPacman}{\propername{Ms Pacman}}
+\newcommand{\NameThisGame}{\propername{Name This Game}}
 \newcommand{\Pong}{\propername{Pong}}
 \newcommand{\Pooyan}{\propername{Pooyan}}
+\newcommand{\PrivateEye}{\propername{Private Eye}}
 \newcommand{\Qbert}{\propername{Q*Bert}}
 \newcommand{\RiverRaid}{\propername{Riverraid}}
+\newcommand{\RobotTank}{\propername{Robot Tank}}
 \newcommand{\Seaquest}{\propername{Seaquest}}
@@
 \newcommand{\Tennis}{\propername{Tennis}}
+\newcommand{\TimePilot}{\propername{Time Pilot}}
+\newcommand{\Tutankham}{\propername{Tutankham}}
+\newcommand{\UpNDown}{\propername{Up And Down}}
 \newcommand{\Venture}{\propername{Venture}}
 \newcommand{\VideoPinball}{\propername{Video Pinball}}
+\newcommand{\WizardOfWor}{\propername{Wizard Of Wor}}
 \newcommand{\Zaxxon}{\propername{Zaxxon}}
@@
 
-Last, \IW(1) and \BFS\ select an action every $5$ frames, while \UCT\  selects 
+Last, \IW(1) and \BFS\  in the lookahead select an action every $5$ frames, while \UCT\  selects 
 an action in every frame.  This means that in order to explore a branch 
@@
 
-More interestingly,  \IW(1) outscores \UCT\  in $17$ of the \numgames\ games,
-while \BFS\ outscores \UCT\ in $18$.
+More interestingly,  \IW(1) outscores \UCT\  in $27$ of the \numgames\ games,
+while \BFS\ outscores \UCT\ in $21$.
 %\footnote{These numbers can actually
 %improve further as 6  games for \IW(1) and 4 games for \BFS\ are still running.}
-On the other, \UCT\ does better than $\IW(1)$ and \BFS\ in $9$ and $11$ games respectively.
-The relative performance between $\IW(1)$ and \BFS\ is closer with $\IW(1)$ being  the
-best of the two in $12$ games, and  \BFS\  in $11$. In terms of the number of games 
-where an algorithm is the overall winner, the situation is  even: 
-$\IW(1)$ is the overall best in $12$ games, \BFS\ in $11$ games, and \UCT\ in $12$ games. 
-Also, \BRFS\ is best in $1$ game (\Centipede), while the other three algorithms
-are tied in another game (\Pong).
+On the other, \UCT\ does better than $\IW(1)$ and \BFS\ in $15$ and $17$ games respectively.
+The relative performance between $\IW(1)$ and \BFS\ makes $\IW(1)$  the
+best of the two in $22$ games, and  \BFS\  in $12$. In terms of the number of games 
+where an algorithm is the overall winner, $\IW(1)$  domainates the others: 
+$\IW(1)$ is the overall best in $23$ games, \BFS\ in $11$ games, and \UCT\ in $18$ games. 
+Also, \BRFS\ is best in $2$ games (\Centipede, tied up in \Boxing), while the other three algorithms
+are tied in another 2 games (\Pong, \Boxing), and $\IW(1)$ and \BFS\ are tied in \TimePilot.
 
@@
 highway filled with traffic. Every time the chicken gets across, there is one
-unit of reward. If the chicken is hit by a car, it has to start again. 
+unit of reward. If the chicken is hit by a car, it goes back some lanes. 
 In \Freeway, only $12$ out of $18$ possible actions have an effect:

### 2014-11-19 icaps15-paper/paper.tex
changed 35 to 48 games in abstract

@@
 first state in the search  to make true some subset of $i$ atoms. 
-The  empirical results over 35 games suggest that the performance 
+The  empirical results over \numgames\ games suggest that the performance 
 of \IW\ with the  $k$ parameter fixed to $1$, i.e., $\IW(1)$, is at the level of the state of the art represented by 

### 2014-11-19 icaps15-paper/paper.tex
submited version to icaps

@@
 Our experimental setup  follows theirs except that a maximum budget of  $150,000$ simulated frames 
-is applied to \textit{all the algorithms}. \UCT\ uses this budget by running
+is applied to $\IW(1)$, \BFS, and \UCT. \UCT\ uses this budget by running
  $500$   rollouts of depth $300$. % \footnote{In \UCT,  $500$ base policy rollouts
@@
 for $\IW(1)$ and \BFS\ that may therefore not be exploiting all the 
-options afforded by the budget. We hope to investigate the implications of
-this gap in the future.
+options afforded by the budget. We'll look at this further in the future.
 
@@
 %improve further as 6  games for \IW(1) and 4 games for \BFS\ are still running.}
-On the other, \UCT\ does better than $\IW(1)$ and \BFS\ in $15$ and $17$ games respectively.
+On the other hand, \UCT\ does better than $\IW(1)$ and \BFS\ in $15$ and $17$ games respectively.
 The relative performance between $\IW(1)$ and \BFS\ makes $\IW(1)$  the
@@
 are deep but can be reached by  repeating the same action.
-This is the case of \Freeway, where a chicken has to run across a ten lane
-highway filled with traffic. Every time the chicken gets across, there is one
+This is the case of \Freeway, where a chicken has to run to the top of the screen across a ten lane
+highway filled with traffic. Every time the chicken gets across (starting at the bottom), there is one
 unit of reward. If the chicken is hit by a car, it goes back some lanes. 
-In \Freeway, only $12$ out of $18$ possible actions have an effect:
-$6$ actions in moving the chicken up one lane (up-right, up-left, up-fire, up-right-fire, up-left-fire), 
-$6$ in moving the chicken down one lane (down-right, down-left, down-fire, down-right-fire, down-left-fire), 
+In \Freeway, only $12$ out of the $18$ possible actions have an effect:
+$6$ actions move the chicken up (up-right, up-left, up-fire, up-right-fire, up-left-fire), 
+$6$ actions move the chicken down (down-right, down-left, down-fire, down-right-fire, down-left-fire), 
 and $6$ actions do nothing.  \emph{Perturb}  does well in this domain 
-when the selected fixed action is one of the $6$ that moves the chicken up.  As noted 
+when the selected fixed action moves the chicken up.  As noted 
 in Table~\ref{performance} and seen in the provided video, \UCT\ does not manage to take the chicken
-across the highway at all. The reason for \UCT\ to perform erratically 
-is that the roll-outs do not manage to collect any reward. A sequence of actions 
-with at least $48$~\footnote{One needs to move up the chicken up for at least $4$ seconds ($240$ frames) in order to get across the 
-highway.} actions embedded that move the chicken up are required in order 
-to collect a reward. Since the probability for the roll-out to choose an action that moves the chicken up 
-is $1/3$, with the parameters chosen by BNVB, the likelihood of \UCT\ generating such a sequence of actions 
-that \emph{also} avoids incoming traffic in a timely fashion is rather low.
-%% the previous number was not correct; no-op actions do not count, and thus needs
-%%% 200 more move up actions than move down actions, although this is not accurate either
-%%%% as move down actions that take chicken below the screen count as no-ops too ..
-$\IW(1)$ does not have this limitation and is best in \Freeway.
+across the highway at all. 
+%
+The reason that \UCT\ does not collect any reward is that it needs to move the chicken up
+at least 240 times~\footnote{One needs to move the chicken up for at least $4$ seconds ($240$ frames) in order to get across the 
+highway.} something that is very unlikely in a random exploration. $\IW(1)$ does not have this limitation and is best in \Freeway.
+%% The reason for \UCT\ to perform erratically 
+%% is that the roll-outs do not manage to collect any reward. A sequence of actions 
+%% with at least $48$ actions embedded that move the chicken up are required in order 
+%% to collect a reward. Since the probability for the roll-out to choose an action that moves the chicken up 
+%% is $1/3$, with the parameters chosen by BNVB, the likelihood of \UCT\ generating such a sequence of actions 
+%% that \emph{also} avoids incoming traffic in a timely fashion is rather low.
+%% %% the previous number was not correct; no-op actions do not count, and thus needs
+%% %%% 200 more move up actions than move down actions, although this is not accurate either
+%% %%%% as move down actions that take chicken below the screen count as no-ops too ..
+%% $\IW(1)$ does not have this limitation and is best in \Freeway.
 
@@
 \cite{deep-mind-atari} in the $7$ games considered there, and
-\BFS\ does so in $6$ of the $7$ games. Comparing 
-with the scores reported for the best Reinforcement Learning algorithm
-in \cite{bellemare:jair2013}, we note that either $\IW(1)$ and \BFS\ 
-are outperforming significantly the best learning algorithm, in those games
-where it is reported to perform better than \UCT\ (\Montezuma, \Venture\
-and \Bowling). In \Montezuma\ rewards are very
+\BFS\ does so in $6$ of the $7$ games. Comparing with the scores reported for the reinforcement learning algorithms by BNVB,
+we note that both $IW(1)$ and \BFS\ do much better than the best learning algorithm in
+those games where the learning algorithms outperform \UCT\ like \Montezuma, \Venture\
+and \Bowling.  In \Montezuma\ rewards are very
 sparse, deep, and most of the actions lead to losing a life with
@@
 until there is just one life left, when their play noticeably improves.
-This can be seen in the videos mentioned above, and suggests that
- some simple form of learning would be useful both
-for planners and reinforcement learning approaches.
+This can be seen in the videos mentioned above, and suggest a simple form of learning that would be useful to both planners
+and reinforcement learning algorithms.
 
@@
 maximum reward.  \IW\ may actually  fail to find such paths even when calling \IW($k$) with a high 
-$k$ parameter. Optimality  could be  achieved then  by replacing the breadth-first search
+$k$ parameter. Optimality  could be  achieved by replacing the breadth-first search
 underlying \IW($k$) by Dijkstra's algorithm yet such a move would make the relation between
@@
 $s$ that is a \emph{duplicate} of a state $s'$ that has been previously generated and not pruned, 
-$s'$ is replaced by $s$ if $R(s) > R(s')$. Moreover in such a case, the change in reward 
-must be propagated to the descendants of $R(s')$ that are in memory. This is similar to the change required 
+$s'$ is replaced by $s$ if $R(s) > R(s')$, with the change of reward propagated to the descendants of $s'$
+that are in memory. This is similar to the change required 
 in the  A* search algorithm for preserving optimality when moving from consistent to inconsistent heuristics \cite{pearl:heuristics}. 
@@
 encoded \emph{procedurally} (as in simulators) as long as the structure of the states 
-is known. This suggests to us that the algorithms discussed in the 
-present paper could be applied in a broad variety of settings where only on-line 
-action selection is feasible or suitable.
+is known. 
+%% This suggests to us that the algorithms discussed in the 
+%% present paper could be applied in a broad variety of settings where only on-line 
+%% action selection is feasible or suitable.
 
@@
 discussed. First, by using the RAM state of the Atari console one can consider
-that they are ``cheating'', as they use information that is not readily available to the human player, 
-and second, they do not get better with experience. In the future, we would like to explore 
+that they are ``cheating'', as they use information that is not available to the human player. And second, they do not get better with experience. In the future, we would like to explore 
 whether variations of these algorithms  can be used to play from the information conveyed by the  
- state of screen pixels as opposed to the RAM state, as well as exploring the potential uses of the novel
-ideas presented for learning, in order to improve performance.
+ state of screen pixels as opposed to the RAM state, and the potential uses of the new ideas for learning to improve performance.
 

### 2014-11-19 src/games/supported/MontezumaRevenge.cpp

@@ -40,6 +40,11 @@ void MontezumaRevengeSettings::step(const System& system) {
 
     // update terminal status
     int new_lives = readRam(&system, 0xBA);
+    int m_lives = new_lives;
+    // if( new_lives < 5){
+    // 	//std::cout << "new lives" << new_lives << std::endl;
+    // 	m_reward = -10;
+    // }
     int some_byte = readRam(&system, 0xFE);
     m_terminal = new_lives == 0 && some_byte == 0x60;
 }

### 2014-11-19 src/games/supported/MontezumaRevenge.hpp

@@ -54,6 +54,8 @@ class MontezumaRevengeSettings : public RomSettings {
         bool m_terminal;
         reward_t m_reward;
         reward_t m_score;
+        reward_t m_lives;
+    
 };
 
 #endif // __MONTEZUMAREVENGE_HPP__

### 2014-11-28 src/ale_interface.hpp
Stella internal RNG has been harnessed
* Moved initialization of Stella's internal RNG to the same function
  we set the seed for the Standard Library RNG

@@ -18,6 +18,7 @@
 
 #include "emucore/FSNode.hxx"
 #include "emucore/OSystem.hxx"
+#include "emucore/Random.hxx"
 #include "os_dependent/SettingsWin32.hxx"
 #include "os_dependent/OSystemWin32.hxx"
 #include "os_dependent/SettingsUNIX.hxx"
@@ -102,12 +103,15 @@ static void createOSystem(int argc, char* argv[],
     // seed random number generator
     if (theOSystem->settings().getString("random_seed") == "time") {
         cerr << "Random Seed: Time" << endl;
+	Random randGen;
         srand((unsigned)time(0));
         //srand48((unsigned)time(0));
     } else {
         int seed = theOSystem->settings().getInt("random_seed");
         assert(seed >= 0);
         cerr << "Random Seed: " << seed << endl;
+	Random::seed(seed);
+
         srand((unsigned)seed);
         //srand48((unsigned)seed);
     }

### 2014-11-28 src/environment/stella_environment.cpp

@@ -57,7 +57,8 @@ StellaEnvironment::StellaEnvironment(OSystem* osystem, RomSettings* settings):
 /** Resets the system to its start state. */
 void StellaEnvironment::reset() {
   // RNG for generating environments
-  Random randGen;
+  // MRJ: This is done now in src/ale_interface.hpp, at function createOSystem()
+  // Random randGen;
   
   m_state.resetEpisodeFrameNumber();
   // Reset the paddles

### 2014-11-30 icaps15-paper/paper.tex
added new table. Still need to change the experimental section. This change is towards the workshop camera ready.
Miquel's final iteration

@@
- 
  \documentclass[letterpaper]{article}
@@
 
-\newcommand{\numgames}{$48$}
+\newcommand{\numgames}{$54$}
 
+\newcommand{\Alien}{\propername{Alien}}
+\newcommand{\Amidar}{\propername{Amidar}}
 \newcommand{\Assault}{\propername{Assault}}
@@
 \newcommand{\CrazyClimber}{\propername{Crazy Climber}}
+\newcommand{\DemonAttack}{\propername{Demon Attack}}
 \newcommand{\DoubleDunk}{\propername{Double Dunk}}
+\newcommand{\ElevatorAction}{\propername{Elevator Action}}
 \newcommand{\Enduro}{\propername{Enduro}}
@@
 \newcommand{\Frostbite}{\propername{Frostbite}}
+\newcommand{\Gopher}{\propername{Gopher}}
 \newcommand{\Gravitar}{\propername{Gravitar}}
 \newcommand{\Hero}{\propername{Hero}}
+\newcommand{\IceHockey}{\propername{Ice Hockey}}
 \newcommand{\JamesBond}{\propername{James Bond}}
@@
 \newcommand{\RiverRaid}{\propername{Riverraid}}
+\newcommand{\RoadRunner}{\propername{Road Runner}}
 \newcommand{\RobotTank}{\propername{Robot Tank}}
 \newcommand{\Seaquest}{\propername{Seaquest}}
+\newcommand{\Skiing}{\propername{Skiing}}
 \newcommand{\SpaceInvaders}{\propername{Space Invaders}}
@@
 /Title (Classical Planning Algorithms on the Atari Video Games)
-/Author ( )}
-% (Nir Lipovetzky, Miquel Ramirez, Hector Geffner)}
+/Author (Nir Lipovetzky, Miquel Ramirez, Hector Geffner)}
 
@@
 
-\author{\textbf{ID:} 52 \ \ ; \ \ \textbf{Keywords:} classical planning, video games}
+% \author{\textbf{ID:} 52 \ \ ; \ \ \textbf{Keywords:} classical planning, video games}
 
-% \author{Nir Lipovetzky \\
-%       The University of Melbourne \\
-%    Melbourne, Australia\\
-%    {\normalsize\url{@unimelb.edu.au}}
-%    \And
-%    Miquel Ramirez \\
-%    Australian National University \\
-%    Canberra, Australia\\
-%   {\normalsize\url{@gmail.com}}
-%    \And
-%    Hector Geffner \\
-%    ICREA \&  Universitat Pompeu Fabra \\
-%    Barcelona, SPAIN \\
-%    {\normalsize\url{@upf.edu}}\thanks{firstname.lastname}
-% }
+ \author{Nir Lipovetzky \\
+       The University of Melbourne \\
+    Melbourne, Australia\\
+%     {\normalsize\url{nir.lipovetzky@unimelb.edu.au}}
+%%% footnote doesn't look nice ; this looks better
+%% although can't use Nir's unimelb address that is too long :-)
+%% 
+    {\small\url{nirlipo@gmail.com}}
+    \And
+    Miquel Ramirez \\
+    Australian National University \\
+    Canberra, Australia\\
+   {\small\url{miquel.ramirez@gmail.com}}
+    \And
+    Hector Geffner \\
+    ICREA \&  Universitat Pompeu Fabra \\
+    Barcelona, SPAIN \\
+    {\small\url{hector.geffner@upf.edu}}
+ %%%%% \thanks{firstname.lastname}
+}
 
@@
 The transition and reward functions do not need to be known \emph{a priori}, yet in that case, the
-state and  reward that results from the application of an action in a state needs to be \emph{observable}.
+state and  reward that results from the application of an action in a state need to be \emph{observable}.
 The task is to compute an action sequence $a_0, \ldots, a_m$ for a large horizon $m$
@@
 
-Formally, the \emph{width} $w(P)$ of a problem $P$ is $i$
-iff $i$ is the minimum positive integer for which
-there is a sequence of tuples (sets) with at most $i$ atoms each, $t_0, t_1, \ldots, t_n$
-such that 1)~$t_0$ is true in the initial state of $P$, 2)~any shortest plan $\pi$  that achieves $t_i$ in $P$
-can be extended into a shortest plan that achieves $t_{i+1}$ by extending   $\pi$ with one action, and
-3)~any shortest plan that achieves $t_n$ is also a shortest plan for $P$.
+Formally, the \emph{width} $w(P)$ of a problem $P$ is $i$ iff $i$ is the minimum positive integer for which
+there is a sequence $t_0, t_1, \ldots, t_n$ of  atom sets  $t_k$  with at most $i$ atoms each, 
+such that 1)~$t_0$ is true in the initial state of $P$, 2)~any shortest plan $\pi$  that achieves $t_k$ in $P$
+can be extended into a shortest plan that achieves $t_{k+1}$ by extending   $\pi$ with one action, and
+3)~any shortest plan that achieves $t_n$ is  a shortest plan for achieving the goal of $P$.
 
@@
 ``trail of  stepping stones''  $t_0, t_1, \ldots, t_n$ to reach the problem goal such that
-A)~these ``stepping stones''  contain at most $i$ atoms each, B)~each  stepping stone $t_{i+1}$ is
-at distance 1 from the previous one $t_i$ when $t_i$ has been reached following the ``trail'',
-C)~the ``trail''  preserves ``optimality''; i.e., no  tuple $t_i$ can be reached in less than $i$ steps.
+A)~these ``stepping stones''  contain at most $i$ atoms each, B)~each  stepping stone $t_{k+1}$ is
+at distance 1 from the previous one $t_k$ when $t_k$ has been reached following the ``trail'',
+C)~the ``trail''  preserves ``optimality''; i.e., no  tuple $t_k$ can be reached in less than $k$ steps.
 
@@
 for problems where a goal state needs to be reached, the notions remain
-relevant  in  optimization problems as well. Indeed, if a  good path is made of
+relevant  in  optimisation problems as well. Indeed, if a  good path is made of
 states $s_i$ each of which has a low width, \IW\ can be made
@@
 The number of nodes \emph{generated} by $\IW(1)$ is $n \times D \times b$ in the worst case,
-where $n$ is the number of variables, $D$ is the size of their domains, and $b$ is the number of
-actions per  state.  This same number in breadth-first search is not linear in $n$
+where $n$ is the number of problem variables, $D$ is the size of their domains, and $b$ is the number of
+actions per  state.  This same number in a  breadth-first search is not linear in $n$
 but exponential. For the Atari games,  $n=128$, $D=256$, and $b=18$, so that the product is
@@
 First, in the breadth-first search underlying $\IW(1)$, we generate the children in
-random order. This makes the executions that results from the
-$\IW(1)$ lookahead less susceptible to fall into loops; a potential
+random order. This makes the executions that result from the
+$\IW(1)$ lookahead less susceptible to be trapped into loops; a potential
 problem in local search algorithms with no memory or learning.
@@
 \input{games-table}
-We tested \IW(1) and \BFS\ over \numgames\
- of the $55$ games considered in
-\cite{bellemare:jair2013}, from now on  abbreviated as BNVB.\footnote{
-Entries marked as  ``--'' in the table are   still running at the time of submission. 
-Each run can take up to $24$ hours of CPU time and there are $10$ runs per game and algorithm. 
-Eventually all of  the  $55$ games will  be included in the evaluation.}
+We tested \IW(1) and \BFS\ over \numgames\  of the $55$ games 
+considered in \cite{bellemare:jair2013}, from now on  abbreviated as BNVB.\footnote{
+We left out  \textsc{Skiing} as the reported figures apparently use a  different reward structure.}
+%% check
+% \footnote{
+% Entries marked as  ``--'' in the table are   still running at the time of submission. 
+% Each run can take up to $24$ hours of CPU time and there are $10$ runs per game and algorithm. 
+% Eventually all of  the  $55$ games will  be included in the evaluation.}
 Table~\ref{performance} shows the performance of $\IW(1)$ and \BFS\
-% , characterized by the average score obtained over all episodes, the standard deviation
-% of the score obtained and time (in seconds) per lookahead,
 in comparison  with breadth-first search (\BRFS) and \UCT.
@@
 Also,  as reported by  BNVB, all of the algorithms 
-reuse the sub-tree from the previous lookahead that is rooted in the selected 
+reuse the frames in the sub-tree of the previous lookahead that is rooted in the selected 
 child, deleting its siblings and their descendants. More precisely, no calls 
@@
 and such reused frames are not discounted from the budget that is thus a bound 
-on the number of ``new'' frames per lookahead. In addition, in $\IW(1)$, the states that 
-are ``reused'' from the previous searches are ignored in the computation of the novelty of 
+on the number of \emph{new}  frames per lookahead. In addition, in $\IW(1)$, the states that 
+are reused  from the previous searches are ignored in the computation of the novelty of 
 new states so that more states can escape pruning. Otherwise, $\IW(1)$ often
@@
 
-Table~\ref{performance} shows that both $\IW(1)$ and \BFS\ generally
- outperform \BRFS, which rarely collects any reward in many domains, as 
- the depth of the \BRFS\ search tree results in a lookahead of $0.3$
+Table~\ref{performance} shows that both $\IW(1)$ and \BFS\ outperform \BRFS, which  rarely collects
+reward in  many domains as  the depth of the \BRFS\ search tree results in a lookahead of $0.3$
 seconds ($20$ frames or $4$ nodes deep). The notable exception to this
 is \Centipede\ where abundant reward can be collected with a shallow
-lookahead. On the other hand, both $\IW(1)$ and \BFS\ reach states that 
-are up to $350$--$1,500$ frames deep ($70$--$260$ nodes or $6$--$22$ seconds), 
-even if $\IW(1)$ does not always use all the simulation frames allocated due to its
-pruning. This behaviour can be observed in games such as
-\Breakout, \CrazyClimber, \Kangaroo, and \Pooyan, where the average
-CPU time for each lookahead is up to $10$ times faster than \BFS. 
-Computation time for \UCT\ and \BRFS\ are  similar to \BFS, as the
-most expensive part of the computation is the  generation of 
-frames through the simulator,  and these three  algorithms 
+lookahead. On the other hand, both $\IW(1)$ and \BFS\ normally 
+reach states that are up to $350$--$1500$ frames deep ($70$--$260$ nodes or 
+$6$--$22$ seconds),  even if $\IW(1)$ does not always use all the simulation frames allocated due to its
+agressive pruning. This can be observed in games such as \Breakout, \CrazyClimber, \Kangaroo, and \Pooyan, where the average
+CPU time for each lookahead is up to $10$ times faster than \BFS.  Computation time for \UCT\ and \BRFS\ are  similar to \BFS, as the
+most expensive part of the computation is the  generation of frames through the simulator,  and these three  algorithms 
 always use  the full budget.
 
-More interestingly,  \IW(1) outscores \UCT\  in $27$ of the \numgames\ games,
-while \BFS\ outscores \UCT\ in $21$.
+More interestingly,  \IW(1) outscores \UCT\  in $31$ of the \numgames\ games,
+while \BFS\ outscores \UCT\ in $26$.
 %\footnote{These numbers can actually
 %improve further as 6  games for \IW(1) and 4 games for \BFS\ are still running.}
-On the other hand, \UCT\ does better than $\IW(1)$ and \BFS\ in $15$ and $17$ games respectively.
+On the other hand, \UCT\ does better than $\IW(1)$ and \BFS\ in $19$ and $25$ games respectively.
 The relative performance between $\IW(1)$ and \BFS\ makes $\IW(1)$  the
-best of the two in $22$ games, and  \BFS\  in $12$. In terms of the number of games 
-where an algorithm is the overall winner, $\IW(1)$  domainates the others: 
-$\IW(1)$ is the overall best in $23$ games, \BFS\ in $11$ games, and \UCT\ in $18$ games. 
+best of the two in $34$ games, and  \BFS\  in $16$. In terms of the number of games 
+where an algorithm is the best, % $\IW(1)$  domainates the others: 
+$\IW(1)$ is  the best in $26$ games, \BFS\ in $13$ games, and \UCT\ in $19$ games. 
 Also, \BRFS\ is best in $2$ games (\Centipede, tied up in \Boxing), while the other three algorithms
-are tied in another 2 games (\Pong, \Boxing), and $\IW(1)$ and \BFS\ are tied in \TimePilot.
+are tied in another 2 games (\Pong, \Boxing). 
+% , and $\IW(1)$ and \BFS\ are tied in \TimePilot.
 
@@
 The reason that \UCT\ does not collect any reward is that it needs to move the chicken up
-at least 240 times~\footnote{One needs to move the chicken up for at least $4$ seconds ($240$ frames) in order to get across the 
+at least 240 times\footnote{One needs to move the chicken up for at least $4$ seconds ($240$ frames) in order to get across the 
 highway.} something that is very unlikely in a random exploration. $\IW(1)$ does not have this limitation and is best in \Freeway.
@@
 \BFS\ does so in $6$ of the $7$ games. Comparing with the scores reported for the reinforcement learning algorithms by BNVB,
-we note that both $IW(1)$ and \BFS\ do much better than the best learning algorithm in
+we note that both $\IW(1)$ and \BFS\ do much better than the best learning algorithm in
 those games where the learning algorithms outperform \UCT\ like \Montezuma, \Venture\
@@
 no immediate penalty or consequence.  All algorithms achieve $0$ score, 
-except for \BFS\ that achieves an average score of $80$, and a score
-of $400$ in one of the runs. This means that even \BFS\  is not able to 
-consistently find rewards in this game. This game and several other games 
-like \Breakout, and \SpaceInvaders\ could be much simpler by adding
+except for \BFS\ that achieves an average score of $540$, and a score
+of $2500$ in one of the runs. This means however that even \BFS\  is not able to 
+consistently find rewards in this game. This game and several others
+like \Breakout\ and \SpaceInvaders\ could be much simpler by adding
 negative rewards for losing a life. We have indeed observed that
@@
 
-Regarding the Atari games, we identify two problems in all the planning methods
+Concerning the Atari games, there are two problems in all the planning methods
 discussed. First, by using the RAM state of the Atari console one can consider
-that they are ``cheating'', as they use information that is not available to the human player. And second, they do not get better with experience. In the future, we would like to explore 
+that they are ``cheating'', as they use information that is not available to the human player. And second, they do not get 
+better with experience. In the future, we would like to explore 
 whether variations of these algorithms  can be used to play from the information conveyed by the  
- state of screen pixels as opposed to the RAM state, and the potential uses of the new ideas for learning to improve performance.
+state of the screen pixels as opposed to the RAM state, and the potential uses of the new ideas for learning,
+such the use of the  novelty measures for shaping rewards.
+
+\medskip
+
+\noindent \textbf{Acknowledgments.} We thank the people involved in the creation and maintenance of the 
+Arcade Learning Environment  that have made this work possible.
 

### 2014-12-01 icaps15-paper/paper.tex
added ack
Fixed numbers

@@
 of nodes generated by $\IW(2)$ in the worst case is  $(n \times D)^2 \times b$, which is
-equal to $1,073,741,824$ which is too large, forcing us to consider only a tiny fraction of
+equal to $19,327,352,832$ which is too large, forcing us to consider only a tiny fraction of
 such states. For classical planning problems, the growth in the number of nodes from
@@
 $(n' \times D')^2 \times b$ states, which for $n'=1024$ and $D=2$
-represent  $36,864$ and $7,5497,472$ states respectively.
+represent  $36,864$ and $75,497,472$ states respectively.
 These are feasible numbers, but we haven't used  this representation 
@@
 
-\input{games-table}
+\begin{table*}
+\centering
+% \addtolength{\tabcolsep}{-2pt}
+% {\scriptsize
+{\small
+\begin{tabular}{@{}l|rr|rr|r|r@{}}
+%% table header
+% \hline\\[-3ex]
+%\hline
+\multicolumn{1}{c|}{}& \multicolumn{2}{c|}{$\IW(1)$}& \multicolumn{2}{c|}{\BFS} & \multicolumn{1}{c|}{\BRFS} &  \multicolumn{1}{c}{\UCT}
+% \\[.2ex]
+\\
+% \hline\\\\[-3ex]
+\hline
+%% table body
+\multicolumn{1}{c|}{Game}  &
+% \multicolumn{1}{c|}{Avg. Score} &
+% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
+% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
+% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c|}{Avg. Time}  &
+% \multicolumn{1}{c}{Avg. Score} & \multicolumn{1}{c}{Avg. Time}\\[.2ex]
+ \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
+ \multicolumn{1}{c}{Score} & \multicolumn{1}{c|}{Time}  &
+ \multicolumn{1}{c|}{Score} &
+ \multicolumn{1}{c}{Score} 
+% \\[.2ex]
+\\[.2ex]
+% \hline\\
+\hline
+\Alien & \mc{1}{\textbf{25634}} & 81 & \mc{1}{12252} & 81 & 784 & 7785 \\ \hline
+\Amidar & \mc{1}{\textbf{1377}} & 28 &\mc{1}{ 1090} & 37 & 5 & 180 \\ \hline
+\Assault & 953 & 18 & 827 & 25 & 414 & \textbf{1512} \\ \hline
+\Asterix & 153400 & 24 & 77200 & 27 & 2136 & \textbf{290700} \\ \hline
+\Asteroids & \mc{1}{\textbf{51338}} & 66 & \mc{1}{22168} & 65 & 3127 & 4661 \\ \hline
+\Atlantis & 159420 & 13 & 154180 & 71 & 30460 & \textbf{193858} \\ \hline
+\BankHeist & \mc{1}{\textbf{717}} & 39 & 362 & 64 & 22 & 498 \\ \hline
+\BattleZone & 11600 & 86 &\mc{1}{ \textbf{330800}} & 87 & 6313 & 70333 \\ \hline
+\BeamRider &\mc{1}{ 9108} & 23 & \mc{1}{\textbf{9298}} & 29 & 694 & 6625 \\ \hline
+\Berzerk & \mc{1}{\textbf{2096}} & 58 & \mc{1}{802} & 73 & 195 & 554 \\ \hline
+\Bowling & \mc{1}{\textbf{69}} & 10 & \mc{1}{50} & 60 & 26 & 25 \\ \hline
+\Boxing & \mc{1}{\textbf{100}} & 15 &\mc{1}{ \textbf{100}} & 22 & \textbf{100} & \textbf{100} \\ \hline
+\Breakout &\mc{1}{ 384} & 4 &\mc{1}{ \textbf{772}} & 39 & 1 & 364 \\ \hline
+\Carnival & \mc{1}{\textbf{6372}} & 16 &\mc{1}{ 5516} & 53 & 950 & 5132 \\ \hline
+\Centipede & 99207 & 39 & 94236 & 67 & \textbf{125123} & 110422 \\ \hline
+\ChopperCommand & 10980 & 76 & 27220 & 73 & 1827 & \textbf{34019} \\ \hline
+\CrazyClimber & 36160 & 4 & 36940 & 58 & 37110 & \textbf{98172} \\ \hline
+\DemonAttack & 20116 & 33 & 16025 & 41 & 443 & \textbf{28159} \\ \hline
+\DoubleDunk & -14 & 41 & 21 & 41 & -19 & \textbf{24} \\ \hline
+\ElevatorAction & 13480 & 26 & 10820 & 27 & 730 & \textbf{18100} \\ \hline
+\Enduro &\mc{1}{ \textbf{500}} & 66 & \mc{1}{359} & 38 & 1 & 286 \\ \hline
+\FishingDerby & 30 & 39 & 6 & 62 & -92 & \textbf{38} \\ \hline
+\Freeway &  \mc{1}{\textbf{31}} & 32 & \mc{1}{23} & 61 & 0 & 0 \\ \hline
+\Frostbite & \mc{1}{902} & 12 &  \mc{1}{\textbf{2672}} & 38 & 137 & 271 \\ \hline
+\Gopher & 18256 & 19 & 15808 & 53 & 1019 & \textbf{20560} \\ \hline
+\Gravitar &\mc{1}{ 3920} & 62 &  \mc{1}{\textbf{5980}} & 62 & 395 & 2850 \\ \hline
+\Hero &  \mc{1}{\textbf{12985}} & 37 & 11524 & 69 & 1324 & 12860 \\ \hline
+\IceHockey &  \mc{1}{\textbf{55}} & 89 &\mc{1}{ 49} & 89 & -9 & 39 \\ \hline
+\JamesBond &  \mc{1}{\textbf{23070}} & 0 & \mc{1}{10080} & 30 & 25 & 330 \\ \hline
+\JourneyEscape & \mc{1}{40080} & 38 &  \mc{1}{\textbf{40600}} & 67 & 1327 & 7683 \\ \hline
+\Kangaroo &  \mc{1}{\textbf{8760}} & 8 &\mc{1}{ 5320} & 31 & 90 & 1990 \\ \hline
+\Krull &  \mc{1}{\textbf{6030}} & 28 & 4884 & 42 & 3089 & 5037 \\ \hline
+\KungFuMaster &  \mc{1}{\textbf{63780}} & 21 & 42180 & 43 & 12127 & 48855 \\ \hline
+\Montezuma & 0 & 14 &  \mc{1}{\textbf{540}} & 39 & 0 & 0 \\ \hline
+\MsPacman & 21695 & 21 & 18927 & 23 & 1709 & \textbf{22336} \\ \hline
+\NameThisGame & 9354 & 14 & 8304 & 25 & 5699 & \textbf{15410} \\ \hline
+\Pong & \mc{1}{\textbf{21}} & 17 &\mc{1}{ \textbf{21}} & 35 & -21 & \textbf{21} \\ \hline
+\Pooyan & 11225 & 8 & 10760 & 16 & 910 & \textbf{17763} \\ \hline
+\PrivateEye & -99 & 18 & \mc{1}{\textbf{2544}} & 44 & 58 & 100 \\ \hline
+\Qbert & 3705 & 11 & 11680 & 35 & 133 & \textbf{17343} \\ \hline
+\RiverRaid & \mc{1}{\textbf{5694}} & 18 &\mc{1}{ 5062} & 37 & 2179 & 4449 \\ \hline
+\RoadRunner & \mc{1}{\textbf{94940}} & 25 &  \mc{1}{68500} &  41  & 245 & 38725 \\ \hline
+\RobotTank &\mc{1}{ \textbf{68}} & 34 &\mc{1}{ 52} & 34 & 2 & 50 \\ \hline
+\Seaquest &\mc{1}{ \textbf{14272}} & 25 &\mc{1}{ 6138} & 33 & 288 & 5132 \\ \hline
+\SpaceInvaders &\mc{1}{ 2877} & 21 & \mc{1}{\textbf{3974}} & 34 & 112 & 2718 \\ \hline
+\StarGunner &\mc{1}{ 1540} & 19 & \mc{1}{\textbf{4660}} & 18 & 1345 & 1207 \\ \hline
+\Tennis &\mc{1}{ \textbf{24}} & 21 & \mc{1}{\textbf{24}} & 36 & -24 & 3 \\ \hline
+\TimePilot & 35000 & 9 & 36180 & 29 & 4064 & \textbf{63855} \\ \hline
+\Tutankham & 172 & 15 & 204 & 34 & 64 & \textbf{226} \\ \hline
+\UpNDown &\mc{1}{ \textbf{110036}} & 12 & 54820 & 14 & 746 & 74474 \\ \hline
+\Venture & \mc{1}{\textbf{1200}} & 22 & \mc{1}{980} & 35 & 0 & 0 \\ \hline
+\VideoPinball & \mc{1}{\textbf{388712}} & 43 & 62075 & 43 & 55567 & 254748 \\ \hline
+\WizardOfWor & \mc{1}{\textbf{121060}} & 25 & 81500 & 27 & 3309 & 105500 \\ \hline
+\Zaxxon & \mc{1}{\textbf{29240}} & 34 & 15680 & 31 & 0 & 22610 \\ \hline
+
+\multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\ \hline
+\# Times Best (\numgames\ games)& \multicolumn{2}{c|}{\textbf{26}} & \multicolumn{2}{c|}{13} & \multicolumn{1}{c|}{1} & \multicolumn{1}{c|}{19} \\ \hline
+\# Times Better than \IW\ (\numgames\ games) & \multicolumn{2}{c|}{--} & \multicolumn{2}{c|}{16} & \multicolumn{1}{c|}{1}  & \multicolumn{1}{c|}{\textbf{19}} \\ \hline
+\# Times Better than \BFS\ (\numgames\ games) & \multicolumn{2}{c|}{\textbf{34}} & \multicolumn{2}{c|}{--} & \multicolumn{1}{c|}{1}  & \multicolumn{1}{c|}{25} \\ \hline
+\# Times Better than \UCT\ (\numgames\ games) & \multicolumn{2}{c|}{\textbf{31}} & \multicolumn{2}{c|}{26} & \multicolumn{1}{c|}{1}  & \multicolumn{1}{c|}{--} \\ \hline
+
+\end{tabular}
+}
+
+ \caption{\small Performance that results form various  lookahead algorithms in \numgames\  Atari games. The algorithms, \BRFS, $\IW(1)$, \BFS, and \UCT, are evaluated
+ over $10$ runs (episodes) for each game. 
+%%% Averages scores  over 5 runs shown with ``random'' seeds $0$ to $4$,  
+%%% this does not apply to UCT and BRFS if taken from JAIR ..
+%% 
+The maximum episode duration is  $18,000$  frames and every algorithm is limited to a lookahead budget of 150,000 simulated frames.
+% , except for UCT limited to 500 rollouts, 300 frames deep instead. 
+Figures for \BRFS\ and \UCT\ taken from \cite{bellemare:jair2013}.
+Average  CPU times per action in seconds, rounded to nearest integer, shown for  $\IW(1)$ and \BFS. 
+Numbers  in bold show best performer in terms of average score, while numbers shaded in light grey show scores  that are better than \UCT's. 
+Bottom part of the table shows pairwise comparisons among the algorithms. 
+% Entries shown as ``--'' are still running at the time of submission. Each run can take up to 24 hours of CPU time. Eventually
+% all 55 games from  \cite{bellemare:jair2013}  to be included.
+}
+\label{performance}
+
+\end{table*}
+
 We tested \IW(1) and \BFS\ over \numgames\  of the $55$ games 
@@
 
-\noindent \textbf{Acknowledgments.} We thank the people involved in the creation and maintenance of the 
-Arcade Learning Environment  that have made this work possible.
+\noindent \textbf{Acknowledgments.} We thank the people involved in
+the creation and maintenance of the Arcade Learning Environment that
+have made this work possible. This research was co-funded by the
+Australian Research Council linkage grant LP11010015.
 

### 2014-12-01 icaps15-paper/paper.tex
Final AAAI Submission

@@
 This is why the average time per action is similar to all the algorithms
-except $\IW(1)$, that due to its aggressive pruning does not always use
+except $\IW(1)$, that due to its pruning does not always use
 the full budget and takes less time per action on average.
@@
 
-Last, \IW(1) and \BFS\  in the lookahead select an action every $5$ frames, while \UCT\  selects 
+Last, in the lookahead, \IW(1) and \BFS\ select an action every $5$ frames, while \UCT\  selects 
 an action in every frame.  This means that in order to explore a branch 
@@
 The reason that \UCT\ does not collect any reward is that it needs to move the chicken up
-at least 240 times\footnote{One needs to move the chicken up for at least $4$ seconds ($240$ frames) in order to get across the 
+at least 240 times\footnote{One needs to move the chicken up for at least $4$ seconds ($240$ frames) in order to get it across the 
 highway.} something that is very unlikely in a random exploration. $\IW(1)$ does not have this limitation and is best in \Freeway.
@@
 the creation and maintenance of the Arcade Learning Environment that
-have made this work possible. This research was co-funded by the
-Australian Research Council linkage grant LP11010015.
+have made this work possible. The research by Nir Lipovetzky is partially
+funded by the Australian Research Council Linkage grant LP11010015.
 

### 2015-02-02 src/agents/UCTNoveltySearchTree.cpp
New changes but algorithm tends to keep in the middle of screen, where all novelty is closer

@@ -17,8 +17,6 @@
 #include "random_tools.h"
 #include "misc_tools.h"
 
-#define UNDEFINED_DEPTH std::numeric_limits<int>::max()
-
 UCTNoveltySearchTree::UCTNoveltySearchTree(RomSettings * rom_settings, Settings &settings,
 			     ActionVect &actions, StellaEnvironment* _env):
 	SearchTree(rom_settings, settings, actions, _env) {
@@ -29,15 +27,16 @@ UCTNoveltySearchTree::UCTNoveltySearchTree(RomSettings * rom_settings, Settings 
 		settings.getFloat("uct_exploration_constant", true);
 
 	uct_novelty_pruning = 
-		settings.getFloat("uct_novelty_pruning", true);
+		settings.getBool("uct_novelty_pruning", true);
 	
 	novelty_depth.resize( RAM_SIZE * 256, UNDEFINED_DEPTH );
-	novelty_reward.resize( uct_search_depth+1 );
-	for(int i = 0; i < uct_search_depth+1; i++)
+
+	novelty_reward.resize( uct_search_depth+2 );
+	for(int i = 0; i < uct_search_depth+2; i++)
 		novelty_reward[i].resize( RAM_SIZE * 256, 0);
 	exploration_weight = 1;
 	reward_weight = 1;
-	novelty_factor = 2;
+	novelty_factor = 1.001;
 }
 
 /* *********************************************************************
@@ -131,7 +130,7 @@ Action UCTNoveltySearchTree::get_best_action(void) {
 	//  for all children of p_root
 	int best_branch;
   
-	// best_branch = get_best_branch((UCTNoveltyTreeNode*)p_root, false);
+	//best_branch = get_best_branch((UCTNoveltyTreeNode*)p_root, false);
 	// Select based on most explored action
 	best_branch = get_most_visited_branch((UCTNoveltyTreeNode*)p_root);
   
@@ -178,8 +177,12 @@ int UCTNoveltySearchTree::single_uct_iteration(void) {
 				}
 			}
 
-			if(all_dups)
-				std::cout << "Node depth:" << node->m_depth << std::endl;
+			// if(all_dups){
+			// 	//std::cout << "Node depth:" << node->m_depth << std::endl;
+			// 	update_values((UCTNoveltyTreeNode*)node, -1);
+			// 	node = p_root;
+			// 	continue;
+			// }
 			assert(!all_dups);
 
[…]

### 2015-02-02 src/agents/UCTNoveltySearchTree.hpp

@@ -90,7 +90,7 @@ class UCTNoveltySearchTree : public SearchTree {
 
 	reward_t get_exploration_bonus( const ALERAM& machine_state, unsigned depth );
 
-	bool is_novel(ALEState &state, Action a, unsigned depth);
+	bool is_local_novel(ALEState &state, Action a, unsigned depth, 	std::vector<int>& local_novelty_depth);
 
 	bool is_novel(UCTNoveltyTreeNode* node);
 

### 2015-02-02 src/agents/UCTNoveltyTreeNode.cpp

@@ -14,6 +14,7 @@
  */
 #include "UCTNoveltyTreeNode.hpp"
 
+
 UCTNoveltyTreeNode::UCTNoveltyTreeNode(TreeNode *parent, ALEState &parentState): 
   TreeNode(parent, parentState),
   visit_count(0),
@@ -29,3 +30,4 @@ UCTNoveltyTreeNode::UCTNoveltyTreeNode(TreeNode *parent, ALEState &parentState,
   sum_returns(0),
   sim_steps(0) {
 }
+

### 2015-02-02 src/agents/UCTNoveltyTreeNode.hpp

@@ -18,28 +18,37 @@
 
 #include "TreeNode.hpp"
 
-class UCTNoveltyTreeNode: public TreeNode  {
-  public:
-    UCTNoveltyTreeNode(TreeNode *parent, ALEState &parentState); 
-    
-    UCTNoveltyTreeNode(TreeNode *parent, ALEState &parentState, 
-      int num_simulate_steps, Action a, SearchTree *tree);	
+#define UNDEFINED_DEPTH std::numeric_limits<int>::max()
 
-    int num_steps() {
-	int num_steps = 0;
-	
-	for (size_t a = 0; a < v_children.size(); a++) {
-	    if (v_children[a]->is_initialized())
-		num_steps += ((UCTNoveltyTreeNode*)v_children[a])->num_steps();
+class UCTNoveltyTreeNode: public TreeNode  {
+public:
+	UCTNoveltyTreeNode(TreeNode *parent, ALEState &parentState); 
+    
+	UCTNoveltyTreeNode(TreeNode *parent, ALEState &parentState, 
+			   int num_simulate_steps, Action a, SearchTree *tree);	
+
+	void init_novelty()
+	{
+		local_novelty_depth.resize( RAM_SIZE * 256, UNDEFINED_DEPTH );
 	}
 	
-	return num_steps + sim_steps;
-    }
+	int num_steps() {
+		int num_steps = 0;
+	
+		for (size_t a = 0; a < v_children.size(); a++) {
+			if (v_children[a]->is_initialized())
+				num_steps += ((UCTNoveltyTreeNode*)v_children[a])->num_steps();
+		}
+	
+		return num_steps + sim_steps;
+	}
     
-    /** Additional UCTNovelty-specific variables */
-    int visit_count; // The number of visits to this node
-    float sum_returns; // The sum of the returns received from this node
-    int sim_steps; // The Simulation of the rollout steps
+	/** Additional UCTNovelty-specific variables */
+	int visit_count; // The number of visits to this node
+	float sum_returns; // The sum of the returns received from this node
+	int sim_steps; // The Simulation of the rollout steps
+
+ 	std::vector<int>      local_novelty_depth;
 };
 
 #endif // __UCT_NOVELTY_TREE_NODE_HPP__

### 2015-02-04 src/agents/UCTNoveltySearchTree.cpp
changing initialization of shaping rewards

@@ -34,9 +34,9 @@ UCTNoveltySearchTree::UCTNoveltySearchTree(RomSettings * rom_settings, Settings 
 	novelty_reward.resize( uct_search_depth+2 );
 	for(int i = 0; i < uct_search_depth+2; i++)
 		novelty_reward[i].resize( RAM_SIZE * 256, 0);
-	exploration_weight = 1;
-	reward_weight = 1;
-	novelty_factor = 1.001;
+	exploration_weight = 1;//.0001;
+	reward_weight = 100000;
+	novelty_factor = 2.0;
 }
 
 /* *********************************************************************
@@ -71,11 +71,20 @@ void UCTNoveltySearchTree::update_tree(void) {
 	for(unsigned i = 0; i < RAM_SIZE * 256; i++)
 		novelty_depth[i] = UNDEFINED_DEPTH;
 
-	if(uct_novelty_pruning)
-		for(int i = 0; i < uct_search_depth+1; i++)
-			for(unsigned j = 0; j < RAM_SIZE * 256; j++)
-				novelty_reward[i][j] = 0;
+	for(unsigned j = 0; j < RAM_SIZE * 256; j++)
+		novelty_reward[ uct_search_depth + 1][j] = 1;		
+
+	for(int i = uct_search_depth; i >= 0; i--){				
+		for(unsigned j = 0; j < RAM_SIZE * 256; j++){								
+			novelty_reward[i][j] = novelty_reward[ i + 1 ][ j ] * novelty_factor;
+		}
+		
+	}
 	
+	
+	
+	get_exploration_bonus( m_env->getRAM(), 0 );
+
 	std::cout << "starting with " << simulation_steps << " simulation steps" <<std::endl;
 	
 	int num_iterations = 0;
@@ -217,21 +226,35 @@ int UCTNoveltySearchTree::single_uct_iteration(void) {
 		assert(!node->is_initialized());
 		
 		
-		node->m_depth = node->p_parent->m_depth + 1;
+		node->m_depth = node->p_parent->m_depth + sim_steps_per_node;
 		if (node->depth() > m_max_depth ) m_max_depth = node->depth();		
 
-		node->init(this, leaf_choice, sim_steps_per_node); 
+		//node->init(this, leaf_choice, sim_steps_per_node); 
+
+		return_t step_return;
+		int steps = simulate_game( (UCTNoveltyTreeNode*) node, leaf_choice, sim_steps_per_node, 
+				  step_return, node->is_terminal, true, true);
[…]

### 2015-02-04 src/agents/UCTNoveltyTreeNode.cpp

@@ -19,6 +19,7 @@ UCTNoveltyTreeNode::UCTNoveltyTreeNode(TreeNode *parent, ALEState &parentState):
   TreeNode(parent, parentState),
   visit_count(0),
   sum_returns(0),
+  return_achieved(0),
   sim_steps(0)
 {
 }
@@ -28,6 +29,7 @@ UCTNoveltyTreeNode::UCTNoveltyTreeNode(TreeNode *parent, ALEState &parentState,
   TreeNode(parent, parentState, tree, a, num_simulate_steps),
   visit_count(0),
   sum_returns(0),
+  return_achieved(0),
   sim_steps(0) {
 }
 

### 2015-02-04 src/agents/UCTNoveltyTreeNode.hpp

@@ -46,6 +46,7 @@ public:
 	/** Additional UCTNovelty-specific variables */
 	int visit_count; // The number of visits to this node
 	float sum_returns; // The sum of the returns received from this node
+	return_t return_achieved; //return from root up to current node
 	int sim_steps; // The Simulation of the rollout steps
 
  	std::vector<int>      local_novelty_depth;

### 2015-02-04 ijcai15-paper/paper.tex
First iteration IJCAI-15

--- 2014-12-01 icaps15-paper/paper.tex
@@
- \documentclass[letterpaper]{article}
-\usepackage{aaai}
+% \documentclass[letterpaper]{article}
+
+\documentclass{article}
+% The file ijcai15.sty is the style file for IJCAI-15 (same as ijcai07.sty).
+\usepackage{ijcai15}
+
+% Use the postscript times font!
 \usepackage{times}
+
+% \usepackage{aaai}
+% \usepackage{times}
+
 \usepackage{helvet}
@@
 \frenchspacing
-\setlength{\pdfpagewidth}{8.5in}
-\setlength{\pdfpageheight}{11in}
+% \setlength{\pdfpagewidth}{8.5in}
+% \setlength{\pdfpageheight}{11in}
 
@@
 %
-\pdfinfo{
-/Title (Classical Planning Algorithms on the Atari Video Games)
-/Author (Nir Lipovetzky, Miquel Ramirez, Hector Geffner)}
+% \pdfinfo{
+% /Title (Classical Planning Algorithms on the Atari Video Games)
+% /Author (Nir Lipovetzky, Miquel Ramirez, Hector Geffner)}
 
@@
 
-\title{Classical Planning Algorithms ¡ on the Atari Video Games}
+\title{Classical Planning with Simulations: \\ Results on the Atari Video Games}
 
-% \author{\textbf{ID:} 52 \ \ ; \ \ \textbf{Keywords:} classical planning, video games}
+\author{\textbf{ID:} \ldots  \ \ ; \ \ \textbf{Keywords:} classical planning, search, video games}
 
- \author{Nir Lipovetzky \\
-       The University of Melbourne \\
-    Melbourne, Australia\\
+% \author{Nir Lipovetzky \\
+%        The University of Melbourne \\
+%     Melbourne, Australia\\
 %     {\normalsize\url{nir.lipovetzky@unimelb.edu.au}}
@@
 %% 
-    {\small\url{nirlipo@gmail.com}}
-    \And
-    Miquel Ramirez \\
-    Australian National University \\
-    Canberra, Australia\\
-   {\small\url{miquel.ramirez@gmail.com}}
-    \And
-    Hector Geffner \\
-    ICREA \&  Universitat Pompeu Fabra \\
-    Barcelona, SPAIN \\
-    {\small\url{hector.geffner@upf.edu}}
+%    {\small\url{nirlipo@gmail.com}}
+%    \And
+%    Miquel Ramirez \\
+%    Australian National University \\
+%    Canberra, Australia\\
+%   {\small\url{miquel.ramirez@gmail.com}}
+%    \And
+%    Hector Geffner \\
+%    ICREA \&  Universitat Pompeu Fabra \\
+%    Barcelona, SPAIN \\
+%    {\small\url{hector.geffner@upf.edu}}
  %%%%% \thanks{firstname.lastname}
-}
-
-%AAAI Press\\
-%Association for the Advancement of Artificial Intelligence\\
-%2275 East Bayshore Road, Suite 160\\
-% Palo Alto, California 94303\\}
+%}
 
@@
 
-%% Pending: randomization, discount factor, iw11 ..
-%% iw2 multivalued expands up to 256^2 n^nodes, while iw2 boolean 16^2 n^2 nodes; can iw2 boolean beat iw1 multivalued?
-%% Important. optimization, need Dikjstra por postprocessing .. see emails
-
 \begin{abstract}
-\begin{quote}
 The Atari 2600 games supported in  the Arcade Learning Environment~\cite{bellemare:jair2013}
 all feature a known initial (RAM) state and actions that have deterministic effects. Classical planners, however,
-cannot be used  for selecting actions  for  two reasons: first,  no  compact PDDL-model of  the games is given,
-and  more importantly,  the action effects and goals are  not known \emph{a priori}. Moreover, in these games there is usually
-no set of goals to be achieved but rewards to be collected. These features do not preclude the use of   classical  algorithms like
-breadth-first search or Dijkstra's algorithm,  but    these methods are not effective over  large state spaces. We thus turn to a
-different class of classical planning algorithms  introduced recently  that perform a \emph{structured exploration} of the state space; namely,
-like breadth-first search and Dijkstra's  algorithm they are ``blind'' and hence  do not require prior knowledge of state transitions,
-costs (rewards) or goals, and yet, like  heuristic search algorithms, they have been shown to be   effective for solving   problems over huge
-state spaces. The simplest such algorithm, called Iterated Width or \IW, consists of  a sequence of calls  $\IW(1)$, $\IW(2)$, \ldots,
-$\IW(k)$ where  $\IW(i)$ is  a breadth-first search in which a state is  pruned when it is not the
-first state in the search  to make true some subset of $i$ atoms. 
-The  empirical results over \numgames\ games suggest that the performance 
-of \IW\ with the  $k$ parameter fixed to $1$, i.e., $\IW(1)$, is at the level of the state of the art represented by 
-\UCT.  A simple best-first variation  of \IW\  that combines exploration and exploitation proves
-to be very competitive as well. 
-\end{quote}
+cannot be used  off-the-shelf as there is no  compact PDDL-model of  the games, and action effects and goals are  not
+known \emph{a priori}. Indeed, there are no explicit goals, and the planner must select actions on-line while interacting 
+with a simulator that returns successor states and rewards. None of this precludes  the use of  blind lookahead algorithms 
+for action selection like breadth-first search or  Dijkstra's yet such methods are not effective over large  state spaces. 
+We thus turn to a different class of  planning methods  introduced recently  that  have been shown to be  effective for solving
+large planning problems but which do not require  prior knowledge of state transitions and costs (rewards) or goals.
+The simplest such algorithm, called Iterated Width or \IW, consists of  a sequence of calls  $\IW(1)$, $\IW(2)$, \ldots,
+where  $\IW(k)$ is  a breadth-first search in which a state is  pruned when it is not the first state in the  search  to make true 
+a subset of $k$ atoms. The  empirical results over \numgames\ Atari games show that  $\IW(1)$ performs at the level of UCT, 
+the state-of-the-art planning method in this domain, and suggest the potential of width-based methods for planning with
+simulators when factored, compact action models are not available.
 \end{abstract}
@@
 have deterministic effects, the problem of selecting the next action
-to be done cannot be addressed with state-of-the-art classical planners \cite{geffner:book}.
-This is because there is  no compact PDDL-like encoding of the domain, and more
-importantly,  the goal to be achieved in each game  is not given. Indeed, there are
-often no goals  but rewards $r(a,s)$ that  depend on the action $a$ and
-the state $s$ in  which  the action is performed, and these rewards are not known.\footnote{Actually, in the Atari games, the rewards $r(a,s)$ depend only
-on the state $f(a,s)$ that results from doing action $a$ in the state $s$.}
-Thus no model of the  goals  or the  rewards can be used to bias the search.
+to be done cannot be addressed with off-the-shelf classical planners 
+\cite{planning:book,geffner:book}. This is because there is  no compact PDDL-like encoding of the domain and 
+the goal to be achieved in each game  is not given. Indeed, there are no goals but rewards, and 
+the planner must select actions on-line while interacting with a simulator that just
+returns successor states and rewards.
+%\footnote{Actually, in the Atari games, the rewards $r(a,s)$ depend only
+% on the state $f(a,s)$ that results from doing action $a$ in the state $s$.}
 
@@
 true when the value of the $i$-th byte in the state vector $s$ is $x_j$.
+The  empirical results over \numgames\ Atari games show that  $\IW(1)$ performs at the level of UCT, 
+the state-of-the-art planning method in this domain, and suggest the potential of width-based methods for planning with
+simulators when factored, compact action models are not available.
 
+\Omit{
 While a normal, complete breadth-first search runs in time that is exponential in the total number of variables $X_i$,
@@
 additive reward measure  to be optimized, like in the Atari games, is direct.
+}
 
@@
 
-The paper is organized as follows. We review the iterated width  algorithm and where it comes from,
-look at  simple   variations of the algorithm that appear to be convenient  for the Atari
-games, present and analyze the  experimental results, and discuss related and future work.
+The paper is organized as follows. We review the iterated width  algorithm and its properties,
+look at the    variations of the algorithm that we used in the Atari games, present and analyze the  
+experimental results, and discuss related and future work.
 
@@
 
-For classical planning, the termination condition is the achievement of the goal. In the on-line setting,
+For classical planning, the termination condition is the achievement of the goal. In the \emph{on-line setting,}
 as in the Atari games, the termination condition is given by a time window or a maximum number of generated nodes.
-The  best path found by \IW \ is the path  that has a  maximum accumulated reward.
-The accumulated reward $R(s)$  of a state $s$ reached in an iteration of \IW \ is determined by the  unique parent state $s'$
+The  best path found by \IW \ is then the  path  that has a  maximum accumulated reward. The accumulated reward $R(s)$  
+of a state $s$ reached in an iteration of \IW \ is determined by the  unique parent state $s'$
 and action $a$ leading to $s$ from $s'$ as $R(s) = R(s') + r(a,s')$. The best state is the state $s$ with maximum reward
 $R(s)$ generated but not pruned by \IW, and the best path is the one that leads to the state $s$ from the current state.
-The action selected to be done next is  the first action along  such a path.
+The action selected  in the on-line setting  is  the first action along  such a path. This action is then executed
+and the process repeats from the resulting state.
 
@@
 %%% the  duplicate states as well.
-
-\begin{table}[t!]
-\begin{center}
-%\resizebox{3in}{!}{
-{\scriptsize
-% \begin{tabular}{r@{}lrrrr@{}}
-\begin{tabular}{llcccc}
-%% table header
-\hline\\[-1.3ex]
-%\toprule
-\# & \multicolumn{1}{c}{Domain} & \multicolumn{1}{c}{I} & \multicolumn{1}{c}{$\IW(1)$} & \multicolumn{1}{c}{$\IW(2)$} & \multicolumn{1}{c}{Neither} \\[.2ex]
-
-\hline\\[-1.3ex]
-%\hline
-%% table body
-1. & 8puzzle	&400	&55\%	&45\%	&0\%\\
-2. & Barman	&232	&9\%	&0\%	&91\%\\
-3. & Blocks World	&598	&26\%	&74\%	&0\%\\
-4. &Cybersecure	&86	&65\%	&0\%	&35\%\\
-% 5. &Depots	&189	&11\%	&66\%	&23\%\\
-% 6. & Driver	&259	&45\%	&55\%	&0\%\\
-% 7. & Elevators	&510	&0\%	&100\%	&0\%\\
-% 8. & Ferry	&650	&36\%	&64\%	&0\%\\
-% 9. & Floortile	&538	&96\%	&4\%	&0\%\\
-% 10. & Freecell	&76	&8\%	&92\%	&0\%\\
-% 11. & Grid	&19	&5\%	&84\%	&11\%\\
-% 12. & Gripper	&1275	&0\%	&100\%	&0\%\\
-% 13. & Logistics	&249	&18\%	&82\%	&0\%\\
-% 14. & Miconic	&650	&0\%	&100\%	&0\%\\
-% 15. & Mprime	&43	&5\%	&95\%	&0\%\\
-% 16. & Mystery	&30	&7\%	&93\%	&0\%\\
-% 17. & NoMystery	&210	&0\%	&100\%	&0\%\\
-% 18. & OpenStacks	&630	&0\%	&0\%	&100\%\\
-% 19. & OpenStacksIPC6	&1230	&5\%	&16\%	&79\%\\
-\ldots & \ldots & \ldots & \ldots & \ldots & \\
-20. & ParcPrinter	&975	&85\%	&15\%	&0\%\\
-21. & Parking	&540	&77\%	&23\%	&0\%\\
-22. & Pegsol	&964	&92\%	&8\%	&0\%\\
-23. & Pipes-NonTan	&259	&44\%	&56\%	&0\%\\
-24. & Pipes-Tan	&369	&59\%	&37\%	&3\%\\
-25. & PSRsmall	&316	&92\%	&0\%	&8\%\\
-26. & Rovers	&488	&47\%	&53\%	&0\%\\
-27. & Satellite	&308	&11\%	&89\%	&0\%\\
-28. & Scanalyzer	&624	&100\%	&0\%	&0\%\\
-29. & Sokoban	&153	&37\%	&36\%	&27\%\\
-30. & Storage	&240	&100\%	&0\%	&0\%\\
-31. & Tidybot	&84	&12\%	&39\%	&49\%\\
-32. & Tpp	&315	&0\%	&92\%	&8\%\\
-33. & Transport	&330	&0\%	&100\%	&0\%\\
-34. & Trucks	&345	&0\%	&100\%	&0\%\\
-35. & Visitall	&21859	&100\%	&0\%	&0\%\\
-36. & Woodworking	&1659	&100\%	&0\%	&0\%\\
-37. & Zeno	&219	&21\%	&79\%	&0\%\\[.2ex]
-%\hline
-\hline\\[-1.3ex]
-& Total/Avgs&37921	&37.0\%	&51.3\%	&11.7\%\\
-\hline
-\end{tabular}
-}
-
-\vskip .1cm
-
-{\scriptsize
-\begin{tabular}{@{}ccccc@{}}
-\hline\\[-1.3ex]
-\multicolumn{1}{c}{\# Instances}  &\multicolumn{1}{c}{\IW} & \multicolumn{1}{c}{\ID} & \multicolumn{1}{c}{\BRFS} &\multicolumn{1}{c}{\textit{GBFS +} $h_{add}$} \\[.2ex]
-\hline\\[-1.3ex]
-37921  &34627 &9010	&8762 &34849 \\
-\hline
-%\hline
-\end{tabular}
-}
-
-\end{center}
-
-\caption{\small
-\emph{Top:} Number of classical planning instances per domain  and percentages solved by $\IW(1)$, $\IW(2)$, or neither.
-Problems  obtained by splitting benchmarks  with $N$ atomic goals into $N$ problems with atomic goals.
-\emph{Bottom:} Number of instances solved by \IW\ in comparison with  Iterative Deepening (\ID),
-Breadth-First Search (\BRFS),   and Greedy Best First Search (\textit{GBFS}) guided by  additive heuristic.
-Time and  memory outs  after 30 minutes or 2 GB. Table from \cite{nir:ecai2012}.
-}
-\label{table1}
-\end{table}
-%%%%
-%%%
 
@@
 but   unlike these algorithms, it uses the factored representation of the states in terms of variables to structure the
-search in a different way. This structured exploration  has proved to be very
-effective over classical planning benchmark domains  when goals are  single atoms.\footnote{
+search. This structured exploration  has proved to be very effective over classical planning benchmark domains  when goals are  single atoms.\footnote{
 Any conjunctive goal can be mapped into a single dummy atomic goal by adding an action that
 achieves the dummy goal and that has the original conjunctive goal as a precondition.
-Yet, this changes the definition of the domain.}
-Table~\ref{table1} from \cite{nir:ecai2012} shows
-the percentage of instances that are  solved by  the first iteration of \IW, i.e.  $\IW(1)$, by the second iteration
-$\IW(2)$, and by neither one.  These are instances that  have been obtained from the existing benchmarks
-by splitting problems with $N$ atomic goals, into $N$ problems with one atomic goal each. As the table shows,
-$37\%$ of the 37921 instances are solved by $\IW(1)$ while $51.3\%$ are  solved by $\IW(2)$. Since $\IW(k)$
-runs in time that is exponential in $k$, this mean that almost $90\%$ of the 37,921 instances are  solved in time that
-is either linear or quadratic in the number of problem variables, which in these encoding are all \emph{boolean}.
-Furthermore, when the performance of \IW\   is compared with  breadth-first search and iterative deepening,
-on the one hand,  and  with a  Greedy Best First Search guided by the additive heuristic $h_{add}$
-\cite{bonet:aij-hsp} on the other (this algorithm is similar
-to the one used by  the FF planner when hill climbing search fails \cite{hoffmann:ff}),
-it turns out that ``blind'' \IW\ solves as many problems as the informed search, 34,627 vs. 34,849, far ahead
-of the other blind algorithms \BRFS\ and \ID\  that solve 9,010 and 8,762 problems each. This is shown in the bottom
-part of  Table~\ref{table1}. Moreover, \IW\ is  faster and results in shorter plans than a heuristic search algorithm
-\cite{nir:ecai2012}.
+Yet, this changes the definition of the domain.} For example, $37\%$ of the 37,921 problems
+considered  in \cite{nir:ecai2012} are solved by $\IW(1)$ while $51.3\%$ are  solved by $\IW(2)$.
+These are instances obtained from 37 benchmark domains by splitting problems with $N$ atomic
+goals into $N$ problems with one atomic goal each.  Since $\IW(k)$ runs in time that is exponential in $k$, 
+this mean that almost $90\%$ of the 37,921 instances are  solved in time that is either linear or quadratic in the number of problem variables, 
+which in such encodings are all \emph{boolean}. Furthermore, when the performance of \IW\   is compared with  
+a  Greedy Best First Search guided by the additive heuristic $h_{add}$, it turns out that ``blind'' \IW\ solves as many problems as the informed 
+search, 34,627 vs. 34,849, far ahead of  other blind search  algorithms like \BRFS\ and \ID\  that solve 9,010 and 8,762 problems each. 
+Moreover, \IW\ is  faster and results in shorter plans than in the heuristic search.
 
-The min $k$ value for which $\IW(k)$ solves a problem  is bounded and small in most of
+The min $k$ value for which $\IW(k)$ solves a problem  is indeed bounded and small in most of
 these instances. This is  actually no  accident and  has a \emph{theoretical explanation.} Lipovetzky and Geffner define
@@
 
+\Omit{
 One way to understand this definition is that the problem  width $w(P)$ is at most  $i$ if  there is a
@@
 C)~the ``trail''  preserves ``optimality''; i.e., no  tuple $t_k$ can be reached in less than $k$ steps.
+}
 
@@
 \section{Experimental Results}
+
+We tested \IW(1) and \BFS\ over \numgames\  of the $55$ different games 
+considered in \cite{bellemare:jair2013}, from now on  abbreviated as BNVB.\footnote{
+We left out  \textsc{Skiing} as the reported figures apparently use a  different reward structure.}
+%% check
+% \footnote{
+% Entries marked as  ``--'' in the table are   still running at the time of submission. 
+% Each run can take up to $24$ hours of CPU time and there are $10$ runs per game and algorithm. 
+% Eventually all of  the  $55$ games will  be included in the evaluation.}
+The two  algorithms were used to play the  games in the \emph{on-line planning
+setting} supported by ALE where we will compare them with the planning algorithms 
+considered by BNVB; namely, breadth-first search and \UCT. ALE supports also a \emph{learning setting} 
+where the goal is to learn controllers that map  states into actions without
+doing any lookahead. Algorithms across  the two settings are thus not directly comparable
+as they compute different things. Learning controllers appears as a more challenging problem
+and it is thus not surprising that planning algorithms like UCT tend to achieve a higher score than 
+learning algorithms. In addition, the learning algorithms reported by BNVB tend to use 
+the state of the screen pixels, while the planning algorithms,  use the state
+of the RAM memory. It is not clear however whether one input representation is
+more challenging than the other; for the learning algorithms, BNVB mention
+that the results tend to be better for the screen inputs.
+
+Table~\ref{performance} shows the performance of $\IW(1)$ and \BFS\
+in comparison  with breadth-first search (\BRFS) and \UCT.
+Videos of selected games played by $\IW(1)$, \BFS, and \UCT\ can be 
+seen in Youtube.\footnote{\url{http://www.youtube.com/playlist?list=PLXpQcXUQ_CwenUazUivhXyYvjuS6KQOI0}.}
+The discount factor used by  all the algorithms is $\gamma=0.995$.
+The scores reported for \BRFS\ and \UCT\ are taken from BNVB.  
+Our experimental setup  follows theirs except that a maximum budget of  $150,000$ simulated frames 
+is applied to $\IW(1)$, \BFS, and \UCT. \UCT\ uses this budget by running
+ $500$   rollouts of depth $300$. % \footnote{In \UCT,  $500$ base policy rollouts
+% of depth $300$ is equivalent to a budget of $150,000$ simulated frames.}.
+The bound on the number of simulated frames is like a bound on
+lookahead time, as most of the time in the lookahead  is spent
+in calls to the emulator for computing the next RAM state. 
+This is why the average time per action is similar to all the algorithms
+except $\IW(1)$, that due to its pruning does not always use
+the full budget and takes less time per action on average.
+
+Also,  as reported by  BNVB, all of the algorithms 
+reuse the frames in the sub-tree of the previous lookahead that is rooted in the selected 
+child, deleting its siblings and their descendants. More precisely, no calls 
+to the emulator are done for transitions that are cached in that sub-tree, 
+and such reused frames are not discounted from the budget that is thus a bound 
+on the number of \emph{new}  frames per lookahead. In addition, in $\IW(1)$, the states that 
+are reused  from the previous searches are ignored in the computation of the novelty of 
+new states so that more states can escape pruning. Otherwise, $\IW(1)$ often
+uses a fraction of the  budget.  This is not needed in \BFS\ which does 
+no pruning.  $\IW(1)$ and \BFS\ are  limited to  search up to a depth of
+$m=1,500$ frames and up to $m=150,000$ frames per root branch. This is
+to avoid the search from  going too deep or being too committed to a
+single root action. 
+% Similarly UCT defines a limit of $m=300$ frames for each rollout. 
+% Note that UCT is more likely to have a performance
+% loss if the depth is increased, as the number of rollouts would
+% decrease in order to maintain the same budget of $150,000$ simulated
+% frames.
+
+Last, in the lookahead, \IW(1) and \BFS\ select an action every $5$ frames, while \UCT\  selects 
+an action in every frame.  This means that in order to explore a branch 
+$300$ frames deep, UCT gets to choose $300$ actions, while \IW($1$) and \BFS\ 
+get to choose $60$ actions, both however using the same $300$ frames from the
+budget. For this, we  followed the setup  of \BRFS\ in BNVB 
+that also selects actions every $5$ frames, matching the behavior
+of the emulator that requests an action also every $5$ frames. Since the 
+lookahead budget is given by a maximum number of (new) frames, and the time 
+is mostly taken by calls to the emulator, this may not be the best choice  
+for $\IW(1)$ and \BFS\ that may therefore not be exploiting all the 
+options afforded by the budget. We'll look at this further in the future.
+
+\Omit{
+The last difference inherited from BNVB between all algorithms
+and \UCT\, is that \UCT\ selects an action randomly in the rollout every
+single frame, while the others select an action every 5 frames, and
+the remain ones are $no-op$, as is done in the simulation
+environment. Therefore, a node in all algorithms represent $5$ frames,
+but in UCT represents $1$ frame. As a result, a UCT rollout of depth
+$300$ frames involve the selection of 300 actions, decreasing the
+budget of simulated frames by $300$, while the other algorithms need a
+branch of $1500$ frames to select $300$ actions, decreasing the budget
+by $1500$ frames instead.
+}
+
+Table~\ref{performance} shows that both $\IW(1)$ and \BFS\ outperform \BRFS, which  rarely collects
+reward in  many domains as  the depth of the \BRFS\ search tree results in a lookahead of $0.3$
+seconds ($20$ frames or $4$ nodes deep). The notable exception to this
+is \Centipede\ where abundant reward can be collected with a shallow
+lookahead. On the other hand, both $\IW(1)$ and \BFS\ normally 
+reach states that are up to $350$--$1500$ frames deep ($70$--$260$ nodes or 
+$6$--$22$ seconds),  even if $\IW(1)$ does not always use all the simulation frames allocated due to its
+agressive pruning. This can be observed in games such as \Breakout, \CrazyClimber, \Kangaroo, and \Pooyan, where the average
+CPU time for each lookahead is up to $10$ times faster than \BFS.  Computation time for \UCT\ and \BRFS\ are  similar to \BFS, as the
+most expensive part of the computation is the  generation of frames through the simulator,  and these three  algorithms 
+always use  the full budget.
+
+More interestingly,  \IW(1) outscores \UCT\  in $31$ of the \numgames\ games,
+while \BFS\ outscores \UCT\ in $26$.
+%\footnote{These numbers can actually
+%improve further as 6  games for \IW(1) and 4 games for \BFS\ are still running.}
+On the other hand, \UCT\ does better than $\IW(1)$ and \BFS\ in $19$ and $25$ games respectively.
+The relative performance between $\IW(1)$ and \BFS\ makes $\IW(1)$  the
+best of the two in $34$ games, and  \BFS\  in $16$. In terms of the number of games 
+where an algorithm is the best, % $\IW(1)$  domainates the others: 
+$\IW(1)$ is  the best in $26$ games, \BFS\ in $13$ games, and \UCT\ in $19$ games. 
+Also, \BRFS\ is best in $2$ games (\Centipede, tied up in \Boxing), while the other three algorithms
+are tied in another 2 games (\Pong, \Boxing). 
+% , and $\IW(1)$ and \BFS\ are tied in \TimePilot.
+
+\Omit{As mentioned in the
+previous section, the order in which $\IW(1)$ and \BFS\ generate the
+children of an expanded node is randomized, something which has a
+potential effect on the novelty measures of nodes, and hence in the
+search graph. We found that this randomization helps, in particular in
+\Qbert, where not randomizing results in $\IW(1)$ being trapped into
+loops sometimes.}
 
@@
  over $10$ runs (episodes) for each game. 
-%%% Averages scores  over 5 runs shown with ``random'' seeds $0$ to $4$,  
-%%% this does not apply to UCT and BRFS if taken from JAIR ..
-%% 
 The maximum episode duration is  $18,000$  frames and every algorithm is limited to a lookahead budget of 150,000 simulated frames.
-% , except for UCT limited to 500 rollouts, 300 frames deep instead. 
-Figures for \BRFS\ and \UCT\ taken from \cite{bellemare:jair2013}.
-Average  CPU times per action in seconds, rounded to nearest integer, shown for  $\IW(1)$ and \BFS. 
+Figures for \BRFS\ and \UCT\ taken from Bellemare et al. Average  CPU times per action in seconds, rounded to nearest integer, shown for  $\IW(1)$ and \BFS. 
 Numbers  in bold show best performer in terms of average score, while numbers shaded in light grey show scores  that are better than \UCT's. 
-Bottom part of the table shows pairwise comparisons among the algorithms. 
-% Entries shown as ``--'' are still running at the time of submission. Each run can take up to 24 hours of CPU time. Eventually
-% all 55 games from  \cite{bellemare:jair2013}  to be included.
-}
+Bottom part of the table shows pairwise comparisons among the algorithms.}
 \label{performance}
-
 \end{table*}
-
-We tested \IW(1) and \BFS\ over \numgames\  of the $55$ games 
-considered in \cite{bellemare:jair2013}, from now on  abbreviated as BNVB.\footnote{
-We left out  \textsc{Skiing} as the reported figures apparently use a  different reward structure.}
-%% check
-% \footnote{
-% Entries marked as  ``--'' in the table are   still running at the time of submission. 
-% Each run can take up to $24$ hours of CPU time and there are $10$ runs per game and algorithm. 
-% Eventually all of  the  $55$ games will  be included in the evaluation.}
-Table~\ref{performance} shows the performance of $\IW(1)$ and \BFS\
-in comparison  with breadth-first search (\BRFS) and \UCT.
-Videos of selected games played by $\IW(1)$, \BFS, and \UCT\ can be 
-seen in Youtube.\footnote{\url{http://www.youtube.com/playlist?list=PLXpQcXUQ_CwenUazUivhXyYvjuS6KQOI0}.}
-The discount factor used by  all the algorithms is $\gamma=0.995$.
-The scores reported for \BRFS\ and \UCT\ are taken from BNVB.  
-Our experimental setup  follows theirs except that a maximum budget of  $150,000$ simulated frames 
-is applied to $\IW(1)$, \BFS, and \UCT. \UCT\ uses this budget by running
- $500$   rollouts of depth $300$. % \footnote{In \UCT,  $500$ base policy rollouts
-% of depth $300$ is equivalent to a budget of $150,000$ simulated frames.}.
-The bound on the number of simulated frames is like a bound on
-lookahead time, as most of the time in the lookahead  is spent
-in calls to the emulator for computing the next RAM state. 
-This is why the average time per action is similar to all the algorithms
-except $\IW(1)$, that due to its pruning does not always use
-the full budget and takes less time per action on average.
-
-Also,  as reported by  BNVB, all of the algorithms 
-reuse the frames in the sub-tree of the previous lookahead that is rooted in the selected 
-child, deleting its siblings and their descendants. More precisely, no calls 
-to the emulator are done for transitions that are cached in that sub-tree, 
-and such reused frames are not discounted from the budget that is thus a bound 
-on the number of \emph{new}  frames per lookahead. In addition, in $\IW(1)$, the states that 
-are reused  from the previous searches are ignored in the computation of the novelty of 
-new states so that more states can escape pruning. Otherwise, $\IW(1)$ often
-uses a fraction of the  budget.  This is not needed in \BFS\ which does 
-no pruning.  $\IW(1)$ and \BFS\ are  limited to  search up to a depth of
-$m=1,500$ frames and up to $m=150,000$ frames per root branch. This is
-to avoid the search from  going too deep or being too committed to a
-single root action. 
-% Similarly UCT defines a limit of $m=300$ frames for each rollout. 
-% Note that UCT is more likely to have a performance
-% loss if the depth is increased, as the number of rollouts would
-% decrease in order to maintain the same budget of $150,000$ simulated
-% frames.
-
-Last, in the lookahead, \IW(1) and \BFS\ select an action every $5$ frames, while \UCT\  selects 
-an action in every frame.  This means that in order to explore a branch 
-$300$ frames deep, UCT gets to choose $300$ actions, while \IW($1$) and \BFS\ 
-get to choose $60$ actions, both however using the same $300$ frames from the
-budget. For this, we  followed the setup  of \BRFS\ in BNVB 
-that also selects actions every $5$ frames, matching the behavior
-of the emulator that requests an action also every $5$ frames. Since the 
-lookahead budget is given by a maximum number of (new) frames, and the time 
-is mostly taken by calls to the emulator, this may not be the best choice  
-for $\IW(1)$ and \BFS\ that may therefore not be exploiting all the 
-options afforded by the budget. We'll look at this further in the future.
-
-\Omit{
-The last difference inherited from BNVB between all algorithms
-and \UCT\, is that \UCT\ selects an action randomly in the rollout every
-single frame, while the others select an action every 5 frames, and
-the remain ones are $no-op$, as is done in the simulation
-environment. Therefore, a node in all algorithms represent $5$ frames,
-but in UCT represents $1$ frame. As a result, a UCT rollout of depth
-$300$ frames involve the selection of 300 actions, decreasing the
-budget of simulated frames by $300$, while the other algorithms need a
-branch of $1500$ frames to select $300$ actions, decreasing the budget
-by $1500$ frames instead.
-}
-
-Table~\ref{performance} shows that both $\IW(1)$ and \BFS\ outperform \BRFS, which  rarely collects
-reward in  many domains as  the depth of the \BRFS\ search tree results in a lookahead of $0.3$
-seconds ($20$ frames or $4$ nodes deep). The notable exception to this
-is \Centipede\ where abundant reward can be collected with a shallow
-lookahead. On the other hand, both $\IW(1)$ and \BFS\ normally 
-reach states that are up to $350$--$1500$ frames deep ($70$--$260$ nodes or 
-$6$--$22$ seconds),  even if $\IW(1)$ does not always use all the simulation frames allocated due to its
-agressive pruning. This can be observed in games such as \Breakout, \CrazyClimber, \Kangaroo, and \Pooyan, where the average
-CPU time for each lookahead is up to $10$ times faster than \BFS.  Computation time for \UCT\ and \BRFS\ are  similar to \BFS, as the
-most expensive part of the computation is the  generation of frames through the simulator,  and these three  algorithms 
-always use  the full budget.
-
-More interestingly,  \IW(1) outscores \UCT\  in $31$ of the \numgames\ games,
-while \BFS\ outscores \UCT\ in $26$.
-%\footnote{These numbers can actually
-%improve further as 6  games for \IW(1) and 4 games for \BFS\ are still running.}
-On the other hand, \UCT\ does better than $\IW(1)$ and \BFS\ in $19$ and $25$ games respectively.
-The relative performance between $\IW(1)$ and \BFS\ makes $\IW(1)$  the
-best of the two in $34$ games, and  \BFS\  in $16$. In terms of the number of games 
-where an algorithm is the best, % $\IW(1)$  domainates the others: 
-$\IW(1)$ is  the best in $26$ games, \BFS\ in $13$ games, and \UCT\ in $19$ games. 
-Also, \BRFS\ is best in $2$ games (\Centipede, tied up in \Boxing), while the other three algorithms
-are tied in another 2 games (\Pong, \Boxing). 
-% , and $\IW(1)$ and \BFS\ are tied in \TimePilot.
-
-\Omit{As mentioned in the
-previous section, the order in which $\IW(1)$ and \BFS\ generate the
-children of an expanded node is randomized, something which has a
-potential effect on the novelty measures of nodes, and hence in the
-search graph. We found that this randomization helps, in particular in
-\Qbert, where not randomizing results in $\IW(1)$ being trapped into
-loops sometimes.}
 
@@
 
-\section{Summary and Future Work}
+\section{Summary}
 
@@
 that originate in work in classical planning,  can be used to play  the Atari video
-games where they  achieve  state-of-the-art performance. This level of play 
-is the result of a structured exploration of the state space that manages to combine 
-the scope of blind search algorithms with the efficiency of heuristic search methods. 
-There are several  issues worth studying, we  mention two of them.
-Regarding planning, the results show the potential of width-based algorithms
-for a broader class of planning problems. As shown, indeed, these algorithms
-do not require PDDL encodings and can deal with state successor functions
-encoded \emph{procedurally} (as in simulators) as long as the structure of the states 
-is known. 
+games where they  achieve  state-of-the-art performance. The  results also  suggest 
+more generally the potential of width-based methods for planning with simulators when factored, 
+compact action models are not available. In this sense, the scope of these planning methods 
+is broader than those of heuristic-search planning methods that require propositional
+encodings of actions and goals, and with suitable extensions, 
+may potentially  approach the scope of MCTS-methods like UCT that work on simulators as well.
+
+\newpage
+
 %% This suggests to us that the algorithms discussed in the 
@@
 
+\Omit{
 Concerning the Atari games, there are two problems in all the planning methods
@@
 funded by the Australian Research Council Linkage grant LP11010015.
+}
 
@@
 
-\bibliographystyle{aaai}
+\bibliographystyle{named}
 \bibliography{control}

### 2015-02-04 ijcai15-paper/make-tarball.sh
Helpful script for packing up stuff

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+tar jcvf ijcai_15.tar.bz2 *.tex *.bib *.bst

### 2015-02-05 ijcai15-paper/paper.tex
Typo fixed in Table caption
Tweaks to comments about width-based vs. RL

@@
 
- \caption{\small Performance that results form various  lookahead algorithms in \numgames\  Atari games. The algorithms, \BRFS, $\IW(1)$, \BFS, and \UCT, are evaluated
+ \caption{\small Performance that results from various  lookahead algorithms in \numgames\  Atari games. The algorithms, \BRFS, $\IW(1)$, \BFS, and \UCT, are evaluated
  over $10$ runs (episodes) for each game. 
@@
 
-$\IW(1)$ also outperforms the best learning algorithm in
+$\IW(1)$ obtains better scores than the best learning algorithm in
 \cite{deep-mind-atari} in the $7$ games considered there, and
-\BFS\ does so in $6$ of the $7$ games. Comparing with the scores reported for the reinforcement learning algorithms by BNVB,
+\BFS\ does so in $6$ of the $7$ games. Comparing with the scores reported for the reinforcement learning algorithms in BNVB,
 we note that both $\IW(1)$ and \BFS\ do much better than the best learning algorithm in
-those games where the learning algorithms outperform \UCT\ like \Montezuma, \Venture\
-and \Bowling.  In \Montezuma\ rewards are very
-sparse, deep, and most of the actions lead to losing a life with
-no immediate penalty or consequence.  All algorithms achieve $0$ score, 
+those games where the learning algorithms outperform \UCT\, namely, \Montezuma, \Venture\
+and \Bowling. We take this as evidence that $\IW(1)$ and \BFS\ are as at least as good as learning
+algorithms at finding rewards in games where \UCT\ is not very effective.
+
+For instance, in \Montezuma\ rewards are very sparse, deep, and most of the actions lead to losing a life with
+no immediate penalty or consequence.  In our experiments, all algorithms achieve $0$ score, 
 except for \BFS\ that achieves an average score of $540$, and a score
-of $2500$ in one of the runs. This means however that even \BFS\  is not able to 
+of $2,500$ in one of the runs. This means however that even \BFS\  is not able to 
 consistently find rewards in this game. This game and several others

### 2015-02-05 ijcai15-paper/paper.tex
Macro for MCTS

@@
 \newcommand{\UCT}{{\textsc{UCT}}}
+\newcommand{\MCTS}{{\textsc{MCTS}}}
 
@@
 encodings of actions and goals, and with suitable extensions, 
-may potentially  approach the scope of MCTS-methods like UCT that work on simulators as well.
+may potentially  approach the scope of \MCTS\ methods like \UCT\ that work on simulators as well.
 

### 2015-02-05 src/agents/UCTNoveltySearchTree.cpp
latest changes on shaping

@@ -84,8 +84,13 @@ void UCTNoveltySearchTree::update_tree(void) {
 	
 	
 	get_exploration_bonus( m_env->getRAM(), 0 );
+	
+	float mc_return;
+	do_monte_carlo((UCTNoveltyTreeNode*)p_root, uct_search_depth, mc_return);
 
 	std::cout << "starting with " << simulation_steps << " simulation steps" <<std::endl;
+	
+	p_root->m_depth = 0;
 	
 	int num_iterations = 0;
 	while (true) {
@@ -237,10 +242,10 @@ int UCTNoveltySearchTree::single_uct_iteration(void) {
 		if( node->p_parent != NULL)
 			((UCTNoveltyTreeNode*)node)->return_achieved = ((UCTNoveltyTreeNode*)node->p_parent)->return_achieved;
 		
-
 		((UCTNoveltyTreeNode*)node)->return_achieved += step_return;		
 		node->initialized = true;
-		
+		node->act = leaf_choice;
+
 		//((UCTNoveltyTreeNode*)node)->init_novelty(); 
 		
 
@@ -253,8 +258,8 @@ int UCTNoveltySearchTree::single_uct_iteration(void) {
 		// else
 		// std::cout << "Node depth:" << node->m_depth << "action: " << action_to_string( leaf_choice ) << "is Novel"<< std::endl;
 
-		// if( ((UCTNoveltyTreeNode*)node)->return_achieved > 0)
-		// 	std::cout << "Node depth:" << node->m_depth << "action: " << action_to_string( leaf_choice )  << " return: " << ((UCTNoveltyTreeNode*)node)->return_achieved << std::endl;
+		//if( ((UCTNoveltyTreeNode*)node)->return_achieved > 0)
+		// 	std::cout << "\t New Node depth:" << node->m_depth << "action: " << action_to_string( leaf_choice )  << " return: " << ((UCTNoveltyTreeNode*)node)->return_achieved << std::endl;
 		// Before declaring ourselves done, ensure that this is not a duplicate
 		//  of another action
 		if (ignore_duplicates)
@@ -273,6 +278,16 @@ int UCTNoveltySearchTree::single_uct_iteration(void) {
   
 	int mc_steps = uct_search_depth - node_depth; 
   
+	//if( ((UCTNoveltyTreeNode*)node)->return_achieved > 0)
+		// {
[…]

### 2015-02-09 ijcai15-paper/paper.tex
nir comments

@@
 \usepackage{times}
+
+\usepackage{todonotes}
+\newcommand{\nir}{\todo[inline,color=orange!10]}
+\newcommand{\hg}{\todo[color=blue!10]}
+\newcommand{\mrj}{\todo[inline,color=green!10]}
 
@@
 
-\title{Classical Planning with Simulations: \\ Results on the Atari Video Games}
+\title{Classical Planning with Simulators: \\ Results on the Atari Video Games}
 
@@
 as a  \emph{net-benefit planning problem} \cite{coles2012survey,keyder:jair09} with unknown state transitions  and  rewards. 
-Namely, we seek  an action sequence $a_0, a_1, \ldots, a_m$ from the  current  state $s_0$ that generates
+\Omit{Namely}
+\nir{While the \emph{reinforcement learning} problem seeks to synthesize a controller as a policy/function that maximizes 
+the reward of any trajectory}, we seek  an action sequence $a_0, a_1, \ldots, a_m$ from the  current  state $s_0$ that generates
 a state sequence $s_0, s_1, \ldots, s_{m+1}$ with  maximum  total reward $\sum_{i=0}^{m} r(a_i,s_i)$,

### 2015-02-09 ijcai15-paper/paper.tex
Miguel's Iteration #1
* Added Nir's sentence, with a few changes
* Fixed the bleeding into the References page following the above by
  trimming the space between the bullet points describing IW (via
  the package enumitem).
* Hunted a couple cases we were using the English British/Australian
  spelling for words ending on 'ze' (optmize vs. optimise, analyze vs.
  analyze, etc.).

@@
 \usepackage{todonotes}
-\newcommand{\nir}{\todo[inline,color=orange!10]}
+\newcommand{\nir}{\todo[color=orange!10]}
 \newcommand{\hg}{\todo[color=blue!10]}
@@
 % \setlength{\pdfpageheight}{11in}
+\usepackage{enumitem}
+%\setlist{nosep} % or \setlist{noitemsep} to leave space around whole list
 
@@
 
-\author{\textbf{ID:} \ldots  \ \ ; \ \ \textbf{Keywords:} classical planning, search, video games}
+\author{\textbf{ID:} 320  \ \ ; \ \ \textbf{Keywords:} classical planning, search, video games}
 
@@
 as a  \emph{net-benefit planning problem} \cite{coles2012survey,keyder:jair09} with unknown state transitions  and  rewards. 
-\Omit{Namely}
-\nir{While the \emph{reinforcement learning} problem seeks to synthesize a controller as a policy/function that maximizes 
-the reward of any trajectory}, we seek  an action sequence $a_0, a_1, \ldots, a_m$ from the  current  state $s_0$ that generates
+%\Omit{Namely} \Omit{While the \emph{reinforcement learning} problem seeks to synthesize a controller as a policy/function that maximizes 
+%the reward of any trajectory\nir{NIR remark}}
+While in the former a policy mapping every state $s$ to an action $a$ that maximizes expected reward is sought, in the latter one 
+seeks an action sequence $a_0, a_1, \ldots, a_m$ from the  \emph{current}  state $s_0$ that generates
 a state sequence $s_0, s_1, \ldots, s_{m+1}$ with  maximum  total reward $\sum_{i=0}^{m} r(a_i,s_i)$,
@@
 
-The limitation of the   basic blind search  methods is that they are not effective  over large  spaces, neither
+The limitation of the   basic blind search  methods is that they are not effective  over large state spaces, neither
 for solving problems off-line, nor for guiding  a lookahead search for solving problems on-line.
@@
 
-\begin{itemize}
+\begin{itemize}[noitemsep]
 \item The \emph{novelty of a newly generate state $s$  in a  search algorithm}
@@
 
-\item  $\IW(i)$ is a breadth-first search that prunes newly generated states when their novelty measure
+\item  $\IW(i)$ is a breadth-first search that prunes newly generated states when their novelty \Omit{measure}
 is greater than $i$.
@@
 
+\medskip
 For classical planning, the termination condition is the achievement of the goal. In the \emph{on-line setting,}
@@
 for problems where a goal state needs to be reached, the notions remain
-relevant  in  optimisation problems as well. Indeed, if a  good path is made of
+relevant  in  optimization problems as well. Indeed, if a  good path is made of
 states $s_i$ each of which has a low width, \IW\ can be made
@@
 as they compute different things. Learning controllers appears as a more challenging problem
-and it is thus not surprising that planning algorithms like UCT tend to achieve a higher score than 
+and it is thus not surprising that planning algorithms like \UCT\ tend to achieve a higher score than 
 learning algorithms. In addition, the learning algorithms reported by BNVB tend to use 
@@
 environment. Therefore, a node in all algorithms represent $5$ frames,
-but in UCT represents $1$ frame. As a result, a UCT rollout of depth
+but in UCT represents $1$ frame. As a result, a \UCT\ rollout of depth
 $300$ frames involve the selection of 300 actions, decreasing the
@@
 
- \caption{\small Performance that results from various  lookahead algorithms in \numgames\  Atari games. The algorithms, \BRFS, $\IW(1)$, \BFS, and \UCT, are evaluated
+ \caption{\small Performance that results from various  lookahead algorithms in \numgames\  Atari 2600 games. The algorithms, \BRFS, $\IW(1)$, \BFS, and \UCT, are evaluated
  over $10$ runs (episodes) for each game. 
@@
 breadth-first search. These novelty measures make use of the factored representation of the states for
-providing a structure to the search: states that have width 1 are explored first in linear time,
-then states that have  width 2  are explored in quadratic time, and so on. In classical planning problems
+providing a structure to the search: states that have width $1$ are explored first in linear time,
+then states that have  width $2$ are explored in quadratic time, and so on. In classical planning problems
 with atomic goals, this way of organizing the search  pays off both theoretically and practically.
@@
 the locations that are  reachable with it, and the locations reachable with
-the  other controllers in the population (a diversity measure). The novelty measure used by \IW, on the other hand, is domain-independent and it
+the  other controllers in the population (a diversity measure). The novelty measure used by 
+\IW, on the other hand, is domain-independent and it
 is determined by the structure of the states as captured by the problem variables.

### 2015-02-10 ijcai15-paper/paper.tex
Removed space around item list
Hector changes

@@
 % \setlength{\pdfpageheight}{11in}
-\usepackage{enumitem}
-%\setlist{nosep} % or \setlist{noitemsep} to leave space around whole list
+
+% \usepackage{enumitem}
+% \setlist{nosep} % or \setlist{noitemsep} to leave space around whole list
+% HG: better not play with spacing .. doesn't look as nice
 
@@
 
-\author{\textbf{ID:} 320  \ \ ; \ \ \textbf{Keywords:} classical planning, search, video games}
+\author{\textbf{ID:} 320  \ \ ; \ \ \textbf{Keywords:} classical planning, search, video games, application}
 
@@
 
-The action selection problem in the Atari games can be seen as a \emph{reinforcement learning} problem
+The action selection problem in the Atari games can be addressed  as a \emph{reinforcement learning} problem
 \cite{sutton:book} over a deterministic MDP where the state transitions and rewards are not known, or alternatively,
 as a  \emph{net-benefit planning problem} \cite{coles2012survey,keyder:jair09} with unknown state transitions  and  rewards. 
-%\Omit{Namely} \Omit{While the \emph{reinforcement learning} problem seeks to synthesize a controller as a policy/function that maximizes 
-%the reward of any trajectory\nir{NIR remark}}
-While in the former a policy mapping every state $s$ to an action $a$ that maximizes expected reward is sought, in the latter one 
-seeks an action sequence $a_0, a_1, \ldots, a_m$ from the  \emph{current}  state $s_0$ that generates
-a state sequence $s_0, s_1, \ldots, s_{m+1}$ with  maximum  total reward $\sum_{i=0}^{m} r(a_i,s_i)$,
-where $m$ is a sufficiently large planning horizon, and the rewards $r(a_i,s_i)$ and the  state-transitions $s_{i+1}=f(a_i,s_i)$
-are known only after  action $a_i$ is applied in the state $s_i$.
+ALE supports the two settings: an \emph{on-line plannning setting} where  actions are selected
+after a lookahead, and a \emph{learning setting}  that  must produce   controllers for mapping  states into
+actions reactively without any lookahead. In this work we pursue the on-line planning approach where the lookahead
+seeks  action sequences with  maximum  reward up to a certain horizon.
 
@@
 
-\begin{itemize}[noitemsep]
+\begin{itemize}
 \item The \emph{novelty of a newly generate state $s$  in a  search algorithm}
@@
 and so on.
-
 \item  $\IW(i)$ is a breadth-first search that prunes newly generated states when their novelty \Omit{measure}
 is greater than $i$.
-
 \item $\IW$  calls $\IW(i)$ sequentially for $i=1,2,\ldots$ until a termination condition is reached,
@@
 goals into $N$ problems with one atomic goal each.  Since $\IW(k)$ runs in time that is exponential in $k$, 
-this mean that almost $90\%$ of the 37,921 instances are  solved in time that is either linear or quadratic in the number of problem variables, 
+this means that almost $90\%$ of the 37,921 instances are  solved in time that is either linear or quadratic in the number of problem variables, 
 which in such encodings are all \emph{boolean}. Furthermore, when the performance of \IW\   is compared with  

### 2015-02-11 ijcai15-paper/paper.tex
typos fixed by nir

@@
 no pruning.  $\IW(1)$ and \BFS\ are  limited to  search up to a depth of
-$m=1,500$ frames and up to $m=150,000$ frames per root branch. This is
+$1,500$ frames and up to $150,000$ frames per root branch. This is
 to avoid the search from  going too deep or being too committed to a
@@
 
-$\IW(1)$ obtains better scores than the best learning algorithm in
+$\IW(1)$ obtains better scores than the best learning algorithm
 \cite{deep-mind-atari} in the $7$ games considered there, and

### 2015-04-30 ijcai15-paper/paper.tex
Mods for Camera Ready
* Added UniMelb cluster data
* Added data on the UCT experiments
* Added authors and affiliations

@@
 
-\author{\textbf{ID:} 320  \ \ ; \ \ \textbf{Keywords:} classical planning, search, video games, application}
+%\author{\textbf{ID:} 320  \ \ ; \ \ \textbf{Keywords:} classical planning, search, video games, application}
 
-% \author{Nir Lipovetzky \\
-%        The University of Melbourne \\
-%     Melbourne, Australia\\
-%     {\normalsize\url{nir.lipovetzky@unimelb.edu.au}}
-%%% footnote doesn't look nice ; this looks better
-%% although can't use Nir's unimelb address that is too long :-)
-%% 
-%    {\small\url{nirlipo@gmail.com}}
-%    \And
-%    Miquel Ramirez \\
-%    Australian National University \\
-%    Canberra, Australia\\
-%   {\small\url{miquel.ramirez@gmail.com}}
-%    \And
-%    Hector Geffner \\
-%    ICREA \&  Universitat Pompeu Fabra \\
-%    Barcelona, SPAIN \\
-%    {\small\url{hector.geffner@upf.edu}}
- %%%%% \thanks{firstname.lastname}
-%}
+ \author{Nir Lipovetzky \\
+        {\small The University of Melbourne }\\
+     {\small Melbourne, Australia}\\
+     {\footnotesize\url{nir.lipovetzky@unimelb.edu.au}}
+%% footnote doesn't look nice ; this looks better
+% although can't use Nir's unimelb address that is too long :-)
+% 
+%    {\tiny\url{nirlipo@gmail.com}}
+    \And
+    Miquel Ramirez \\
+    {\small Australian National University} \\
+    {\small Canberra, Australia}\\
+   {\footnotesize\url{miquel.ramirez@anu.edu.au}}
+    \And
+    Hector Geffner \\
+    {\small ICREA \& Universitat Pompeu Fabra }\\
+    {\small Barcelona, Spain }\\
+    {\footnotesize\url{hector.geffner@upf.edu}}
+ %%%% \thanks{firstname.lastname}
+}
 
@@
 more challenging than the other; for the learning algorithms, BNVB mention
-that the results tend to be better for the screen inputs.
+that the results tend to be better for the screen inputs. Experiments were
+run on a high-performance computation cluster run by the University of Melbourne, where each computing node consists of a $6$-core Intel Xeon 
+E5-2440, with $2.4$ GHz clock speed, with 64 GBytes of RAM installed.
 
@@
 for $\IW(1)$ and \BFS\ that may therefore not be exploiting all the 
-options afforded by the budget. We'll look at this further in the future.
+options afforded by the budget. Interestingly, when \UCT\ is limited to one
+action each $5$ frames, its performance is reduced by up to a $50$\% in
+games where it performs very well (\CrazyClimber), and does not a difference in those games where it does very poorly (\Freeway). We'll look at this further in the future.
 
@@
 
-\newpage
-
 %% This suggests to us that the algorithms discussed in the 

### 2015-04-30 ijcai15-paper/paper.tex
Changes updated. need to cut down to 6 pages

@@
 Last, in the lookahead, \IW(1) and \BFS\ select an action every $5$ frames, while \UCT\  selects 
-an action in every frame.  This means that in order to explore a branch 
+an action every frame.  This means that in order to explore a branch 
 $300$ frames deep, UCT gets to choose $300$ actions, while \IW($1$) and \BFS\ 
@@
 options afforded by the budget. Interestingly, when \UCT\ is limited to one
-action each $5$ frames, its performance is reduced by up to a $50$\% in
-games where it performs very well (\CrazyClimber), and does not a difference in those games where it does very poorly (\Freeway). We'll look at this further in the future.
+action every $5$ frames, its performance is reduced by up to a $50$\% in
+games where it performs very well (\CrazyClimber), and does not improve in those games where it performs very poorly (\Freeway).
 

### 2015-04-30 ijcai15-paper/paper.tex
last changes

@@
 
-\usepackage{helvet}
-\usepackage{courier}
+%\usepackage{helvet}
+%\usepackage{courier}
 \usepackage{url}
@@
 \begin{abstract}
-The Atari 2600 games supported in  the Arcade Learning Environment~\cite{bellemare:jair2013}
-all feature a known initial (RAM) state and actions that have deterministic effects. Classical planners, however,
-cannot be used  off-the-shelf as there is no  compact PDDL-model of  the games, and action effects and goals are  not
-known \emph{a priori}. Indeed, there are no explicit goals, and the planner must select actions on-line while interacting 
-with a simulator that returns successor states and rewards. None of this precludes  the use of  blind lookahead algorithms 
-for action selection like breadth-first search or  Dijkstra's yet such methods are not effective over large  state spaces. 
-We thus turn to a different class of  planning methods  introduced recently  that  have been shown to be  effective for solving
-large planning problems but which do not require  prior knowledge of state transitions and costs (rewards) or goals.
-The simplest such algorithm, called Iterated Width or \IW, consists of  a sequence of calls  $\IW(1)$, $\IW(2)$, \ldots,
-where  $\IW(k)$ is  a breadth-first search in which a state is  pruned when it is not the first state in the  search  to make true 
-a subset of $k$ atoms. The  empirical results over \numgames\ Atari games show that  $\IW(1)$ performs at the level of UCT, 
-the state-of-the-art planning method in this domain, and suggest the potential of width-based methods for planning with
-simulators when factored, compact action models are not available.
+The Atari 2600 games supported in the Arcade Learning
+Environment~\cite{bellemare:jair2013} all feature a known initial
+(RAM) state and actions that have deterministic effects. Classical
+planners, however, cannot be used off-the-shelf as there is no compact
+PDDL-model of the games, and action effects and goals are not known
+\emph{a priori}. Indeed, there are no explicit goals, and the planner
+must select actions on-line while interacting with a simulator that
+returns successor states and rewards. None of this precludes the use
+of blind lookahead algorithms for action selection like breadth-first
+search or Dijkstra's yet such methods are not effective over large
+state spaces.  We thus turn to a different class of
+planning methods introduced recently that have been shown to be
+effective for solving large planning problems but which do not require
+prior knowledge of state transitions and costs (rewards) or goals. The
+empirical results over \numgames\ Atari games show that the simplest
+such algorithm performs at the level of UCT, the state-of-the-art
+planning method in this domain, and suggest the potential of
+width-based methods for planning with simulators when factored,
+compact action models are not available.
+
+%% The Atari 2600 games supported in the Arcade Learning
+%% Environment~\cite{bellemare:jair2013} all feature a known initial
+%% (RAM) state and actions that have deterministic effects. Classical
+%% planners, however, cannot be used off-the-shelf as there is no compact
+%% PDDL-model of the games, and action effects and goals are not known
+%% \emph{a priori}. Indeed, there are no explicit goals, and the planner
+%% must select actions on-line while interacting with a simulator that
+%% returns successor states and rewards. None of this precludes the use
+%% of blind lookahead algorithms for action selection like breadth-first
+%% search or Dijkstra's yet such methods are not effective over large
+%% state spaces.  We thus turn to a different class of planning methods
+%% introduced recently that have been shown to be effective for solving
+%% large planning problems but which do not require prior knowledge of
+%% state transitions and costs (rewards) or goals.  The simplest such
+%% algorithm, called Iterated Width or \IW, consists of a sequence of
+%% calls $\IW(1)$, $\IW(2)$, \ldots, where $\IW(k)$ is a breadth-first
+%% search in which a state is pruned when it is not the first state in
+%% the search to make true a subset of $k$ atoms. The empirical results
+%% over \numgames\ Atari games show that $\IW(1)$ performs at the level
+%% of UCT, the state-of-the-art planning method in this domain, and
+%% suggest the potential of width-based methods for planning with
+%% simulators when factored, compact action models are not available.
+
 \end{abstract}
@@
 \cite{planning:book,geffner:book}. This is because there is  no compact PDDL-like encoding of the domain and 
-the goal to be achieved in each game  is not given. Indeed, there are no goals but rewards, and 
+the goal to be achieved in each game  is not given, precluding the automatic derivation of heuristic functions and other powerful inferences. Indeed, there are no goals but rewards, and 
 the planner must select actions on-line while interacting with a simulator that just
@@
 that the results tend to be better for the screen inputs. Experiments were
-run on a high-performance computation cluster run by the University of Melbourne, where each computing node consists of a $6$-core Intel Xeon 
+run on a cluster, where each computing node consists of a $6$-core Intel Xeon 
 E5-2440, with $2.4$ GHz clock speed, with 64 GBytes of RAM installed.
@@
 
-\noindent \textbf{Acknowledgments.} We thank the people involved in
+\noindent }
+
+\section*{Acknowledgments} We thank the people involved in
 the creation and maintenance of the Arcade Learning Environment that
@@
 funded by the Australian Research Council Linkage grant LP11010015.
-}
 

### 2015-04-30 ijcai15-paper/paper.tex
camera ready ijcai

--- 2015-04-30 ijcai15-paper/paper.tex
@@
 \begin{document}
-% The file aaai.sty is the style file for AAAI Press
-% proceedings, working notes, and technical reports.
-%
-% \pdfinfo{
-% /Title (Classical Planning Algorithms on the Atari Video Games)
-% /Author (Nir Lipovetzky, Miquel Ramirez, Hector Geffner)}
+
+\pdfinfo{
+/Title (Classical Planning with Simulators: Results on the Atari Video Games)
+/Author (Nir Lipovetzky, Miquel Ramirez, Hector Geffner)}
 
@@
  \author{Nir Lipovetzky \\
-        {\small The University of Melbourne }\\
-     {\small Melbourne, Australia}\\
+        { The University of Melbourne }\\
+     { Melbourne, Australia}\\
      {\footnotesize\url{nir.lipovetzky@unimelb.edu.au}}
@@
     Miquel Ramirez \\
-    {\small Australian National University} \\
-    {\small Canberra, Australia}\\
+    { Australian National University} \\
+    { Canberra, Australia}\\
    {\footnotesize\url{miquel.ramirez@anu.edu.au}}
@@
     Hector Geffner \\
-    {\small ICREA \& Universitat Pompeu Fabra }\\
-    {\small Barcelona, Spain }\\
+    { ICREA \& U. Pompeu Fabra }\\
+    { Barcelona, Spain }\\
     {\footnotesize\url{hector.geffner@upf.edu}}
@@
 \begin{abstract}
-The Atari 2600 games supported in  the Arcade Learning Environment~\cite{bellemare:jair2013}
-all feature a known initial (RAM) state and actions that have deterministic effects. Classical planners, however,
-cannot be used  off-the-shelf as there is no  compact PDDL-model of  the games, and action effects and goals are  not
-known \emph{a priori}. Indeed, there are no explicit goals, and the planner must select actions on-line while interacting 
-with a simulator that returns successor states and rewards. None of this precludes  the use of  blind lookahead algorithms 
-for action selection like breadth-first search or  Dijkstra's yet such methods are not effective over large  state spaces. 
-We thus turn to a different class of  planning methods  introduced recently  that  have been shown to be  effective for solving
-large planning problems but which do not require  prior knowledge of state transitions and costs (rewards) or goals.
-The simplest such algorithm, called Iterated Width or \IW, consists of  a sequence of calls  $\IW(1)$, $\IW(2)$, \ldots,
-where  $\IW(k)$ is  a breadth-first search in which a state is  pruned when it is not the first state in the  search  to make true 
-a subset of $k$ atoms. The  empirical results over \numgames\ Atari games show that  $\IW(1)$ performs at the level of UCT, 
-the state-of-the-art planning method in this domain, and suggest the potential of width-based methods for planning with
-simulators when factored, compact action models are not available.
+The Atari 2600 games supported in the Arcade Learning
+Environment~\cite{bellemare:jair2013} all feature a known initial
+(RAM) state and actions that have deterministic effects. Classical
+planners, however, cannot be used off-the-shelf as there is no compact
+PDDL-model of the games, and action effects and goals are not known
+\emph{a priori}. Indeed, there are no explicit goals, and the planner
+must select actions on-line while interacting with a simulator that
+returns successor states and rewards. None of this precludes the use
+of blind lookahead algorithms for action selection like breadth-first
+search or Dijkstra's yet such methods are not effective over large
+state spaces.  We thus turn to a different class of
+planning methods introduced recently that have been shown to be
+effective for solving large planning problems but which do not require
+prior knowledge of state transitions, costs (rewards) or goals. The
+empirical results over \numgames\ Atari games show that the simplest
+such algorithm performs at the level of UCT, the state-of-the-art
+planning method in this domain, and suggest the potential of
+width-based methods for planning with simulators when factored,
+compact action models are not available.
+
 \end{abstract}
@@
 \cite{planning:book,geffner:book}. This is because there is  no compact PDDL-like encoding of the domain and 
-the goal to be achieved in each game  is not given. Indeed, there are no goals but rewards, and 
+the goal to be achieved in each game  is not given, precluding the automatic derivation of heuristic functions and other inferences. Indeed, there are no goals but rewards, and 
 the planner must select actions on-line while interacting with a simulator that just
@@
 after a lookahead, and a \emph{learning setting}  that  must produce   controllers for mapping  states into
-actions reactively without any lookahead. In this work we pursue the on-line planning approach where the lookahead
-seeks  action sequences with  maximum  reward up to a certain horizon.
+actions reactively without any lookahead. In this work, we are interested in the on-line planning setting.
 
@@
 The paper is organized as follows. We review the iterated width  algorithm and its properties,
-look at the    variations of the algorithm that we used in the Atari games, present and analyze the  
-experimental results, and discuss related and future work.
+look at the    variations of the algorithm that we used in the Atari games, and present the experimental results.
 
@@
 achieves the dummy goal and that has the original conjunctive goal as a precondition.
-Yet, this changes the definition of the domain.} For example, $37\%$ of the 37,921 problems
+Yet, this mapping changes the definition of the domain.} For example, $37\%$ of the 37,921 problems
 considered  in \cite{nir:ecai2012} are solved by $\IW(1)$ while $51.3\%$ are  solved by $\IW(2)$.
@@
 
-The number of nodes \emph{generated} by $\IW(1)$ is $n \times D \times b$ in the worst case,
-where $n$ is the number of problem variables, $D$ is the size of their domains, and $b$ is the number of
-actions per  state.  This same number in a  breadth-first search is not linear in $n$
-but exponential. For the Atari games,  $n=128$, $D=256$, and $b=18$, so that the product is
-equal to $589,824$, which is large but feasible. On the other hand, the number
-of nodes generated by $\IW(2)$ in the worst case is  $(n \times D)^2 \times b$, which is
-equal to $19,327,352,832$ which is too large, forcing us to consider only a tiny fraction of
-such states. For classical planning problems, the growth in the number of nodes from
-$\IW(1)$ to $\IW(2)$ is not that large, as the variables are boolean. Indeed, we could have
-taken the state vector for the Atari games as a vector of 1024 boolean variables,
-and apply these algorithms to that representation. The results would be different.
-In such a case, $\IW(1)$ and $\IW(2)$  would generate $n' \times D' \times b$ and
-$(n' \times D')^2 \times b$ states, which for $n'=1024$ and $D=2$
-represent  $36,864$ and $75,497,472$ states respectively.
-These are feasible numbers, but we haven't used  this representation 
-under the assumption that the correlations among the bits in each one of
-the 128 words of the state vectors are meaningful.
-In summary, from the basic \IW\ algorithm we are testing only the first
-call $\IW(1)$.
+The number of nodes \emph{generated} by $\IW(1)$ is $n \times D \times
+b$ in the worst case, where $n$ is the number of problem variables,
+$D$ is the size of their domains, and $b$ is the number of actions per
+state.  This same number in a breadth-first search is not linear in
+$n$ but exponential. For the Atari games, $n=128$, $D=256$, and
+$b=18$, so that the product is equal to $589,824$, which is large but
+feasible. On the other hand, the number of nodes generated by $\IW(2)$
+in the worst case is $(n \times D)^2 \times b$, which is equal to
+$19,327,352,832$ which is too large, forcing us to consider only a
+tiny fraction of such states. For classical planning problems, the
+growth in the number of nodes from $\IW(1)$ to $\IW(2)$ is not that
+large, as the variables are boolean. Indeed, we could have taken the
+state vector for the Atari games as a vector of 1024 boolean
+variables, and apply these algorithms to that representation. The
+number of atoms would indeed be much smaller, and both IW(1) and IW(2)
+would run faster then. However by ignoring the correlations among bits
+in each one of the 128 words, the results would be weaker.
 
@@
 the state of the screen pixels, while the planning algorithms,  use the state
-of the RAM memory. It is not clear however whether one input representation is
-more challenging than the other; for the learning algorithms, BNVB mention
+of the RAM memory. It is not clear however whether the use of one input representation is
+more challenging than the use of the other. For the learning algorithms, BNVB mention
 that the results tend to be better for the screen inputs. Experiments were
-run on a high-performance computation cluster run by the University of Melbourne, where each computing node consists of a $6$-core Intel Xeon 
+run on a cluster, where each computing node consists of a $6$-core Intel Xeon 
 E5-2440, with $2.4$ GHz clock speed, with 64 GBytes of RAM installed.
@@
 
-Last, in the lookahead, \IW(1) and \BFS\ select an action every $5$ frames, while \UCT\  selects 
-an action every frame.  This means that in order to explore a branch 
-$300$ frames deep, UCT gets to choose $300$ actions, while \IW($1$) and \BFS\ 
-get to choose $60$ actions, both however using the same $300$ frames from the
-budget. For this, we  followed the setup  of \BRFS\ in BNVB 
-that also selects actions every $5$ frames, matching the behavior
-of the emulator that requests an action also every $5$ frames. Since the 
-lookahead budget is given by a maximum number of (new) frames, and the time 
-is mostly taken by calls to the emulator, this may not be the best choice  
-for $\IW(1)$ and \BFS\ that may therefore not be exploiting all the 
-options afforded by the budget. Interestingly, when \UCT\ is limited to one
-action every $5$ frames, its performance is reduced by up to a $50$\% in
-games where it performs very well (\CrazyClimber), and does not improve in those games where it performs very poorly (\Freeway).
+Last, in the lookahead, \IW(1) and \BFS\ select an action every $5$
+frames, while \UCT\ selects an action every frame.  This means that in
+order to explore a branch $300$ frames deep, UCT gets to choose $300$
+actions, while \IW($1$) and \BFS\ get to choose $60$ actions, both
+however using the same $300$ frames from the budget. For this, we
+followed the setup of \BRFS\ in BNVB that also selects actions every
+$5$ frames, matching the behavior of the emulator that requests an
+action also every $5$ frames. Since the lookahead budget is given by a
+maximum number of (new) frames, and the time is mostly taken by calls
+to the emulator, this may not be the best choice for $\IW(1)$ and
+\BFS\ that may therefore not be exploiting all the options afforded by
+the budget. Interestingly, when \UCT\ is limited to one action every
+$5$ frames, its performance is reduced by up to a $50$\% in games
+where it performs very well (\CrazyClimber), and does not appear to
+improve in those games where it performs very poorly (\Freeway).
 
@@
 
-\noindent \textbf{Acknowledgments.} We thank the people involved in
-the creation and maintenance of the Arcade Learning Environment that
-have made this work possible. The research by Nir Lipovetzky is partially
+\noindent }
+
+\section*{Acknowledgments} We thank the creators of the ALE. The research by Nir Lipovetzky is partially
 funded by the Australian Research Council Linkage grant LP11010015.
-}
 

### 2015-04-30 hsdip-15-paper/paper.tex
added HSDIP formatted as IJCAI submission
hsdip camera ready

--- 2015-02-11 ijcai15-paper/paper.tex
@@
 % The file ijcai15.sty is the style file for IJCAI-15 (same as ijcai07.sty).
-\usepackage{ijcai15}
+\usepackage{aaai}
 
@@
 
-\author{\textbf{ID:} 320  \ \ ; \ \ \textbf{Keywords:} classical planning, search, video games, application}
+%\author{ \textbf{Keywords:} classical planning, search, video games, application}
 
-% \author{Nir Lipovetzky \\
-%        The University of Melbourne \\
-%     Melbourne, Australia\\
-%     {\normalsize\url{nir.lipovetzky@unimelb.edu.au}}
-%%% footnote doesn't look nice ; this looks better
-%% although can't use Nir's unimelb address that is too long :-)
-%% 
-%    {\small\url{nirlipo@gmail.com}}
-%    \And
-%    Miquel Ramirez \\
-%    Australian National University \\
-%    Canberra, Australia\\
-%   {\small\url{miquel.ramirez@gmail.com}}
-%    \And
-%    Hector Geffner \\
-%    ICREA \&  Universitat Pompeu Fabra \\
-%    Barcelona, SPAIN \\
-%    {\small\url{hector.geffner@upf.edu}}
- %%%%% \thanks{firstname.lastname}
-%}
+ \author{Nir Lipovetzky \\
+        { The University of Melbourne }\\
+     { Melbourne, Australia}\\
+     {\footnotesize\url{nir.lipovetzky@unimelb.edu.au}}
+%% footnote doesn't look nice ; this looks better
+% although can't use Nir's unimelb address that is too long :-)
+% 
+%    {\tiny\url{nirlipo@gmail.com}}
+    \And
+    Miquel Ramirez \\
+    { Australian National University} \\
+    { Canberra, Australia}\\
+   {\footnotesize\url{miquel.ramirez@anu.edu.au}}
+    \And
+    Hector Geffner \\
+    { ICREA \& U. Pompeu Fabra }\\
+    { Barcelona, Spain }\\
+    {\footnotesize\url{hector.geffner@upf.edu}}
+ %%%% \thanks{firstname.lastname}
+}
 
 \maketitle
+\begin{abstract}
+The Atari 2600 games supported in the Arcade Learning
+Environment~\cite{bellemare:jair2013} all feature a known initial
+(RAM) state and actions that have deterministic effects. Classical
+planners, however, cannot be used off-the-shelf as there is no compact
+PDDL-model of the games, and action effects and goals are not known
+\emph{a priori}. Indeed, there are no explicit goals, and the planner
+must select actions on-line while interacting with a simulator that
+returns successor states and rewards. None of this precludes the use
+of blind lookahead algorithms for action selection like breadth-first
+search or Dijkstra's yet such methods are not effective over large
+state spaces.  We thus turn to a different class of
+planning methods introduced recently that have been shown to be
+effective for solving large planning problems but which do not require
+prior knowledge of state transitions, costs (rewards) or goals. The
+empirical results over \numgames\ Atari games show that the simplest
+such algorithm performs at the level of UCT, the state-of-the-art
+planning method in this domain, and suggest the potential of
+width-based methods for planning with simulators when factored,
+compact action models are not available.
 
-\begin{abstract}
-The Atari 2600 games supported in  the Arcade Learning Environment~\cite{bellemare:jair2013}
-all feature a known initial (RAM) state and actions that have deterministic effects. Classical planners, however,
-cannot be used  off-the-shelf as there is no  compact PDDL-model of  the games, and action effects and goals are  not
-known \emph{a priori}. Indeed, there are no explicit goals, and the planner must select actions on-line while interacting 
-with a simulator that returns successor states and rewards. None of this precludes  the use of  blind lookahead algorithms 
-for action selection like breadth-first search or  Dijkstra's yet such methods are not effective over large  state spaces. 
-We thus turn to a different class of  planning methods  introduced recently  that  have been shown to be  effective for solving
-large planning problems but which do not require  prior knowledge of state transitions and costs (rewards) or goals.
-The simplest such algorithm, called Iterated Width or \IW, consists of  a sequence of calls  $\IW(1)$, $\IW(2)$, \ldots,
-where  $\IW(k)$ is  a breadth-first search in which a state is  pruned when it is not the first state in the  search  to make true 
-a subset of $k$ atoms. The  empirical results over \numgames\ Atari games show that  $\IW(1)$ performs at the level of UCT, 
-the state-of-the-art planning method in this domain, and suggest the potential of width-based methods for planning with
-simulators when factored, compact action models are not available.
 \end{abstract}
 
+\noindent \emph{A version of this short paper has been accepted at IJCAI \cite{lipo:ijcai15}}
 \section{Introduction}
@@
 the state of the screen pixels, while the planning algorithms,  use the state
-of the RAM memory. It is not clear however whether one input representation is
-more challenging than the other; for the learning algorithms, BNVB mention
-that the results tend to be better for the screen inputs.
+of the RAM memory. It is not clear however whether the use of one input representation is
+more challenging than the use of the other. For the learning algorithms, BNVB mention
+that the results tend to be better for the screen inputs. Experiments were
+run on a cluster, where each computing node consists of a $6$-core Intel Xeon 
+E5-2440, with $2.4$ GHz clock speed, with 64 GBytes of RAM installed.
 
@@
 
-Last, in the lookahead, \IW(1) and \BFS\ select an action every $5$ frames, while \UCT\  selects 
-an action in every frame.  This means that in order to explore a branch 
-$300$ frames deep, UCT gets to choose $300$ actions, while \IW($1$) and \BFS\ 
-get to choose $60$ actions, both however using the same $300$ frames from the
-budget. For this, we  followed the setup  of \BRFS\ in BNVB 
-that also selects actions every $5$ frames, matching the behavior
-of the emulator that requests an action also every $5$ frames. Since the 
-lookahead budget is given by a maximum number of (new) frames, and the time 
-is mostly taken by calls to the emulator, this may not be the best choice  
-for $\IW(1)$ and \BFS\ that may therefore not be exploiting all the 
-options afforded by the budget. We'll look at this further in the future.
+Last, in the lookahead, \IW(1) and \BFS\ select an action every $5$
+frames, while \UCT\ selects an action every frame.  This means that in
+order to explore a branch $300$ frames deep, UCT gets to choose $300$
+actions, while \IW($1$) and \BFS\ get to choose $60$ actions, both
+however using the same $300$ frames from the budget. For this, we
+followed the setup of \BRFS\ in BNVB that also selects actions every
+$5$ frames, matching the behavior of the emulator that requests an
+action also every $5$ frames. Since the lookahead budget is given by a
+maximum number of (new) frames, and the time is mostly taken by calls
+to the emulator, this may not be the best choice for $\IW(1)$ and
+\BFS\ that may therefore not be exploiting all the options afforded by
+the budget. Interestingly, when \UCT\ is limited to one action every
+$5$ frames, its performance is reduced by up to a $50$\% in games
+where it performs very well (\CrazyClimber), and does not appear to
+improve in those games where it performs very poorly (\Freeway).
 
@@
 
-\noindent \textbf{Acknowledgments.} We thank the people involved in
-the creation and maintenance of the Arcade Learning Environment that
-have made this work possible. The research by Nir Lipovetzky is partially
-funded by the Australian Research Council Linkage grant LP11010015.
-}
+\noindent }
+
+\section*{Acknowledgments} 
+We thank the people who created the ALE.  Nir Lipovetzky is partially
+funded by the ARC Linkage grant LP11010015.
 
